### 2017-10-20 papers/main.tex

\documentclass{llncs}

\usepackage{makeidx}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{indentfirst}
\usepackage{verbatim}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{url}
\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{comment}

\sloppy

\def\transarrow{\xrightarrow}
\newcommand{\setarrow}[1]{\def\transarrow{#1}}

\newcommand{\trule}[2]{\frac{#1}{#2}}
\newcommand{\crule}[3]{\frac{#1}{#2},\;{#3}}
\newcommand{\withenv}[2]{{#1}\vdash{#2}}
\newcommand{\trans}[3]{{#1}\transarrow{#2}{#3}}
\newcommand{\ctrans}[4]{{#1}\transarrow{#2}{#3},\;{#4}}
\newcommand{\llang}[1]{\mbox{\lstinline[mathescape]|#1|}}
\newcommand{\pair}[2]{\inbr{{#1}\mid{#2}}}
\newcommand{\inbr}[1]{\left<{#1}\right>}
\newcommand{\highlight}[1]{\color{red}{#1}}
\newcommand{\ruleno}[1]{\eqno[\scriptsize\textsc{#1}]}
\newcommand{\inmath}[1]{\mbox{$#1$}}
\newcommand{\lfp}[1]{fix_{#1}}
\newcommand{\gfp}[1]{Fix_{#1}}
\newcommand{\vsep}{\vspace{-2mm}}
\newcommand{\supp}[1]{\scriptsize{#1}}
\newcommand{\G}{\mathfrak G}
\newcommand{\sembr}[1]{\llbracket{#1}\rrbracket}
\newcommand{\cd}[1]{\texttt{#1}}
\newcommand{\miniKanren}{\texttt{miniKanren}\xspace}
\newcommand{\ocanren}{\texttt{ocanren}\xspace}

\lstdefinelanguage{ocanren}{
keywords={fresh, let, in, match, with, when, class, type,
object, method, of, rec, repeat, until, while, not, do, done, as, val, inherit,
new, module, sig, deriving, datatype, struct, if, then, else, open, private, virtual, include, success, failure},
sensitive=true,
commentstyle=\small\itshape\ttfamily,
keywordstyle=\ttfamily\underbar,
identifierstyle=\ttfamily,
basewidth={0.5em,0.5em},
columns=fixed,
fontadjust=true,
literate={fun}{{$\lambda$}}1 {->}{{$\to$}}3 {===}{{$\equiv$}}1 {=/=}{{$\not\equiv$}}1 {|>}{{$\triangleright$}}3 {/\\}{{$\wedge$}}2 {\\/}{{$\vee$}}2 {^}{{$\uparrow$}}1,
morecomment=[s]{(*}{*)}
}

\lstset{
mathescape=true,
basicstyle=\small,
identifierstyle=\ttfamily,
keywordstyle=\bfseries,
commentstyle=\scriptsize\rmfamily,
basewidth={0.5em,0.5em},
fontadjust=true,
language=ocanren
}

\usepackage{letltxmacro}
\newcommand*{\SavedLstInline}{}
\LetLtxMacro\SavedLstInline\lstinline
\DeclareRobustCommand*{\lstinline}{%
  \ifmmode
    \let\SavedBGroup\bgroup
    \def\bgroup{%
      \let\bgroup\SavedBGroup
      \hbox\bgroup
    }%
  \fi
  \SavedLstInline
}

\begin{document}

\mainmatter

\title{MiniKanren Semantics for Programs Transformation}

\maketitle

\section{Syntax}

\begin{figure}
\centering 
{\bf Supplementary syntax categories:}
$$
\begin{array}{rcll}
  \mathcal{X} &=& x_1,\, x_2,\dots & \mbox{\supp(syntactic variables)} \\
  \mathcal{S} &=& \sigma_1,\, \sigma_2,\dots & \mbox{\supp(semantic variables)} \\
  \mathcal{R}^n &=& r^n_1, r^n_2, \dots &\mbox{\supp{(predicate names indexed by the arity n)}}
\end{array}
$$

{\bf Terms:}
$$
\begin{array}{rcll}
  \mathcal{T(V)} &=& \mathcal{V} \,|\, C^n \, t_1,\dots,t_n & \mbox{\supp(terms parameterized by variable type)} \\
  \mathcal{T_X} &=& \mathcal{T(X)} & \mbox{\supp(terms which contain only syntactic variables)} \\
  \mathcal{T_S} &=& \mathcal{T(S)} & \mbox{\supp(terms which contain only syntactic variables)} 
\end{array}
$$

{\bf Goals:}
$$
\begin{array}{rcll}
  \mathcal{G} &=& t_1 \equiv t_2 & \mbox{\supp(unification)} \\
              &|& g_1 \wedge g_2 & \mbox{\supp(conjunction)} \\
              &|& g_1 \vee   g_2 & \mbox{\supp(disjunction)} \\
              &|&\lstinline|fresh $x$ $\;g$| & \mbox{\supp{(fresh logical variable binder)}}\\
              &|& \mathcal{R}^{n} \, t_1 ,\dots, t_n & \mbox{\supp{(relations)}} \\
\end{array}
$$

{\bf Definitions:}
$$
\begin{array}{rcll}
  \mathcal{D}^n &=& \lambda x_1, \dots, x_n. g & \mbox{\supp(relation definition)}
\end{array}
$$

{\bf States:}
$$
\begin{array}{rcll}
  I &=& \mathcal{X} \rightarrow \mathcal{T_S} & \mbox{\supp(syntax variables interpretation)} \\
  \Sigma &=& \mathcal{S} \rightarrow \mathcal{T_S} & \mbox{\supp(substitutions)} \\
  \Delta &=& 2^\mathcal{S} & \mbox{\supp(the set of used semantic variables)} \\
  \mathfrak{S} &=& I \times \Sigma \times \Delta & \mbox{\supp(states)} \\
  \Gamma &=& \mathcal{R}^n \rightarrow \mathcal{D}^n & \\
  \Gamma &\vdash& \mathfrak{s} \xrightarrow{g} \mathfrak{\bar{s}} & \mbox{\supp(semantics of the goal g)}
\end{array} 
$$

\end{figure}

\newpage
\section{Semantics}

\begin{figure}
\centering 
$$
\withenv{\Gamma}{ \ctrans{(\iota,\sigma,\delta)}{t_1 \equiv t_2}{[]}{mgu \, (\iota \, t_1) \sigma \,\, (\iota \, t_2) \sigma = \bot} } 
$$

$$
\withenv{\Gamma}{ \trans{(\iota,\sigma,\delta)}{t_1 \equiv t_2}{(\iota, mgu \, (\iota \, t_1) \sigma \; (\iota \, t_2) \sigma , \delta)}}	
$$

$$
\crule{\withenv{\Gamma}{\trans{(\iota', \sigma, \delta')}{g}{\mathfrak{\bar{s}}}}}
      {\withenv{\Gamma}{\trans{(\iota, \sigma, \delta)}{\lstinline|fresh $\;x$ $\;g$|}{\mathfrak{\bar{s}}}}}
      {\iota' = \iota[x \leftarrow s], \delta' = \delta \cup \{x\}, x \notin \delta}
$$

$$
\crule{\withenv{\Gamma}{\trans{(\iota',\sigma,\delta)}{g}{\mathfrak{\bar{s}}}}}
      {\withenv{\Gamma}{\trans{(\iota,\sigma,\delta)}{r^n \, t_1 \dots t_n}{\mathfrak{\bar{s}}}}}
      {\iota' = \iota[x_i \leftarrow t_i \iota], \Gamma \, r^n = \lambda x_1 \dots x_n. g}
$$

$$
\begin{array}{rcll}
  \square \iota &:& \mathcal{T_X} \rightarrow \mathcal{T_S} & \\
  x \iota &=& \iota (x) & \\
  \mathcal{C}^n \, t_1 \dots t_n &=& \mathcal{C}^n \, (t_1 \iota) \dots (t_n \iota)
\end{array} 
$$

$$
\trule{\withenv{\Gamma}{\trans{\mathfrak{s}}{g_1}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{s}}{g_2}{\mathfrak{\bar{s_2}}}}}
      {\withenv{\Gamma}{\trans{\mathfrak{s}}{g_1 \vee g_2}{\mathfrak{\bar{s_1}} \oplus \mathfrak{\bar{s_2}}}}}
$$

$$
\trule{\withenv{\Gamma}{\trans{(\iota,\sigma,\delta)}{g_1}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{\bar{s_1}}}{g_2 \iota}{\mathfrak{\bar{s_2}}}}}
      {\withenv{\Gamma}{\trans{(\iota,\sigma,\delta)}{g_1 \wedge g_2}{\mathfrak{\bar{s_2}}}}}
$$

$$
\trule{\withenv{\Gamma}{\trans{\mathfrak{s}}{g}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{\bar{s}}}{g}{\mathfrak{\bar{s_2}}}}}
      {\withenv{\Gamma}{\trans{\mathfrak{s}:\mathfrak{\bar{s}}}{g}{\mathfrak{\bar{s_1}} \oplus \mathfrak{\bar{s_2}}}}}
$$

\end{figure}

\end{document}

### 2017-10-20 src/Driving.hs
Split

@@ -103,6 +103,21 @@ substitute s (g1 :/\: g2 ) = substitute s g1 :/\: substitute s g2
 substitute s (g1 :\/: g2 ) = substitute s g1 :\/: substitute s g2
 substitute s (Invoke f as) = Invoke f $ map (E.substitute s) as
 
+weakCouple :: (G S, G S) -> Bool
+weakCouple gs =
+  case gs of
+    (_ :=:  _  , _ :=:  _  ) -> True
+    (_ :/\: _  , _ :/\: _  ) -> True
+    (_ :\/: _  , _ :\/: _  ) -> True
+    (Invoke _ _, Invoke _ _) -> True
+    _                        -> False
+
+split :: [G S] -> [G S] -> ([G S], [G S])
+split gs1 gs2 = 
+  case elemIndex False $ map weakCouple $ zip gs1 gs2 of
+    Nothing -> splitAt (length gs2) gs1
+    Just i  -> splitAt (i-1)        gs1
+
 invoke :: Stack -> E.Gamma -> E.Sigma -> G S -> [G S] -> Tree 
 invoke cs (p, i, d) s (Invoke f as') conjs = 
   let (_, fs, g) = p f in
@@ -115,7 +130,12 @@ invoke cs (p, i, d) s (Invoke f as') conjs =
           if length conjs == length conjs' 
           then let (msg, s1, s2, d') = generalizeGoals d (Invoke f as' : conjs) (Invoke g bs : conjs') in
                (Gen s1 (invoke ((f, as', conjs):cs) (p, i, d') s msg conjs))
-          else undefined {- !!! split -}
+          else if length conjs' < length conjs 
+               then let cs'           = (f, as', conjs):cs in
+                    let (left, right) = split (Invoke f as' : conjs) (Invoke g bs : conjs') in
+                    Split (eval cs' (p, i, d) s left)
+                          (eval cs' (p, i, d) s right)
+               else error "Wow..."
         Nothing -> let (g', env') = E.pre_eval (p, i, d) g in
                    eval ((f, as', map (substitute s) conjs) : cs) env' s (g':conjs)  
 

### 2017-10-20 src/PrintTree.hs
Merging

@@ -43,7 +43,6 @@ import Data.GraphViz.Attributes.Complete (
   zVal,
   focus
   )
-import System.Mem.StableName
 import Syntax
 import qualified Eval as E
 
@@ -87,15 +86,7 @@ label tree =
     label' t@(PrintTree.Fresh _ _ ch)   i ns es = addChild    i (show t) ns es ch
     label' t@(Gen _ ch)                 i ns es = addChild    i (show t) ns es ch
     label' t@(Or ch1 ch2)               i ns es = addChildren i (show t) ns es ch1 ch2
-{-
-data Tree  = 
-  Fail                         | 
-  Success E.Sigma              | 
-  Or      Tree Tree            | 
-  Fresh   X S Tree             | 
-  Rename  String [Ts] Renaming |
-  Gen     Generalizer Tree -- deriving Show
--}
+
 instance Show (Tree) where 
   show Fail = "_|_"
   show (Success s) = ("S\n" ++ show s)

### 2017-10-20 src/Syntax.hs

@@ -1,3 +1,5 @@
+{-# LANGUAGE TypeSynonymInstances #-}
+{-# LANGUAGE FlexibleInstances #-}
 
 
 module Syntax where
@@ -11,8 +13,8 @@ data Term v = V v | C String [Term v] deriving (Eq, Ord) --, Show)
 type Tx     = Term X
 type Ts     = Term S
 
---instance Show S where 
---  show i = "_." ++ show i
+instance {-# OVERLAPPING #-} Show S where 
+  show i = "_." ++ show i
 
 instance Show a => Show (Term a) where
   show (V v) = show v

### 2018-03-14 papers/main.tex
MiniKanren CPD draft

@@
 
-\title{MiniKanren Semantics for Programs Transformation}
+\title{Program Transformations for miniKanren}
 
 \maketitle
+
+\section{Conjunctive Partial Deduction for MiniKanren}
+
+Conjunctive partial deduction is done in two steps: driving and residualization. 
+The driving step constructs a directed graph which resembles a partial SLDNF tree for Prolog programs. Residual program is constructed from this graph in the straightest way possible.
+
+The graph is almost a tree in the sense that cycles are only possible in one case: the goal in the current node is a renaming of a goal encountered before.
+The graph is refered to as the driving tree henceforth.
+
+The driving tree is constructed in the process of driving a goal. Each node contains a current goal and a current substition and is related closely to the evaluation step which produced it.
+
+There are several types of nodes in the driving tree: 
+\begin{enumerate}
+  \item \emph{Fail} is a leaf node. It corresponds to the refutation of some goal.
+  \item \emph{Success} is a leaf node. It contains a computed answer for the goal
+  \item \emph{Or} node means the current goal is a disjuction: $g_1 \vee   g_2$. The children nodes are the results of driving of goals $g_1$ and $g_2$. 
+  \item \emph{Call} node correspond to the conjunction of relation calls: $g_1 \wedge g_2 \wedge \dots \wedge g_n$, where $g_i = R^k_i t_1 \dots t_k$.
+It has a single child node which is produced by an unfolding of every call in the conjunction: $g_1' \wedge g_2' \wedge \dots \wedge g_n'$, where $g_i' = unfold \ g_i$.
+  \item \emph{Rename} node is a leaf node which means there is a node among the ancestors which is a renaming of the current goal. The index of the ancestor node is stored in the current node. 
+  \item \emph{Gen} node corresponds to a result of generalization.
+If there is an ancestor node with a goal which is embedded in the current goal, and the goals contain the same number of conjuncts, an msg for them is computed and stored in the \emph{Gen} node.
+  \item \emph{Split} node means that the ancestor embedded node contains less conjuncts that the current one. In this case, the conjunction is splitted into two which are driven independently. 
+\end{enumerate}
+
+First, a miniKanren goal is partially evaluated to a conjunction of relation calls.
+
+\begin{enumerate}
+  \item If there is a disjunction, the disjuncts are driven separatly. By distribution, we can get rid of disjunctions in any goal. 
+  \item All fresh variables are lifted to a higher level and are made into semantic variables with a pre-evaluation step.
+  \item Every unification is evaluated into a substitution which goes along with the goal while driving. A fail to unify means there is a refutation and a \emph{Fail} node is created. 
+  \item If after these steps a goal is an empty list of conjunctions, a \emph{Success} node is created with a substitution computed. 
+\end{enumerate}
+
+When the current goal is the conjunction of relation calls, one of the following cases takes place: 
+\begin{enumerate}
+  \item If the current goal is a renaming of some goal encountered before, then a \emph{Renaming} leaf node is created since there is no need in driving the goal further. 
+  \item If the current goal is not a renaming of some ancestor goal, check it for embedding of some ancestor goal. For an embedding, one of two options is possible:
+  \begin{enumerate}
+    \item The number of conjuncts is the same, which means the ``larger'' conjunction is larger in the subterms.
+In this case, an msg is computed for the current and the ancestor conjunction and then the msg driven further.
+    \item The current conjunction has more conjuncts than the ancestor conjunction.
+In this case, the conjunction is split into two using the \emph{split} funciton.
+Let $g_1 \wedge \dots \wedge g_n \trianglelefteq h_1 \wedge \dots \wedge h_k, n < k$.
+In this case, the conjunction $h_1 \wedge \dots \wedge h_k$ is split into two subconjunctions $h_1 \wedge \dots \wedge h_j$ and $h_{j+1} \wedge \dots \wedge h_k$, such that $g_i \trianglelefteq h_i$ for each $i \in [1..j]$ and $g_{j+1} \ntrianglelefteq h_{j+1}$.
+The two subconjunctions are then driven independently.
+  \end{enumerate}     
+  \item If the current goal is neither renaming nor embedding of any ancestor, it is driven further. 
+\end{enumerate}
+
+Embedding we use currently is the classic one which ignores the variables by making any variable embed into any other variable: $X \trianglelefteq Y$ for any variables $X, Y$.
+The embedding on the lists of conjuncts demands the heads of the conjuncts to be calls of the same relation.
+
+$$
+\crule{}
+      {X \trianglelefteq Y}
+      {variable \ embedding}
+$$
+
+$$
+\crule{\exists i \in [1..n]. X \trianglelefteq t_i}
+      {X \trianglelefteq C^n t_1 \dots t_n}
+      {term \ diving}
+$$
+
+$$
+\crule{\forall i \in [1..n]. \ t_i \trianglelefteq u_i}
+      {C^n t_1 \dots t_n \trianglelefteq C^n u_1 \dots u_n}
+      {term \ coupling}
+$$
+
+$$
+\crule{\forall i \in [1..n]. \ t_i \trianglelefteq u_i}
+      {R^n t_1 \dots t_n \trianglelefteq R^n u_1 \dots u_n}
+      {relation \ call}
+$$
+
+$$
+\crule{R_g^n t_1 \dots t_n \trianglelefteq R_h^n u_1 \dots u_n, g_2 \wedge \dots \wedge g_n \trianglelefteq h_2 \wedge \dots \wedge h_n}
+      {R_g^n t_1 \dots t_n \wedge g_2 \wedge \dots \wedge g_n \trianglelefteq R_h^n u_1 \dots u_n \wedge \dots \wedge h_n}
+      {conjunction \ coupling}
+$$
+
+$$
+\crule{g_1 \wedge \dots \wedge g_n \trianglelefteq h_2 \wedge \dots \wedge h_n}
+      {g_1 \wedge \dots \wedge g_n \trianglelefteq h_1 \wedge h_2 \wedge \dots \wedge h_n}
+      {conjunction \ diving}
+$$
+
+You proposed the different embedding before, the one named ``New homeomorphic embedding'', which keeps track of the variables. It is different in that a variable is only embedded into itself:
+
+$$
+\crule{}
+      {X \trianglelefteq X}
+      {variable \ embedding}
+$$
+
+And the top-level embedding checks, that there is a substitution which makes the goal to be embedded into the other goal by the classic homeomorphic embedding:
+
+$$
+\crule{\exists \sigma. A \sigma \trianglelefteq B}
+      {A {\trianglelefteq}^{subst} B}
+      {toplevel}
+$$
+
+We found a program which fails to terminate with the new embedding. It is \verb!smallesto! --- a relation between a list, its smallest element and the rest of the elements.
+
+In prolog syntax it is defined as the following:
+
+\begin{verbatim}
+leo(0,Y,true).
+leo(succ(X),0,false).
+leo(succ(X),succ(Y),Z) <- leo(X,Y,Z).
+
+gto(0,Y,false).
+gto(succ(X),0,true).
+gto(succ(X),succ(Y),Z) <- gto(X,Y,Z).
+
+minmaxo(X,Y,X,Y) <- leo(X,Y,true).
+minmaxo(X,Y,Y,X) <- gto(X,Y,true).
+
+smallesto([S],S,[]).
+smallesto([H|T],S,[M|Tt]) <- minmaxo(H,Ss,S,M), smallesto(T,Ss,Tt).
+\end{verbatim}
+
+Several branches in the driving tree of \verb!smallesto x y z! with \verb!x!, \verb!y! and 
+\verb!z! fresh have the following pattern:
+
+\begin{verbatim}
+0) smallesto(V0, V1, V2)
+1) minmaxo(V3, V5, V1, V7) /\ smallesto(V4, V5, V6)
+2) leo(V3, V5, true) /\ minmaxo(V8, V10, V5, V12]) /\ smallesto(V9, V10, V11)
+3) gto(V8, V10, true) /\ minmaxo(V19, V21, V10, V23) /\ smallesto(V20, V21, V22)
+4) gto(V31,V32, true) /\ leo(succ(V32), V21, true) /\ 
+       minmaxo(V33, V35, V21, V37])) /\ smallesto(V34, V35, V36])
+5) ... 
+\end{verbatim}
+
+In this program, goals grow to the left while driving.
+Our embedding demands conjunctions to be coupled, which means that the first conjuncts should be calls of the same relation.
+Because of that, only the goal 4) can possibly be embedded into the goal 5).
+But, since the original embedding relation does not ignore variables, they are not.
+See, the thrid argument of \verb!minmaxo!
+in the goal 4) is \verb!V10!
+--- the same variable as the second in \verb!gto!.
+Unfortunatelly, the goal 5) the second variable of \verb!gto!
+is \verb!V32!
+while the third one of \verb!minmaxo!
+is \verb!V21!.
+In the goal 5) \verb!V21!
+is also present in the \verb!leo!
+relation --- and this link between the variables make it impossible for the driving to terminate on this branch.
+
+So, it seems that proposed new homeomorphic embedding does not ensure termination. Why is it even a wqo?
+
+Another problem is that $split$ only splits a goal in two subgoals, and if each of the relation calls involved unfold in a conjunciton of at least two relation calls, goals are going to grow faster, than $split$ splits.
+
+If we have the following program:
+
+\begin{verbatim}
+a(X) <- b(X), c(X)
+b(X) <- a(X), b(X), c(X)
+c(X) <- a(X), b(X), c(X)
+\end{verbatim}
+
+Driving of \verb!a(X)! goes like this:
+
+\begin{verbatim}
+1)  a(X)
+2)  b(X) /\ c(X) -- nothing is embedded into this goal
+3)  a(X) /\ b(X) /\ c(X) /\ a(X) /\ b(X) /\ c(X) -- 1) is embedded into 2); splitting
+4a) a(X) -- first subconjuction after split, renaming of 1)
+4b) b(X) /\ c(X) /\ a(X) /\ b(X) /\ c(X) -- second subconjunction after split
+5b) a(X) /\ b(X) /\ c(X) /\ a(X) /\ b(X) /\ c(X) /\ b(X) /\ c(X) /\ 
+         a(X) /\ b(X) /\ c(X) /\ a(X) /\ b(X) /\ c(X) 
+...
+\end{verbatim}
+
+This can probably be fixed by some restriction on unfolding of relation calls, but I do not see a good way to do it.
+Before, we tried to unconditionally unfold only the first relation in the conjunction, and the following conjunctions are unfolded only if the unfolding is deterministic.
+This approach seems viable, but still quite heuristic, especially if we do not fix a left-to-right execution strategy.
+
+The last problem we encountered is an enourmous size of the residualized programs.
+The biggest problem there is that we introduce dozens of definitions of the same relation.
+There can be a simple relation (as \verb!gto!
+or \verb!leo!) which gets to a renaming after one unfolding, but it occurs in different branches of the driving tree, it gets residualized into the same program each time it occurs.
+For a relational sorting program, the size of generated program is about 10000 lines.
+
+\newpage
 

### 2018-03-14 papers/main.tex
Typos

@@
 
-Conjunctive partial deduction is done in two steps: driving and residualization. 
+The conjunctive partial deduction is done in two steps: driving and residualization. 
 The driving step constructs a directed graph which resembles a partial SLDNF tree for Prolog programs. Residual program is constructed from this graph in the straightest way possible.
@@
 The graph is almost a tree in the sense that cycles are only possible in one case: the goal in the current node is a renaming of a goal encountered before.
-The graph is refered to as the driving tree henceforth.
+The graph is referred to as the driving tree henceforth.
 
-The driving tree is constructed in the process of driving a goal. Each node contains a current goal and a current substition and is related closely to the evaluation step which produced it.
+The driving tree is constructed in the process of driving a goal. Each node contains a current goal and a current substitution and is related closely to the evaluation step which produced it.
 
@@
   \item \emph{Success} is a leaf node. It contains a computed answer for the goal
-  \item \emph{Or} node means the current goal is a disjuction: $g_1 \vee   g_2$. The children nodes are the results of driving of goals $g_1$ and $g_2$. 
-  \item \emph{Call} node correspond to the conjunction of relation calls: $g_1 \wedge g_2 \wedge \dots \wedge g_n$, where $g_i = R^k_i t_1 \dots t_k$.
+  \item \emph{Or} node means the current goal is a disjunction: $g_1 \vee   g_2$. The children nodes are the results of driving of goals $g_1$ and $g_2$. 
+  \item \emph{Call} node corresponds to the conjunction of relation calls: $g_1 \wedge g_2 \wedge \dots \wedge g_n$, where $g_i = R^k_i t_1 \dots t_k$.
 It has a single child node which is produced by an unfolding of every call in the conjunction: $g_1' \wedge g_2' \wedge \dots \wedge g_n'$, where $g_i' = unfold \ g_i$.
@@
 If there is an ancestor node with a goal which is embedded in the current goal, and the goals contain the same number of conjuncts, an msg for them is computed and stored in the \emph{Gen} node.
-  \item \emph{Split} node means that the ancestor embedded node contains less conjuncts that the current one. In this case, the conjunction is splitted into two which are driven independently. 
+  \item \emph{Split} node means that the ancestor embedded node contains fewer conjuncts than the current one. In this case, the conjunction is split into two which are driven independently. 
 \end{enumerate}
@@
 \begin{enumerate}
-  \item If there is a disjunction, the disjuncts are driven separatly. By distribution, we can get rid of disjunctions in any goal. 
+  \item If there is a disjunction, the disjuncts are driven separately. By distribution, we can get rid of disjunctions in any goal. 
   \item All fresh variables are lifted to a higher level and are made into semantic variables with a pre-evaluation step.
@@
 \begin{enumerate}
-  \item If the current goal is a renaming of some goal encountered before, then a \emph{Renaming} leaf node is created since there is no need in driving the goal further. 
+  \item If the current goal is a renaming of some goal encountered before, then a \emph{Renaming} leaf node is created since there is no need for driving the goal further. 
   \item If the current goal is not a renaming of some ancestor goal, check it for embedding of some ancestor goal. For an embedding, one of two options is possible:
@@
     \item The number of conjuncts is the same, which means the ``larger'' conjunction is larger in the subterms.
-In this case, an msg is computed for the current and the ancestor conjunction and then the msg driven further.
+In this case, an msg is computed for the current and the ancestor conjunction and then the msg is driven further.
     \item The current conjunction has more conjuncts than the ancestor conjunction.
-In this case, the conjunction is split into two using the \emph{split} funciton.
+In this case, the conjunction is split into two using the \emph{split} function.
 Let $g_1 \wedge \dots \wedge g_n \trianglelefteq h_1 \wedge \dots \wedge h_k, n < k$.
@@
 But, since the original embedding relation does not ignore variables, they are not.
-See, the thrid argument of \verb!minmaxo!
+See, the third argument of \verb!minmaxo!
 in the goal 4) is \verb!V10!
 --- the same variable as the second in \verb!gto!.
-Unfortunatelly, the goal 5) the second variable of \verb!gto!
+Unfortunately, the goal 5) the second variable of \verb!gto!
 is \verb!V32!
@@
 
-Another problem is that $split$ only splits a goal in two subgoals, and if each of the relation calls involved unfold in a conjunciton of at least two relation calls, goals are going to grow faster, than $split$ splits.
+Another problem is that $split$ only splits a goal into two subgoals, and if each of the relation calls involved unfold in a conjunction of at least two relation calls, goals are going to grow faster, than $split$ splits.
 
@@
 
-The last problem we encountered is an enourmous size of the residualized programs.
-The biggest problem there is that we introduce dozens of definitions of the same relation.
+The last problem we encountered is an enormous size of the residualized programs.
+The biggest problem is that we introduce dozens of definitions of the same relation.
 There can be a simple relation (as \verb!gto!
 or \verb!leo!) which gets to a renaming after one unfolding, but it occurs in different branches of the driving tree, it gets residualized into the same program each time it occurs.
-For a relational sorting program, the size of generated program is about 10000 lines.
+For a relational sorting program, the size of the generated program is about 10000 lines.
 
@@
 $$
-\trule{\withenv{\Gamma}{\trans{\mathfrak{s}}{g_1}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{s}}{g_2}{\mathfrak{\bar{s_2}}}}}
+\trule{\withenv{\Gamma}{\trans{\mathfrak{s}}{g_1}{\mathfrak{\bar{s_1}}}} \;\;\;\; \withenv{\Gamma}{\trans{\mathfrak{s}}{g_2}{\mathfrak{•}\bar{s_2}}}}
       {\withenv{\Gamma}{\trans{\mathfrak{s}}{g_1 \vee g_2}{\mathfrak{\bar{s_1}} \oplus \mathfrak{\bar{s_2}}}}}

### 2018-03-14 papers/main.tex
Text

@@
 
-The conjunctive partial deduction is done in two steps: driving and residualization. 
-The driving step constructs a directed graph which resembles a partial SLDNF tree for Prolog programs. Residual program is constructed from this graph in the straightest way possible.
+The conjunctive partial deduction is performed in two steps: driving and residualization. 
+The driving step constructs a directed graph, which resembles a partial SLDNF tree for Prolog programs. Residual program is constructed from this graph in the straightest way possible.
 
-The graph is almost a tree in the sense that cycles are only possible in one case: the goal in the current node is a renaming of a goal encountered before.
+The graph is almost a tree in the sense that cycles are only possible in one case: the goal in the current node is a renaming of a goal, encountered before.
 The graph is referred to as the driving tree henceforth.
 
-The driving tree is constructed in the process of driving a goal. Each node contains a current goal and a current substitution and is related closely to the evaluation step which produced it.
+The driving tree is constructed in the process of driving a goal. Each node contains a current goal and a current substitution, and is closely related to the evaluation step, which produced it.
 
 There are several types of nodes in the driving tree:
+
 \begin{enumerate}
   \item \emph{Fail} is a leaf node. It corresponds to the refutation of some goal.
-  \item \emph{Success} is a leaf node. It contains a computed answer for the goal
-  \item \emph{Or} node means the current goal is a disjunction: $g_1 \vee   g_2$. The children nodes are the results of driving of goals $g_1$ and $g_2$. 
-  \item \emph{Call} node corresponds to the conjunction of relation calls: $g_1 \wedge g_2 \wedge \dots \wedge g_n$, where $g_i = R^k_i t_1 \dots t_k$.
-It has a single child node which is produced by an unfolding of every call in the conjunction: $g_1' \wedge g_2' \wedge \dots \wedge g_n'$, where $g_i' = unfold \ g_i$.
+  \item \emph{Success} is a leaf node. It contains a computed answer for the goal.
+  \item \emph{Or} node means, that the current goal is a disjunction $g_1 \vee   g_2$. The children nodes are the results of driving of goals $g_1$ and $g_2$. 
+  \item \emph{Call} node corresponds to the conjunction of relation calls $g_1 \wedge g_2 \wedge \dots \wedge g_n$, where $g_i = R^k_i t_1 \dots t_k$.
+It has a single child node, which is produced by unfolding of every call in the conjunction $g_1' \wedge g_2' \wedge \dots \wedge g_n'$, where $g_i' = unfold \ g_i$.
   \item \emph{Rename} node is a leaf node which means there is a node among the ancestors which is a renaming of the current goal. The index of the ancestor node is stored in the current node. 
   \item \emph{Gen} node corresponds to a result of generalization.
-If there is an ancestor node with a goal which is embedded in the current goal, and the goals contain the same number of conjuncts, an msg for them is computed and stored in the \emph{Gen} node.
-  \item \emph{Split} node means that the ancestor embedded node contains fewer conjuncts than the current one. In this case, the conjunction is split into two which are driven independently. 
+If there is an ancestor node with a goal, which is embedded in the current goal, and the goals contain the same number of conjuncts, the msg for them is computed and stored in the \emph{Gen} node.
+  \item \emph{Split} node means, that the embedded node contains fewer conjuncts, than the current one.
+In this case, the conjunction is split into two subconjunctions, which are driven independently.
 \end{enumerate}
@@
 \begin{enumerate}
-  \item If there is a disjunction, the disjuncts are driven separately. By distribution, we can get rid of disjunctions in any goal. 
-  \item All fresh variables are lifted to a higher level and are made into semantic variables with a pre-evaluation step.
-  \item Every unification is evaluated into a substitution which goes along with the goal while driving. A fail to unify means there is a refutation and a \emph{Fail} node is created. 
-  \item If after these steps a goal is an empty list of conjunctions, a \emph{Success} node is created with a substitution computed. 
+  \item If there is a disjunction, the disjuncts are driven separately. Using the distribution law, we can get rid of disjunctions in any goal by generating \emph{Or}-nodes with no information loss. 
+  \item All ``fresh'' constructs are combined together and lifted to the topmost level in every definition; thus, a relational definition body contains no more, than one top-level fresh construct, which encloses a freshless body.
+This purely syntactic transformation makes the driving easier, since we do not need to interpret the ``fresh'' construct in an arbitrary position.
+  \item Every unification is evaluated into a substitution, which is threaded along with the goal while driving. A fail to unify means refutation and thus a \emph{Fail} node is created. 
+  \item If after these steps a goal is an empty list of conjunctions, a \emph{Success} node is created with the substitution computed. 
 \end{enumerate}
 
-When the current goal is the conjunction of relation calls, one of the following cases takes place: 
+When the current goal is the conjunction of relational calls, one of the following cases takes place:
+
 \begin{enumerate}
   \item If the current goal is a renaming of some goal encountered before, then a \emph{Renaming} leaf node is created since there is no need for driving the goal further. 
-  \item If the current goal is not a renaming of some ancestor goal, check it for embedding of some ancestor goal. For an embedding, one of two options is possible:
+  \item If the current goal is not a renaming of some ancestor goal, we check it for embedding of some ancestor goal. For an embedding, one of two options is possible:
   \begin{enumerate}
-    \item The number of conjuncts is the same, which means the ``larger'' conjunction is larger in the subterms.
-In this case, an msg is computed for the current and the ancestor conjunction and then the msg is driven further.
-    \item The current conjunction has more conjuncts than the ancestor conjunction.
-In this case, the conjunction is split into two using the \emph{split} function.
+    \item The number of conjuncts is the same, which means the subterms of the ``larger'' conjunction embed subterms of the ``smaller''. In this case, the msg is computed and is driven further. 
+    \item The current conjunction has more conjuncts, than the ancestor conjunction.
+In this case, the conjunction is split into two subconjunctions using the \emph{split} function.
 Let $g_1 \wedge \dots \wedge g_n \trianglelefteq h_1 \wedge \dots \wedge h_k, n < k$.
 In this case, the conjunction $h_1 \wedge \dots \wedge h_k$ is split into two subconjunctions $h_1 \wedge \dots \wedge h_j$ and $h_{j+1} \wedge \dots \wedge h_k$, such that $g_i \trianglelefteq h_i$ for each $i \in [1..j]$ and $g_{j+1} \ntrianglelefteq h_{j+1}$.
-The two subconjunctions are then driven independently.
+The two subconjunctions are then driven independently and the results are connected by a \emph{Split} node.
   \end{enumerate}     
@@
 
-Embedding we use currently is the classic one which ignores the variables by making any variable embed into any other variable: $X \trianglelefteq Y$ for any variables $X, Y$.
-The embedding on the lists of conjuncts demands the heads of the conjuncts to be calls of the same relation.
+Embedding we use currently is the classic one, which ignores the variables by making any variable embed into any other variable: $X \trianglelefteq Y$ for any variables $X, Y$.
+The embedding on the lists of conjuncts requires the heads of the conjuncts to be calls of the same relation.
 
@@
 
-You proposed the different embedding before, the one named ``New homeomorphic embedding'', which keeps track of the variables. It is different in that a variable is only embedded into itself:
+      You proposed a different embedding before, the one named ``new homeomorphic embedding'' ($\trianglelefteq^\wedge$), which discriminates the variables in such a way, that a variable is only embedded into itself.
+
+\begin{comment}
 
@@
 
-And the top-level embedding checks, that there is a substitution which makes the goal to be embedded into the other goal by the classic homeomorphic embedding:
+      The top-level embedding checks, that there is a substitution, which makes the goal to be embedded into the other goal by the classic homeomorphic embedding:
 
@@
 $$
+\end{comment}
 
-We found a program which fails to terminate with the new embedding. It is \verb!smallesto! --- a relation between a list, its smallest element and the rest of the elements.
+However, we have found a program, whose driving fails to terminate with the new embedding. It is \verb!smallesto!~--- a relation between a list, its smallest element, and the rest of the elements.
 
-In prolog syntax it is defined as the following:
+In Prolog syntax it can be defined as follows:
 
@@
 
-Several branches in the driving tree of \verb!smallesto x y z! with \verb!x!, \verb!y! and 
-\verb!z! fresh have the following pattern:
+Several branches in the driving tree of \verb!smallesto(x, y, z)! with \verb!x!, \verb!y! and 
+\verb!z! fresh variables have the following pattern:
 
@@
 
-In this program, goals grow to the left while driving.
-Our embedding demands conjunctions to be coupled, which means that the first conjuncts should be calls of the same relation.
-Because of that, only the goal 4) can possibly be embedded into the goal 5).
+In this program, the goals grow ``to the left'' during the driving.
+Our embedding requires conjunctions to be coupled, which means, that their first conjuncts should be the calls of the same relation.
+Because of that, only the goal 3) can possibly be embedded into the goal 4).
 But, since the original embedding relation does not ignore variables, they are not.
-See, the third argument of \verb!minmaxo!
-in the goal 4) is \verb!V10!
---- the same variable as the second in \verb!gto!.
-Unfortunately, the goal 5) the second variable of \verb!gto!
-is \verb!V32!
-while the third one of \verb!minmaxo!
+Indeed, the third argument of \verb!minmaxo!
+in the goal 3) is \verb!V10!~--- the same variable as the second in \verb!gto!.
+Unfortunately, in the goal 4) the second variable of \verb!gto!
+is \verb!V32!, while the third one of \verb!minmaxo!
 is \verb!V21!.
-In the goal 5) \verb!V21!
+In the goal 4) \verb!V21!
 is also present in the \verb!leo!
-relation --- and this link between the variables make it impossible for the driving to terminate on this branch.
+relation~--- and this link between the variables makes it impossible for the driving to terminate in this branch.
 
-So, it seems that proposed new homeomorphic embedding does not ensure termination. Why is it even a wqo?
+%So, it seems, that the proposed new homeomorphic embedding does not ensure termination. Why is it even a wqo?
 
-Another problem is that $split$ only splits a goal into two subgoals, and if each of the relation calls involved unfold in a conjunction of at least two relation calls, goals are going to grow faster, than $split$ splits.
+Another problem is that the $split$ only splits a goal into two subgoals, and if each of the relational calls involved unfolds in a conjunction of at least two relation calls, the goals are going to grow faster, than the $split$ splits.
 
@@
 
-This can probably be fixed by some restriction on unfolding of relation calls, but I do not see a good way to do it.
+This can probably be fixed by some restriction on unfolding of relational calls, but we do not see a good way to do it.
 Before, we tried to unconditionally unfold only the first relation in the conjunction, and the following conjunctions are unfolded only if the unfolding is deterministic.
@@
 There can be a simple relation (as \verb!gto!
-or \verb!leo!) which gets to a renaming after one unfolding, but it occurs in different branches of the driving tree, it gets residualized into the same program each time it occurs.
-For a relational sorting program, the size of the generated program is about 10000 lines.
+or \verb!leo!), which gets to a renaming after one unfolding, but it occurs in different branches of the driving tree, and it is residualized multiple times.
+For a relational sorting program, the size of the residual program is about 10000 lines.
+
+The rest of the text contains the formal description of the syntax and semantics of the language we work with.
 

### 2018-03-14 papers/main.tex
Typos

@@
   \item \emph{Or} node means, that the current goal is a disjunction $g_1 \vee   g_2$. The children nodes are the results of driving of goals $g_1$ and $g_2$. 
-  \item \emph{Call} node corresponds to the conjunction of relation calls $g_1 \wedge g_2 \wedge \dots \wedge g_n$, where $g_i = R^k_i t_1 \dots t_k$.
+  \item \emph{Call} node corresponds to the conjunction of relational calls $g_1 \wedge g_2 \wedge \dots \wedge g_n$, where $g_i = R^k_i t_1 \dots t_k$.
 It has a single child node, which is produced by unfolding of every call in the conjunction $g_1' \wedge g_2' \wedge \dots \wedge g_n'$, where $g_i' = unfold \ g_i$.
@@
 
-First, a miniKanren goal is partially evaluated to a conjunction of relation calls.
+First, a miniKanren goal is partially evaluated to a conjunction of relational calls.
 
@@
 
-Another problem is that the $split$ only splits a goal into two subgoals, and if each of the relational calls involved unfolds in a conjunction of at least two relation calls, the goals are going to grow faster, than the $split$ splits.
+Another problem is that the $split$ only splits a goal into two subgoals, and if each of the relational calls involved unfolds in a conjunction of at least two relational calls, the goals are going to grow faster, than the $split$ splits.
 

### 2020-02-07 papers/specialization/SEIM-2020.tex
Template for the specialization paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\newcommand{\miniKanren}{\textsc{miniKanren}}

\title{Особенности Специализации Реляционных Программ}

\author{\IEEEauthorblockN{1\textsuperscript{st} Мария Куклина}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Екатерина Вербицкая}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}

}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
Специализаия, реляционное программирование, суперкомпиляция, частичная дедукция 
\end{IEEEkeywords}

\section{Введение}

Вводная про реляционное программирование. (чем полезно в народном хозяйстве)

Вводная про \miniKanren. Как выглядят программы, про запуск в разных направлениях.

Про решение задач поиска силами распознавателя на \miniKanren.~\cite{lozov2019relational}

Про необходимость улучшать специализаторы.

В этой статье мы делаем то-сё, пятое-десятое, измеряем, сравниваем и описываем, что делать дальше.

\section{Специализация}

Специализация (или частичные вычисления) это вот это.~\cite{jones1993partial}

Есть много разных вариантов специализаторов, в том числе суперкомпиляция~\cite{soerensen1996positive} для функциональных языков и конъюнктивная частичная дедукция~\cite{de1999conjunctive} для логических.

\subsection{Суперкомпиляция}

\subsection{Конъюнктивная частичная дедукция}

\section{Подходы для специализации \miniKanren}

Мы адаптировали конъюнктивную частичную дедукцию и суперкомпиляцию для \miniKanren, при этом для каждой из них сделали много разных модификаций.

В этом разделе мы их опишем и сравним.

\subsection{Подход 1}

\subsection{Подход 2}

\subsection{...}

\subsection{Подход n}

\subsection{Сравнение подходов и выводы}

\section{Заключение}

Итого, мы реализовали разные подходы, сравнили их и получили следующие результаты.

\section*{Acknowledgment}

Выражаем благодарность Дмитрию Юрьевичу Булычеву и Даниилу Андреевичу Березуну за плодотворные дискуссии и конструктивную критику.

\bibliographystyle{IEEEtran.bst}
\bibliography{IEEEbib.bib}
\end{document}

### 2020-02-07 papers/bta/SEIM-2020.tex
Template for the BTA paper

--- 2020-02-07 papers/specialization/SEIM-2020.tex
@@
 \newcommand{\miniKanren}{\textsc{miniKanren}}
+\newcommand{\mercury}{\textsc{mercury}}
 
-\title{Особенности Специализации Реляционных Программ}
+\title{Анализ времени связывания для реляционных программ}
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Мария Куклина}
+\author{\IEEEauthorblockN{1\textsuperscript{st} Ирина Артемьева}
 \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
@@
 \begin{IEEEkeywords}
-Специализаия, реляционное программирование, суперкомпиляция, частичная дедукция 
+Реляционное программирование, анализ времени связывания
 \end{IEEEkeywords}
@@
 
-Про необходимость улучшать специализаторы.
+Про то, что даже имея хороший специализатор~\cite{jones1993partial}, сложно достичь супер-быстрых программ, поэтому мы решили транслировать в функциональные.
+
+\miniKanren --- язык хитрый, и трансляция в функциональный язык должна осуществляться с учетом направления. При этом направление влияет в том числе на порядок вычислений.
+
+Для определения порядка связывания имен в программах давно придумали анализ времени связывания.
+В \mercury он давно есть~\cite{vanhoof2004binding}, а для \miniKanren его что-то не сделали, но мы исправили это недоразумение.
 
@@
 
-\section{Специализация}
+\section{\miniKanren}
 
-Специализация (или частичные вычисления) это вот это.~\cite{jones1993partial}
+\miniKanren это такой язык.
 
-Есть много разных вариантов специализаторов, в том числе суперкомпиляция~\cite{soerensen1996positive} для функциональных языков и конъюнктивная частичная дедукция~\cite{de1999conjunctive} для логических.
+Вот его конструкции, вот так он исполняется (в разные стороны по-разному!).
 
-\subsection{Суперкомпиляция}
+Что нужно для трансляции его в функциональный язык (с выводом: было бы неплохо иметь bta).
 
-\subsection{Конъюнктивная частичная дедукция}
+\section{Анализ времени связывания для \miniKanren}
 
-\section{Подходы для специализации \miniKanren}
+Вот так мы его делаем.
 
-Мы адаптировали конъюнктивную частичную дедукцию и суперкомпиляцию для \miniKanren, при этом для каждой из них сделали много разных модификаций.
-
-В этом разделе мы их опишем и сравним.
-
-\subsection{Подход 1}
-
-\subsection{Подход 2}
-
-\subsection{...}
-
-\subsection{Подход n}
-
-\subsection{Сравнение подходов и выводы}
+Вот поэтому считаем, что это работает. Вот тут семантика для \miniKanren~\cite{rozplokhas2019certified}.
 
@@
 
-Итого, мы реализовали разные подходы, сравнили их и получили следующие результаты.
+Итого, мы разработали анализ времени связывания для \miniKanren.
+
+Доказали, что он работает.
+
+В будущем еще и в транслятор вставим, вот житуха начнется-то.
 

### 2020-02-27 papers/TEASE-LP/example.tex
Draft of the TEASE-LP submission.

\documentclass[submission,copyright,creativecommons]{eptcs}
\providecommand{\event}{TEASE-LP 2020} % Name of the event you are submitting to
\usepackage{underscore}           % Only needed if you use pdflatex.
\usepackage{amsmath}

\title{Binding-Time Analysis for \textsc{miniKanren}}
\author{Ekaterina Verbitskaia
\institute{JetBrains Research\\
Saint Petersburg, Russia}
\email{kajigor@gmail.com}
\and
Irina Artemeva
\institute{IFMO University\\
Saint Petersburg, Russia}
\email{irina-pluralia@rambler.ru}
\and
Dmitri Boulytchev
\institute{SPbSU\\
Saint Petersburg, Russia}
\email{dboulytchev@gmail.com}
}
\def\titlerunning{Binding-Time Analysis for \miniKanren{}}
\def\authorrunning{E. Verbitskaia, I. Artemeva \& D. Boulytchev}

\newcommand{\miniKanren}{\textsc{miniKanren}}

\begin{document}

\maketitle

\begin{abstract}
We present a binding-time analysis algorithm for \miniKanren{}.
It is capable to determine the order in which names within a program are bound and can be used to facilitate specialization and as a step of translation into a functional language.
\end{abstract}

\section{Introduction}

\miniKanren{} is a family of domain-specific languages for pure \emph{relational} programming.
The core of \miniKanren{} is very small and can be implemented in few lines of the host language~\cite{hemann2013ukanren}.
A program in \miniKanren{} consists of a \emph{goal} and a set of \emph{relation definitions} with fixed arity.
A goal is either a \emph{disjunction} of two goals, a \emph{conjunction} of two goals, \emph{invocation of a relation} or a \emph{unification of two terms}.
A term is either a \emph{variable} or a \emph{constructor} of arity $n$ applied to $n$ terms.
All free variables within a goal are brought up into scope with a $\underline{fresh}$ operator.
The abstract syntax of the language is presented below.
We denote a goal with $G$, a term with $T$, a variable with $V$, a name of the relation with arity $n$ with $R^n$ and a name of the constructor with arity $n$ with $C^n$.
\begin{align*}
  G &:= G \vee G \mid G \wedge G \mid \underline{call} \ R^n \ [T_1, \dots, T_n] \mid T \equiv T \mid \underline{fresh} \ V \ G  \\
  T &:= V \mid \underline{cons} \ C^n \  [T_1, \dots, T_n]
\end{align*}

A program in \miniKanren{} can be run in different directions: given some of its arguments, it computes all the possible values of the rest of the arguments.
When the relation is run with only the last argument unknown, we say it is run in the \emph{forward direction}, otherwise we call it running in the \emph{backward direction}.
The search employed in \miniKanren{} is complete, so all possible answers will be computed eventually, albeit it may take a long time.
In reality, running time is often unpredictable and depends on the direction.
There are different approaches to tackling this problem: using a divergence test to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, employing specialization~\cite{lozov2019relational}, even rewriting a program by hand introducing some redundancy\footnote{Example of a program which can be rewritten by hand to improve running time in both direction: \url{https://github.com/JetBrains-Research/OCanren/blob/master/regression/test002sort.ml}}.

It has been shown in~\cite{lozov2019relational} that online conjunctive partial deduction is capable of improving running time of a program, but we believe that offline specialization may provide better results.
Offline specializers employ some static analysis before the specialization step.
The most prominent of them is \emph{binding-time analysis}.
This analysis is used to determine which data is available statically, and which --- dynamically.
Having this annotations, specialization algorithm can ignore all dynamic data and only symbolically execute a program with respect to the static data.
This usually leads to the more precise and powerful specialization.

The authors are also working on translation of the \miniKanren{} code into a functional programming language.
In translation, it is crucial to know the order in which variables of a program are bound.
We see binding-time analysis as a suitable technique for determining this order.

\section{Binding-Time Analysis for \miniKanren{}}

A program in miniKanren is a goal accompanied with a set of relation definitions.

The goal of bta is to determine the order in which variables are getting bound within a given goal and all relations called within it.

We will consider a goal in \miniKanren{} to be a disjunction of conjunctions of relations invocations or unifications.
It is always possible to transform any \miniKanren{} goal in this normal form.

We annotate each variable in a goal with either an $undef$ annotation or a natural number.

We start with a goal in which input parameters are annotated with $0$, while all other variables are annotated with $undef$.

Then we run a the following procedure until a fix-point is reached.

Every disjunct is treated seperately from other disjuncts, since they do not influence each other.

Within a conjunction, there can be either a unification or an invocation.
Whenever a new annotation is assigned to some variable use within a conjunction is propagated onto other uses of the same variable.

First we annotate unifications with some partial annotations.
When there are no unifications which can be annotated, we annotate invocations of the current relation.
Then --- all other invocations.
Any of the later annotations can create new partially annotated unifications, so they have to be treated again, in the order they have appeared.

Annotating unification works like this.
There are several possible cases to consider.
The annotations are written in superscripts, $t[x_0, \dots, x_k]$ denotes a term with free variables $x_0, \dots, x_k$.

\begin{itemize}
  \item The unification of an unannotated variable with a term: $x^{undef} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$, where all free variables in a term are annotated with numbers. Then we annotate $x$ with a $1+max\{i_0,\dots,i_k\}$.
  \item The unification of an annotated variable with a term, in which some variables are not annotated: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$. Here all unannotated variables within the term are annotated with $1+n$.
  \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ Name \ [t_0^{i_0}, \dots, t_k^{i_k}] \equiv \underline{cons} \ Name \ [s_0^{j_0}, \dots, s_k^{j_k}]$. This is equivalent to the conjunction of the unifications $t_l^{i_l} \equiv s_l^{j_l}$ and is treated accordingly.
  \item Two cases symmetrical to the first two.
  \item If a unification does not conform to any of these cases, do not annotate it, leaving it for later.
\end{itemize}

When annotating invocations, there are possible several cases.

\begin{itemize}
  \item All variables used within an invocation are annotated with $undef$. Annotation in this case is impossible, proceed to the next case.
  \item All variables are annotated. No annotation is needed -- proceed to the next case.
  \item We have already annotated the relation with the same in the consistent direction. Then all $undef$ variables get the annotation $1+n$, where $n$ is the maximum annotation of the annotated variables.
  \item We have not encountered a consistent invocation. Then we try to annotate the body of the corresponding relation.
\end{itemize}

\nocite{*}
\bibliographystyle{eptcs}
\bibliography{generic}
\end{document}

### 2020-02-28 papers/TEASE-LP/example.tex
TEASE-LP edits

@@
 \newcommand{\miniKanren}{\textsc{miniKanren}}
+\newcommand{\prolog}{\textsc{Prolog}}
+\newcommand{\mercury}{\textsc{mercury}}
 
@@
 It has been shown in~\cite{lozov2019relational} that online conjunctive partial deduction is capable of improving running time of a program, but we believe that offline specialization may provide better results.
-Offline specializers employ some static analysis before the specialization step.
+Offline specializers employ a static analysis before the specialization step.
 The most prominent of them is \emph{binding-time analysis}.
-This analysis is used to determine which data is available statically, and which --- dynamically.
-Having this annotations, specialization algorithm can ignore all dynamic data and only symbolically execute a program with respect to the static data.
-This usually leads to the more precise and powerful specialization.
+It is used to determine which data is available statically, and which --- dynamically.
+Having this annotations, specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
+This usually leads to more precise and powerful specialization.
 
@@
 
+In the next section we present a binding-time analysis algorithm for \miniKanren{}.
+
 \section{Binding-Time Analysis for \miniKanren{}}
 
-A program in miniKanren is a goal accompanied with a set of relation definitions.
+A binding-time analysis starts by annotating a given goal.
+Some free variables of the goal are marked with $0$ which denotes they are the \emph{input} or \emph{static} variables, others are annotated with $undef$.
+The aim is to determine the order in which the $undef$ variables are bound within the goal as well as to annotate all relation definitions called.
+Existing binding-time analysis for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
+Natural numbers can serve as annotations~\cite{Thiemann1997AUF} and are better suited to our problem.
+Thus we use natural numbers, $0$ and $undef$ for annotations.
 
-The goal of bta is to determine the order in which variables are getting bound within a given goal and all relations called within it.
+We assume that all goals are in \emph{normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level:
+$$G = \underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ T_1 \dots T_k \mid T_1 \equiv T_2)$$
 
-We will consider a goal in \miniKanren{} to be a disjunction of conjunctions of relations invocations or unifications.
-It is always possible to transform any \miniKanren{} goal in this normal form.
+The algorithm is to execute the following steps until a fix-point is reached.
+Since disjuncts do not influence each other, we treat every disjunct in isolation.
+A conjunct is either a unification or a relation call.
+At each step, a unification suitable for annotation is selected.
+If there are no unifications to annotate, then we select a relation call.
+Whenever a new annotation is assigned to a variable, all other uses of the same variable within the conjunction get the same annotation.
 
-We annotate each variable in a goal with either an $undef$ annotation or a natural number.
+Unifications are annotated if there is enough information.
+There are several possible cases described below.
+We write annotations in superscripts; $t[x_0, \dots, x_k]$ denotes a term with free variables $x_0, \dots, x_k$.
+\begin{itemize}
+  \item The unification of an unannotated variable with a term in which all free variables are annotated with numbers: $x^{undef} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
+In this case we annotate $x$ with $1+max\{i_0,\dots,i_k\}$.
+  \item The unification of an annotated variable with a term, in which some variables are not annotated: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
+Here all unannotated variables of the term are annotated with $1+n$.
+  \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0^{i_0}, \dots, t_k^{i_k}] \equiv \underline{cons} \ C^k \ [s_0^{j_0}, \dots, s_k^{j_k}]$.
+This is equivalent to the conjunction of the unifications $t_l^{i_l} \equiv s_l^{j_l}$ and is treated accordingly.
+  \item Two cases symmetrical to the first two, in which a term is unified with a variable.
+\end{itemize}
 
-We start with a goal in which input parameters are annotated with $0$, while all other variables are annotated with $undef$.
+If a unification does not conform to any of these cases, it should not be annotated at the current step.
+Relation calls are only annotated when all unifications have been considered.
+A relation call in which all variables have not been annotated (are marked with $undef$) as well as those already annotated with some positive numbers do not need to be considered.
+Thus relation calls in which only some of the variables are annotated with positive numbers are to be annotated.
 
-Then we run a the following procedure until a fix-point is reached.
-
-Every disjunct is treated seperately from other disjuncts, since they do not influence each other.
-
-Within a conjunction, there can be either a unification or an invocation.
-Whenever a new annotation is assigned to some variable use within a conjunction is propagated onto other uses of the same variable.
-
-First we annotate unifications with some partial annotations.
-When there are no unifications which can be annotated, we annotate invocations of the current relation.
-Then --- all other invocations.
-Any of the later annotations can create new partially annotated unifications, so they have to be treated again, in the order they have appeared.
-
-Annotating unification works like this.
-There are several possible cases to consider.
-The annotations are written in superscripts, $t[x_0, \dots, x_k]$ denotes a term with free variables $x_0, \dots, x_k$.
+We say that two relation calls have a \emph{consistent} direction if they call a relation with the same name and
 
 \begin{itemize}
-  \item The unification of an unannotated variable with a term: $x^{undef} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$, where all free variables in a term are annotated with numbers.
-Then we annotate $x$ with a $1+max\{i_0,\dots,i_k\}$.
-  \item The unification of an annotated variable with a term, in which some variables are not annotated: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
-Here all unannotated variables within the term are annotated with $1+n$.
-  \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ Name \ [t_0^{i_0}, \dots, t_k^{i_k}] \equiv \underline{cons} \ Name \ [s_0^{j_0}, \dots, s_k^{j_k}]$.
-This is equivalent to the conjunction of the unifications $t_l^{i_l} \equiv s_l^{j_l}$ and is treated accordingly.
-  \item Two cases symmetrical to the first two.
-  \item If a unification does not conform to any of these cases, do not annotate it, leaving it for later.
-\end{itemize}
-
-When annotating invocations, there are possible several cases.
-
-\begin{itemize}
-  \item All variables used within an invocation are annotated with $undef$. Annotation in this case is impossible, proceed to the next case.
-  \item All variables are annotated. No annotation is needed -- proceed to the next case.
-  \item We have already annotated the relation with the same in the consistent direction.
+  \item
+   We have already annotated the relation with the same name in the consistent direction.
 Then all $undef$ variables get the annotation $1+n$, where $n$ is the maximum annotation of the annotated variables.

### 2020-02-29 papers/TEASE-LP/example.tex
TEASE-LP more edits

@@
 
-\miniKanren{} is a family of domain-specific languages for pure \emph{relational} programming.
+\miniKanren{}\footnote{\miniKanren{} language web site: \url{http://minikanren.org}} is a family of domain-specific languages for pure \emph{relational} programming.
 The core of \miniKanren{} is very small and can be implemented in few lines of the host language~\cite{hemann2013ukanren}.
@@
 In reality, running time is often unpredictable and depends on the direction.
-There are different approaches to tackling this problem: using a divergence test to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, employing specialization~\cite{lozov2019relational}, even rewriting a program by hand introducing some redundancy\footnote{Example of a program which can be rewritten by hand to improve running time in both direction: \url{https://github.com/JetBrains-Research/OCanren/blob/master/regression/test002sort.ml}}.
+There are different approaches to tackling this problem: using a divergence test to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, employing specialization~\cite{lozov2019relational}, even rewriting a program by hand introducing some redundancy.
+% \footnote{Example of a program which can be rewritten by hand to improve running time in both direction: \url{https://github.com/JetBrains-Research/OCanren/blob/master/regression/test002sort.ml}}.
 
@@
 Here all unannotated variables of the term are annotated with $1+n$.
-  \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0^{i_0}, \dots, t_k^{i_k}] \equiv \underline{cons} \ C^k \ [s_0^{j_0}, \dots, s_k^{j_k}]$.
-This is equivalent to the conjunction of the unifications $t_l^{i_l} \equiv s_l^{j_l}$ and is treated accordingly.
+  \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0, \dots, t_k] \equiv \underline{cons} \ C^k \ [s_0, \dots, s_k]$.
+This is equivalent to the conjunction of the unifications $\bigwedge_{i=0}^{k} t_i \equiv s_i$ and is treated accordingly.
   \item Two cases symmetrical to the first two, in which a term is unified with a variable.
@@
 
-We say that two relation calls have a \emph{consistent} direction if they call a relation with the same name and
+We say that two calls to the same relation have a \emph{consistent} direction if their free variables which are annotated with numbers are ordered in the same way.
+More formally: relation call $R^n [v_1^{i_1}, \dots, v_k^{i_k}]$ has a direction consistent with $R^n [u_1^{j_1}, \dots, u_k^{j_k}]$, if all variables with numerical annotations are ordered in the same way: $v_{l_1}^{i_{l_1}} \leq \dots \leq v_{l_m}^{i_{l_m}}$ and $u_{l_1}^{j_{l_1}} \leq \dots \leq u_{l_m}^{j_{l_m}}$.
 
-\begin{itemize}
-  \item
-   We have already annotated the relation with the same name in the consistent direction.
-Then all $undef$ variables get the annotation $1+n$, where $n$ is the maximum annotation of the annotated variables.
-  \item We have not encountered a consistent invocation. Then we try to annotate the body of the corresponding relation.
-\end{itemize}
+If the current conjunct has a consistent direction with some relation call annotated before, then we use that relation call to annotate the current.
+
+Only a relation call with the direction which has not been encountered before should be annotated.
+It is done by considering its body: variables annotated within the call keep their annotations in the body, which is annotated by  the described algorithm.
+
+This algorithm terminates since there are finitely many relation calls to consider, and each relation call has finitely many possible annotations.
 

### 2020-03-03 papers/TEASE-LP/example.tex
TEASE-LP: edits

@@
 
-\begin{abstract}
-We present a binding-time analysis algorithm for \miniKanren{}.
-It is capable to determine the order in which names within a program are bound and can be used to facilitate specialization and as a step of translation into a functional language.
-\end{abstract}
+% \begin{abstract}
+% We present a binding-time analysis algorithm for \miniKanren{}.
+% It is capable to determine the order in which names within a program are bound and can be used to facilitate specialization and as a step of conversion into a functional language.
+% \end{abstract}
 
-\section{Introduction}
-
-\miniKanren{}\footnote{\miniKanren{} language web site: \url{http://minikanren.org}} is a family of domain-specific languages for pure \emph{relational} programming.
-The core of \miniKanren{} is very small and can be implemented in few lines of the host language~\cite{hemann2013ukanren}.
-A program in \miniKanren{} consists of a \emph{goal} and a set of \emph{relation definitions} with fixed arity.
-A goal is either a \emph{disjunction} of two goals, a \emph{conjunction} of two goals, \emph{invocation of a relation} or a \emph{unification of two terms}.
-A term is either a \emph{variable} or a \emph{constructor} of arity $n$ applied to $n$ terms.
-All free variables within a goal are brought up into scope with a $\underline{fresh}$ operator.
-The abstract syntax of the language is presented below.
-We denote a goal with $G$, a term with $T$, a variable with $V$, a name of the relation with arity $n$ with $R^n$ and a name of the constructor with arity $n$ with $C^n$.
-\begin{align*}
-  G &:= G \vee G \mid G \wedge G \mid \underline{call} \ R^n \ [T_1, \dots, T_n] \mid T \equiv T \mid \underline{fresh} \ V \ G  \\
-  T &:= V \mid \underline{cons} \ C^n \  [T_1, \dots, T_n]
-\end{align*}
-
+\miniKanren{}\footnote{\miniKanren{} language web site: \url{http://minikanren.org}} is a family of domain-specific languages for pure \emph{relational} programming which core can be implemented in a few lines of the host language~\cite{hemann2013ukanren}.
 A program in \miniKanren{} can be run in different directions: given some of its arguments, it computes all the possible values of the rest of the arguments.
 When the relation is run with only the last argument unknown, we say it is run in the \emph{forward direction}, otherwise we call it running in the \emph{backward direction}.
+A single relational specification creates a multitude of directions, each solving distinctive problems.
+
 The search employed in \miniKanren{} is complete, so all possible answers will be computed eventually, albeit it may take a long time.
-In reality, running time is often unpredictable and depends on the direction.
-There are different approaches to tackling this problem: using a divergence test to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, employing specialization~\cite{lozov2019relational}, even rewriting a program by hand introducing some redundancy.
-% \footnote{Example of a program which can be rewritten by hand to improve running time in both direction: \url{https://github.com/JetBrains-Research/OCanren/blob/master/regression/test002sort.ml}}.
+In reality, the running time depends on the direction and is highly unpredictable.
+This is where the promise of \miniKanren{} falls short: one has to write the relational specification with the specific direction in mind for it to be efficient.
+There are several attempts to tackle this problem in an automatic fashion: one way is to use a \emph{divergence test} to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, the other is to employ \emph{specialization}~\cite{lozov2019relational}, while the one we have been working recently is to converse a relational specification into a function which runs fast in the given direction.
 
-It has been shown in~\cite{lozov2019relational} that online conjunctive partial deduction is capable of improving running time of a program, but we believe that offline specialization may provide better results.
+Given a relational specification equipped with a certain direction, the functional conversion constructs a function in which unifications are mapped either into a pattern matching or are used in let-bindings to compute the intermediate results.
+Disjunctions give rise to branches in pattern matchings within a function definition, while conjunctions mean sequential computations.
+The nondeterministic computations are modeled with lists.
+The most nontrivial part of the functional conversion is to determine the order in which to bind each variable within the function body.
+
+It has been shown in~\cite{lozov2019relational} that an online conjunctive partial deduction is capable of improving running time of a program in a given direction, but we believe that offline specialization may provide better results.
 Offline specializers employ a static analysis before the specialization step.
-The most prominent of them is \emph{binding-time analysis}.
-It is used to determine which data is available statically, and which --- dynamically.
-Having this annotations, specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
+The most prominent of them is the \emph{binding-time analysis}.
+It is used to determine which data is available statically, and which~--- dynamically.
+Having this annotations, a specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
 This usually leads to more precise and powerful specialization.
 
-The authors are also working on translation of the \miniKanren{} code into a functional programming language.
-In translation, it is crucial to know the order in which variables of a program are bound.
-We see binding-time analysis as a suitable technique for determining this order.
+We believe that a variation of a binding-time analysis can serve as a solution to figuring out the order of variable bindings necessary for functional conversion as well as be a part of an offline specializer.
+The goal of a binding-time analysis is to identify at which step of a computation each variable gets its value.
+This process is called annotations: each variable gets an annotation from the domain of \emph{binding-time}.
+Existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
+Natural numbers serve as annotations in~\cite{Thiemann1997AUF} and are better suited to our problem.
 
-In the next section we present a binding-time analysis algorithm for \miniKanren{}.
+The binding-time analysis for \miniKanren{} should annotate a given goal as well as determine suitable directions for all relation calls involved in the result computation.
+We start by marking some free variables within the goal with $0$ which denotes they are the \emph{input} or \emph{static} variables, while all others are annotated with $undef$.
+This is the way we specify a direction for the goal.
 
-\section{Binding-Time Analysis for \miniKanren{}}
+We assume that all goals are in \emph{normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level.
+Relation calls and constructors can only have variables as their arguments.
+Unification should have a single variable on the left-hand side with the exception of unifying two constructors of arity $0$.
+$$G = \underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T \mid \underline{cons} \ C^0 \equiv \underline{cons} \ C^0)$$
 
-A binding-time analysis starts by annotating a given goal.
-Some free variables of the goal are marked with $0$ which denotes they are the \emph{input} or \emph{static} variables, others are annotated with $undef$.
-The aim is to determine the order in which the $undef$ variables are bound within the goal as well as to annotate all relation definitions called.
-Existing binding-time analysis for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
-Natural numbers can serve as annotations~\cite{Thiemann1997AUF} and are better suited to our problem.
-Thus we use natural numbers, $0$ and $undef$ for annotations.
-
-We assume that all goals are in \emph{normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level:
-$$G = \underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ T_1 \dots T_k \mid T_1 \equiv T_2)$$
-
-The algorithm is to execute the following steps until a fix-point is reached.
+The algorithm is to execute the following steps until a fixpoint is reached.
 Since disjuncts do not influence each other, we treat every disjunct in isolation.
@@
 Here all unannotated variables of the term are annotated with $1+n$.
-  \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0, \dots, t_k] \equiv \underline{cons} \ C^k \ [s_0, \dots, s_k]$.
+  % \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0, \dots, t_k] \equiv \underline{cons} \ C^k \ [s_0, \dots, s_k]$.
 This is equivalent to the conjunction of the unifications $\bigwedge_{i=0}^{k} t_i \equiv s_i$ and is treated accordingly.
-  \item Two cases symmetrical to the first two, in which a term is unified with a variable.
+  % \item Two cases symmetrical to the first two, in which a term is unified with a variable.
 \end{itemize}
@@
 We say that two calls to the same relation have a \emph{consistent} direction if their free variables which are annotated with numbers are ordered in the same way.
-More formally: relation call $R^n [v_1^{i_1}, \dots, v_k^{i_k}]$ has a direction consistent with $R^n [u_1^{j_1}, \dots, u_k^{j_k}]$, if all variables with numerical annotations are ordered in the same way: $v_{l_1}^{i_{l_1}} \leq \dots \leq v_{l_m}^{i_{l_m}}$ and $u_{l_1}^{j_{l_1}} \leq \dots \leq u_{l_m}^{j_{l_m}}$.
+More formally: relation call $R^n \ v_1^{i_1} \dots \ v_k^{i_k}$ has a direction consistent with $R^n \ u_1^{j_1} \dots \ u_k^{j_k}$, if all variables with numerical annotations are ordered in the same way: $v_{l_1}^{i_{l_1}} \leq \dots \leq v_{l_m}^{i_{l_m}}$ and $u_{l_1}^{j_{l_1}} \leq \dots \leq u_{l_m}^{j_{l_m}}$.
 

### 2020-03-04 papers/TEASE-LP/example.tex
TEASE-LP edits

@@
 Irina Artemeva
-\institute{IFMO University\\
+\institute{ITMO University\\
 Saint Petersburg, Russia}
@@
 \and
-Dmitri Boulytchev
-\institute{SPbSU\\
+Daniil Berezun
+\institute{JetBrains Research\\
 Saint Petersburg, Russia}
-\email{dboulytchev@gmail.com}
+\email{daniil.berezun@jetbrains.com }
 }
 \def\titlerunning{Binding-Time Analysis for \miniKanren{}}
-\def\authorrunning{E. Verbitskaia, I. Artemeva \& D. Boulytchev}
+\def\authorrunning{E. Verbitskaia, I. Artemeva \& D. Berezun}
 
@@
 
-\miniKanren{}\footnote{\miniKanren{} language web site: \url{http://minikanren.org}} is a family of domain-specific languages for pure \emph{relational} programming which core can be implemented in a few lines of the host language~\cite{hemann2013ukanren}.
+\miniKanren{}\footnote{\miniKanren{} language web site: \url{http://minikanren.org}} is a family of domain-specific languages for pure \emph{relational} programming the core of which can be implemented in a few lines of the host language~\cite{hemann2013ukanren}.
 A program in \miniKanren{} can be run in different directions: given some of its arguments, it computes all the possible values of the rest of the arguments.
 When the relation is run with only the last argument unknown, we say it is run in the \emph{forward direction}, otherwise we call it running in the \emph{backward direction}.
-A single relational specification creates a multitude of directions, each solving distinctive problems.
+A single relational specification creates a multitude of directions, each solving a distinctive problem.
 
@@
 This is where the promise of \miniKanren{} falls short: one has to write the relational specification with the specific direction in mind for it to be efficient.
-There are several attempts to tackle this problem in an automatic fashion: one way is to use a \emph{divergence test} to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, the other is to employ \emph{specialization}~\cite{lozov2019relational}, while the one we have been working recently is to converse a relational specification into a function which runs fast in the given direction.
+There are several attempts to tackle this problem in an automatic fashion: one way is to use a \emph{divergence test} to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, the other is to employ \emph{specialization}~\cite{lozov2019relational}, while the one we have been working on recently is to converse a relational specification into a function which runs fast in the given direction.
 
@@
 
-It has been shown in~\cite{lozov2019relational} that an online conjunctive partial deduction is capable of improving running time of a program in a given direction, but we believe that offline specialization may provide better results.
+It has been shown in~\cite{lozov2019relational} that an online conjunctive partial deduction is capable of improving running time of a program in a given direction, but we believe that offline specialization may provide even better results in the context of relational interpreters.
 Offline specializers employ a static analysis before the specialization step.
-The most prominent of them is the \emph{binding-time analysis}.
+The most prominent of them is \emph{binding-time analysis}.
 It is used to determine which data is available statically, and which~--- dynamically.
-Having this annotations, a specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
-This usually leads to more precise and powerful specialization.
+Having these annotations, a specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
+This usually leads to a more precise and powerful specialization.
 
 We believe that a variation of a binding-time analysis can serve as a solution to figuring out the order of variable bindings necessary for functional conversion as well as be a part of an offline specializer.
-The goal of a binding-time analysis is to identify at which step of a computation each variable gets its value.
-This process is called annotations: each variable gets an annotation from the domain of \emph{binding-time}.
+The goal of binding-time analysis is to identify at which step of a computation each variable gets its value.
+This process is called annotation: each variable gets an annotation from the domain of \emph{binding-time}.
 Existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
@@
 The binding-time analysis for \miniKanren{} should annotate a given goal as well as determine suitable directions for all relation calls involved in the result computation.
-We start by marking some free variables within the goal with $0$ which denotes they are the \emph{input} or \emph{static} variables, while all others are annotated with $undef$.
+We start by marking some free variables within the goal with $0$ which means they are the \emph{input} or \emph{static} variables, while all others are annotated with $undef$.
 This is the way we specify a direction for the goal.
 
-We assume that all goals are in \emph{normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level.
+We assume that all goals are in \emph{canonical normal form}: $\underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$.
+A goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level.
 Relation calls and constructors can only have variables as their arguments.
 Unification should have a single variable on the left-hand side with the exception of unifying two constructors of arity $0$.
-$$G = \underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T \mid \underline{cons} \ C^0 \equiv \underline{cons} \ C^0)$$
+We leave unifications of the zero-arity constructors out, since they can either be evaluated to a failure or to a success without extending the substitution which can be determined statically.
+Any core \miniKanren{} program can be converted to this form.
 
@@
 Unifications are annotated if there is enough information.
-There are several possible cases described below.
+There are two possible cases described below.
 We write annotations in superscripts; $t[x_0, \dots, x_k]$ denotes a term with free variables $x_0, \dots, x_k$.
 \begin{itemize}
-  \item The unification of an unannotated variable with a term in which all free variables are annotated with numbers: $x^{undef} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
+  \item The unification of a variable annotated with $undef$ with a term in which all free variables are annotated with numbers: $x^{undef} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
 In this case we annotate $x$ with $1+max\{i_0,\dots,i_k\}$.
-  \item The unification of an annotated variable with a term, in which some variables are not annotated: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
-Here all unannotated variables of the term are annotated with $1+n$.
+  \item The unification of a variable annotated with a number with a term, some variables of which are annotated with $undef$: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
+Here all $undef$ variables are annotated with $1+n$.
   % \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0, \dots, t_k] \equiv \underline{cons} \ C^k \ [s_0, \dots, s_k]$.
@@
 Relation calls are only annotated when all unifications have been considered.
-A relation call in which all variables have not been annotated (are marked with $undef$) as well as those already annotated with some positive numbers do not need to be considered.
+A relation call in which all variables are annotated with $undef$ as well as those annotated with only positive numbers do not need to be considered.
 Thus relation calls in which only some of the variables are annotated with positive numbers are to be annotated.
@@
 We say that two calls to the same relation have a \emph{consistent} direction if their free variables which are annotated with numbers are ordered in the same way.
-More formally: relation call $R^n \ v_1^{i_1} \dots \ v_k^{i_k}$ has a direction consistent with $R^n \ u_1^{j_1} \dots \ u_k^{j_k}$, if all variables with numerical annotations are ordered in the same way: $v_{l_1}^{i_{l_1}} \leq \dots \leq v_{l_m}^{i_{l_m}}$ and $u_{l_1}^{j_{l_1}} \leq \dots \leq u_{l_m}^{j_{l_m}}$.
+More formally: a relation call $R^n \ v_1^{i_1} \dots \ v_k^{i_k}$ has a direction consistent with $R^n \ u_1^{j_1} \dots \ u_k^{j_k}$, if all variables with numerical annotations are ordered in the same way: $v_{l_1}^{i_{l_1}} \leq \dots \leq v_{l_m}^{i_{l_m}}$ and $u_{l_1}^{j_{l_1}} \leq \dots \leq u_{l_m}^{j_{l_m}}$.
+For example, relation calls $R^3 \ x^{undef} \ y^{0} \ z^{0}$, $R^3 \ x^{undef} \ y^{1} \ z^{2}$ and $R^3 x^{undef} y^{undef} z^{3}$ all have consistent directions, while the direction of $R^3 \ x^{1} \ y^{2} \ z^{undef}$ is not consistent with the direction of $R^3 \ x^{undef} \ y^{1} \ z^{2}$.
 
-If the current conjunct has a consistent direction with some relation call annotated before, then we use that relation call to annotate the current.
-
+If the current conjunct has a consistent direction with some relation call annotated before, then there is no need to explore this direction again.
+We use the previous call to annotate the current one.
 Only a relation call with the direction which has not been encountered before should be annotated.
 It is done by considering its body: variables annotated within the call keep their annotations in the body, which is annotated by the described algorithm.
+All annotated relation definitions are stored while algorithm runs, so they can be reused if necessary.
 
-This algorithm terminates since there are finitely many relation calls to consider, and each relation call has finitely many possible annotations.
+We presented a binding-time analysis algorithm for \miniKanren{} that determines the order in which variables are bound.
+The algorithm terminates since there are finitely many different relation calls to consider, and each relation call has finitely many possible annotations.
+Unfortunately, some variables can remain $undef$ after the algorithm is finished.
+This happens when there are relation calls which influence the direction of each other.
+Currently we choose the leftmost relation call to annotate first; while another, less efficient but more fair, solution may be to consider all possible combinations of directions.
+In our experience, this problem rarely arises in the context of relational interpreters, but we should further investigate the class of such programs and develop a better strategy of dealing with them.
 

### 2020-03-04 papers/TEASE-LP/example.tex
TEASE-LP edits

@@
 
-It has been shown in~\cite{lozov2019relational} that an online conjunctive partial deduction is capable of improving running time of a program in a given direction, but we believe that offline specialization may provide even better results in the context of relational interpreters.
-Offline specializers employ a static analysis before the specialization step.
-The most prominent of them is \emph{binding-time analysis}.
-It is used to determine which data is available statically, and which~--- dynamically.
-Having these annotations, a specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
-This usually leads to a more precise and powerful specialization.
+In compilation and specialization to determine at what time the values of the variables can be computed, \emph{binding-time analysis}, or bta, is used.
+Typically, bta divides all computations into two stages: \emph{compile-time}, or \emph{static}, and \emph{execution time}, or \emph{dynamic}.
+One approach to bta is to explicitly annotate program variables and computations by equipping them with their binding times.
+In logic programming there is no predefined execution direction --- no order in which conjunctions should be considered ---
+which complicates the problem of identifying variables binding times.
+It has been shown~\cite{lozov2019relational} that online conjunctive partial deduction may improve program running time in a given direction.
+We are investigating the applicability of offline specialization methods in the context of functional conversion.
+Offline approaches are known~\cite{jones1993partial} to be safe, helpful in the design and implementation of partial evaluation projects, and are more efficient than the online ones, which is crucial in the context of logic programming.
+The existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
+Natural numbers serve as annotations in~\cite{Thiemann1997AUF} and are better suited to our problem.
 
-We believe that a variation of a binding-time analysis can serve as a solution to figuring out the order of variable bindings necessary for functional conversion as well as be a part of an offline specializer.
-The goal of binding-time analysis is to identify at which step of a computation each variable gets its value.
-This process is called annotation: each variable gets an annotation from the domain of \emph{binding-time}.
-Existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
-Natural numbers serve as annotations in~\cite{Thiemann1997AUF} and are better suited to our problem.
+% It has been shown in~\cite{lozov2019relational} that an online conjunctive partial deduction is capable of improving running time of a program in a given direction, but we believe that offline specialization may provide even better results in the context of relational interpreters.
+% Offline specializers employ a static analysis before the specialization step.
+% The most prominent of them is \emph{binding-time analysis}.
+% It is used to determine which data is available statically, and which~--- dynamically.
+% Having these annotations, a specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
+% This usually leads to a more precise and powerful specialization.
+
+% We believe that a variation of a binding-time analysis can serve as a solution to figuring out the order of variable bindings necessary for functional conversion as well as be a part of an offline specializer.
+% The goal of binding-time analysis is to identify at which step of a computation each variable gets its value.
+% This process is called annotation: each variable gets an annotation from the domain of \emph{binding-time}.
+% Existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
+% Natural numbers serve as annotations in~\cite{Thiemann1997AUF} and are better suited to our problem.
 
@@
 
-We assume that all goals are in \emph{canonical normal form}: $\underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$.
-A goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level.
+We assume that all goals are in \emph{canonical normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level.
 Relation calls and constructors can only have variables as their arguments.
@@
 Any core \miniKanren{} program can be converted to this form.
+$$Goal = \underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$$
 

### 2020-03-04 papers/TEASE-LP/example.tex
TEASE-LP: small edits

@@
 \newcommand{\mercury}{\textsc{mercury}}
+\newcommand\undef{unde\!f}
+\newcommand\fresh{f\!resh}
 
@@
 
-In compilation and specialization to determine at what time the values of the variables can be computed, \emph{binding-time analysis}, or bta, is used.
+\emph{Binding-time analysis}, or bta, is used in compilation and specialization to determine at what time the values of the variables can be computed.
 Typically, bta divides all computations into two stages: \emph{compile-time}, or \emph{static}, and \emph{execution time}, or \emph{dynamic}.
-One approach to bta is to explicitly annotate program variables and computations by equipping them with their binding times.
+One approach to bta is to explicitly annotate program variables and computations by equipping them with their \emph{binding times}.
 In logic programming there is no predefined execution direction --- no order in which conjunctions should be considered ---
@@
 We are investigating the applicability of offline specialization methods in the context of functional conversion.
-Offline approaches are known~\cite{jones1993partial} to be safe, helpful in the design and implementation of partial evaluation projects, and are more efficient than the online ones, which is crucial in the context of logic programming.
+Offline approaches are known~\cite{jones1993partial} to be safe, helpful in the design and implementation of partial evaluation projects, and more efficient than the online ones, which is crucial in the context of logic programming.
 The existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
@@
 The binding-time analysis for \miniKanren{} should annotate a given goal as well as determine suitable directions for all relation calls involved in the result computation.
-We start by marking some free variables within the goal with $0$ which means they are the \emph{input} or \emph{static} variables, while all others are annotated with $undef$.
+We start by marking some free variables within the goal with $0$ which means they are the \emph{input} or \emph{static} variables, while all others are annotated with $\undef{}$.
 This is the way we specify a direction for the goal.
 
-We assume that all goals are in \emph{canonical normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{fresh}$ operator on the top level.
+We assume that all goals are in \emph{canonical normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{\fresh{}}$ operator on the top level.
 Relation calls and constructors can only have variables as their arguments.
 Unification should have a single variable on the left-hand side with the exception of unifying two constructors of arity $0$.
-We leave unifications of the zero-arity constructors out, since they can either be evaluated to a failure or to a success without extending the substitution which can be determined statically.
+We leave out unifications of the zero-arity constructors, since they can either evaluate to a failure or to a success without extending the substitution which can be determined statically.
 Any core \miniKanren{} program can be converted to this form.
-$$Goal = \underline{fresh} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$$
+$$Goal = \underline{\fresh{}} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$$
 
@@
 \begin{itemize}
-  \item The unification of a variable annotated with $undef$ with a term in which all free variables are annotated with numbers: $x^{undef} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
+  \item The unification of a variable annotated with $\undef{}$ with a term in which all free variables are annotated with numbers: $x^{\undef{}} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
 In this case we annotate $x$ with $1+max\{i_0,\dots,i_k\}$.
-  \item The unification of a variable annotated with a number with a term, some variables of which are annotated with $undef$: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
-Here all $undef$ variables are annotated with $1+n$.
+  \item The unification of a variable annotated with a number with a term, some variables of which are annotated with $\undef{}$: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
+Here all $\undef{}$ variables are annotated with $1+n$.
   % \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0, \dots, t_k] \equiv \underline{cons} \ C^k \ [s_0, \dots, s_k]$.
@@
 Relation calls are only annotated when all unifications have been considered.
-A relation call in which all variables are annotated with $undef$ as well as those annotated with only positive numbers do not need to be considered.
+A relation call in which all variables are annotated with $\undef{}$ as well as those annotated with only positive numbers do not need to be considered.
 Thus relation calls in which only some of the variables are annotated with positive numbers are to be annotated.
@@
 More formally: a relation call $R^n \ v_1^{i_1} \dots \ v_k^{i_k}$ has a direction consistent with $R^n \ u_1^{j_1} \dots \ u_k^{j_k}$, if all variables with numerical annotations are ordered in the same way: $v_{l_1}^{i_{l_1}} \leq \dots \leq v_{l_m}^{i_{l_m}}$ and $u_{l_1}^{j_{l_1}} \leq \dots \leq u_{l_m}^{j_{l_m}}$.
-For example, relation calls $R^3 \ x^{undef} \ y^{0} \ z^{0}$, $R^3 \ x^{undef} \ y^{1} \ z^{2}$ and $R^3 x^{undef} y^{undef} z^{3}$ all have consistent directions, while the direction of $R^3 \ x^{1} \ y^{2} \ z^{undef}$ is not consistent with the direction of $R^3 \ x^{undef} \ y^{1} \ z^{2}$.
+For example, relation calls $R^3 \ x^{\undef{}} \ y^{0} \ z^{0}$, $R^3 \ x^{\undef{}} \ y^{1} \ z^{2}$ and $R^3 x^{\undef{}} y^{\undef{}} z^{3}$ all have consistent directions, while the direction of $R^3 \ x^{1} \ y^{2} \ z^{\undef{}}$ is not consistent with the direction of $R^3 \ x^{\undef{}} \ y^{1} \ z^{2}$.
 
@@
 The algorithm terminates since there are finitely many different relation calls to consider, and each relation call has finitely many possible annotations.
-Unfortunately, some variables can remain $undef$ after the algorithm is finished.
+Unfortunately, some variables can remain $\undef{}$ after the algorithm is finished.
 This happens when there are relation calls which influence the direction of each other.

### 2020-03-04 papers/TEASE-LP/example.tex
TEASE-LP: final edits

@@
 The search employed in \miniKanren{} is complete, so all possible answers will be computed eventually, albeit it may take a long time.
-In reality, the running time depends on the direction and is highly unpredictable.
+In fact, the running time depends on the direction and is highly unpredictable.
 This is where the promise of \miniKanren{} falls short: one has to write the relational specification with the specific direction in mind for it to be efficient.
@@
 
-Given a relational specification equipped with a certain direction, the functional conversion constructs a function in which unifications are mapped either into a pattern matching or are used in let-bindings to compute the intermediate results.
+Given a relational specification equipped with a certain direction, the \emph{functional conversion} constructs a function in which unifications are mapped either into a pattern matching or are used in let-bindings to compute the intermediate results.
 Disjunctions give rise to branches in pattern matchings within a function definition, while conjunctions mean sequential computations.
@@
 One approach to bta is to explicitly annotate program variables and computations by equipping them with their \emph{binding times}.
-In logic programming there is no predefined execution direction --- no order in which conjunctions should be considered ---
+In relational programming there is no predefined execution direction --- no order in which conjunctions should be considered ---
 which complicates the problem of identifying variables binding times.
@@
 We are investigating the applicability of offline specialization methods in the context of functional conversion.
-Offline approaches are known~\cite{jones1993partial} to be safe, helpful in the design and implementation of partial evaluation projects, and more efficient than the online ones, which is crucial in the context of logic programming.
+Offline approaches are known~\cite{jones1993partial} to be safe, helpful in the design and implementation of partial evaluation projects, and more efficient than the online ones, which is crucial in the context of relational programming.
 The existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
@@
 The binding-time analysis for \miniKanren{} should annotate a given goal as well as determine suitable directions for all relation calls involved in the result computation.
-We start by marking some free variables within the goal with $0$ which means they are the \emph{input} or \emph{static} variables, while all others are annotated with $\undef{}$.
+We start by annotating some free variables within the goal with $0$ which means they are the \emph{input} or \emph{static} variables, while all others are annotated with $\undef{}$.
 This is the way we specify a direction for the goal.
@@
 We assume that all goals are in \emph{canonical normal form}: a goal is a disjunction of conjunctions of either relation calls or unifications; all free variables are brought up into scope by the $\underline{\fresh{}}$ operator on the top level.
-Relation calls and constructors can only have variables as their arguments.
-Unification should have a single variable on the left-hand side with the exception of unifying two constructors of arity $0$.
+Relation calls ($R^k$) and constructors can only have variables ($V$) as their arguments.
+Unification should have a single variable on the left-hand side and a term ($T$) on the right-hand side, with the exception of unifying two constructors of arity $0$.
 We leave out unifications of the zero-arity constructors, since they can either evaluate to a failure or to a success without extending the substitution which can be determined statically.
@@
 
-The algorithm is to execute the following steps until a fixpoint is reached.
+The bta algorithm is to execute the following steps until a fixpoint is reached.
 Since disjuncts do not influence each other, we treat every disjunct in isolation.
-A conjunct is either a unification or a relation call.
-At each step, a unification suitable for annotation is selected.
+At each step, a unification suitable for annotation is selected within the conjunction.
 If there are no unifications to annotate, then we select a relation call.
-Whenever a new annotation is assigned to a variable, all other uses of the same variable within the conjunction get the same annotation.
+Whenever a new annotation is assigned to a variable, all other occurrences of this variable within the conjunction get the same annotation.
 
@@
   \item The unification of a variable annotated with a number with a term, some variables of which are annotated with $\undef{}$: $x^{n} \equiv t[y_0^{i_0},\dots, y_k^{i_k}]$.
-Here all $\undef{}$ variables are annotated with $1+n$.
+Here all $\undef{}$ variables are to be annotated with $1+n$.
   % \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0, \dots, t_k] \equiv \underline{cons} \ C^k \ [s_0, \dots, s_k]$.
@@
 Relation calls are only annotated when all unifications have been considered.
-A relation call in which all variables are annotated with $\undef{}$ as well as those annotated with only positive numbers do not need to be considered.
-Thus relation calls in which only some of the variables are annotated with positive numbers are to be annotated.
+A relation call in which all variables are annotated with $\undef{}$ as well as those with numerical annotations do not need to be considered.
+Thus relation calls in which only some of the variables are annotated with numbers should be examined further.
 
-We say that two calls to the same relation have a \emph{consistent} direction if their free variables which are annotated with numbers are ordered in the same way.
-More formally: a relation call $R^n \ v_1^{i_1} \dots \ v_k^{i_k}$ has a direction consistent with $R^n \ u_1^{j_1} \dots \ u_k^{j_k}$, if all variables with numerical annotations are ordered in the same way: $v_{l_1}^{i_{l_1}} \leq \dots \leq v_{l_m}^{i_{l_m}}$ and $u_{l_1}^{j_{l_1}} \leq \dots \leq u_{l_m}^{j_{l_m}}$.
+We say that two calls to the same relation have a \emph{consistent} direction if the annotations order free variables in the same way.
 For example, relation calls $R^3 \ x^{\undef{}} \ y^{0} \ z^{0}$, $R^3 \ x^{\undef{}} \ y^{1} \ z^{2}$ and $R^3 x^{\undef{}} y^{\undef{}} z^{3}$ all have consistent directions, while the direction of $R^3 \ x^{1} \ y^{2} \ z^{\undef{}}$ is not consistent with the direction of $R^3 \ x^{\undef{}} \ y^{1} \ z^{2}$.
 
-If the current conjunct has a consistent direction with some relation call annotated before, then there is no need to explore this direction again.
+If at the current step the conjunct under consideration has a consistent direction with some relation call annotated before, then there is no need to explore this direction again.
 We use the previous call to annotate the current one.
@@
 It is done by considering its body: variables annotated within the call keep their annotations in the body, which is annotated by the described algorithm.
-All annotated relation definitions are stored while algorithm runs, so they can be reused if necessary.
+All annotated relation definitions are memoized while algorithm runs, so they can be reused if necessary.
 

### 2020-05-18 papers/TEASE-LP/example.tex
Unify

@@
 
-\miniKanren{}\footnote{\miniKanren{} language web site: \url{http://minikanren.org}} is a family of domain-specific languages for pure \emph{relational} programming the core of which can be implemented in a few lines of the host language~\cite{hemann2013ukanren}.
+\miniKanren{}\footnote{\miniKanren{} language web site: \url{http://minikanren.org}} is a family of domain-specific languages for pure \emph{relational} programming, the core of which can be implemented in a few lines of the host language~\cite{hemann2013ukanren}.
 A program in \miniKanren{} can be run in different directions: given some of its arguments, it computes all the possible values of the rest of the arguments.
@@
 This is where the promise of \miniKanren{} falls short: one has to write the relational specification with the specific direction in mind for it to be efficient.
-There are several attempts to tackle this problem in an automatic fashion: one way is to use a \emph{divergence test} to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, the other is to employ \emph{specialization}~\cite{lozov2019relational}, while the one we have been working on recently is to converse a relational specification into a function which runs fast in the given direction.
+There are several attempts to tackle this problem in an automatic fashion: one way is to use a \emph{divergence test} to stop execution of definitely diverging computations~\cite{rozplokhas2018improving}, the other is to employ \emph{specialization}~\cite{lozov2019relational}, while the one we have been working on recently is to convert a relational specification into a function that runs fast in the given direction.
 
-Given a relational specification equipped with a certain direction, the \emph{functional conversion} constructs a function in which unifications are mapped either into a pattern matching or are used in let-bindings to compute the intermediate results.
-Disjunctions give rise to branches in pattern matchings within a function definition, while conjunctions mean sequential computations.
-The nondeterministic computations are modeled with lists.
+Given a relational specification equipped with a certain direction, \emph{functional conversion} constructs a function in which unifications are mapped either into pattern matching or are used in let-bindings to compute the intermediate results.
+Disjunctions give rise to branches in pattern matchings within a function definition, while conjunctions represent sequential computations.
+% The nondeterministic computations are modeled with lists.
+Lists are used to model multitude of possible answers which are available by the use of disjunctions.
 The most nontrivial part of the functional conversion is to determine the order in which to bind each variable within the function body.
 
-\emph{Binding-time analysis}, or bta, is used in compilation and specialization to determine at what time the values of the variables can be computed.
-Typically, bta divides all computations into two stages: \emph{compile-time}, or \emph{static}, and \emph{execution time}, or \emph{dynamic}.
-One approach to bta is to explicitly annotate program variables and computations by equipping them with their \emph{binding times}.
+\emph{Binding-time analysis}, or BTA, is used in compilation and specialization to determine at what time the values of the variables can be computed.
+Typically, BTA divides all computations into two stages: \emph{compile-time}, or \emph{static}, and \emph{execution time}, or \emph{dynamic}.
+One approach to BTA is to explicitly annotate program variables and computations by equipping them with their \emph{binding times}.
 In relational programming there is no predefined execution direction --- no order in which conjunctions should be considered ---
@@
 Natural numbers serve as annotations in~\cite{Thiemann1997AUF} and are better suited to our problem.
-
-% It has been shown in~\cite{lozov2019relational} that an online conjunctive partial deduction is capable of improving running time of a program in a given direction, but we believe that offline specialization may provide even better results in the context of relational interpreters.
-% Offline specializers employ a static analysis before the specialization step.
-% The most prominent of them is \emph{binding-time analysis}.
-% It is used to determine which data is available statically, and which~--- dynamically.
-% Having these annotations, a specialization algorithm can ignore dynamic data and only symbolically execute a program with respect to the static data.
-% This usually leads to a more precise and powerful specialization.
-
-% We believe that a variation of a binding-time analysis can serve as a solution to figuring out the order of variable bindings necessary for functional conversion as well as be a part of an offline specializer.
-% The goal of binding-time analysis is to identify at which step of a computation each variable gets its value.
-% This process is called annotation: each variable gets an annotation from the domain of \emph{binding-time}.
-% Existing binding-time analyses for pure \prolog{}~\cite{craig2004fully} and \mercury{}~\cite{vanhoof2004binding} only distinguish between static and dynamic variables, which does not help to work out the order.
-% Natural numbers serve as annotations in~\cite{Thiemann1997AUF} and are better suited to our problem.
 
@@
 Any core \miniKanren{} program can be converted to this form.
-$$Goal = \underline{\fresh{}} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$$
+We use the following representation of the goal in canonical form:
+$$Goal = \underline{\fresh{}} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$$.
 
-The bta algorithm is to execute the following steps until a fixpoint is reached.
+The BTA algorithm executes the following steps until a fixpoint is reached.
 Since disjuncts do not influence each other, we treat every disjunct in isolation.
@@
 Here all $\undef{}$ variables are to be annotated with $1+n$.
-  % \item The unification of two constructors with the same name and the same number of arguments: $\underline{cons} \ C^k \ [t_0, \dots, t_k] \equiv \underline{cons} \ C^k \ [s_0, \dots, s_k]$.
-This is equivalent to the conjunction of the unifications $\bigwedge_{i=0}^{k} t_i \equiv s_i$ and is treated accordingly.
-  % \item Two cases symmetrical to the first two, in which a term is unified with a variable.
 \end{itemize}
@@
 
-We say that two calls to the same relation have a \emph{consistent} direction if the annotations order free variables in the same way.
-For example, relation calls $R^3 \ x^{\undef{}} \ y^{0} \ z^{0}$, $R^3 \ x^{\undef{}} \ y^{1} \ z^{2}$ and $R^3 x^{\undef{}} y^{\undef{}} z^{3}$ all have consistent directions, while the direction of $R^3 \ x^{1} \ y^{2} \ z^{\undef{}}$ is not consistent with the direction of $R^3 \ x^{\undef{}} \ y^{1} \ z^{2}$.
+Let us define what it means for two calls to the same relation to have  \emph{consistent} direction.
+We~start by mapping all numerical annotations to $0$.
+Considering $\undef{}$ annotation to be less than $0$, we then compare the annotations of variables of one relation call to the annotations of the respective variable of the other.
+If~all variable pairs are ordered in the same way, then they have consistent direction.
+For~example, the relation calls $R^3 \ x^{\undef{}} \ y^{0} \ z^{0}$ and $R^3 \ x^{\undef{}} \ y^{1} \ z^{2}$ have consistent direction, since they become equal after mapping annotations to $0$.
+The relation call $R^3 x^{\undef{}} y^{\undef{}} z^{3}$ also have consistent direction with the call $R^3 \ x^{\undef{}} \ y^{0} \ z^{0}$ since after mapping to $0$, the first call becomes $R^3 x^{\undef{}} y^{\undef{}} z^{0}$, and annotations of all variables but one are the same.
+Whereas the direction of $R^3 \ x^{0} \ y^{0} \ z^{\undef{}}$ is not consistent with the direction of $R^3 \ x^{\undef{}} \ y^{0} \ z^{0}$: annotation of $x$ in the first call is greater than in the second ($0 > \undef{}$) while the opposite holds for the variable $z$.
 
@@
 All annotated relation definitions are memoized while algorithm runs, so they can be reused if necessary.
+
+If all variables of the input relation call are specified as dynamic, our algorithm is unable to apply any of the rules described.
+The same issue arises when there are dynamic variables which unify only with $\undef{}$ fresh variables or constructors over them.
+Considering the way functional conversion should behave in this circumstance helps to accommodate it in the BTA algorithm.
+Whenever a fresh variable is part of the answer, \miniKanren{} reifies it as a symbolic variable which stands for all possible values.
+It is reasonable to reperesent this with a lazy generator when doing functional conversion.
+When dealing with this direction, BTA first attempts annotation as described and if there are $\undef{}$ variables left, they are treated as if they were unified with a generator available statically.
 

### 2020-05-18 papers/miniKanren/ocanren/bottles/build.sh

@@ -1,2 +1,2 @@
-ocamlfind c -rectypes -o unify_run -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg bottles.ml bottles_run.ml
+ocamlfind c -rectypes -o bottles -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg general.ml tester.ml bottles.ml bottles_trans.ml bottles_defer.ml bottles_fun.ml bottles_run.ml
 

### 2020-05-18 papers/miniKanren/ocanren/bottles/temp/build.sh

@@ -1,2 +1,0 @@
-ocamlfind c -rectypes -o bottles -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg general.ml tester.ml bottles.ml bottles_trans.ml bottles_fun.ml bottles_run.ml
-

### 2020-05-18 papers/miniKanren/ocanren/last/build.sh

@@ -1,2 +1,2 @@
-ocamlfind c -rectypes -o peval -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg peval.ml
+ocamlfind c -rectypes -o run -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml ecce.ml original_fresh.ml original.ml trans_bottles.ml trans_cfa.ml trans_cfhu.ml trans_cfnp.ml trans_defer.ml trans_global.ml trans.ml run.ml
 

### 2020-05-18 papers/miniKanren/ocanren/last/clean.sh

@@ -1,1 +1,1 @@
-rm -f *.cmi *.cmo peval
+rm -f *.cmi *.cmo run

### 2020-05-18 papers/miniKanren/ocanren/plain/build.sh

@@ -1,2 +1,2 @@
-ocamlfind c -rectypes -o peval -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg peval.ml
+ocamlfind c -rectypes -o run -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml ecce.ml trans.ml run.ml
 

### 2020-05-18 papers/miniKanren/ocanren/unify/build.sh

@@ -1,2 +1,2 @@
-ocamlfind c -rectypes -o unify_run -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg unify_fun.ml unify_rel.ml unify_spec.ml unify_run.ml
+ocamlfind c -rectypes -o run -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml spec.ml trans.ml run.ml
 

### 2020-05-18 papers/miniKanren/ocanren/unify/clean.sh

@@ -1,0 +1,1 @@
+rm -f *.cmo *.cmi run

### 2020-05-18 src/OCanrenize.hs

@@ -23,7 +23,7 @@ instance {- OCanren v => -} OCanren (Term X) where
   -- ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "(%s %% %s)" (ocanren h) (ocanren t)
   -- ocanren (C "%"    [h,t]) = printf "(%s %% %s)" (ocanren h) (ocanren t)
   ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "Std.(%%) %s %s" (ocanren h) (ocanren t)
-  ocanren (C "%" [h, t]) = printf "Std.(%%) %s %s" (ocanren h) (ocanren t)
+  ocanren (C "%" [h, t]) = printf "Std.(%%) (%s) (%s)" (ocanren h) (ocanren t)
   ocanren (C "O" []) = "Std.Nat.zero"
   ocanren (C "S" [x]) = printf "Std.Nat.succ (%s)" (ocanren x)
 
@@ -65,6 +65,17 @@ instance {-OCanren v =>-} OCanren (G X) where
 --ocanren (Invoke f ts) = printf "(print_string \"%s\\n\";%s)" (f ++ concat [' ' : ocanren t | t <- ts]) (f ++ concat [' ' : ocanren t | t <- ts])
   ocanren (Invoke f ts) = printf "(%s %s)" f (printArgs $ map ocanren ts)
   ocanren (Let (Def n as b) g) = printf "let rec %s %s = %s in %s" n (printArgs as) (ocanren b) (ocanren g)
+
+
+-- instance {-OCanren v =>-} OCanren (G X) where
+-- --ocanren (t1 :=:  t2)  = printf "(print_string \"%s === %s\\n\"; %s === %s)" (ocanren t1) (ocanren t2) (ocanren t1) (ocanren t2)
+--   ocanren (t1 :=:  t2)  = printf "(%s === %s)" (ocanren t1) (ocanren t2)
+--   ocanren (g1 :/\: g2)  = printf "defer (%s &&& %s)" (ocanren g1) (ocanren g2)
+--   ocanren (g1 :\/: g2)  = printf "defer (%s ||| %s)" (ocanren g1) (ocanren g2)
+--   ocanren (Fresh x g )  = let (names, goal) = freshVars [x] g in printf "(fresh ((%s)) (%s))" (printArgs names) (ocanren goal)
[…]

### 2020-05-18 test/auto/Test/NonConjunctive.hs

@@ -51,7 +51,7 @@ runProp = do
     runNc (-1) "prop__" prop''
     -- runNc (-1) "prop1___" prop1'''
     -- runNc (-1) "prop2___" prop2'''
-    -- runNc (-1) "propPlain" propPlain
+    runNc (-1) "propPlain" propPlain
     -- runNc (-1) "propPlain_" propPlain'
     -- runNc (-1) "prop__1" prop''1
     -- runNc (-1) "prop__2" prop''2
@@ -97,7 +97,7 @@ runPath = do
     runNc (-1) "path" Program.Path.query1
 
 unit_nonConjunctiveTest = do
-  -- runUnify
+  runUnify
   -- runPath
   runProp
   runBottles

### 2020-05-21 papers/miniKanren/text/conclusion.tex
Draft for miniKanren workshop

\section{Conclusion}

We compared some of the specialization techniques for \mk{}.

There seem to be no one good technique.

### 2020-05-21 papers/miniKanren/text/intro.tex
Draft for miniKanren workshop

\section{Introduction}

The \mk{} language family is nice.

Unpredictable running time for different directions.

Specialization sometimes helps, sometimes does not, sometimes makes everything worse.

Control issues, blah blah blah

This problem also appear in supercompilation (partial evaluation, specialization) of functional and imperative programming languages.

<Here should be review of problems in specialization of functional languages>

We came up with the new specialization algorithm, which seems nice, but is as unpredictable as the others.

Here are some numbers...

Contribution is the following
\begin{itemize}
  \item explanation of specialization
  \item new specialization algorithm
  \item why some programs behave worse after specialization
\end{itemize}

### 2020-05-21 papers/miniKanren/text/nonconj.tex
Draft for miniKanren workshop

\section{Non-conjunctive Partial Deduction}

We came up with the idea.

Here we'll briefly describe it.

First we build a tree, than we generate a residual program from it. Residualization is more or less trivial, once the tree is built.

Building the tree is... Well... Complicated.

There are issues with when to unfold, when to stop.

There are a bunch of heuristics for it.

\subsection{Unfolding}

Unfolding is when a call to a relation is substituted by its body.

The most important question is when to unfold.

To much unfolding is sometimes is even worse than not enough unfolding.

\subsection{Heuristic}

First we unfold those conjuncts which are static.

Then --- deterministic.

Then those which are less branching.

The last to be unfolded are those calls, which unfold to a substitution.

### 2020-05-21 papers/miniKanren/text/intro.tex
about transformers disagv

@@
 This problem also appear in supercompilation (partial evaluation, specialization) of functional and imperative programming languages.
+
+While the aim of all meta-computation techniques is residual program efficiency, they rarely consider the residual program efficiency from the target language evaluator point of view.
+The main optimization source is precomputing all possible intermediate and statically-known semantics steps at transformation-time while other criteria like residual program size or possible optimizations and execution cost of different language constructions by target language evaluator are usually out of consideration~\cite{jonesbook}.
+It is known that supercompilation may adversely affect GHC optimizations yielding standalone compilation more powerful~\cite{SCBE,TCES} and cause code explosion~\cite{SCHC}.
+Moreover, caused by transformer complexity it may be hard to predict real program speedup in each case on a bunch of examples even disregarding problems above.
+Wherein, for example, for partial evaluation, an apogee of useless is the use of all static variables in a dynamic context only while it is known to be usefull for some specific form of language processors~\cite{jonesbook,bulyonkov84}.
+Anyway, the formalization problem of cases for useful transformers application is poorly studied.
 

### 2020-05-21 papers/miniKanren/text/evaluation.tex
miniKanren workshop draft

\section{Evaluation}

In the evaluation we compared the performance of our CPD, Ecce and the new non-conjunctive partial evaluator.
We have also implemented the branching heuristic instead of the deterministic one in the CPD, just to be sure, it is not a cureall.

We used the following 4 programs to test our specializers on.
\begin{itemize}
  \item propositional evaluator
  \begin{itemize}
    \item Compicated bool expression in the last conjunctive
    \item Table implementation of the boolean operation, placed as a first conjunct
  \end{itemize}
  \item searching for a unifier of two terms
  \item searching for paths of a specific length in a graph
\end{itemize}

The first one demonstrates the drawback of the approach implemented in ECCE: left-to-right deterministic unfolding.
The boolean expression which narrows the first conjunts is the last, so ECCE does not unfold it until too late.

The second one should be easier for this approach.

The unification is a difficult problem, specialization of which is of importance.

### 2020-05-21 papers/miniKanren/text/nonconj.tex
miniKanren workshop draft

@@
 
-We came up with the idea.
+In this section we will describe a novel approach to specialization of relational programs.
+This approach draws inspiration from both conjunctive partial deduction and supercompilation.
+The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.
 
-Here we'll briefly describe it.
+The high-level idea behind the algorithm is to select a relation call to unfold by using a heuristic which decides if the call can narrow down the answer set.
+A driving tree is constructed for the selected call in isolation.
+The leaves of the computed tree are examined.
+If all leaves are either computed substitutions or are calls to some relations accompanied with non-empty substitutions, then the leaves are collected and are put back into the root conjunction instead of the examined call.
+According to denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
 
-First we build a tree, than we generate a residual program from it. Residualization is more or less trivial, once the tree is built.
+A driving process creates a process tree, from which a residual program is later created.
+The nodes of process tree include a \emph{configuraion} which describes the state of program evaluation at some point.
+In our case configuration is a conjunction of relation calls.
+The substitution computed at each step is also stored in the tree node, although it is not included into the configuration.
 
-Building the tree is... Well... Complicated.
+Each time we examine a conjunction of calls, we \emph{split} them into separate nodes which are driven independently from each other.
+Among the relation calls we select one which is according to the heuristic is likely to narrow down the answer set.
+If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected.
 
-There are issues with when to unfold, when to stop.
+This process does create branchings whenever a disjunction is examined.
+At each step we make sure that we do not start driving a conjunction which we have already examined.
+To do this, we check if the current conjunction is renaming of any other configuration in the tree.
+If it is, then we create a special node which then is residualized into a call to the suitable relation.
 
-There are a bunch of heuristics for it.
+We decided not to do generalization in this approach.
+The generalization is used in supercompilation and partial deduction to ensure termination at the same time as some degree of specialization.
+The generalization of two terms is usually a \emph{most-specific generalization}.
+Generalization is used to abstract away some information computed during driving.
+In conjunctive partial deduction generalization is modified to support treating of conjunctions.
+The generalization selects subconjuctions of two conjuncts which are similar (call to the same relation and their arguments have similar shape and distribution).
+For the subconjunctions selected a most-specific generalization is computed.
+
+In our approach we only do splitting of a conjunction into individual relation calls.
+This makes any program with an accumulating parameter to be a problem.
+Sometimes when there is a need to do a proper generalization, it is in reality just an instance of some other goal within the tree and we can simply create a call there.
+Otherwise we are unable to meaningfully specialize such goal, but we can always just include the initial program in the residual program and call the corresponding relation.
 
@@
 
-Unfolding is when a call to a relation is substituted by its body.
+Unfolding is a process of substitution of some relation call by its body with simultaneous computation of unifications.
+To unfold a relation call we do the following steps.
+First, the formal arguments of a relation are substituted for the actual arguments of the call in the body.
+All fresh variables get instantiated.
+The body is transformed into a canonical form (disjunction of conjunctions of either calls or unifications).
+All unifications are computed.
+Those disjuncts in which unifications fails are removed.
+Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
 
 The most important question is when to unfold.
-
-To much unfolding is sometimes is even worse than not enough unfolding.
+Unfortunately, too much unfolding is sometimes is even worse than not enough unfolding.
+There is a fine edge between those.
+This problem is mentioned in the (CONTROL PAPER).
+We believe that the following heuristic provides a reasonable control.
 
@@
 
+The intuition behind the heuristic is to find those calls which are safe to unfold.
+We deem every static conjunct (non-recursive) to be safe because they never lead to growth in the number of conjunctions.
+Those calls which unfold deterministically, meaning there is only one disjunct in the unfolded relation, are also considered to be safe.
+
+The other more complicated case is when there are less disjuncts than there can possibly be.
+This signifies that at least one branch of computations is gotten rid of.
+
+The final heuristic selects the first conjunct which suites either of the following cases.
 First we unfold those conjuncts which are static.
+Then --- deterministic.
+Then those which are less branching.
+The last to be unfolded are those calls, which unfold to a substitution with not conjunction.
 
-Then --- deterministic.
+\subsection{Residualization}
 
-Then those which are less branching.
+Residualization is quite straightforward.
+A branching in the process tree becomes a disjunction.
+A split node becomes a conjunction.
+Computed substitution is residualized as a conjunction of unifications.
+A renaming node is just a call to a relation.
+Relations are created for configurations on which leaf nodes are renamed.
 
-The last to be unfolded are those calls, which unfold to a substitution.
+One other thing is that when some configuration is occurred within the tree which is an instance of a configuration for which a new relation is created, then we just create a call.

### 2020-05-22 papers/miniKanren/text/evaluation.tex
WIP

@@
 
-In the evaluation we compared the performance of our CPD, Ecce and the new non-conjunctive partial evaluator.
-We have also implemented the branching heuristic instead of the deterministic one in the CPD, just to be sure, it is not a cureall.
+In our study we compared the performance of our implementation of CPD, ECCE and the new non-conjunctive partial evaluator.
+We have also implemented the branching heuristic instead of the deterministic one in the CPD, just to ensure, it is not a cureall.
 
-We used the following 4 programs to test our specializers on.
+We used the following 4 programs to test the specializers on.
 \begin{itemize}
-  \item propositional evaluator
-  \begin{itemize}
-    \item Compicated bool expression in the last conjunctive
-    \item Table implementation of the boolean operation, placed as a first conjunct
-  \end{itemize}
-  \item searching for a unifier of two terms
-  \item searching for paths of a specific length in a graph
+  \item Two implementations of an evaluator of logic formulas.
+  \item A program to compute a unifier of two terms.
+  \item A program to search for paths of a specific length in a graph
 \end{itemize}
 
-The first one demonstrates the drawback of the approach implemented in ECCE: left-to-right deterministic unfolding.
-The boolean expression which narrows the first conjunts is the last, so ECCE does not unfold it until too late.
+All these relations are relational interpreters.
+Their last argument is a boolean value which is $true$ if the other arguments are in relation and $false$ otherwise.
 
-The second one should be easier for this approach.
+\subsection{Evaluator of Logic Formulas}
 
-The unification is a difficult problem, specialization of which is of importance.
+The relation $eval^o$ describes evaluation of a subset of first-order logic formulas in a given substitution.
+$eval^o$ has 3 arguments.
+The first argument is a list of boolean values which serves as a substitution.
+The $i$-th value of the list is the value of the $i$-th variable.
+
+The second argument is a formula with the following abstract syntax.
+A formula is either a \emph{variable} represented with a Peano number, a \emph{negation} of a formula, a \emph{conjunction} of two formulas or a \emph{disjunction} of two formulas.
+
+The third argument is a boolean value which is $true$ if the given formula in the given substition is true and $false$ otherwise.
+
+One possible implementation of the evaluator in the syntax of \oc{} is presented in listing~\ref{eval:last}.
+Here the relation \lstinline{elem$^o subst v res$} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
+The relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} encode corresponding boolean operations.
+
+\begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b]
+  let rec eval$^o$_last subst fm res = conde [
+    fresh (x y z v w) (
+      (fm === var v /\ elem$^o$ subst v res);
+      (fm === conj x y /\ eval$^o$_last st x v /\ eval$^o$_last st y w /\ and$^o$ v w res);
+      (fm === disj x y /\ eval$^o$_last st x v /\ eval$^o$_last st y w /\ or$^o$ v w res);
+      (fm === neg x /\ eval$^o$_last st x v /\ not$^o$ v res))]
+\end{lstlisting}
+
+Note, that the calls to boolean relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} are placed last within each conjunction.
+This poses a challenge to the CPD-based specializers.
+Conjunctive partial deduction unfolds relation calls from left to right, so when specializing this relation for running backwards (i.e.
+considering the goal \lstinline{eval$^o$ subst fm ^true}), it fails to propagate the direction data onto recursive calls of \lstinline{eval$^o$}.
+It leads to over-unfolding, big residual programs and less than optimal performance.
+
+Moving boolean operations to the left, as shown in listing~\ref{eval:fst}, we get a program which is easier for CPD to specialize.
+
+\begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b]
+  let rec eval$^o$_plain subst fm res = conde [
+    fresh (x y z v w) (
+      (fm === var v /\ elem$^o$ subst v res);
+      (fm === conj x y /\ and$^o$_table v w res /\ eval$^o$_plain st x v /\ eval$^o$_plain st y w);
+      (fm === disj x y /\ or$^o$_table v w res /\ eval$^o$_plain st x v /\ eval$^o$_plain st y w);
+      (fm === neg x /\ not$^o$_table v res /\ eval$^o$_plain st x v))]
+  \end{lstlisting}
+
+The next complication for partial deduction is how much sequential unfoldings should be performed to  restrict other calls in a conjunction in a meaningful way.
+In this example the way in which boolean relations are implemented affect the specialization result.
+The simplest way to implement these relations is with a table as shown in listing~\ref{not:table}
+
+\begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b]
+  let not$^o$_table x y = conde [
+     (x === ^true /\ y === ^false;
+      x === ^false /\ y === ^true)]
+\end{lstlisting}
+
+The other way to implement boolean operations is via one base boolean relation such as \lstinline{nand$^o$} which is in turn has a table-based implementation.
+
+\begin{lstlisting}[label={not:table}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b]
+  let not$^o$ x y = nand$^o$ x x y
+
+  let or$^o$ x y z = nand$^o$ x x xx /\ nand$^o$ y y yy /\ nand$^o$ xx yy z
+
+  let and$^o$ x y z = nand$^o$ x y xy /\ nand$^o$ xy xy z
+
+  let nand$^o$ a b c = conde [
+    ( a === ^false /\ b === ^false /\ c === ^true );
+    ( a === ^false /\ b === ^true /\ c === ^true );
+    ( a === ^true /\ b === ^false /\ c === ^true );
+    ( a === ^true /\ b === ^true /\ c === ^false )]
+\end{lstlisting}
+
+We considered two implementations of \lstinline{eval$^o$}: \lstinline{eval$^o$_plain} and \lstinline{eval$^o$_last} and studied how specializers behave on them.
+The first one uses table-based boolean operations and places them at the second place in each conjunction.
+The relation \lstinline{eval$^o$_last} employs boolean operations implemented via \lstinline{nand$^o$} and place them at the end of each conjunction.
+These two programs are complete opposites from the standpoint of CPD which the time measurements in table~\ref{tbl:eval} confirmed.
+
+We measured time necessary to generate $10000$ formulas over two variables which evaluate to \lstinline{^true}.
+We compared the result of specialization of the goal \lstinline{eval$^o$ subst fm ^true} by our implementation of CPD, done by ECCE system and the new non-conjunctive partial deduction.
+Since ECCE cannot work directly on \mk{} programs, we first translated them to \pro{} and then back: this translation is purely syntactical.
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c}
+                   & last & plain \\
+  \hline\hline
+  Original         & >60.00s & >60.00s \\
+  \hline
+  Ecce             & 7.12s  & 9.22s \\
+  \hline
+  CPD              & 31.31s & 5.46s \\
+  \hline
+  Non-CPD          & 4.99s  & 5.05s \\
+  \hline
+  Branches         & 17.21s  & 6.17s \\
+  \hline
+  \end{tabular}
+
+  \caption{Searching for 10000 true logic formulas}
+  \label{tbl:eval}
+\end{table}
+
+\subsection{Search for a Unifier}
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c}
+
+    & running time \\
+  \hline\hline
+  Original         & >300.00s \\
+  \hline
+  CPD              & 2.35s \\
+  \hline
+  Non-CPD          & 14.90s \\
+  \hline
+  Branches         & >300.00s \\
+  \hline
+  \end{tabular}
+
+  \caption{Searching for a unifier for terms f(X, X, g(Z, t)) and f(g(p, L), Y, Y)}
+  \label{tbl:unify}
+\end{table}
+
+The unification of two terms $t$ and $u$ is searching for a substitution $\theta$ such that $t \theta = u \theta$, $\theta$ is called a unifier.
+We search for any unifier, not necessarily most specific.
+The details on this benchmark can be found in the earlier paper CITE.
+This example demonstrates how too much unfolding can be introduced with the non-conjunctive partial deduction.
+
+\subsection{Search for Paths in a Graph}
+
+Here we search for 5 paths in a graph.
+The details on this benchmark can be found in the earlier paper CITE.
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c}
+
+    & running time \\
+  \hline\hline
+  Original         & 19.86s \\
+  \hline
+  CPD              & 4.66s \\
+  \hline
+  Non-CPD          & 3.00s \\
+  \hline
+  \end{tabular}
+
+  \caption{Searching for paths in a graph}
+  \label{tbl:unify}
+\end{table}

### 2020-05-22 papers/miniKanren/text/pd.tex
WIP

\section{Conjunctive Partial Deduction}

Partial deduction is a specialization technique for logic languages.

Conjunctive partial deduction is a modification of it, in which conjunctions of atoms are considered as a whole.

An adaptation of conjunctive partial deduction to \mk{} was described in an earlier paper.

Conjunctive partial deduction has its problems.
It is designed for \pro{} which fixes the order in which atoms are evaluated.
While driving, CPD consideres atoms from left to right which leads to the following problem.
If an atom which can restrict the answer set for other atoms is last in the conjunction, then before it gets unfolded, CPD would unfold all the others which are calls for free variables.
This leads to over-unfolding.

\mk{} is a language in which the set of answers does not depend on the order of calls within conjunction.
The order only affects the running time.
It would be lovely, if the specialization technique for \mk{} chooses the best order of conjuncts.
By blindly implementing CPD, we fail to do so.
This is why we created a novel specialization approach which we called non-conjunctive partial deduction.

### 2020-05-22 papers/miniKanren/text/abstract.tex
Introduction

\begin{abstract}
  We explore partial deduction for \mk{}: a specialization technique aimed at improving the performance of a relation in the given direction.
  We describe a novel approach to specialization of \mk{} based on partial deduction and supercompilation.
  On several examples, we demonstrate issues which arise during partial deduction.
\end{abstract}

### 2020-05-22 papers/miniKanren/text/related.tex
Introduction

--- 2020-05-21 papers/miniKanren/text/intro.tex
@@
-\section{Introduction}
-
-The \mk{} language family is nice.
-
-Unpredictable running time for different directions.
-
-Specialization sometimes helps, sometimes does not, sometimes makes everything worse.
-
-Control issues, blah blah blah
-
-This problem also appear in supercompilation (partial evaluation, specialization) of functional and imperative programming languages.
+\section{Related Works}
 
@@
 Anyway, the formalization problem of cases for useful transformers application is poorly studied.
-
-<Here should be review of problems in specialization of functional languages>
-
-We came up with the new specialization algorithm, which seems nice, but is as unpredictable as the others.
-
-Here are some numbers...
-
-Contribution is the following
-\begin{itemize}
-  \item explanation of specialization
-  \item new specialization algorithm
-  \item why some programs behave worse after specialization
-\end{itemize}

### 2020-05-22 papers/miniKanren/text/intro.tex

--- 2020-05-21 papers/miniKanren/text/intro.tex
@@
 
-The \mk{} language family is nice.
+The core feature of the family of relational programming languages \mk{}\footnote{\mk{} language web site: \url{http://minikanren.org}} is their ability to run a program in different directions.
+Having specified a relation for adding two numbers, one can also compute the substraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
+By running a relational interpreter \emph{backwards} one can obtain a \emph{solver}~\cite{lozov2019relational} thus solving a much more complicated problem.
 
-Unpredictable running time for different directions.
+The search employed in \mk{} is complete which means that every answer will be found, although it may take a long time.
+The promise of \mk{} falls short when speaking of performance.
+The running time of a program in \mk{} is highly unpredictable and varies greatly for different directions.
+What is even worse, it depends on the order of the relation calls within a program.
+One order can be good for one direction, but slow the computation drastically in the different direction.
 
-Specialization sometimes helps, sometimes does not, sometimes makes everything worse.
+Specialization or partial evaluation~\cite{jonesbook} is a technique aimed at improving performance of a program given some information about it beforehand.
+It may either be a known value of some argument, its structure (i.e. the length of an input list) or, in case of relational program, --- the direction in which it is intended to be run.
+An earlier paper~\cite{lozov2019relational} showed that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk{} programs.
+Unfortunately, it may also not affect the running time of a program or even make it slower.
 
-Control issues, blah blah blah
+Control issues in partial deduction of logic programming language \pro{} have been studied before~\cite{leuschel2002logic}.
+The ideas described there are aimed at left-to-right evaluation strategy of \pro{}.
+Since the search in \mk{} is complete, it is safe to reorder come relation calls within the goal for better performance.
+While sometimes conjunctive partial deduction gives great perforamnce boost, sometimes it does not behave as well as it could have been.
 
-This problem also appear in supercompilation (partial evaluation, specialization) of functional and imperative programming languages.
-
-<Here should be review of problems in specialization of functional languages>
-
-We came up with the new specialization algorithm, which seems nice, but is as unpredictable as the others.
-
-Here are some numbers...
-
-Contribution is the following
-\begin{itemize}
-  \item explanation of specialization
-  \item new specialization algorithm
-  \item why some programs behave worse after specialization
-\end{itemize}
+In this paper we show on examples some issues which face conjunctive partial deduction.
+We also describe a novel approach to partial deduction of a relational programming language \mk{}.
+We compare it to the existing specialization algorithms on several programs and discuss why some \mk{} programs run slower after specialization.

### 2020-05-22 papers/miniKanren/text/related.tex
Related work

--- 2020-05-21 papers/miniKanren/text/intro.tex
@@
-\section{Introduction}
+\section{Related Work}
 
-The \mk{} language family is nice.
+Specialization is an attractive technique aimed to improve the performance of a program if some of its arguments are known statically.
+It is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation} and many more.
+While supercompilation generally improves the behavior of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in general case.
 
-Unpredictable running time for different directions.
+What is worse they rarely consider the residual program efficiency from the point of view of the target language evaluator.
+The main optimization source is computing in advance all possible intermediate and statically-known semantics steps at transformation-time.
+Other criteria like the size of the generated program or possible optimizations and execution cost of different language constructions by the target language evaluator are usually out of consideration~\cite{jonesbook}.
+It is known that supercompilation may adversely affect GHC optimizations yielding standalone compilation more powerful~\cite{SCBE,TCES} and cause code explosion~\cite{SCHC}.
+Moreover, it may be hard to predict the real speedup of any given program on concrete examples even disregarding problems above because of the complexity of the transformation algorithm.
+The worst case for partial evaluation is when all static variables are used in a dynamic context, and there is a lot of advice on how to implement a partial evaluator as well as a target program so that specialization really improved its performance~\cite{jonesbook,bulyonkov84}.
+There is lack of research in determining the classes of programs which transformers would definitely speed up.
 
-Specialization sometimes helps, sometimes does not, sometimes makes everything worse.
+Conjunctive partial deduction makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro{}.
+CPD constructs a tree which models goal evaluation and resembles SLD-NF tree.
+After the tree is constructed, a residual program is generated from it.
+The specialization is done in two levels of control: the local control determines the shape of the residual programs, while the global control ensures that every relation which can be called in the residual program is indeed defined.
+The leaves of local control trees become nodes of the global control tree.
+CPD analyses these nodes at the global level and runs local control for all those which are new.
 
-Control issues, blah blah blah
+At the local level CPD examines a conjunction of atoms by considering each atom one-by-one from left to right.
+An atom is \emph{unfolded} if it is deemed safe.
+When an atom is unfolded, a clause which head can be unified with the atom is found, and a new node is added into the tree where the atom in the conjunction is replaced with the body of that clause.
+If there are more than one suitable head, then several branches are added into the tree which correspond to the disjunction in the residualized program.
+An adaptation of CPD for the \mk{} programming language is described in~\cite{lozov2019relational}.
 
-This problem also appear in supercompilation (partial evaluation, specialization) of functional and imperative programming languages.
-
-<Here should be review of problems in specialization of functional languages>
-
-We came up with the new specialization algorithm, which seems nice, but is as unpredictable as the others.
-
-Here are some numbers...
-
-Contribution is the following
-\begin{itemize}
-  \item explanation of specialization
-  \item new specialization algorithm
-  \item why some programs behave worse after specialization
-\end{itemize}
+The most well-behaved strategy of local control in CPD for \pro{} is \emph{deterministic unfolding}.
+An atom is unfolded only if only one suitable clause head exist for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
+This means that if a non-deterministic atom is the left-most within a conjunction, it is most likely to be unfoldled and introduce a lot of
+We believe this is the core problem with CPD which limits its power when applied to \mk{}.
+The strategy of unfolding atoms from left to right is reasonable in the context of \pro{} because it mimics the way it executes.
+But it often leads to larger global control trees and, as a result, bigger less efficient programs.
+The evaluation result of a \mk{} program does not depend on the order of atoms (relation calls) within a conjunction, thus we believe a better result can be achieved by selecting a relation call which can restrict the number of branches in the tree.
+We describe our approach which implements this idea in the next section.

### 2020-05-22 papers/miniKanren/text/nonconj.tex
add pseudocode algo-1
add some refs to fig 1

@@
 \section{Non-conjunctive Partial Deduction}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.7\textwidth]{algo_pseudo-crop.pdf}
+  \caption{\db{Non-conjunctive Partial Deduction Pseudo Code}}
+  \label{fig:ncpd-pseudo}
+\end{figure}
 
@@
 The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.
+\db{The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}\footnote{
+    \db{For the sake of breavity and clarity, we provide functions \texttt{drive\_disj} and \texttt{drive\_conj} describing driving processes for disjunction and conjunction correspondly.
+Driving itself if a trivial combination of presented functions (line 2)}}}.
 
-The high-level idea behind the algorithm is to select a relation call to unfold by using a heuristic which decides if the call can narrow down the answer set.
-A driving tree is constructed for the selected call in isolation.
+The high-level idea behind the algorithm is to select a relation call to unfold by using a heuristic which decides if the call can narrow down the answer set \db{(line 15)}.
+A driving tree is constructed for the selected call in isolation \db{(line 16)}.
 The leaves of the computed tree are examined.
-If all leaves are either computed substitutions or are calls to some relations accompanied with non-empty substitutions, then the leaves are collected and are put back into the root conjunction instead of the examined call.
+If all leaves are either computed substitutions or are calls to some relations accompanied with non-empty substitutions, then the leaves are collected and are put back into the root conjunction instead of the examined call \db{(lines 19,20)}.
 According to denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.

### 2020-05-22 papers/miniKanren/text/nonconj.tex
add refs to lines of pseudo code

@@
 \db{The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}\footnote{
-    \db{For the sake of breavity and clarity, we provide functions \texttt{drive\_disj} and \texttt{drive\_conj} describing driving processes for disjunction and conjunction correspondly.
-Driving itself if a trivial combination of presented functions (line 2)}}}.
+    \db{For the sake of brevity and clarity, we provide functions \texttt{drive\_disj} and \texttt{drive\_conj} describing driving processes for disjunction and conjunction correspondingly.
+      Driving itself is a trivial combination of presented functions (line 2)}}}.
 
@@
 Each time we examine a conjunction of calls, we \emph{split} them into separate nodes which are driven independently from each other.
-Among the relation calls we select one which is according to the heuristic is likely to narrow down the answer set.
-If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected.
+Among the relation calls we select one which is according to the heuristic is likely to narrow down the answer set \db{(line 15)}.
+If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected \db{(line 22)}.
 
-This process does create branchings whenever a disjunction is examined.
+This process does create branchings whenever a disjunction is examined \db{(lines 4--6)}.
 At each step we make sure that we do not start driving a conjunction which we have already examined.
-To do this, we check if the current conjunction is renaming of any other configuration in the tree.
+To do this, we check if the current conjunction is renaming of any other configuration in the tree \db{(line 12)}.
 If it is, then we create a special node which then is residualized into a call to the suitable relation.
@@
 This makes any program with an accumulating parameter to be a problem.
-Sometimes when there is a need to do a proper generalization, it is in reality just an instance of some other goal within the tree and we can simply create a call there.
+Sometimes when there is a need to do a proper generalization, it is in reality just an instance of some other goal within the tree and we can simply create a call there \db{(line 13)}.
 Otherwise we are unable to meaningfully specialize such goal, but we can always just include the initial program in the residual program and call the corresponding relation.

### 2020-05-22 papers/miniKanren/text/conclusion.tex
Evaluation and conclusion

@@
 
-We compared some of the specialization techniques for \mk{}.
-
-There seem to be no one good technique.
+In this paper we discussed some issues which arise in partial deduction of a relational programming language \mk{}.
+We presented a novel approach to partial deduction which uses a heuristic to select the most suitable relation call to unfold at each step of driving.
+We compared this approach to the earlier implementation of conjunctive partial deduction.
+Our evaluation showed that there is still not one good technique which definitively speeds up every relational program.

### 2020-05-22 papers/miniKanren/text/evaluation.tex
Evaluation and conclusion

--- 2020-05-21 papers/miniKanren/text/evaluation.tex
@@
 
-In the evaluation we compared the performance of our CPD, Ecce and the new non-conjunctive partial evaluator.
-We have also implemented the branching heuristic instead of the deterministic one in the CPD, just to be sure, it is not a cureall.
+In our study we compared the CPD adaptation for \mk{} and the new non-conjunctive partial deduction.
+We have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check wether it can improve the quality of the specialization.
 
-We used the following 4 programs to test our specializers on.
+We used the following 4 programs to test the specializers on.
 \begin{itemize}
-  \item propositional evaluator
-  \begin{itemize}
-    \item Compicated bool expression in the last conjunctive
-    \item Table implementation of the boolean operation, placed as a first conjunct
-  \end{itemize}
-  \item searching for a unifier of two terms
-  \item searching for paths of a specific length in a graph
+  \item Two implementations of an evaluator of logic formulas.
+  \item A program to compute a unifier of two terms.
+  \item A program to search for paths of a specific length in a graph.
 \end{itemize}
 
-The first one demonstrates the drawback of the approach implemented in ECCE: left-to-right deterministic unfolding.
-The boolean expression which narrows the first conjunts is the last, so ECCE does not unfold it until too late.
+The last two relations are described in~\cite{lozov2019relational} thus we will not describe them here.
 
-The second one should be easier for this approach.
+% All these relations are relational interpreters.
+% Their last argument is a boolean value which is $true$ if the other arguments are in relation and $false$ otherwise.
 
-The unification is a difficult problem, specialization of which is of importance.
+\subsection{Evaluator of Logic Formulas}
+
+The relation \lstinline{eval$^o$} describes evaluation of a subset of first-order logic formulas in a given substitution.
+It has 3 arguments.
+The first argument is a list of boolean values which serves as a substitution.
+The $i$-th value of the substitution is the value of the $i$-th variable.
+The second argument is a formula with the following abstract syntax.
+A formula is either a \emph{variable} represented with a Peano number, a \emph{negation} of a formula, a \emph{conjunction} of two formulas or a \emph{disjunction} of two formulas.
+The third argument is the value of the formula in the given substitution.
+
+In this paper we specialize the relation to synthesize formulas which evaluate to \lstinline{^true}.
+To do so we run the specializer for the goal with the last argument fixed to \lstinline{^true}, while the first two arguments remain free variables.
+Depending on the way the \lstinline{eval$^o$} is implemented, different specializers generate significantly different residual programs.
+
+\subsubsection{The Order of Relation Calls}
+
+One possible implementation of the evaluator in the syntax of \oc{} is presented in listing~\ref{eval:last}.
+In this implementation the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
+The relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} encode corresponding boolean operations.
+
+\begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b]
+  let rec eval$^o$ subst fm res = conde [
+    fresh (x y z v w) (
+      (fm === var v /\ elem$^o$ subst v res);
+      (fm === conj x y /\ eval$^o$ st x v /\ eval$^o$ st y w /\ and$^o$ v w res);
+      (fm === disj x y /\ eval$^o$ st x v /\ eval$^o$ st y w /\ or$^o$ v w res);
+      (fm === neg x /\ eval$^o$ st x v /\ not$^o$ v res))]
+\end{lstlisting}
+
+Note, that the calls to boolean relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} are placed last within each conjunction.
+This poses a challenge to the CPD-based specializers.
+Conjunctive partial deduction unfolds relation calls from left to right, so when specializing this relation for running backwards (i.e.
+considering the goal \lstinline{eval$^o$ subst fm ^true}), it fails to propagate the direction data onto recursive calls of \lstinline{eval$^o$}.
+Knowing that \lstinline{res} is \lstinline{^true}, we can conclude that in the call \lstinline{and$^o$ v w res} variables \lstinline{v} and \lstinline{w} have to be \lstinline{^true} as well.
+There are three possible options for these variables in the call \lstinline{or$^o$ v w res} and one for the call \lstinline{not$^o$}.
+These variables are used in recursive calls of \lstinline{eval$^o$} and thus restrict the result of driving them.
+CPD fails to recognize this, and thus unfold recursive calls of \lstinline{eval$^o$} applied to fresh variables.
+It leads to over-unfolding, big residual programs and less than optimal performance.
+
+The non-conjunctive partial deduction will first unfold those calls which are selected with the heuristic.
+Since exploring boolean operations makes more sense, they are unfolded before recursive calls of \lstinline{eval$^o$}.
+The way non-conjunctive partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in listing~\ref{eval:fst}.
+This program is easier for CPD to transform which demonstrates how unequal is the behavior of CPD for the similar programs.
+
+\begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b]
+  let rec eval$^o$ subst fm res = conde [
+    fresh (x y z v w) (
+      (fm === var v /\ elem$^o$ subst v res);
+      (fm === conj x y /\ and$^o$_table v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
+      (fm === disj x y /\ or$^o$_table v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
+      (fm === neg x /\ not$^o$_table v res /\ eval$^o$ st x v))]
+  \end{lstlisting}
+
+\subsubsection{Unfolding of Complex Relations}
+
+Depending on the way a relation is implemented, it may take different number of driving steps to reach the point when any useful information is derived through its unfolding.
+Partial deduction tries to unfold every relation call unless it is unsafe, but not all relation calls serve to restrict the search space and thus should be unfolded.
+In the implementation of \lstinline{eval$^o$} boolean operations can effectively restrict variables within the conjunctions and should be unfolded until they do.
+But depending on the way they are implemented, different number of driving steps should be performed to do so.
+The simplest way to implement these relations is with a table as demonstrated with the implementation of \lstinline{not$^o$} in listing~\ref{not:table}.
+It is enough to unfold such relation calls once to derive useful information about variables.
+
+\begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b]
+  let not$^o$_table x y = conde [
+     (x === ^true /\ y === ^false;
+      x === ^false /\ y === ^true)]
+\end{lstlisting}
+
+The other way to implement boolean operations is via one basic boolean relation such as \lstinline{nand$^o$} which is in turn has a table-based implementation (see listing~\ref{not:nando}).
+It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
+Non-conjunctive partial deduction drives the selected call until it derives useful substitutions for the variables involved while CPD with deterministic unfolding may fail to do so.
+
+\begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b]
+  let not$^o$ x y = nand$^o$ x x y
+
+  let or$^o$ x y z = nand$^o$ x x xx /\ nand$^o$ y y yy /\ nand$^o$ xx yy z
+
+  let and$^o$ x y z = nand$^o$ x y xy /\ nand$^o$ xy xy z
+
+  let nand$^o$ a b c = conde [
+    ( a === ^false /\ b === ^false /\ c === ^true );
+    ( a === ^false /\ b === ^true /\ c === ^true );
+    ( a === ^true /\ b === ^false /\ c === ^true );
+    ( a === ^true /\ b === ^true /\ c === ^false )]
+\end{lstlisting}
+
+\subsection{Evaluation Results}
+
+In our study we considered two implementations of \lstinline{eval$^o$}: one we call \lstinline{plain} and the other --- \lstinline{last} and compared how specializers behave on them.
+The \lstinline{plain} relation uses table-based boolean operations and places them further to the left in each conjunction.
+The relation \lstinline{last} employs boolean operations implemented via \lstinline{nand$^o$} and place them at the end of each conjunction.
+These two programs are complete opposites from the standpoint of CPD.
+
+We measured time necessary to generate $10000$ formulas over two variables which evaluate to \lstinline{^true}.
+We compared the results of specialization of the goal \lstinline{eval$^o$ subst fm ^true} by our implementation of CPD, the new non-conjunctive partial deduction, and the CPD modified with the branching heuristic.
+Our evaluation confirmed that CPD behaves very differently on these two implementations of the same relation: the running time of the specialized \lstinline{last} relation is almost 6 times as high as the running time of the specialized \lstinline{plain} relation.
+The running time of two programs generated with our novel non-conjunctive partial deduction is very close and it is a little bit better than the best by CPD.
+CPD with the branching heuristic still gives different quality transformations.
+The results are shown in table~\ref{tbl:eval}.
+
+Besides evaluator of logic formulas we also run the transformers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
+These two relations are described in paper~\cite{lozov2019relational} so we will not go into too much details here.
+
+The \lstinline{unify} relation was executed to find 5 unifiers of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
+The original \mk{} program fails to terminate on this goal in 300 seconds.
+On this example the most performant is the program generated by CPD (2.35 seconds) while the program generated by adding branching heuristic also fails to terminate in 300 seconds.
+The non-conjunctive partial deduction shows some improvement with the residual program terminated within 15 seconds.
+While driving this program, non-conjunctive partial deduction does too much unfolding which negatively impact the running time as compared to CPD.
+
+The last test executed \lstinline{isPath} relation to search for 5 paths in the graph with 20 vertices and 30 edges.
+On this program non-conjunctive partial deduction showed better transformation results than CPD, although the difference is not that drastic.
+
+All evaluation results are presented in the table~\ref{tbl:eval}.
+Each column correspond to the relation being run as described above.
+The row marked ``Original'' contains the running time of the original \mk{} relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
+
+\begin{table}
+  \centering
+  \begin{tabular}{c|c|c||c||c}
+                   & last & plain & unify & isPath \\
+  \hline\hline
+  Original         & >60.00s & >60.00s & >300.00s & 19.86s \\
+  \hline
+  CPD              & 31.31s & 5.46s & 2.35s & 4.66s \\
+  \hline
+  Non-CPD          & 4.99s  & 5.05s & 14.90s & 3.00s \\
+  \hline
+  Branching        & 17.21s  & 6.17s & >300.00s & N/A \\
+  \hline
+  \end{tabular}
+
+  \caption{Evaluation results}
+  \label{tbl:eval}
+\end{table}
+
+% \subsection{Search for a Unifier}
+
+% % \begin{table}
+% %   \centering
+% %   \begin{tabular}{c|c}
+
+% %     & running time \\
+% %   \hline\hline
+% %   Original         & >300.00s \\
+% %   \hline
+% %   CPD              & 2.35s \\
+% %   \hline
+% %   Non-CPD          & 14.90s \\
+% %   \hline
+% %   Branches         & >300.00s \\
+% %   \hline
+% %   \end{tabular}
+
+% %   \caption{Searching for a unifier for terms f(X, X, g(Z, t)) and f(g(p, L), Y, Y)}
+% %   \label{tbl:unify}
+% % \end{table}
+
+% The unification of two terms $t$ and $u$ is searching for a substitution $\theta$ such that $t \theta = u \theta$, $\theta$ is called a unifier.
+% We search for any unifier, not necessarily most specific.
+% The details on this benchmark can be found in the earlier paper CITE.
+% This example demonstrates how too much unfolding can be introduced with the non-conjunctive partial deduction.
+
+% \subsection{Search for Paths in a Graph}
+
+% Here we search for 5 paths in a graph.
+% The details on this benchmark can be found in the earlier paper CITE.
+
+% % \begin{table}
+% %   \centering
+% %   \begin{tabular}{c|c}
+
+% %     & running time \\
+% %   \hline\hline
+% %   Original         & 19.86s \\
+% %   \hline
+% %   CPD              & 4.66s \\
+% %   \hline
+% %   Non-CPD          & 3.00s \\
+% %   \hline
+% %   \end{tabular}
+
+% %   \caption{Searching for paths in a graph}
+% %   \label{tbl:unify}
+% % \end{table}

### 2020-05-22 papers/miniKanren/text/nonconj.tex

@@
+\newcommand{\code}[1]{\texttt{#1}}
+
 \section{Non-conjunctive Partial Deduction}
@@
 \db{The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}\footnote{
-    \db{For the sake of brevity and clarity, we provide functions \texttt{drive\_disj} and \texttt{drive\_conj} describing driving processes for disjunction and conjunction correspondingly.
-      Driving itself is a trivial combination of presented functions (line 2)}}}.
+    \db{For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} describing driving processes for disjunction and conjunction correspondingly.
+      Driving itself is a trivial combination of presented functions (line 2).}}}.
+
+\db{
+Hereafter, we consider all goals and relation bodies to be in \emph{canonical normal form}~---~a disjunction of conjunctions of either calls or unifications.
+Moreover, we assume all fresh variables to be instantiated and all unifications to be computed on each step.
+Those disjuncts in which unifications fails are removed.
+Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
+Any miniKanren term can be trivially transformed to the described form\footnote{
+  \db{On Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization;
+    The code is ommited for brevity.}
+}.}
 
@@
 To unfold a relation call we do the following steps.
+\db{TODO: the following has alredy mentioned above.}
 First, the formal arguments of a relation are substituted for the actual arguments of the call in the body.

### 2020-05-22 papers/miniKanren/text/evaluation.tex
align and wrap in figure* all lstlistings

@@
 
-\begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b]
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.8\textwidth}
+    \begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b, frame=tb]
   let rec eval$^o$ subst fm res = conde [
@@
     \end{lstlisting}
+  \end{minipage}
+\end{figure*}
 
@@
 
-\begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b]
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.85\textwidth}
+    \begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b, frame=tb]
   let rec eval$^o$ subst fm res = conde [
@@
 
-\begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b]
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.45\textwidth}
+    \begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b, frame=tb]
   let not$^o$_table x y = conde [
@@
 
-\begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b]
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.7\textwidth}
+    \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b, frame=tb]
   let not$^o$ x y = nand$^o$ x x y

### 2020-05-22 papers/miniKanren/text/nonconj.tex
propose a small extension for Unfolding subsection

@@
 The most important question is when to unfold.
-Unfortunately, too much unfolding is sometimes is even worse than not enough unfolding.
+\db{
+  {\color{blue}PROPOSED EXTENSION:}, and when to stop unfolding even if the relation call seems to be good for specialization.
+  Usually, unfolding in functional programming languages specialization, as well as inlining in imperative one, is considered to be safe from a residual program efficiency point of view.
+  It may only lead to code explosion or code duplication which is mostly left to a target program compiler optimization or even is out of consideration at all if specializer is considered as a standalone tool~\cite{jonesbook}.
+  Unfortunately, this is not the case for the specialization of relational programming language.
+  Unlike functional and imperative, in relational programming language unfolding may easily affect the target program efficiency.
+  It is crucial not to unfold too much since, for example, extra unifications or even repeated computations generated by optimized relation calls and further propagated to neighbouring conjunction may appear.
+}
+\db{{\color{blue}Unfortunately $\to$ }That is}, too much unfolding is sometimes is even worse than not enough unfolding.
 There is a fine edge between those.

### 2020-05-23 papers/miniKanren/text/nonconj.tex
Nonconj

@@
 
-In this section we will describe a novel approach to specialization of relational programs.
+In this section we describe a novel approach to specialization of relational programs.
 This approach draws inspiration from both conjunctive partial deduction and supercompilation.
 The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.
-\db{The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}\footnote{
-    \db{For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} describing driving processes for disjunction and conjunction correspondingly.
-      Driving itself is a trivial combination of presented functions (line 2).}}}.
 
-\db{
-Hereafter, we consider all goals and relation bodies to be in \emph{canonical normal form}~---~a disjunction of conjunctions of either calls or unifications.
-Moreover, we assume all fresh variables to be instantiated and all unifications to be computed on each step.
-Those disjuncts in which unifications fails are removed.
-Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
-Any miniKanren term can be trivially transformed to the described form\footnote{
-  \db{On Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization;
-    The code is ommited for brevity.}
-}.}
-
-The high-level idea behind the algorithm is to select a relation call to unfold by using a heuristic which decides if the call can narrow down the answer set \db{(line 15)}.
-A driving tree is constructed for the selected call in isolation \db{(line 16)}.
-The leaves of the computed tree are examined.
-If all leaves are either computed substitutions or are calls to some relations accompanied with non-empty substitutions, then the leaves are collected and are put back into the root conjunction instead of the examined call \db{(lines 19,20)}.
-According to denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
+The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}.
+For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} which describe how to process disjunctions and conjunctions respectively.
+Driving itself is a trivial combination of presented functions (line 2).
 
 A driving process creates a process tree, from which a residual program is later created.
+The process tree is meant to mimic the execution of the input program.
 The nodes of process tree include a \emph{configuraion} which describes the state of program evaluation at some point.
@@
 
-Each time we examine a conjunction of calls, we \emph{split} them into separate nodes which are driven independently from each other.
-Among the relation calls we select one which is according to the heuristic is likely to narrow down the answer set \db{(line 15)}.
-If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected \db{(line 22)}.
+Hereafter, we consider all goals and relation bodies to be in \emph{canonical normal form}~---~a disjunction of conjunctions of either calls or unifications.
+Moreover, we assume all fresh variables to be introduced into scope and all unifications to be computed at each step.
+Those disjuncts in which unifications fail are removed.
+Other disjuncts take form of a conjunction of relation calls accompanied with a substitution computed from unificaitons.
+Any miniKanren term can be trivially transformed to the described form.
+In Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization.
+The code is omitted for brevity.
 
-This process does create branchings whenever a disjunction is examined \db{(lines 4--6)}.
+% The very first step of driving a conjunction is to apply a substitution to the variables in relation calls (line 10).
+
+There are several core ideas behind this algorithm.
+The first is to select an arbitrary relation to unfold, not necessarily the leftmost which is safe.
+The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time.
+If those contradictions are found, then they are exposed by considering the conjunction as a whole and substituting the result of the call unfolding back into the conjunction the call was selected from thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD.
+Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined.
+
+When the heuristic selects a call to unfold (line 15), a process tree is constructed for the selected call \emph{in isolation} (line 16).
+The leaves of the computed tree are examined.
+If all leaves are either computed substitutions or are renamings of some relations accompanied with non-empty substitutions, then the leaves are collected and replace the considered call in the root conjunction (lines 19,20).
+If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected (line 22).
+According to the denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
+
+% Each time we examine a conjunction of calls, we \emph{split} them into separate nodes which are driven independently from each other.
+% Among the relation calls we select one which is according to the heuristic is likely to narrow down the answer set \db{(line 15)}.
+% If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected \db{(line 22)}.
+
+This process creates branchings whenever a disjunction is examined lines 4--6).
 At each step we make sure that we do not start driving a conjunction which we have already examined.
-To do this, we check if the current conjunction is renaming of any other configuration in the tree \db{(line 12)}.
-If it is, then we create a special node which then is residualized into a call to the suitable relation.
+To do this, we check if the current conjunction is renaming of any other configuration in the tree (line 12).
+If it is, then we create a special node which then is residualized into a call to the corresponding relation.
+This is known as \emph{folding} in supercompilation.
 
-We decided not to do generalization in this approach.
-The generalization is used in supercompilation and partial deduction to ensure termination at the same time as some degree of specialization.
-The generalization of two terms is usually a \emph{most-specific generalization}.
-Generalization is used to abstract away some information computed during driving.
-In conjunctive partial deduction generalization is modified to support treating of conjunctions.
-The generalization selects subconjuctions of two conjuncts which are similar (call to the same relation and their arguments have similar shape and distribution).
-For the subconjunctions selected a most-specific generalization is computed.
+We decided not to perform generalization in this approach in the same fashion as CPD or supercompilation does.
+Our conjunctions are always splitted into individual calls and are joined back together if it is sensible.
+If the need for generalization arises, i.e. embedding is detected, then we immediately stop driving this conjunction (line 13).
+When redisualizatin such conjunction, we just generate a conjunction of calls to the input program before specialization.
 
-In our approach we only do splitting of a conjunction into individual relation calls.
-This makes any program with an accumulating parameter to be a problem.
-Sometimes when there is a need to do a proper generalization, it is in reality just an instance of some other goal within the tree and we can simply create a call there \db{(line 13)}.
-Otherwise we are unable to meaningfully specialize such goal, but we can always just include the initial program in the residual program and call the corresponding relation.
+% The generalization is used in supercompilation and partial deduction to ensure termination at the same time as some degree of specialization.
+% The generalization of two terms is usually a \emph{most-specific generalization}.
+% Generalization is used to abstract away some information computed during driving.
+% In conjunctive partial deduction generalization is modified to support treating of conjunctions.
+% The generalization selects subconjuctions of two conjuncts which are similar (call to the same relation and their arguments have similar shape and distribution).
+% For the subconjunctions selected a most-specific generalization is computed.
+
+% In our approach we only do splitting of a conjunction into individual relation calls.
+% This makes any program with an accumulating parameter to be a problem.
+% Sometimes when there is a need to do a proper generalization, it is in reality just an instance of some other goal within the tree and we can simply create a call there \db{(line 13)}.
+% Otherwise we are unable to meaningfully specialize such goal, but we can always just include the initial program in the residual program and call the corresponding relation.
 

### 2020-05-23 papers/miniKanren/text/nonconj.tex
fix line numbers corrsponding to Fig 1

@@
-\newcommand{\code}[1]{\texttt{#1}}
+./\newcommand{\code}[1]{\texttt{#1}}
 
@@
   \centering
-  \includegraphics[width=0.7\textwidth]{algo_pseudo-crop.pdf}
+  \includegraphics[width=0.85\textwidth]{algo_pseudo-crop.pdf}
   \caption{\db{Non-conjunctive Partial Deduction Pseudo Code}}
@@
 The first is to select an arbitrary relation to unfold, not necessarily the leftmost which is safe.
-The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time.
-If those contradictions are found, then they are exposed by considering the conjunction as a whole and substituting the result of the call unfolding back into the conjunction the call was selected from thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD.
-Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined.
+The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time \db{(line 14; \code{heuristically\_select\_a\_call} stands for heurictics combination, the exact code is omitted)}.
+If those contradictions are found, then they are exposed by considering the conjunction as a whole and substituting the result of the call unfolding back into the conjunction the call was selected from thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD \db{(lines 15--23)}.
+Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined \db{(line 25)}.
 
-When the heuristic selects a call to unfold (line 15), a process tree is constructed for the selected call \emph{in isolation} (line 16).
+When the heuristic selects a call to unfold (lines 14--15), a process tree is constructed for the selected call \emph{in isolation} (line 17).
 The leaves of the computed tree are examined.
-If all leaves are either computed substitutions or are renamings of some relations accompanied with non-empty substitutions, then the leaves are collected and replace the considered call in the root conjunction (lines 19,20).
-If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected (line 22).
+If all leaves are either computed substitutions or are renamings of some relations accompanied with non-empty substitutions, then the leaves are collected and replace the considered call in the root conjunction (lines \db{20--21}).
+If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected (line \db{23}).
 According to the denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
@@
 
-This process creates branchings whenever a disjunction is examined lines 4--6).
+This process creates branchings whenever a disjunction is examined lines 4--\db{5}).
 At each step we make sure that we do not start driving a conjunction which we have already examined.
-To do this, we check if the current conjunction is renaming of any other configuration in the tree (line 12).
+To do this, we check if the current conjunction is renaming of any other configuration in the tree (line \db{11}).
 If it is, then we create a special node which then is residualized into a call to the corresponding relation.
@@
 Our conjunctions are always splitted into individual calls and are joined back together if it is sensible.
-If the need for generalization arises, i.e. embedding is detected, then we immediately stop driving this conjunction (line 13).
+If the need for generalization arises, i.e. embedding is detected, then we immediately stop driving this conjunction (line \db{12}).
 When redisualizatin such conjunction, we just generate a conjunction of calls to the input program before specialization.

### 2020-05-23 papers/miniKanren/text/nonconj.tex
Unfolding

--- 2020-05-23 papers/miniKanren/text/nonconj.tex
@@
 
-Unfolding is a process of substitution of some relation call by its body with simultaneous computation of unifications.
-To unfold a relation call we do the following steps.
-\db{TODO: the following has alredy mentioned above.}
-First, the formal arguments of a relation are substituted for the actual arguments of the call in the body.
-All fresh variables get instantiated.
-The body is transformed into a canonical form (disjunction of conjunctions of either calls or unifications).
-All unifications are computed.
-Those disjuncts in which unifications fails are removed.
-Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
+Unfolding in our case is done by substitution of some relation call by its body with simultaneous normalization and computation of unifications.
+% To unfold a relation call we do the following steps.
+% \db{TODO: the following has alredy mentioned above.}
+% First, the formal arguments of a relation are substituted for the actual arguments of the call in the body.
+% All fresh variables get instantiated.
+% The body is transformed into a canonical form (disjunction of conjunctions of either calls or unifications).
+% All unifications are computed.
+% Those disjuncts in which unifications fails are removed.
+% Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
+The unfolding itself is straightforward however it is not always clear what to unfold and when to \emph{stop} unfolding.
+Unfolding in functional programming languages specialization, as well as inlining in imperative one, is usually considered to be safe from a residual program efficiency point of view.
+It may only lead to code explosion or code duplication which is mostly left to a target program compiler optimization or even is out of consideration at all if a specializer is considered as a standalone tool~\cite{jonesbook}.
 
-The most important question is when to unfold.
-\db{
-  {\color{blue}PROPOSED EXTENSION:}, and when to stop unfolding even if the relation call seems to be good for specialization.
-  Usually, unfolding in functional programming languages specialization, as well as inlining in imperative one, is considered to be safe from a residual program efficiency point of view.
-  It may only lead to code explosion or code duplication which is mostly left to a target program compiler optimization or even is out of consideration at all if specializer is considered as a standalone tool~\cite{jonesbook}.
 Unfortunately, this is not the case for the specialization of relational programming language.
-  Unlike functional and imperative, in relational programming language unfolding may easily affect the target program efficiency.
-  It is crucial not to unfold too much since, for example, extra unifications or even repeated computations generated by optimized relation calls and further propagated to neighbouring conjunction may appear.
-}
-\db{{\color{blue}Unfortunately $\to$ }That is}, too much unfolding is sometimes is even worse than not enough unfolding.
-There is a fine edge between those.
-This problem is mentioned in the (CONTROL PAPER).
-We believe that the following heuristic provides a reasonable control.
+Unlike functional and imperative, in logic and relational programming language unfolding may easily affect the target program efficiency~\cite{leuschel2002logic}.
+Unfolding too much may create extra unifications, which is by itself a costly operation, or even introduce duplicated computations by propagating the unfolding results onto neighboring conjuncts.
+
+There is a fine edge between too much unfolding and not enough unfolding.
+The former may be even worse than the latter.
+We believe that the following heuristic provides a reasonable approach to controling unfolding.
 

### 2020-05-23 papers/miniKanren/text/related.tex
Nonconj
add parapraph about SC

@@
 It is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation} and many more.
+
+\db{
+  The heart of supercompilation-based techniques is \emph{driving}~---~a symbolic execution of a program through all possible execution paths.
+  The result of driving is so-called \emph{process tree} where nodes correspond to \emph{configurations} presenting computation state, for example, a term in case of pure functional programming languages and each path in the tree~--~to some concrete program execution.
+  The two main sources for supercompilation optimizations are aggressive information propagation about variable values, equalities, and disabilities and precomputing of all deterministic semantic evaluation steps, i.e.
+combining of consecutive process tree nodes with no branching, aka.
+\emph{deforestation}~\cite{TODO}.
+  When the tree is constructed, the resulting program can be extracted from the process tree, this process is called \emph{residualization} and resulting program is called \emph{residual}.
+  Of cause, process tree is possibly infinite.
+  In order to ensure termination in supercompilation \emph{whistles}, heuristics to identify possibly infinite branches, are used.
+  If the whistle signalled during some branch construction, the process is forced to be stoped.
+  Then iether required source code will be copied to the residual program or process tree will be forced to fold in a \emph{process graph}.
+  The main instrument ro perform such a folding is generalization.
+  When two dangerously similar nodes are met, for example, two consecutive recursive calls of a function with accumulating parameter, and the first is not an instance of the second, then one may construct a new, generalized, node such that both of original nodes are instances of the last one, then one of the initial nodes is replaced with the generalized one.
+  There are several ways to ensure process correctness and termination, the most familiar and popular one is \emph{homeomorphic embedding}~\cite{TODO}.
+}
+
 While supercompilation generally improves the behavior of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in general case.
-
 What is worse they rarely consider the residual program efficiency from the point of view of the target language evaluator.
@@
 
-Conjunctive partial deduction makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro{}.
+Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro{}.
 CPD constructs a tree which models goal evaluation and resembles SLD-NF tree.

### 2020-05-23 papers/miniKanren/text/figures/heuristic.tex
Heuristic

\documentclass[12pt,a4paper]{report}
\usepackage[lmargin=3.81cm,tmargin=2.54cm,rmargin=2.54cm,bmargin=2.52cm]{geometry}
\usepackage{multirow}
\usepackage{color,xcolor}
\definecolor{darkblue}{rgb}{0,0,.75}
\usepackage{listings} %iclude code in your document
\definecolor{dkgreen}{rgb}{0,0.6,0}

\lstloadlanguages{Matlab} %use listings with Matlab for Pseudocode
\lstnewenvironment{PseudoCode}[1][]
{\lstset{basicstyle=\scriptsize,
    keywordstyle=\bfseries\underbar,
    literate=
      {:=}{$\leftarrow{}$}{1}
      {==}{$={}$}{1},
    numbers=left,
    morekeywords={ifd,define,where,let,in,error,case,of,while,do,otherwise,
      read,goto,return,build-cons,build-cst,build-if,if,then,else,do,switch,of},
    morestring=[b]",
    morestring=[s][]{\#\{}{\}},
    deletestring=[b]{"},
    alsoletter={-},
    frame=l,
    keywordstyle = [2]{\textit},
    morekeywords = [2]{lookup,initial_code,extend,reduce,eval,rest,first_command,first_instruction,new_tail,first_symbol},
    showspaces=false,
    showstringspaces=false,
    stringstyle=\color{blue},
    identifierstyle=,
    escapechar=@*,
    xleftmargin=.04\textwidth,#1}}
{}

\usepackage{stmaryrd}
\begin{document}

\thispagestyle{empty}
\begin{figure}[h]
\begin{PseudoCode}[moredelim={[is][stringstyle]{@@}{@@}},
  moredelim={[is][\color{dkgreen}\bfseries]{@!}{@!}},
  moredelim={[is][\color{red}\bfseries]{!!}{!!}}]
heuristically_select_a_call :: Conjunction @*$\to$@* Maybe Call
heuristically_select_a_call C =
  find heuristic C

heuristic :: Call -> Bool
heuristic r =
  isStatic r || isDeterministic r || isLessBranching r
\end{PseudoCode}
\end{figure}

\end{document}

### 2020-05-23 papers/miniKanren/text/nonconj.tex
rem
Heuristic

@@
-./\newcommand{\code}[1]{\texttt{#1}}
+\newcommand{\code}[1]{\texttt{#1}}
 
@@
 
-The intuition behind the heuristic is to find those calls which are safe to unfold.
+This heuristic is aimed at selecting the relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
+The unsafe unfolding leads to uncontrollable increase in the number of relation calls in a conjunction.
+It is best to first unfold those relation calls which can be fully computed up to substitutions.
+
 We deem every static conjunct (non-recursive) to be safe because they never lead to growth in the number of conjunctions.
@@
 
-The other more complicated case is when there are less disjuncts than there can possibly be.
-This signifies that at least one branch of computations is gotten rid of.
+The other part of heuristic is less trivial, we call it the branching heuristic.
+It considers the case when the unfolded relation contains less disjuncts than it could possibly have.
+This means that we found some contradiction, some computations were gotten rid of, and thus the answer set was restricted, which is desirable when unfolding.
+To compute this heuristic we precompute the maximum possible number of disjuncts in each relation and compare this number with the number of disjuncts when unfolding a concrete relation call.
+Maximum number of disjuncts is computed by unfolding the body of the relation in which all relation calls were replaced by a unification which always succeeds.
 
-The final heuristic selects the first conjunct which suites either of the following cases.
-First we unfold those conjuncts which are static.
-Then --- deterministic.
-Then those which are less branching.
-The last to be unfolded are those calls, which unfold to a substitution with not conjunction.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.65\textwidth]{heuristic-crop.pdf}
+  \caption{Heuristic selection pseudocode}
+  \label{fig:heu-pseudo}
+\end{figure}
 
-\subsection{Residualization}
+%  more complicated case is when there are less disjuncts than there can possibly be.
+% This signifies that at least one branch of computations is gotten rid of.
 
-Residualization is quite straightforward.
-A branching in the process tree becomes a disjunction.
-A split node becomes a conjunction.
-Computed substitution is residualized as a conjunction of unifications.
-A renaming node is just a call to a relation.
-Relations are created for configurations on which leaf nodes are renamed.
+The pseudocode describing heuristic is shown in fig.~\ref{fig:heu-pseudo}.
+Selecting a good relation call can fail (line 1).
+The implementation works such that we first select those relation calls which are static, and only if there are none, we proceed to look at deterministic unfoldings and then we search for those which are less branching.
+This heuristic provides a good balance in unfolding.
+% The final heuristic selects the first conjunct which suites either of the following cases.
+% First we unfold those conjuncts which are static.
+% Then --- deterministic.
+% Then those which are less branching.
+% The last to be unfolded are those calls, which unfold to a substitution with not conjunction.
 
-One other thing is that when some configuration is occurred within the tree which is an instance of a configuration for which a new relation is created, then we just create a call.
+% \subsection{Residualization}
+
+% Residualization is quite straightforward.
+% A branching in the process tree becomes a disjunction.
+% A split node becomes a conjunction.
+% Computed substitution is residualized as a conjunction of unifications.
+% A renaming node is just a call to a relation.
+% Relations are created for configurations on which leaf nodes are renamed.
+
+% One other thing is that when some configuration is occurred within the tree which is an instance of a configuration for which a new relation is created, then we just create a call.

### 2020-05-23 papers/miniKanren/text/nonconj.tex
add ref to fig.2

@@
 The first is to select an arbitrary relation to unfold, not necessarily the leftmost which is safe.
-The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time \db{(line 14; \code{heuristically\_select\_a\_call} stands for heurictics combination, the exact code is omitted)}.
+The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time \db{(line 14; \code{heuristically\_select\_a\_call} stands for heurictics combination, see section~\ref{sec:heurictic} for details)}.
 If those contradictions are found, then they are exposed by considering the conjunction as a whole and substituting the result of the call unfolding back into the conjunction the call was selected from thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD \db{(lines 15--23)}.
@@
 \subsection{Heuristic}
+\label{sec:heurictic}
 

### 2020-05-23 papers/miniKanren/text/related.tex
SC

--- 2020-05-22 papers/miniKanren/text/related.tex
@@
 It is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation} and many more.
+
+\db{
+  The heart of supercompilation-based techniques is \emph{driving}~---~a symbolic execution of a program through all possible execution paths.
+  The result of driving is so-called \emph{process tree} where nodes correspond to \emph{configurations} presenting computation state, for example, a term in case of pure functional programming languages.
+  Each path in the tree corresponds to some concrete program execution.
+  The two main sources for supercompilation optimizations are aggressive information propagation about variable values, equalities, and disequalities and precomputing of all deterministic semantic evaluation steps, i.e.
+combining of consecutive process tree nodes with no branching, aka.
+\emph{deforestation}~\cite{TODO}.
+  When the tree is constructed, the resulting, or \emph{residual}, program can be extracted from the process tree by the process called \emph{residualization}.
+  Of course, process tree can contain infinite branches.
+  \emph{Whistles}, heuristics to identify possibly infinite branches, are used to ensure termination in supercompilation.
+  If the whistle signalled during some branch construction, then something is done to ensure termination.
+  The most common approaches include either stopping driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or folding the process tree to fold in a \emph{process graph}.
+  The main instrument to perform such a folding is \emph{generalization}, i.e. abstracting away some computed data about the current term which makes folding possible.
+  One example of the source of infinite branches is consecutive recursive call to the same function with an accumulating parameter: by unfolding such a call further we can only increase the term size and lead to nontermination.
+  The accumulating parameter can be removed by replacing the call with its generalization.
+  There are several ways to ensure process correctness and termination, the most common being \emph{homeomorphic embedding}~\cite{TODO} used as a whistle and most-specific generalization of terms.
+  % When two dangerously similar nodes are encountered,
+  % For example, two consecutive recursive calls of a function with accumulating parameter, and the first is not an instance of the second, then one may construct a new, generalized, node such that both of original nodes are instances of the last one, then one of the initial nodes is replaced with the generalized one.
+}
+
 While supercompilation generally improves the behavior of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in general case.
-
 What is worse they rarely consider the residual program efficiency from the point of view of the target language evaluator.

### 2020-05-23 papers/miniKanren/text/nonconj.tex

@@
 \section{Non-conjunctive Partial Deduction}
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[width=0.85\textwidth]{algo_pseudo-crop.pdf}
-  \caption{\db{Non-conjunctive Partial Deduction Pseudo Code}}
-  \label{fig:ncpd-pseudo}
-\end{figure}
 
@@
 
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=0.85\textwidth]{algo_pseudo-crop.pdf}
+  \caption{\db{Non-conjunctive Partial Deduction Pseudo Code}}
+  \label{fig:ncpd-pseudo}
+\end{figure}
+
 % The very first step of driving a conjunction is to apply a substitution to the variables in relation calls (line 10).
@@
 The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time \db{(line 14; \code{heuristically\_select\_a\_call} stands for heurictics combination, see section~\ref{sec:heurictic} for details)}.
-If those contradictions are found, then they are exposed by considering the conjunction as a whole and substituting the result of the call unfolding back into the conjunction the call was selected from thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD \db{(lines 15--23)}.
-Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined \db{(line 25)}.
+If those contradictions are found, then they are exposed by considering the conjunction as a whole and substituting the result of the call unfolding back into the conjunction the call was selected from thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD \db{(lines 15--22)}.
+Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined \db{(line 23)}.
 
-When the heuristic selects a call to unfold (lines 14--15), a process tree is constructed for the selected call \emph{in isolation} (line 17).
+When the heuristic selects a call to unfold (line \db{15}), a process tree is constructed for the selected call \emph{in isolation} (line \db{16}).
 The leaves of the computed tree are examined.
-If all leaves are either computed substitutions or are renamings of some relations accompanied with non-empty substitutions, then the leaves are collected and replace the considered call in the root conjunction (lines \db{20--21}).
-If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected (line \db{23}).
+If all leaves are either computed substitutions or are renamings of some relations accompanied with non-empty substitutions, then the leaves are collected and replace the considered call in the root conjunction (lines \db{19--20}).
+If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected (line \db{22}).
 According to the denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
@@
   \centering
-  \includegraphics[width=0.65\textwidth]{heuristic-crop.pdf}
+  \includegraphics[width=0.75\textwidth]{heuristic-crop.pdf}
   \caption{Heuristic selection pseudocode}

### 2020-05-23 papers/miniKanren/text/intro.tex
Edits to intro

@@
 Having specified a relation for adding two numbers, one can also compute the substraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
-By running a relational interpreter \emph{backwards} one can obtain a \emph{solver}~\cite{lozov2019relational} thus solving a much more complicated problem.
+Program synthesis can be done by running \emph{backwards} a relational interpreter for some language.
+In general, it is possible to create a solver for a recognizer by translating it into \mk{} and running in the appropriate direction~\cite{lozov2019relational}.
 
@@
 What is even worse, it depends on the order of the relation calls within a program.
-One order can be good for one direction, but slow the computation drastically in the different direction.
+One order can be good for one direction, but slow down the computation drastically in the other direction.
 
@@
 Since the search in \mk{} is complete, it is safe to reorder come relation calls within the goal for better performance.
-While sometimes conjunctive partial deduction gives great perforamnce boost, sometimes it does not behave as well as it could have been.
+While sometimes conjunctive partial deduction gives great performance boost, sometimes it does not behave as well as it could have.
 
-In this paper we show on examples some issues which face conjunctive partial deduction.
+In this paper we show on examples some issues which conjunctive partial deduction faces.
 We also describe a novel approach to partial deduction of a relational programming language \mk{}.

### 2020-05-23 papers/miniKanren/text/related.tex
Related works

@@
 
-\db{
 The heart of supercompilation-based techniques is \emph{driving}~---~a symbolic execution of a program through all possible execution paths.
-  The result of driving is so-called \emph{process tree} where nodes correspond to \emph{configurations} presenting computation state, for example, a term in case of pure functional programming languages.
+The result of driving is so-called \emph{process tree} where nodes correspond to \emph{configurations} which present computation state, for example, a term in case of pure functional programming languages.
 Each path in the tree corresponds to some concrete program execution.
 The two main sources for supercompilation optimizations are aggressive information propagation about variable values, equalities, and disequalities and precomputing of all deterministic semantic evaluation steps, i.e.
-combining of consecutive process tree nodes with no branching, aka.
-\emph{deforestation}~\cite{TODO}.
+combining of consecutive process tree nodes with no branching, also known as \emph{deforestation}~\cite{TODO}.
 When the tree is constructed, the resulting, or \emph{residual}, program can be extracted from the process tree by the process called \emph{residualization}.
 Of course, process tree can contain infinite branches.
-  \emph{Whistles}, heuristics to identify possibly infinite branches, are used to ensure termination in supercompilation.
-  If the whistle signalled during some branch construction, then something is done to ensure termination.
+\emph{Whistles} --- heuristics to identify possibly infinite branches --- are used to ensure termination in supercompilation.
+If the whistle signalled during the construction of some branch, then something should be done to ensure termination.
 The most common approaches include either stopping driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or folding the process tree to fold in a \emph{process graph}.
 The main instrument to perform such a folding is \emph{generalization}, i.e. abstracting away some computed data about the current term which makes folding possible.
-  One example of the source of infinite branches is consecutive recursive call to the same function with an accumulating parameter: by unfolding such a call further we can only increase the term size and lead to nontermination.
+One source of infinite branches is consecutive recursive calls to the same function with an accumulating parameter: by unfolding such a call further one can only increase the term size which leads to nontermination.
 The accumulating parameter can be removed by replacing the call with its generalization.
@@
 % For example, two consecutive recursive calls of a function with accumulating parameter, and the first is not an instance of the second, then one may construct a new, generalized, node such that both of original nodes are instances of the last one, then one of the initial nodes is replaced with the generalized one.
-}
 

### 2020-05-23 papers/miniKanren/text/related.tex
add refs

@@
 The two main sources for supercompilation optimizations are aggressive information propagation about variable values, equalities, and disequalities and precomputing of all deterministic semantic evaluation steps, i.e.
-combining of consecutive process tree nodes with no branching, also known as \emph{deforestation}~\cite{TODO}.
+combining of consecutive process tree nodes with no branching, also known as \emph{deforestation}~\cite{deforestation}.
 When the tree is constructed, the resulting, or \emph{residual}, program can be extracted from the process tree by the process called \emph{residualization}.
@@
 The accumulating parameter can be removed by replacing the call with its generalization.
-There are several ways to ensure process correctness and termination, the most common being \emph{homeomorphic embedding}~\cite{TODO} used as a whistle and most-specific generalization of terms.
+There are several ways to ensure process correctness and termination, the most common being \emph{homeomorphic embedding}~\cite{Higman52,Kruskal60} used as a whistle and most-specific generalization of terms.
 % When two dangerously similar nodes are encountered,

### 2020-05-23 papers/miniKanren/text/related.tex
Related works done

--- 2020-05-23 papers/miniKanren/text/related.tex
@@
 While supercompilation generally improves the behavior of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in general case.
-What is worse they rarely consider the residual program efficiency from the point of view of the target language evaluator.
+What is worse, they rarely consider the residual program efficiency from the point of view of the target language evaluator.
 The main optimization source is computing in advance all possible intermediate and statically-known semantics steps at transformation-time.
@@
 Moreover, it may be hard to predict the real speedup of any given program on concrete examples even disregarding problems above because of the complexity of the transformation algorithm.
-The worst case for partial evaluation is when all static variables are used in a dynamic context, and there is a lot of advice on how to implement a partial evaluator as well as a target program so that specialization really improved its performance~\cite{jonesbook,bulyonkov84}.
+The worst case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization really improved its performance~\cite{jonesbook,bulyonkov84}.
 There is lack of research in determining the classes of programs which transformers would definitely speed up.
@@
 Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro{}.
-CPD constructs a tree which models goal evaluation and resembles SLD-NF tree.
-After the tree is constructed, a residual program is generated from it.
+CPD constructs a tree which models goal evaluation and is similar to SLD-NF tree, then a residual program is generated from this tree.
+Partial deduction itself resembles driving in supercompilation~\cite{gluck1994partial}.
 The specialization is done in two levels of control: the local control determines the shape of the residual programs, while the global control ensures that every relation which can be called in the residual program is indeed defined.
@@
 At the local level CPD examines a conjunction of atoms by considering each atom one-by-one from left to right.
-An atom is \emph{unfolded} if it is deemed safe.
-When an atom is unfolded, a clause which head can be unified with the atom is found, and a new node is added into the tree where the atom in the conjunction is replaced with the body of that clause.
+An atom is \emph{unfolded} if it is deemed safe, i.e. a whistle based on homeomorphic does not signal for the atom.
+When an atom is unfolded, a clause whose head can be unified with the atom is found, and a new node is added into the tree where the atom in the conjunction is replaced with the body of that clause.
 If there are more than one suitable head, then several branches are added into the tree which correspond to the disjunction in the residualized program.
@@
 
-The most well-behaved strategy of local control in CPD for \pro{} is \emph{deterministic unfolding}.
+The most well-behaved strategy of local control in CPD for \pro{} is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
 An atom is unfolded only if only one suitable clause head exist for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
-This means that if a non-deterministic atom is the left-most within a conjunction, it is most likely to be unfoldled and introduce a lot of
+This means that if a non-deterministic atom is the leftmost within a conjunction, it is most likely to be unfoldled and introduce many new relation calls within the conjunction.
 We believe this is the core problem with CPD which limits its power when applied to \mk{}.
-The strategy of unfolding atoms from left to right is reasonable in the context of \pro{} because it mimics the way it executes.
-But it often leads to larger global control trees and, as a result, bigger less efficient programs.
+The strategy of unfolding atoms from left to right is reasonable in the context of \pro{} because it mimics the way programs in \pro{} execute.
+But it often leads to larger global control trees and, as a result, bigger, less efficient programs.
 The evaluation result of a \mk{} program does not depend on the order of atoms (relation calls) within a conjunction, thus we believe a better result can be achieved by selecting a relation call which can restrict the number of branches in the tree.

### 2020-05-23 papers/miniKanren/text/related.tex
remove extra "fold"-word

@@
 If the whistle signalled during the construction of some branch, then something should be done to ensure termination.
-The most common approaches include either stopping driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or folding the process tree to fold in a \emph{process graph}.
+The most common approaches include either stopping driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or folding the process tree to a \emph{process graph}.
 The main instrument to perform such a folding is \emph{generalization}, i.e. abstracting away some computed data about the current term which makes folding possible.

### 2020-05-23 papers/miniKanren/text/nonconj.tex
Nonconj done

@@
 Those disjuncts in which unifications fail are removed.
-Other disjuncts take form of a conjunction of relation calls accompanied with a substitution computed from unificaitons.
-Any miniKanren term can be trivially transformed to the described form.
+Other disjuncts take form of a possibly empty conjunction of relation calls accompanied with a substitution computed from unificaitons.
+Any \mk{} term can be trivially transformed to the described form.
 In Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization.
@@
 The first is to select an arbitrary relation to unfold, not necessarily the leftmost which is safe.
-The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time \db{(line 14; \code{heuristically\_select\_a\_call} stands for heurictics combination, see section~\ref{sec:heurictic} for details)}.
-If those contradictions are found, then they are exposed by considering the conjunction as a whole and substituting the result of the call unfolding back into the conjunction the call was selected from thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD \db{(lines 15--22)}.
-Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined \db{(line 23)}.
+The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time (line 14; \code{heuristically\_select\_a\_call} stands for heurictics combination, see section~\ref{sec:heurictic} for details).
+If those contradictions are found, then they are exposed by considering the conjunction as a whole and replacing the selected relation call with the result of its unfolding thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD (lines 15--22).
+Joining instead of splitting is why we call our transformer \emph{non-conjunctive} partial deduction.
+Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined (line 23).
 
-When the heuristic selects a call to unfold (line \db{15}), a process tree is constructed for the selected call \emph{in isolation} (line \db{16}).
+When the heuristic selects a call to unfold (line 15), a process tree is constructed for the selected call \emph{in isolation} (line 16).
 The leaves of the computed tree are examined.
-If all leaves are either computed substitutions or are renamings of some relations accompanied with non-empty substitutions, then the leaves are collected and replace the considered call in the root conjunction (lines \db{19--20}).
-If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls withing the conjunction and the next suitable call is selected (line \db{22}).
+If all leaves are either computed substitutions or are instances of some relations accompanied with non-empty substitutions, then the leaves are collected and each of them replace the considered call in the root conjunction (lines 19--20).
+If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls within the conjunction, instead the next suitable call is selected (line 22).
 According to the denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
@@
 
-This process creates branchings whenever a disjunction is examined lines 4--\db{5}).
+This process creates branchings whenever a disjunction is examined (lines 4--5).
 At each step we make sure that we do not start driving a conjunction which we have already examined.
-To do this, we check if the current conjunction is renaming of any other configuration in the tree (line \db{11}).
-If it is, then we create a special node which then is residualized into a call to the corresponding relation.
-This is known as \emph{folding} in supercompilation.
+To do this, we check if the current conjunction is a renaming of any other configuration in the tree (line 11).
+If it is, then we fold the tree by creating a special node which then is residualized into a call to the corresponding relation.
 
 We decided not to perform generalization in this approach in the same fashion as CPD or supercompilation does.
-Our conjunctions are always splitted into individual calls and are joined back together if it is sensible.
-If the need for generalization arises, i.e. embedding is detected, then we immediately stop driving this conjunction (line \db{12}).
-When redisualizatin such conjunction, we just generate a conjunction of calls to the input program before specialization.
+Our conjunctions are always splitted into individual calls and are joined back together only if it is sensible.
+If the need for generalization arises, i.e. homeomorphic embedding of conjunctions~\cite{de1999conjunctive} is detected, then we immediately stop driving this conjunction (line 12).
+When redisualizating such conjunction, we just generate a conjunction of calls to the input program before specialization.
 
@@
 The unfolding itself is straightforward however it is not always clear what to unfold and when to \emph{stop} unfolding.
-Unfolding in functional programming languages specialization, as well as inlining in imperative one, is usually considered to be safe from a residual program efficiency point of view.
+Unfolding in functional programming languages specialization, as well as inlining in imperative one, is usually considered to be safe from the point of view of residual program efficiency.
 It may only lead to code explosion or code duplication which is mostly left to a target program compiler optimization or even is out of consideration at all if a specializer is considered as a standalone tool~\cite{jonesbook}.
@@
 The former may be even worse than the latter.
-We believe that the following heuristic provides a reasonable approach to controling unfolding.
+We believe that the following heuristic provides a reasonable approach to unfolding control.
 
@@
 
-We deem every static conjunct (non-recursive) to be safe because they never lead to growth in the number of conjunctions.
+We deem every static (non-recursive) conjunct to be safe because they never lead to growth in the number of conjunctions.
 Those calls which unfold deterministically, meaning there is only one disjunct in the unfolded relation, are also considered to be safe.
 
-The other part of heuristic is less trivial, we call it the branching heuristic.
-It considers the case when the unfolded relation contains less disjuncts than it could possibly have.
+Those relation calls which are neither static nor deterministic are examined with what we call the \emph{less-branching} heuristic.
+It identifies the case when the unfolded relation contains less disjuncts than it could possibly have.
 This means that we found some contradiction, some computations were gotten rid of, and thus the answer set was restricted, which is desirable when unfolding.
@@
 Selecting a good relation call can fail (line 1).
-The implementation works such that we first select those relation calls which are static, and only if there are none, we proceed to look at deterministic unfoldings and then we search for those which are less branching.
-This heuristic provides a good balance in unfolding.
+The implementation works such that we first select those relation calls which are static, and only if there are none, we proceed to consider deterministic unfoldings and then we search for those which are less branching.
+We believe this heuristic provides a good balance in unfolding.
 % The final heuristic selects the first conjunct which suites either of the following cases.

### 2020-05-23 papers/miniKanren/text/nonconj.tex
spell- and grammarly- checked

@@
 
-In this section we describe a novel approach to specialization of relational programs.
+In this section, we describe a novel approach to relational programs specialization.
 This approach draws inspiration from both conjunctive partial deduction and supercompilation.
-The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve performance of the input programs.
+The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve the performance of the input programs.
 
-The algorithm pseudocode is shown on Fig.~\ref{fig:ncpd-pseudo}.
+The algorithm pseudocode is shown in Fig.~\ref{fig:ncpd-pseudo}.
 For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} which describe how to process disjunctions and conjunctions respectively.
@@
 The process tree is meant to mimic the execution of the input program.
-The nodes of process tree include a \emph{configuraion} which describes the state of program evaluation at some point.
+The nodes of the process tree include a \emph{configuration} which describes the state of program evaluation at some point.
 In our case configuration is a conjunction of relation calls.
-The substitution computed at each step is also stored in the tree node, although it is not included into the configuration.
+The substitution computed at each step is also stored in the tree node, although it is not included in the configuration.
 
 Hereafter, we consider all goals and relation bodies to be in \emph{canonical normal form}~---~a disjunction of conjunctions of either calls or unifications.
-Moreover, we assume all fresh variables to be introduced into scope and all unifications to be computed at each step.
+Moreover, we assume all fresh variables to be introduced into the scope and all unifications to be computed at each step.
 Those disjuncts in which unifications fail are removed.
-Other disjuncts take form of a possibly empty conjunction of relation calls accompanied with a substitution computed from unificaitons.
-Any \mk{} term can be trivially transformed to the described form.
+Other disjuncts take the form of possibly empty conjunction of relation calls accompanied with a substitution computed from unifications.
+Any \mk{} term can be trivially transformed into the described form.
 In Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization.
@@
 The first is to select an arbitrary relation to unfold, not necessarily the leftmost which is safe.
-The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time (line 14; \code{heuristically\_select\_a\_call} stands for heurictics combination, see section~\ref{sec:heurictic} for details).
+The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time (line 14; \code{heuristically\_select\_a\_call} stands for heuristics combination, see section~\ref{sec:heurictic} for details).
 If those contradictions are found, then they are exposed by considering the conjunction as a whole and replacing the selected relation call with the result of its unfolding thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD (lines 15--22).
 Joining instead of splitting is why we call our transformer \emph{non-conjunctive} partial deduction.
-Finally, if the heuristic fails to select a potentially good call, then the conjunction is splitted into individual calls which are driven in isolation and are never joined (line 23).
+Finally, if the heuristic fails to select a potentially good call, then the conjunction is split into individual calls which are driven in isolation and are never joined (line 23).
 
@@
 The leaves of the computed tree are examined.
-If all leaves are either computed substitutions or are instances of some relations accompanied with non-empty substitutions, then the leaves are collected and each of them replace the considered call in the root conjunction (lines 19--20).
-If the selected call does not suit the criteria, the results of its unfolding is not propagated onto other relation calls within the conjunction, instead the next suitable call is selected (line 22).
+If all leaves are either computed substitutions or are instances of some relations accompanied with non-empty substitutions, then the leaves are collected and each of them replaces the considered call in the root conjunction (lines 19--20).
+If the selected call does not suit the criteria, the results of its unfolding are not propagated onto other relation calls within the conjunction, instead, the next suitable call is selected (line 22).
 According to the denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
@@
 This process creates branchings whenever a disjunction is examined (lines 4--5).
-At each step we make sure that we do not start driving a conjunction which we have already examined.
+At each step, we make sure that we do not start driving conjunction which we have already examined.
 To do this, we check if the current conjunction is a renaming of any other configuration in the tree (line 11).
@@
 We decided not to perform generalization in this approach in the same fashion as CPD or supercompilation does.
-Our conjunctions are always splitted into individual calls and are joined back together only if it is sensible.
+Our conjunctions are always split into individual calls and are joined back together only if it is sensible.
 If the need for generalization arises, i.e. homeomorphic embedding of conjunctions~\cite{de1999conjunctive} is detected, then we immediately stop driving this conjunction (line 12).
-When redisualizating such conjunction, we just generate a conjunction of calls to the input program before specialization.
+When residualizing such conjunction, we just generate a conjunction of calls to the input program before specialization.
 
@@
 The unfolding itself is straightforward however it is not always clear what to unfold and when to \emph{stop} unfolding.
-Unfolding in functional programming languages specialization, as well as inlining in imperative one, is usually considered to be safe from the point of view of residual program efficiency.
+Unfolding in functional programming languages specialization, as well as inlining in imperative one, is usually considered to be safe from the residual program efficiency point of view.
 It may only lead to code explosion or code duplication which is mostly left to a target program compiler optimization or even is out of consideration at all if a specializer is considered as a standalone tool~\cite{jonesbook}.
@@
 Unlike functional and imperative, in logic and relational programming language unfolding may easily affect the target program efficiency~\cite{leuschel2002logic}.
-Unfolding too much may create extra unifications, which is by itself a costly operation, or even introduce duplicated computations by propagating the unfolding results onto neighboring conjuncts.
+Unfolding too much may create extra unifications, which is by itself a costly operation, or even introduce duplicated computations by propagating the unfolding results onto neighbouring conjuncts.
 
 There is a fine edge between too much unfolding and not enough unfolding.
-The former may be even worse than the latter.
+The former is maybe even worse than the latter.
 We believe that the following heuristic provides a reasonable approach to unfolding control.
@@
 This heuristic is aimed at selecting the relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
-The unsafe unfolding leads to uncontrollable increase in the number of relation calls in a conjunction.
+The unsafe unfolding leads to an uncontrollable increase in the number of relation calls in a conjunction.
 It is best to first unfold those relation calls which can be fully computed up to substitutions.
@@
 Those relation calls which are neither static nor deterministic are examined with what we call the \emph{less-branching} heuristic.
-It identifies the case when the unfolded relation contains less disjuncts than it could possibly have.
+It identifies the case when the unfolded relation contains fewer disjuncts than it could possibly have.
 This means that we found some contradiction, some computations were gotten rid of, and thus the answer set was restricted, which is desirable when unfolding.
 To compute this heuristic we precompute the maximum possible number of disjuncts in each relation and compare this number with the number of disjuncts when unfolding a concrete relation call.
-Maximum number of disjuncts is computed by unfolding the body of the relation in which all relation calls were replaced by a unification which always succeeds.
+The maximum number of disjuncts is computed by unfolding the body of the relation in which all relation calls were replaced by a unification which always succeeds.
 

### 2020-05-23 papers/miniKanren/text/evaluation.tex

@@
 
-In this paper we specialize the relation to synthesize formulas which evaluate to \lstinline{^true}.
-To do so we run the specializer for the goal with the last argument fixed to \lstinline{^true}, while the first two arguments remain free variables.
+All examples of \mk{} relations in this paper are written in \oc{}\footnote{\oc{}: statically typed \mk{} embedding in \ocaml{}.
+The repository of the project: \url{https://github.com/JetBrains-Research/OCanren}} syntax.
+We specialize the \lstinline{eval$^o$} relation to synthesize formulas which evaluate to \lstinline{^true}\footnote{An arrow lifts ordinary values to the logic domain}.
+To do so, we run the specializer for the goal with the last argument fixed to \lstinline{^true}, while the first two arguments remain free variables.
 Depending on the way the \lstinline{eval$^o$} is implemented, different specializers generate significantly different residual programs.
@@
 One possible implementation of the evaluator in the syntax of \oc{} is presented in listing~\ref{eval:last}.
-In this implementation the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
+Here the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
 The relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} encode corresponding boolean operations.
@@
 These variables are used in recursive calls of \lstinline{eval$^o$} and thus restrict the result of driving them.
-CPD fails to recognize this, and thus unfold recursive calls of \lstinline{eval$^o$} applied to fresh variables.
+CPD fails to recognize this, and thus unfolds recursive calls of \lstinline{eval$^o$} applied to fresh variables.
 It leads to over-unfolding, big residual programs and less than optimal performance.
 
-The non-conjunctive partial deduction will first unfold those calls which are selected with the heuristic.
+The non-conjunctive partial deduction first unfolds those calls which are selected with the heuristic.
 Since exploring boolean operations makes more sense, they are unfolded before recursive calls of \lstinline{eval$^o$}.
@@
       (fm === var v /\ elem$^o$ subst v res);
-      (fm === conj x y /\ and$^o$_table v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
-      (fm === disj x y /\ or$^o$_table v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
-      (fm === neg x /\ not$^o$_table v res /\ eval$^o$ st x v))]
+      (fm === conj x y /\ and$^o$ v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
+      (fm === disj x y /\ or$^o$ v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
+      (fm === neg x /\ not$^o$ v res /\ eval$^o$ st x v))]
     \end{lstlisting}
@@
 In the implementation of \lstinline{eval$^o$} boolean operations can effectively restrict variables within the conjunctions and should be unfolded until they do.
-But depending on the way they are implemented, different number of driving steps should be performed to do so.
+But depending on the way they are implemented, different number of driving steps should be performed for that.
 The simplest way to implement these relations is with a table as demonstrated with the implementation of \lstinline{not$^o$} in listing~\ref{not:table}.
@@
     \begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b, frame=tb]
-  let not$^o$_table x y = conde [
+  let not$^o$ x y = conde [
      (x === ^true /\ y === ^false;
@@
 
-In our study we considered two implementations of \lstinline{eval$^o$}: one we call \lstinline{plain} and the other --- \lstinline{last} and compared how specializers behave on them.
+In our study we considered two implementations of \lstinline{eval$^o$}, one we call \lstinline{plain} and the other --- \lstinline{last}, and compared how specializers behave on them.
 The \lstinline{plain} relation uses table-based boolean operations and places them further to the left in each conjunction.
@@
 The running time of two programs generated with our novel non-conjunctive partial deduction is very close and it is a little bit better than the best by CPD.
-CPD with the branching heuristic still gives different quality transformations.
+CPD with the branching heuristic still constructs residual programs of different quality.
 The results are shown in table~\ref{tbl:eval}.
@@
 
-The \lstinline{unify} relation was executed to find 5 unifiers of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
+The \lstinline{unify} relation was executed to find a unifier of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
 The original \mk{} program fails to terminate on this goal in 300 seconds.
@@
   \centering
-  \begin{tabular}{c|c|c||c||c}
+  \begin{tabular}{c||c|c||c||c}
                    & last & plain & unify & isPath \\

### 2020-05-28 papers/TEASE-LP/example.tex
Minor edit

@@
 We use the following representation of the goal in canonical form:
-$$Goal = \underline{\fresh{}} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$$.
+$$Goal = \underline{\fresh{}} \ V_1 \dots V_n \bigvee \bigwedge (\underline{call} \ R^k \ V_1 \dots V_k \mid V \equiv T)$$
 

### 2020-05-28 papers/miniKanren/ocanren/rep/build.sh
Rep: benchmarks

@@ -1,0 +1,1 @@
+ocamlfind c -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml rep20.ml rep1.ml rep50NormExtra.ml rep50Norm.ml rep75Norm.ml rep50.ml rep75.ml rep100.ml rep100Norm.ml run.ml

### 2020-05-28 src/NonConjunctive/Residualization.hs

@@ -2,9 +2,11 @@
 
 module NonConjunctive.Residualization where
 
+import           Control.Applicative   ((<|>))
 import qualified CPD.LocalControl      as LC
 import qualified CPD.Residualization   as CpdR
-import           Data.List             (sort, (\\), nub, find, intercalate, union)
+import           Data.List             (find, intercalate, nub, sort, union,
+                                        (\\))
 import           Data.Maybe            (catMaybes, fromJust, fromMaybe)
 import           Debug.Trace           (trace, traceM)
 import           Embed                 (isInst)
@@ -13,8 +15,8 @@ import           NonConjunctive.Unfold
 import qualified Residualize           as Res
 import           Syntax
 import           Text.Printf           (printf)
-import           Util.Miscellaneous    (fst3, snd3, trd3, show')
-import Control.Applicative ((<|>))
+import           Unfold                (normalize)
+import           Util.Miscellaneous    (fst3, show', snd3, trd3)
 
 topLevel :: Program -> Program
 topLevel input =
@@ -99,7 +101,7 @@ findNode v tree =
 
 nontrivial :: NCTree -> Bool
 nontrivial (Leaf _ _ _ _) = False
-nontrivial _ = True
+nontrivial _              = True
 
 nodeContent (Or _ (LC.Descend goal _) _) = Just goal
 nodeContent (Conj _ goal _)              = Just goal
@@ -116,7 +118,9 @@ generateDef defs invocations ((gs, n, args), tree) =
 generateGoalFromTree :: CpdR.Definitions -> [([G S], G S)] -> NCTree -> [S] -> G X
 generateGoalFromTree definitions invocations tree args =
     case go args True tree of
-      Just goal -> Res.vident <$> goal
+      Just goal ->
+        let normalized = goal in --  NonConjunctive.Unfold.disj $ map NonConjunctive.Unfold.conj $ normalize goal in
+        Res.vident <$> normalized
       Nothing -> error $ printf "Failed to generate relation body for %s" (show $ nodeContent tree)
     -- Res.vident <$> (disj (map conj $ filter (not . null) $ go tree))
   where
@@ -159,9 +163,9 @@ generateGoalFromTree definitions invocations tree args =
     go seen r (Prune _ _)     = error "Failed to residualize: Prune node in tree"
 
[…]

### 2020-05-28 src/NonConjunctive/Unfold.hs

@@ -159,6 +159,9 @@ conjToList x          = [x]
 conj :: [G a] -> G a
 conj (a:as) = foldl (:/\:) a as
 
+disj :: [G a] -> G a
+disj (a:as) = foldl (:\/:) a as
+
 globalLimit :: Int
 globalLimit = 8
 

### 2020-05-28 src/OCanrenize.hs

@@ -19,13 +19,13 @@ instance OCanren String where
 instance {- OCanren v => -} OCanren (Term X) where
   -- ocanren (V v@('q':_)) = printf "(Obj.magic %s)" (ocanren v)
   ocanren (V v)        = ocanren v
-  ocanren (C nil _) | map toLower nil == "nil"  = "Std.List.nil ()"
+  ocanren (C nil _) | map toLower nil == "nil"  = "(Std.List.nil ())"
   -- ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "(%s %% %s)" (ocanren h) (ocanren t)
   -- ocanren (C "%"    [h,t]) = printf "(%s %% %s)" (ocanren h) (ocanren t)
-  ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "Std.(%%) %s %s" (ocanren h) (ocanren t)
-  ocanren (C "%" [h, t]) = printf "Std.(%%) (%s) (%s)" (ocanren h) (ocanren t)
+  ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "(Std.(%%) %s %s)" (ocanren h) (ocanren t)
+  ocanren (C "%" [h, t]) = printf "(Std.(%%) (%s) (%s))" (ocanren h) (ocanren t)
   ocanren (C "O" []) = "Std.Nat.zero"
-  ocanren (C "S" [x]) = printf "Std.Nat.succ (%s)" (ocanren x)
+  ocanren (C "S" [x]) = printf "(Std.Nat.succ (%s))" (ocanren x)
 
   ocanren (C "o" []) = "(o ())"
   ocanren (C "s" [x]) = printf "(s (%s))" (ocanren x)
@@ -35,9 +35,9 @@ instance {- OCanren v => -} OCanren (Term X) where
   -- ocanren (C "s" [x]) = printf "Std.Nat.succ (%s)" (ocanren x)
   ocanren (C "z" []) = "Nat.zero"
 
-  ocanren (C "pair" [x,y]) = printf "Pair.pair (%s) (%s)" (ocanren x) (ocanren y)
-  ocanren (C "none" []) = "Option.none ()"
-  ocanren (C "some" [x]) = printf "Option.some (%s)" (ocanren x)
+  ocanren (C "pair" [x,y]) = printf "(Pair.pair (%s) (%s))" (ocanren x) (ocanren y)
+  ocanren (C "none" []) = "(Option.none ())"
[…]

### 2020-05-28 test/auto/Test/NonConjunctive.hs

@@ -44,7 +44,7 @@ runJu l = runTest Nothing (NC.justUnfold l)
 runNc l = runTest Nothing (NC.nonConjunctive l)
 
 runRep = do
-    runJu 20 "rep" rep
+    runJu 100 "rep" rep
   where
     rep = Program Program.Programs.rep $ fresh ["n", "x"] (call "rep" [V "n", V "x"])
 

### 2020-05-28 test/resources/Program/Programs.hs

@@ -311,7 +311,7 @@ repDef =
         (n === zero &&& x === nil) |||
         (fresh ["n'", "x'"]
           (
-            n === succ n' &&& x === (C "a" []) % x' &&& call "rep" [n', x']
+            n === succ n' &&& x === zero % x' &&& call "rep" [n', x']
           )
         )
       )

### 2020-06-29 papers/aplas/related.tex
APLAS draft

--- 2020-05-23 papers/miniKanren/text/related.tex
@@
 
-While supercompilation generally improves the behavior of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in general case.
-What is worse, they rarely consider the residual program efficiency from the point of view of the target language evaluator.
-The main optimization source is computing in advance all possible intermediate and statically-known semantics steps at transformation-time.
-Other criteria like the size of the generated program or possible optimizations and execution cost of different language constructions by the target language evaluator are usually out of consideration~\cite{jonesbook}.
+While supercompilation generally improves the behaviour of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in the general.
+What is worse, they rarely consider the residual program efficiency from the target language evaluator point of view.
+The main optimization source is computing in advance all possible intermediate and statically-known semantics steps at program transformation-time.
+Other criteria, like the size of the generated program or possible optimizations and execution cost of different language constructions by the target language evaluator, are usually out of consideration~\cite{jonesbook}.
 It is known that supercompilation may adversely affect GHC optimizations yielding standalone compilation more powerful~\cite{SCBE,TCES} and cause code explosion~\cite{SCHC}.
 Moreover, it may be hard to predict the real speedup of any given program on concrete examples even disregarding problems above because of the complexity of the transformation algorithm.
-The worst case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization really improved its performance~\cite{jonesbook,bulyonkov84}.
+The worst-case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization indeed improved its performance~\cite{jonesbook,bulyonkov84}.
 There is lack of research in determining the classes of programs which transformers would definitely speed up.
@@
 
-At the local level CPD examines a conjunction of atoms by considering each atom one-by-one from left to right.
+At the local level, CPD examines a conjunction of atoms by considering each atom one-by-one from left to right.
 An atom is \emph{unfolded} if it is deemed safe, i.e. a whistle based on homeomorphic does not signal for the atom.
 When an atom is unfolded, a clause whose head can be unified with the atom is found, and a new node is added into the tree where the atom in the conjunction is replaced with the body of that clause.
-If there are more than one suitable head, then several branches are added into the tree which correspond to the disjunction in the residualized program.
+If there is more than one suitable head, then several branches are added into the tree which corresponds to the disjunction in the residualized program.
 An adaptation of CPD for the \mk{} programming language is described in~\cite{lozov2019relational}.
@@
 The most well-behaved strategy of local control in CPD for \pro{} is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
-An atom is unfolded only if only one suitable clause head exist for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
-This means that if a non-deterministic atom is the leftmost within a conjunction, it is most likely to be unfoldled and introduce many new relation calls within the conjunction.
+An atom is unfolded only if only one suitable clause head exists for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
+This means that if a non-deterministic atom is the leftmost within conjunction, it is most likely to be unfolded and introduce many new relation calls within the conjunction.
 We believe this is the core problem with CPD which limits its power when applied to \mk{}.

### 2020-06-30 papers/aplas/oldevaluation.tex
Evaluation in progress

--- 2020-05-23 papers/miniKanren/text/evaluation.tex
@@
 In our study we compared the CPD adaptation for \mk{} and the new non-conjunctive partial deduction.
-We have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check wether it can improve the quality of the specialization.
+We have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check whether it can improve the quality of the specialization.
 
@@
 
-The relation \lstinline{eval$^o$} describes evaluation of a subset of first-order logic formulas in a given substitution.
+The relation \lstinline{eval$^o$} describes an evaluation of a subset of first-order logic formulas in a given substitution.
 It has 3 arguments.
@@
 The way non-conjunctive partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in listing~\ref{eval:fst}.
-This program is easier for CPD to transform which demonstrates how unequal is the behavior of CPD for the similar programs.
+This program is easier for CPD to transform which demonstrates how unequal is the behaviour of CPD for similar programs.
 
@@
 
-Depending on the way a relation is implemented, it may take different number of driving steps to reach the point when any useful information is derived through its unfolding.
+Depending on the way a relation is implemented, it may take a different number of driving steps to reach the point when any useful information is derived through its unfolding.
 Partial deduction tries to unfold every relation call unless it is unsafe, but not all relation calls serve to restrict the search space and thus should be unfolded.
 In the implementation of \lstinline{eval$^o$} boolean operations can effectively restrict variables within the conjunctions and should be unfolded until they do.
-But depending on the way they are implemented, different number of driving steps should be performed for that.
+But depending on the way they are implemented, the different number of driving steps should be performed for that.
 The simplest way to implement these relations is with a table as demonstrated with the implementation of \lstinline{not$^o$} in listing~\ref{not:table}.
@@
 
-The other way to implement boolean operations is via one basic boolean relation such as \lstinline{nand$^o$} which is in turn has a table-based implementation (see listing~\ref{not:nando}).
+The other way to implement boolean operations is via one basic boolean relation such as \lstinline{nand$^o$} which is, in turn, has a table-based implementation (see listing~\ref{not:nando}).
 It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
@@
 
-In our study we considered two implementations of \lstinline{eval$^o$}, one we call \lstinline{plain} and the other --- \lstinline{last}, and compared how specializers behave on them.
+In our study, we considered two implementations of \lstinline{eval$^o$}, one we call \lstinline{plain} and the other --- \lstinline{last}, and compared how specializers behave on them.
 The \lstinline{plain} relation uses table-based boolean operations and places them further to the left in each conjunction.
@@
 
-We measured time necessary to generate $10000$ formulas over two variables which evaluate to \lstinline{^true}.
+We measured the time necessary to generate $10000$ formulas over two variables which evaluate to \lstinline{^true}.
 We compared the results of specialization of the goal \lstinline{eval$^o$ subst fm ^true} by our implementation of CPD, the new non-conjunctive partial deduction, and the CPD modified with the branching heuristic.
@@
 Besides evaluator of logic formulas we also run the transformers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
-These two relations are described in paper~\cite{lozov2019relational} so we will not go into too much details here.
+These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
 
@@
 The original \mk{} program fails to terminate on this goal in 300 seconds.
-On this example the most performant is the program generated by CPD (2.35 seconds) while the program generated by adding branching heuristic also fails to terminate in 300 seconds.
+On this example, the most performant is the program generated by CPD (2.35 seconds) while the program generated by adding branching heuristic also fails to terminate in 300 seconds.
 The non-conjunctive partial deduction shows some improvement with the residual program terminated within 15 seconds.
-While driving this program, non-conjunctive partial deduction does too much unfolding which negatively impact the running time as compared to CPD.
+While driving this program, the non-conjunctive partial deduction does too much unfolding which negatively impacts the running time as compared to CPD.
 
 The last test executed \lstinline{isPath} relation to search for 5 paths in the graph with 20 vertices and 30 edges.
-On this program non-conjunctive partial deduction showed better transformation results than CPD, although the difference is not that drastic.
+On this program, non-conjunctive partial deduction showed better transformation results than CPD, although the difference is not that drastic.
 
 All evaluation results are presented in the table~\ref{tbl:eval}.
-Each column correspond to the relation being run as described above.
+Each column corresponds to the relation being run as described above.
 The row marked ``Original'' contains the running time of the original \mk{} relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.

### 2020-06-30 papers/aplas/evaluation.tex
Added ecce running time for unification and path search

--- 2020-06-30 papers/aplas/oldevaluation.tex
@@
 \section{Evaluation}
-
 In our study we compared the CPD adaptation for \mk{} and the new non-conjunctive partial deduction.
@@
 
-We used the following 4 programs to test the specializers on.
+We used the following programs to test the specializers on.
 \begin{itemize}
   \item Two implementations of an evaluator of logic formulas.
+  \item Two implementations of a typechecker for a small expression language.
   \item A program to compute a unifier of two terms.
@@
 
-The last two relations are described in~\cite{lozov2019relational} thus we will not describe them here.
+The last two relations are described in~\cite{lozov2019relational} thus we will only briefly describe them here.
+
+We focused on these particular examples because they are
+
+\begin{itemize}
+  \item Relational interpreters
+  \item Natural
+  \item Observable
+\end{itemize}
 
@@
 
-\subsection{Evaluation Results}
-
+\subsubsection{Evaluation Results}
 In our study, we considered two implementations of \lstinline{eval$^o$}, one we call \lstinline{plain} and the other --- \lstinline{last}, and compared how specializers behave on them.
@@
 These two programs are complete opposites from the standpoint of CPD.
-
-We measured the time necessary to generate $10000$ formulas over two variables which evaluate to \lstinline{^true}.
-We compared the results of specialization of the goal \lstinline{eval$^o$ subst fm ^true} by our implementation of CPD, the new non-conjunctive partial deduction, and the CPD modified with the branching heuristic.
-Our evaluation confirmed that CPD behaves very differently on these two implementations of the same relation: the running time of the specialized \lstinline{last} relation is almost 6 times as high as the running time of the specialized \lstinline{plain} relation.
-The running time of two programs generated with our novel non-conjunctive partial deduction is very close and it is a little bit better than the best by CPD.
-CPD with the branching heuristic still constructs residual programs of different quality.
-The results are shown in table~\ref{tbl:eval}.
-
-Besides evaluator of logic formulas we also run the transformers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
-These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
-
-The \lstinline{unify} relation was executed to find a unifier of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
-The original \mk{} program fails to terminate on this goal in 300 seconds.
-On this example, the most performant is the program generated by CPD (2.35 seconds) while the program generated by adding branching heuristic also fails to terminate in 300 seconds.
-The non-conjunctive partial deduction shows some improvement with the residual program terminated within 15 seconds.
-While driving this program, the non-conjunctive partial deduction does too much unfolding which negatively impacts the running time as compared to CPD.
-
-The last test executed \lstinline{isPath} relation to search for 5 paths in the graph with 20 vertices and 30 edges.
-On this program, non-conjunctive partial deduction showed better transformation results than CPD, although the difference is not that drastic.
-
-All evaluation results are presented in the table~\ref{tbl:eval}.
-Each column corresponds to the relation being run as described above.
-The row marked ``Original'' contains the running time of the original \mk{} relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
+We measured the time necessary to generate $1000$ formulas over two variables which evaluate to \lstinline{^true} (averaged over 10 runs).
 
@@
   \centering
-  \begin{tabular}{c||c|c||c||c}
-                   & last & plain & unify & isPath \\
+  \begin{tabular}{c||c||c}
+                   & last    & plain  \\
   \hline\hline
-  Original         & >60.00s & >60.00s & >300.00s & 19.86s \\
+  Original         & 0.695s  & 1.426s \\
   \hline
-  CPD              & 31.31s & 5.46s & 2.35s & 4.66s \\
+  CPD              & 1.721s  & 0.168s \\
   \hline
-  Non-CPD          & 4.99s  & 5.05s & 14.90s & 3.00s \\
+  Ecce             & 0.205s  & 0.264s \\
   \hline
-  Branching        & 17.21s  & 6.17s & >300.00s & N/A \\
+  Non-CPD          & 0.132s  & 0.137s \\
   \hline
+  Branching        & 0.468s  & 0.179s \\
   \end{tabular}
@@
 
-% \subsection{Search for a Unifier}
+\subsection{Unification and Path Search}
 
-% % \begin{table}
-% %   \centering
-% %   \begin{tabular}{c|c}
+Besides evaluator of logic formulas we also run the transformers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
+These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
 
-% %     & running time \\
-% %   \hline\hline
-% %   Original         & >300.00s \\
-% %   \hline
-% %   CPD              & 2.35s \\
-% %   \hline
-% %   Non-CPD          & 14.90s \\
-% %   \hline
-% %   Branches         & >300.00s \\
-% %   \hline
-% %   \end{tabular}
+The \lstinline{unify} relation was executed to find a unifier of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
+The original \mk{} program fails to terminate on this goal in 30 seconds.
+On this example, the most performant is the program generated by CPD (0.352 seconds) while the program generated by adding branching heuristic also fails to terminate in 30 seconds.
+The non-conjunctive partial deduction shows some improvement with the residual program terminated within 1.947 seconds.
+While driving this program, the non-conjunctive partial deduction does too much unfolding which negatively impacts the running time as compared to CPD.
 
-% %   \caption{Searching for a unifier for terms f(X, X, g(Z, t)) and f(g(p, L), Y, Y)}
-% %   \label{tbl:unify}
-% % \end{table}
+The last test executed \lstinline{isPath} relation to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges.
+On this program, non-conjunctive partial deduction showed better transformation results than CPD, although the difference is not that drastic.
 
-% The unification of two terms $t$ and $u$ is searching for a substitution $\theta$ such that $t \theta = u \theta$, $\theta$ is called a unifier.
-% We search for any unifier, not necessarily most specific.
-% The details on this benchmark can be found in the earlier paper CITE.
-% This example demonstrates how too much unfolding can be introduced with the non-conjunctive partial deduction.
+All evaluation results are presented in the table~\ref{tbl:unify}.
+Each column corresponds to the relation being run as described above.
+The row marked ``Original'' contains the running time of the original \mk{} relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
 
-% \subsection{Search for Paths in a Graph}
+\begin{table}
+  \centering
+  \begin{tabular}{c||c|c|c||c}
+                   & \multicolumn{3}{c}{unify} & isPath \\
+  \hline
+                   & first  & second & third   &        \\
+  \hline\hline
+  Original         & $0.247*10^{-4}s$ & $>30s$ (oum?) & $>30s$ (oum?) & $23.743s$ \\
+  \hline
+  CPD              & $0.081*10^{-4}s$ & $0.017s$ & $0.343s$  & $2.999s$  \\
+  \hline
+  Non-CPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  & $1.786s$  \\
+  \hline
+  Ecce             & $0.128*10^{-4}s$ & $0.026s$ & $1.340s$  & $1.292s$ \\
+  \hline
+  Branching        & $0.205*10^{-4}s$ & $0.101s$ & $>30s$ (oum?) & N/A \\
+  \hline
+  \end{tabular}
 
-% Here we search for 5 paths in a graph.
-% The details on this benchmark can be found in the earlier paper CITE.
+  \caption{Evaluation results of unification and path search}
+  \label{tbl:unify}
+\end{table}
 
-% % \begin{table}
-% %   \centering
-% %   \begin{tabular}{c|c}
+\subsection{Typechecker-Term Generator}
 
-% %     & running time \\
-% %   \hline\hline
-% %   Original         & 19.86s \\
-% %   \hline
-% %   CPD              & 4.66s \\
-% %   \hline
-% %   Non-CPD          & 3.00s \\
-% %   \hline
-% %   \end{tabular}
+This relation implements a typechecker for a tiny language. Being executed in the backward direction it serves as a generator of terms of the given type.
 
-% %   \caption{Searching for paths in a graph}
-% %   \label{tbl:unify}
-% % \end{table}
+Its abstract syntax is the following.
+
+% \begin{align*}
+%   &type \ term = \\
+%   &| \ BConst \ of \ Bool \\
+%   &| \ IConst \ of \ Int \\
+%   &| \ Var \ of \ Int \\
+%   &| \ Plus \ of \ term * term \\
+%   &| \ Mult \ of \ term * term \\
+%   &| \ Eq \ of \ term * term \\
+%   &| \ Lt \ of \ term * term \\
+%   &| \ Let \ of \ term * term \\
+%   &| \ If \ of \ term * term * term
+% \end{align*}
+
+\begin{align*}
+  &type \ term = \\
+  &| \ BConst \ of \ Bool \\
+  &| \ IConst \ of \ Int \\
+  &| \ Var \ of \ Int \\
+  &| \ term + term \\
+  &| \ term * term \\
+  &| \ term = term \\
+  &| \ term < term \\
+  &| \ \underline{let} \ term \ \underline{in} \ term \\
+  &| \ \underline{if} \ term \ \underline{then} \ term \ \underline{else} \ term
+\end{align*}
+
+We use de Bruijn indices for variables.
+
+Typing rules are the following
+
+\begin{table}
+  \setlength{\tabcolsep}{0.5cm}
+  \centering
+  \begin{tabular}{c c}
+    \infer[]{\Gamma \vdash IConst \ i : Int}{} &
+    \infer[]{\Gamma \vdash BConst \ b : Bool}{} \vspace{0.5cm} \\
+
+    \infer[]{\Gamma \vdash t + s : Int}{\Gamma \vdash t : Int, \Gamma \vdash  s : Int} &
+    \infer[]{\Gamma \vdash t * s : Int}{\Gamma \vdash t : Int, \Gamma \vdash  s : Int} \vspace{0.5cm} \\
+
+    \infer[]{\Gamma \vdash t = s : Bool}{\Gamma \vdash t : \tau, \Gamma \vdash  s : \tau} &
+    \infer[]{\Gamma \vdash t < s : Bool}{\Gamma \vdash t : Int, \Gamma \vdash  s : Int} \vspace{0.5cm} \\
+
+    \infer[]{\Gamma \vdash \underline{let} \ v \ b : \tau}{\Gamma \vdash v : \tau_v, \ (\tau_v :: \Gamma) \vdash b : \tau} &
+    \infer[\Gamma \lbrack v \rbrack = \tau]{\Gamma \vdash Var \ v : \tau}{} \vspace{0.5cm} \\
+
+    \multicolumn{2}{c}{
+      \infer[]{\Gamma \vdash \underline{if} \ c \ \underline{then} \ t \ \underline{else} \ s : \tau}{\Gamma \vdash c : Bool, \Gamma \vdash t : \tau, \Gamma \vdash s : \tau}
+    }
+
+  \end{tabular}
+\end{table}
+
+Implementation 1 is a straightforward implementation in a relational programming language.
+
+Implementation 2 is obtained by translating into \mk{} an implementation of the typing rules \ocaml{}.
+
+We measured the time needed to generate 100 formulas (averaged over 10 iterations).
+
+\begin{table}
+  \centering
+  \begin{tabular}{c||c||c}
+              & Implementation 1 & Implementation 2 \\
+  \hline\hline
+  Original    & 0.057s & 0.464s\\
+  \hline
+  Non-CPD     & 0.154s & 0.448s \\
+  \hline
+  Ecce        & 2.543s  & 1.518s \\
+  \end{tabular}
+
+  \caption{Running time of generating 100 closed terms of type Int}
+  \label{tbl:eval}
+\end{table}

### 2020-06-30 papers/miniKanren/ocanren/path/build.sh

@@ -1,1 +1,1 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml spec.ml trans.ml len.ml cpd.ml run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml spec.ml trans.ml len.ml cpd.ml ecce.ml run.ml

### 2020-06-30 papers/miniKanren/ocanren/path/ecce.pl

@@ -1,0 +1,53 @@
+
+/* Specialised program generated by ECCE 2.0 */
+/* PD Goal: A */
+/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
+/* Transformation time: 120 ms */
+/* Unfolding time: 60 ms */
+/* Post-Processing time: 120 ms */
+
+/* Specialised Predicates: 
+isPath__1(A,B) :- isPath(A,B,true).
+elem_conj__2(A,B,C,D,E,F,G) :- elem(pair(A,B),cons(C,D),true), isPath(cons(B,E),cons(F,G),true).
+eqPair_conj__3(A,B,C,D,E,F,G,H,I) :- eqPair(pair(A,B),pair(C,D),false), elem(pair(A,B),cons(E,F),true), isPath(cons(B,G),cons(H,I),true).
+eqNat__4(A,B) :- eqNat(A,B,true).
+eqNat__5(A,B) :- eqNat(A,B,false).
+eqNat__6(A,B) :- eqNat(A,B,C1).
+isPath__7(A,B,C,D) :- isPath(cons(A,B),cons(C,D),true).
+*/
+
+isPath(A,B,true) :- 
+    isPath__1(A,B).
+isPath__1(nil,A).
+isPath__1(cons(A,nil),B).
+isPath__1(cons(A,cons(B,C)),cons(D,E)) :- 
+    elem_conj__2(A,B,D,E,C,D,E).
+elem_conj__2(A,B,pair(C,D),E,F,G,H) :- 
+    eqNat__4(A,C), 
+    eqNat__4(B,D), 
+    isPath__7(B,F,G,H).
+elem_conj__2(A,B,pair(C,D),cons(E,F),G,H,I) :- 
+    eqPair_conj__3(A,B,C,D,E,F,G,H,I).
+eqPair_conj__3(A,B,C,D,E,F,G,H,I) :- 
+    eqNat__5(A,C), 
+    eqNat__6(B,D), 
+    elem_conj__2(A,B,E,F,G,H,I).
+eqPair_conj__3(A,B,C,D,E,F,G,H,I) :- 
+    eqNat__4(A,C), 
+    eqNat__5(B,D), 
[…]

### 2020-06-30 papers/miniKanren/ocanren/unify/build.sh

@@ -1,2 +1,2 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml spec.ml trans.ml trans_unf.ml cpd.ml branches.ml run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml spec.ml trans.ml trans_unf.ml cpd.ml branches.ml ecce.ml run.ml
 

### 2020-06-30 papers/miniKanren/ocanren/unify/original.pl

@@ -1,0 +1,48 @@
+eq_nat(z, z, true).
+eq_nat(s(Q39), z, false).
+eq_nat(z, s(Q41), false).
+eq_nat(s(X), s(Y), Q36) :- eq_nat(X, Y, Q36).
+get_term(Var, nil, none).
+get_term(z, cons(Q32, Xs), Q32).
+get_term(s(N), cons(X, Xs), Q32) :- get_term(N, Xs, Q32).
+check_uni(Subst, constr(N1, A1), constr(N2, A2), false) :- eq_nat(N1, N2, false), forall2(Subst, A1, A2, Q13).
+check_uni(Subst, constr(N1, A1), constr(N2, A2), Q13) :- eq_nat(N1, N2, true), forall2(Subst, A1, A2, Q13).
+check_uni(Subst, var_(V), constr(N, A), false) :- get_term(V, Subst, none).
+check_uni(Subst, var_(V), constr(N, A), Q31) :- get_term(V, Subst, some(T)), check_uni(Subst, T, constr(N, A), Q31).
+check_uni(Subst, constr(N, A), var_(V), false) :- get_term(V, Subst, none).
+check_uni(Subst, constr(N, A), var_(V), Q31) :- get_term(V, Subst, some(T)), check_uni(Subst, constr(N, A), T, Q31).
+check_uni(Subst, var_(V1), var_(V2), Q31) :- get_term(V1, Subst, some(T19)), check_uni(Subst, T19, var_(V2), Q31).
+check_uni(Subst, var_(V1), var_(V2), false) :- get_term(V1, Subst, none), get_term(V2, Subst, some(Q28)).
+check_uni(Subst, var_(V1), var_(V2), Q31) :- get_term(V1, Subst, none), get_term(V2, Subst, none), eq_nat(V1, V2, Q31).
+get_term(Var, nil, none).
+get_term(z, cons(Q32, Xs), Q32).
+get_term(s(N), cons(X, Xs), Q32) :- get_term(N, Xs, Q32).
+eq_nat(z, z, true).
+eq_nat(s(Q39), z, false).
+eq_nat(z, s(Q41), false).
+eq_nat(s(X), s(Y), Q36) :- eq_nat(X, Y, Q36).
+forall2(Subst, nil, nil, true).
+forall2(Subst, cons(X, Xs), cons(Y, Ys), false) :- check_uni(Subst, X, Y, false), forall2(Subst, Xs, Ys, Q4).
[…]

### 2020-06-30 papers/miniKanren/ocanren/unify/unify.pl

@@ -1,0 +1,15 @@
+eq_nat(z, z).
+eq_nat(s(Q1), s(Q2)) :- eq_nat(Q1, Q2).
+forall2(Y5, nil, nil).
+forall2(Y5, cons(Q1, Q2), cons(Q3, Q4)) :- _check_uni(Y5, Q1, Q3), forall2(Y5, Q2, Q4).
+_check_uni(Y8, constr(Q1, Q2), constr(Q3, Q4)) :- eq_nat(Q1, Q3), forall2(Y8, Q2, Q4).
+_check_uni(Y8, var_(Q5), constr(Q6, Q7)) :- get_term(Y8, Q5, Q8), _check_uni(Y8, Q8, constr(Q6, Q7)).
+_check_uni(Y8, constr(Q9, Q10), var_(Q11)) :- get_term(Y8, Q11, Q12), _check_uni(Y8, constr(Q9, Q10), Q12).
+_check_uni(Y8, var_(Q13), var_(Q14)) :- get_term(Y8, Q13, Q15), _check_uni(Y8, Q15, var_(Q14)).
+_check_uni(Y8, var_(Q13), var_(Q14)) :- get_termGet_term(Y8, Q13, Q14), eq_nat(Q13, Q14).
+get_term(cons(some(Y13), Q1), z, Y13).
+get_term(cons(Q2, Q1), s(Q3), Y13) :- get_term(Q1, Q3, Y13).
+get_termGet_term(Y14, Y15, Y16) :- _get_term(Y14, Y15), _get_term(Y14, Y16).
+_get_term(nil, Y18).
+_get_term(cons(none, Q1), z).
+_get_term(cons(Q2, Q1), s(Q3)) :- _get_term(Q1, Q3).

### 2020-07-02 app/Main.hs
Added heuristic selection for CPD

@@ -1,22 +1,24 @@
 module Main where
 
+import           CPD.LocalControl       (Heuristic (..))
 import           NonConjunctive.Unfold
 import qualified Program.Bottles
 import qualified Program.Bridge
 import qualified Program.Bridge2
 import qualified Program.Desert
+import qualified Program.L
 import           Program.List           (appendo, maxLengtho, nil, revAcco,
                                          reverso, (%))
+import qualified Program.LLangType
 import           Program.Path
 import           Program.Programs       (doubleAppendo, rep)
 import qualified Program.Prop
 import qualified Program.Sort
 import           Program.Stlc           (evalo)
 import qualified Program.Typing
-import qualified Program.L
-import qualified Program.LLangType
 import qualified Program.Unify
 import           Syntax
+import qualified Transformer.CPD
 import qualified Transformer.JustUnfold
 import qualified Transformer.NonConj
 import qualified Transformer.PrologToMk
@@ -108,15 +110,32 @@ runTyping = do
 
 runL = do
     -- runNc (-1) "llang" Program.L.query
-    runNc (-1) "llang1" Program.L.query1
+    -- runNc (-1) "llang1" Program.L.query1
+    Transformer.CPD.transform "llangBranch" Program.L.query1 Nothing Branching
+    putStrLn "llangBranch done"
+    Transformer.CPD.transform "llangDeterm" Program.L.query1 Nothing Deterministic
+    putStrLn "llangDeterm done"
+
 
 runL' = do
-    runNc (-1) "llangPeter" Program.LLangType.query1
+    -- runNc (-1) "llangPeter" Program.LLangType.query1
+    Transformer.CPD.transform "llangPeterBranch" Program.LLangType.query1 Nothing Branching
+    putStrLn "llangPeterBranch done"
+    Transformer.CPD.transform "llangPeterDeterm" Program.LLangType.query1 Nothing Deterministic
+    putStrLn "llangPeterDeterm done"
+
+runDoubleApp = do
+    runNc (-1) "da" dA
+
 
 main :: IO ()
 main = do
+    -- runDoubleApp
+    -- runL
     -- runL'
-    Transformer.PrologToMk.transform "test/out/nc/llangPeter/ecce.pl"
+    -- runPath
+    -- runL'
+    Transformer.PrologToMk.transform "test/out/nc/da/ideal.pl"
     -- runTyping
     -- runSort
 

### 2020-07-02 src/CPD/GlobalControl.hs

@@ -85,9 +85,8 @@ conjToList (g :/\: h)     = conjToList g ++ conjToList h
 conjToList x@(Invoke _ _) = [x]
 conjToList _              = error "This conjunction is not a list of calls"
 
-
-topLevel :: Program -> (GlobalTree, G S, [S])
-topLevel (Program defs goal) =
+topLevel :: Program -> LC.Heuristic -> (GlobalTree, G S, [S])
+topLevel (Program defs goal) heuristic =
   -- let (goal', defs) = takeOutLets goal in
   let gamma = E.updateDefsInGamma E.env0 defs in
   let (logicGoal, gamma', names) = E.preEval gamma goal in
@@ -102,7 +101,7 @@ topLevel (Program defs goal) =
         -- let newNodes = (delete goal nodes) in
         let newNodes = filter (not . Embed.isVariant goal) nodes in
 
-        let sldTree = LC.sldResolution goal gamma subst newNodes in
+        let sldTree = LC.sldResolution goal gamma subst newNodes heuristic in
         let (substs, bodies) = partition (null . snd3) $ LC.resultants sldTree in
 
         if null bodies

### 2020-07-02 src/CPD/LocalControl.hs

@@ -21,6 +21,8 @@ import           Util.Miscellaneous
 -- trace :: String -> a -> a
 -- trace _ x = x
 
+data Heuristic = Deterministic | Branching
+
 data Descend a = Descend { getCurr :: a, getAncs :: [a] } deriving (Eq)
 
 instance (Show a) => Show (Descend a) where
@@ -60,17 +62,17 @@ instance E.Subst [Descend (G S)] where
   substitute s =
     map $ \(Descend g ancs) -> Descend (E.substitute s g) ancs
 
-sldResolution :: [G S] -> E.Gamma -> E.Sigma -> [[G S]] -> SldTree
-sldResolution goal gamma subst seen  =
+sldResolution :: [G S] -> E.Gamma -> E.Sigma -> [[G S]] -> Heuristic -> SldTree
+sldResolution goal gamma subst seen heuristic =
   -- sldResolutionStep (map (\x -> Descend x Set.empty) goal) gamma subst Set.empty True
   -- trace "\n\nSLDRESOLUTION \n\n" $
-  sldResolutionStep (map (\x -> Descend x []) goal) gamma subst seen True
+  sldResolutionStep (map (\x -> Descend x []) goal) gamma subst seen True heuristic
 
 showList :: Show a => [a] -> String
 showList = unlines . map show
 
-sldResolutionStep :: [DescendGoal] -> E.Gamma -> E.Sigma -> [[G S]] -> Bool -> SldTree
-sldResolutionStep gs env@(p, i, d@(temp:_)) s seen isFirstTime =
+sldResolutionStep :: [DescendGoal] -> E.Gamma -> E.Sigma -> [[G S]] -> Bool -> Heuristic -> SldTree
+sldResolutionStep gs env@(p, i, d@(temp:_)) s seen isFirstTime heuristic =
   let curs = map getCurr gs in
   let prettySeen = showList seen  in
   -- if variantCheck curs seen
@@ -114,14 +116,14 @@ sldResolutionStep gs env@(p, i, d@(temp:_)) s seen isFirstTime =
             [] ->
               Fail
             -- ns | length ns == 1 || isFirstTime -> -- unfold only if it's deterministic or hasn't been unfolded before
-            ns | getMaximumBranches env' g > length ns || isFirstTime ->
+            ns | needsUnfolding heuristic env' g ns isFirstTime -> -- getMaximumBranches env' g > length ns || isFirstTime ->
               Or (map step ns) (Just g) s
               where
                 step (xs, s') =
[…]

### 2020-07-02 src/OCanrenize.hs

@@ -21,7 +21,7 @@ instance {- OCanren v => -} OCanren (Term X) where
   ocanren (C nil _) | map toLower nil == "nil"  = "(Std.List.nil ())"
   -- ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "(%s %% %s)" (ocanren h) (ocanren t)
   -- ocanren (C "%"    [h,t]) = printf "(%s %% %s)" (ocanren h) (ocanren t)
-  ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "(Std.(%%) %s %s)" (ocanren h) (ocanren t)
+  ocanren (C cons [h,t]) | map toLower cons == "cons" = printf "(Std.(%%) (%s) (%s))" (ocanren h) (ocanren t)
   ocanren (C "%" [h, t]) = printf "(Std.(%%) (%s) (%s))" (ocanren h) (ocanren t)
   ocanren (C "O" []) = "Std.Nat.zero"
   ocanren (C "S" [x]) = printf "(Std.Nat.succ (%s))" (ocanren x)
@@ -62,6 +62,8 @@ instance {-OCanren v =>-} OCanren (G X) where
   ocanren (g1 :\/: g2)  = printf "(%s ||| %s)" (ocanren g1) (ocanren g2)
   ocanren (Fresh x g )  = let (names, goal) = freshVars [x] g in printf "(fresh (%s) (%s))" (printArgs names) (ocanren goal)
 --ocanren (Invoke f ts) = printf "(print_string \"%s\\n\";%s)" (f ++ concat [' ' : ocanren t | t <- ts]) (f ++ concat [' ' : ocanren t | t <- ts])
+  ocanren (Invoke "success" []) = "success"
+  ocanren (Invoke "fail" []) = "fail"
   ocanren (Invoke f ts) = printf "(%s %s)" f (printArgs $ map ocanren ts)
   ocanren (Let (Def n as b) g) = printf "let rec %s %s = %s in %s" n (printArgs as) (ocanren b) (ocanren g)
 

### 2020-07-02 src/Transformer/CPD.hs

@@ -1,0 +1,182 @@
+{-# LANGUAGE BangPatterns #-}
+
+module Transformer.CPD where
+
+import           Control.Monad
+import qualified CPD.GlobalControl        as GC
+import           CPD.LocalControl
+import           CPD.Residualization
+import           Data.Foldable            (for_)
+import           Data.List
+import           Data.Maybe
+import qualified Data.Set                 as Set
+import           Debug.Trace
+import qualified Eval                     as E
+import qualified OCanrenize               as OC
+import           Prelude                  hiding (succ)
+import           Printer.Dot
+import           Printer.GlobalTree
+import           Printer.SldTree
+import           Program.Bool
+import qualified Program.Bottles
+import           Program.Bridge
+import qualified Program.Desert
+import           Program.List
+import           Program.LogicInterpreter
+import           Program.Num
+import           Program.Path             hiding (elem)
+import           Program.Programs
+import           Program.Prop
+import           Program.Sample1
+import           Program.Sort
+import           Program.SpecialProp
+import qualified Program.Sudoku4x4
+import           Program.Unify
+import           Purification
+import           Residualize
+import           Syntax
+import           System.Directory
+import           Text.Printf
+import           Unfold
+import           Util.ConjRetriever
+import           Util.Miscellaneous
+import           Embed
+import           System.CPUTime
+import           System.IO
+import           System.Process           (system)
+import           System.TimeIt
+
+transform filename goal env heuristic = do
+    let (tree, logicGoal, names) = GC.topLevel goal heuristic
+    -- traceM ("\n\n NAMES \n\n" ++ show (vident <$> reverse names) )
+    let path = printf "test/out/%s" filename
+    exists <- doesDirectoryExist path
+    if exists
+    then removeDirectoryRecursive path
[…]

### 2020-07-02 papers/aplas/evaluation.tex
Evaluation in progress

@@
 In our study we compared the CPD adaptation for \mk{} and the new non-conjunctive partial deduction.
-We have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check whether it can improve the quality of the specialization.
+For some programs we have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check whether it can improve the quality of the specialization.
+We measured the execution time of each specialized program for some queries and compared it with the execution time of the original program.
 
@@
 \begin{itemize}
+  \item A program to compute a concatenation of three lists.
+  \item A program to compute both the length of the list and its maximal element.
   \item Two implementations of an evaluator of logic formulas.
@@
 
-The last two relations are described in~\cite{lozov2019relational} thus we will only briefly describe them here.
+The last two relations are described in~\cite{lozov2019relational} thus we will only briefly describe them in this paper.
 
-We focused on these particular examples because they are
+We focused on these particular examples because they are examples of relational interpreters which are natural and observable.
 
-\begin{itemize}
-  \item Relational interpreters
-  \item Natural
-  \item Observable
-\end{itemize}
+\subsection{Concatenation of Three lists}
 
-% All these relations are relational interpreters.
-% Their last argument is a boolean value which is $true$ if the other arguments are in relation and $false$ otherwise.
+The relation \lstinline{doubleAppend$^o$} concatenates three lists by a conjunction of two calls of the \lstinline{append$^o$} relation (see fig.~\ref{doubleApp}).
+These two calls share a variable --- an intermediate list \lstinline{ts} --- which should be treated with care during specialization.
+Partial deduction does ignore the sharing of the variables which renders the approach to be unproductive for this kind of relations.
+This is why this relation is particularly well known in the conjunctive partial deduction work.
+
+The first list gets traversed twice to construct the result.
+First, when the intermidiate list \lstinline{ts} is constructed during the first \lstinline{append$^o$} call and then, when \lstinline{ts} is read during the second call.
+This double traversal negatively impacts the execution time.
+
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.8\textwidth}
+    \begin{lstlisting}[label={doubleApp}, caption={Concatenation of three lists}, captionpos=b, frame=tb]
+  let doubleAppend$^o$ xs ys zs res =
+    fresh (ts) (
+      append$^o$ xs ys ts /\ append$^o$ ts zs res)
+
+  let rec append$^o$ xs ys zs = conde [
+    (xs === [] /\ ys === zs);
+    fresh (h t r) (
+      xs === (h % t) /\
+      zs === (h % r) /\
+      append$^o$ t ys r)]
+    \end{lstlisting}
+  \end{minipage}
+\end{figure*}
+
+The better implementation of this relation (see fig.~\ref{doubleApp:etalon}) does not traverse the first list twice.
+This etalon implementation is generated with the conjunctive partial deduction.
+
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.8\textwidth}
+    \begin{lstlisting}[label={doubleApp:etalon}, caption={Etalon implementation of concatenation of three lists}, captionpos=b, frame=tb]
+  let doubleAppend$^o$ xs ys zs res = conde [
+    (xs === [] /\ append$^o$ ys zs res);
+    fresh (h t r) (
+      xs === h % t /\
+      res === h % r /\
+      doubleAppend$^o$ t ys zs r)]
+    \end{lstlisting}
+  \end{minipage}
+\end{figure*}
+
+We run all translated programs in the forward and the backward direction.
+The forward direction \lstinline{doubleAppend$^o$ x y z ?} concatenates three given lists: the first one of length 700, the second and the third --- of length 1.
+When run in the backward direction \lstinline{doubleAppend$^o$ ? ? ? r} , it searches for the first 100 list triples whose concatenation gives the given list of length 700.
+
+There is no significant difference in the execution time in the backward direction between different specialized versions~\ref{tbl:doubleApp}.
+However the execution time in the backward direction differs: the program generated by Ecce shows the same speedup as the etalon program, while non-conjunctive partial deduction slows the program down.
+\todo{WHY?}
+
+\begin{table}
+  \centering
+  \begin{tabular}{c||c||c}
+                   & forward & backward \\
+  \hline\hline
+  Original         & 0.0040s & 0.0195s \\ \hline
+  Ecce             & 0.0031s & 0.0201s \\ \hline
+  Non-CPD          & 0.0055s & 0.0202s \\ \hline
+  Etalon           & 0.0031s & 0.0203s \\
+  \end{tabular}
+  \caption{Evaluation results for doubleAppendo}
+  \label{tbl:doubleApp}
+\end{table}
 

### 2020-07-02 papers/miniKanren/ocanren/da/build.sh
Double appendo example

@@ -1,0 +1,2 @@
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml da.ml ideal.ml ecce.ml run.ml
+

### 2020-07-02 papers/miniKanren/ocanren/da/da.pl

@@ -1,0 +1,3 @@
+doubleAppendo(Y0, Y1, Y2, Y3) :- appendo(Y0, Y1, Q1), appendo(Q1, Y2, Y3).
+appendo(nil, Y6, Y6).
+appendo(cons(Q1, Q2), Y5, cons(Q1, Q3)) :- appendo(Q2, Y5, Q3).

### 2020-07-02 papers/miniKanren/ocanren/da/ecce.pl

@@ -1,0 +1,27 @@
+
+/* Specialised program generated by ECCE 2.0 */
+/* PD Goal: A */
+/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
+/* Transformation time: 20 ms */
+/* Unfolding time: 20 ms */
+/* Post-Processing time: 20 ms */
+
+/* Specialised Predicates: 
+doubleAppendo__1(A,B,C,D) :- doubleAppendo(A,B,C,D).
+appendo_conj__2(A,B,C,D) :- appendo(A,B,E1), appendo(E1,C,D).
+appendo__3(A,B,C) :- appendo(A,B,C).
+*/
+
+doubleAppendo(A,B,C,D) :- 
+    doubleAppendo__1(A,B,C,D).
+doubleAppendo__1(nil,A,B,C) :- 
+    appendo__3(A,B,C).
+doubleAppendo__1(cons(A,B),C,D,cons(A,E)) :- 
+    appendo_conj__2(B,C,D,E).
+appendo_conj__2(nil,A,B,C) :- 
+    appendo__3(A,B,C).
+appendo_conj__2(cons(A,B),C,D,cons(A,E)) :- 
+    appendo_conj__2(B,C,D,E).
+appendo__3(nil,A,A).
+appendo__3(cons(A,B),C,cons(A,D)) :- 
+    appendo__3(B,C,D).

### 2020-07-02 papers/miniKanren/ocanren/da/ideal.pl

@@ -1,0 +1,5 @@
+doubleAppendo([H|T],Y,Z,[H|R]) :- doubleAppendo(T,Y,Z,R).
+doubleAppendo([], Y, Z, R) :- appendo(Y,Z,R).
+
+appendo([H|T], Y, [H|R]) :- appendo(T, Y, R).
+appendo([], Y, Y).

### 2020-07-02 papers/miniKanren/ocanren/llang2/build.sh
CPD for llang typechecker

@@ -1,1 +1,1 @@
-ocamlfind c -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml l.ml l_ocanren.ml original.ml ecce.ml llang.ml llangPeter.ml eccePeter.ml run.ml
+ocamlfind c -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml l.ml l_ocanren.ml original.ml ecce.ml llang.ml llangPeter.ml eccePeter.ml  cpdPeterBranch.ml cpdDeterm.ml cpdPeterDeterm.ml run.ml

### 2020-07-03 papers/aplas/evaluation.tex
Maxlen example

@@
 
-\subsection{Concatenation of Three lists}
+\subsection{Concatenation of Three Lists}
 
@@
 \end{table}
+
+\subsection{Maximum Element and Length of a List}
+
+The relation \lstinline{maxLenght$^o$} computes both the length of the list and its maximum element (see fig.~\ref{cpd:maxandlength}) by concatenation of two relation calls.
+This relation traverses the list \lstinline{xs} twice: first to compute the maximum and then to compute the length of the list.
+It can be transformed in such a way that the list is only traversed once while computing both components of the result simultaneously.
+This transformation is called \emph{tupling} and can be achieved with conjunctive partial deduction.
+
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.8\textwidth}
+\begin{lstlisting}[label={cpd:maxandlength}, caption={Maximum element and length of the list}, captionpos=b, frame=tb]
+  let maxLength$^o$ xs m l = max$^o$ xs m /\ length$^o$ xs l
+  let rec length$^o$ xs l = conde [
+    (xs === [] /\ l === zero);
+    (fresh (h t m) (
+      xs === h % t /\ l === succ m /\ length$^o$ t m))]
+  let max$^o$ xs m = max$_1^o$ xs zero m
+  let rec max$_1^o$ xs n m = conde [
+    (xs === [] /\ m === n);
+    (fresh (h t) (
+      (xs === h % t) /\
+      (conde [
+        (le$^o$ h n ^true /\ max$_1^o$ t n m);
+        (gt$^o$ h n ^true /\ max$_1^o$ t h m)])))]
+  let rec le$^o$ x y b = conde [
+    (x === zero /\ b === ^true);
+    (fresh (x$_1$) (
+      x === succ x$_1$ /\ y === zero /\ b === ^false));
+    (fresh (x$_1$ y$_1$) (
+      x === succ x$_1$ /\ y === succ y$_1$ /\ le$^o$ x$_1$ y$_1$ b))]
+  let rec gt$^o$ x y b = conde [
+    (x === zero /\ b === ^false);
+    (fresh (x$_1$) (
+      x === succ x$_1$ /\ y === zero /\ b === ^false));
+    (fresh (x$_1$ y$_1$) (
+      x === succ x$_1$ /\ y === succ y$_1$ /\ gt$^o$ x$_1$ y$_1$ b))]
+  \end{lstlisting}
+\end{minipage}
+\end{figure*}
+
+The etalon implementation is shown in fig.~\ref{etalon:maxandlength}.
+It uses a recursive relation to compute the length and maximum element simultaneously while traversing the list once.
+
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.8\textwidth}
+\begin{lstlisting}[label={etalon:maxandlength}, caption={Etalon implementation of maxlengtho}, captionpos=b, frame=tb]
+  let maxLength$^o$ xs m l = maxLength$_1^o$ xs m zero l
+  let rec maxLength$_1^o$ xs m n l = conde [
+    (xs === [] /\ m === n /\ l === zero);
+    (fresh (h t l$_1$)
+       (xs === h % t) /\
+       (l === succ l$_1$) /\
+       (conde [
+         (le$^o$ h n ^true /\ maxLength$_1^o$ t m n l);
+         (gt$^o$ h n ^true /\ maxLength$_1^o$ t m h l)]))]
+  \end{lstlisting}
+\end{minipage}
+\end{figure*}
+
+We measured the execution time of \lstinline{maxLength$^o$ lst m l}, where \lstinline{lst} is a list of Peano numbers from $1$ to $100$.
+Note, that the second disjunct in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
+Thus among others we compared two etalon implentations: one includes these disjuncts, while the other removes them (see table~\ref{tbl:maxlen})
+
+\begin{table}
+  \centering
+  \begin{tabular}{c||c}
+                   & [1..100] \\ \hline\hline
+  Original         & 2.286ms  \\ \hline
+  Etalon           & 4.623ms  \\ \hline
+  Etalon removed   & 3.133ms  \\ \hline
+  Ecce             & 1.701ms  \\ \hline
+  Non-CPD          & 2.578ms
+  \end{tabular}
+
+  \caption{Execution time of maxlengtho}
+  \label{tbl:maxlen}
+\end{table}
+
+Surprising enough, the execution time of the etalon program is worse than of the original.
+Ecce succeeds at improving the program performance, while non-conjunctive partial deduction worsens it a little.
+\todo{why}
 

### 2020-07-03 papers/miniKanren/ocanren/maxlen/build.sh

@@ -1,2 +1,2 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml maxlen.ml ideal.ml len.ml run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml maxlen.ml ideal.ml ideal2.ml len.ml ecce.ml run.ml
 

### 2020-07-03 papers/miniKanren/ocanren/maxlen/ecce_basic.pl

@@ -1,0 +1,46 @@
+
+/* Specialised program generated by ECCE 2.0 */
+/* PD Goal: A */
+/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
+/* Transformation time: 80 ms */
+/* Unfolding time: 80 ms */
+/* Post-Processing time: 80 ms */
+
+/* Specialised Predicates:
+maxLengtho__1(A,B,C) :- maxLengtho(A,B,C).
+maxo1_conj__2(A,B,C) :- maxo1(A,o,B), lengtho(A,C).
+maxo1_conj__3(A,B,C,D) :- maxo1(A,s(B),C), lengtho(A,D).
+leo_conj__4(A,B,C,D,E) :- leo(A,s(B),true), maxo1(C,s(B),D), lengtho(C,E).
+gto__5(A,B) :- gto(s(A),B,true).
+leo__6(A,B) :- leo(A,B,true).
+*/
+
+maxLengtho(A,B,C) :-
+  maxLengtho__1(A,B,C).
+maxLengtho__1(nil,o,o).
+maxLengtho__1(cons(o,A),B,s(C)) :-
+  maxo1_conj__2(A,B,C).
+maxLengtho__1(cons(s(A),B),C,s(D)) :-
+  maxo1_conj__3(B,A,C,D).
+maxo1_conj__2(nil,o,o).
+maxo1_conj__2(cons(o,A),B,s(C)) :-
+  maxo1_conj__2(A,B,C).
+maxo1_conj__2(cons(s(A),B),C,s(D)) :-
+  maxo1_conj__3(B,A,C,D).
+maxo1_conj__3(nil,A,s(A),o).
+maxo1_conj__3(cons(A,B),C,D,s(E)) :-
+  leo_conj__4(A,C,B,D,E).
+maxo1_conj__3(cons(s(s(A)),B),C,D,s(E)) :-
+  gto__5(A,C),
+  maxo1_conj__3(B,s(A),D,E).
+leo_conj__4(o,A,B,C,D) :-
+  maxo1_conj__3(B,A,C,D).
+leo_conj__4(s(A),B,C,D,E) :-
+  leo__6(A,B),
+  maxo1_conj__3(C,B,D,E).
[…]

### 2020-07-03 papers/aplas/evaluation.tex
Evaluation draft

@@
   \centering
-  \begin{minipage}{0.8\textwidth}
+  \begin{minipage}{0.6\textwidth}
     \begin{lstlisting}[label={doubleApp}, caption={Concatenation of three lists}, captionpos=b, frame=tb]
@@
   \centering
-  \begin{minipage}{0.8\textwidth}
+  \begin{minipage}{0.6\textwidth}
     \begin{lstlisting}[label={doubleApp:etalon}, caption={Etalon implementation of concatenation of three lists}, captionpos=b, frame=tb]
@@
   \centering
-  \begin{minipage}{0.8\textwidth}
+  \begin{minipage}{0.85\textwidth}
 \begin{lstlisting}[label={cpd:maxandlength}, caption={Maximum element and length of the list}, captionpos=b, frame=tb]
@@
     (x === zero /\ b === ^true);
-    (fresh (x$_1$) (
-      x === succ x$_1$ /\ y === zero /\ b === ^false));
-    (fresh (x$_1$ y$_1$) (
-      x === succ x$_1$ /\ y === succ y$_1$ /\ le$^o$ x$_1$ y$_1$ b))]
+    (fresh (x$_1$) (x === succ x$_1$ /\ y === zero /\ b === ^false));
+    (fresh (x$_1$ y$_1$) (x === succ x$_1$ /\ y === succ y$_1$ /\ le$^o$ x$_1$ y$_1$ b))]
   let rec gt$^o$ x y b = conde [
     (x === zero /\ b === ^false);
-    (fresh (x$_1$) (
-      x === succ x$_1$ /\ y === zero /\ b === ^false));
-    (fresh (x$_1$ y$_1$) (
-      x === succ x$_1$ /\ y === succ y$_1$ /\ gt$^o$ x$_1$ y$_1$ b))]
+    (fresh (x$_1$) (x === succ x$_1$ /\ y === zero /\ b === ^false));
+    (fresh (x$_1$ y$_1$) (x === succ x$_1$ /\ y === succ y$_1$ /\ gt$^o$ x$_1$ y$_1$ b))]
   \end{lstlisting}
@@
   \centering
-  \begin{minipage}{0.8\textwidth}
+  \begin{minipage}{0.7\textwidth}
 \begin{lstlisting}[label={etalon:maxandlength}, caption={Etalon implementation of maxlengtho}, captionpos=b, frame=tb]
@@
   \centering
-  \begin{minipage}{0.8\textwidth}
+  \begin{minipage}{0.95\textwidth}
     \begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b, frame=tb]
@@
   \centering
-  \begin{minipage}{0.85\textwidth}
+  \begin{minipage}{0.95\textwidth}
     \begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b, frame=tb]
@@
   \centering
-  \begin{minipage}{0.7\textwidth}
+  \begin{minipage}{0.85\textwidth}
     \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b, frame=tb]
@@
   \begin{tabular}{c||c|c|c||c}
-                   & \multicolumn{3}{c}{unify} & isPath \\
+                   & \multicolumn{3}{c||}{unify} & isPath \\
   \hline
@@
 
-This relation implements a typechecker for a tiny language. Being executed in the backward direction it serves as a generator of terms of the given type.
-
-Its abstract syntax is the following.
+This relation implements a typechecker for a tiny language.
+Being executed in the backward direction it serves as a generator of terms of the given type.
+The abstract syntax of the language is presented below.
+The variables are represented with de Bruijn indices, thus let-binding does not specify which variable is being bound.
 
@@
 
-We use de Bruijn indices for variables.
-
-Typing rules are the following
+The typing rules are straightforward and are presented below.
+Boolean and integer constants have the corresponding types regardless of the environment.
+Only terms of type integer can be summed up, multiplied or checked for being less or equal.
+Any terms of the same type can be checked for equality.
+If-then-else expression typechecks only if its condition is of type boolean, while both then- and else-branches have the same type.
+An environment $\Gamma$ is an ordered list, in which the $i$-th element is the type of the variable with the $i$-th de Bruijn index.
+To typecheck a let-binding, first, the term being bound is typechecked and is added in the beginning of the environment $\Gamma$, and then the body is typechecked in the context of the new environment.
+Typechecking a variable with the index $i$ boils down to getting a $i$-th element of the list.
 
@@
 
-Implementation 1 is a straightforward implementation in a relational programming language.
+We compared two possible implementations of these typing rules.
+The first one is obtained by unnesting of the functional program as described in~\cite{lozov2019relational}.
+The second program is hand-written in \oc{}.
+Each implementation has been transformed with non-conjunctive partial deducer and by Ecce.
 
-Implementation 2 is obtained by translating into \mk{} an implementation of the typing rules \ocaml{}.
-
-We measured the time needed to generate 100 formulas (averaged over 10 iterations).
+We measured the time needed to generate 100 terms of type integer (averaged over 10 iterations).
+Ecce significantly worsens the execution time for both implementations.
+Non-conjunctive partial deduction improves the first implementation a little bit, while worsening the performance of the second implementation 3 times.
 
@@
   \begin{tabular}{c||c||c}
-              & Implementation 1 & Implementation 2 \\
-  \hline\hline
-  Original    & 0.057s & 0.464s\\
-  \hline
-  Non-CPD     & 0.154s & 0.448s \\
-  \hline
-  Ecce        & 2.543s  & 1.518s \\
+              & Implementation 1 & Implementation 2 \\ \hline\hline
+  Original    & 0.464s           & 0.057s           \\ \hline
+  Non-CPD     & 0.448s           & 0.154s           \\ \hline
+  Ecce        & 1.518s           & 2.543s           \\
   \end{tabular}
@@
 \end{table}
+
+\todo{why}

### 2020-07-03 papers/aplas/evaluation.tex
Evaluation in progress.

@@
 We focused on these particular examples because they are examples of relational interpreters which are natural and observable.
+In this study we only measured the execution time for sample queries.
+The measured time was averaged over multiple runs for every query.
+The queries were run on a laptop running Ubuntu 18.04 with quad core Intel Core i5 2.30GHz CPU and 8 GB of RAM.
+
+\todo{intro sentence.}
+Row Original contains the execution time of the original version of the program before any transformations.
+Row Ecce corresponds to the result of the transformation with Ecce conjunctive partial deduction system.
+Row Non-CPD --- translation by our implementation of nonconjunctive partial deduction.
+Etalon --- the etalon implementation of the same program.
+CPD --- the earlier implementation of the conjunctive partial deduction for \mk{}.
+Brancing --- the earlier implemenctation of CPD for \mk{} with the branching heuristic.
 

### 2020-07-04 papers/aplas/text/intro.tex
Move APLAS draft to another folder

--- 2020-05-23 papers/miniKanren/text/intro.tex
@@
 The core feature of the family of relational programming languages \mk{}\footnote{\mk{} language web site: \url{http://minikanren.org}} is their ability to run a program in different directions.
-Having specified a relation for adding two numbers, one can also compute the substraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
+Having specified a relation for adding two numbers, one can also compute the subtraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
 Program synthesis can be done by running \emph{backwards} a relational interpreter for some language.
@@
 
-Specialization or partial evaluation~\cite{jonesbook} is a technique aimed at improving performance of a program given some information about it beforehand.
-It may either be a known value of some argument, its structure (i.e. the length of an input list) or, in case of relational program, --- the direction in which it is intended to be run.
+Specialization or partial evaluation~\cite{jonesbook} is a technique aimed at improving the performance of a program given some information about it beforehand.
+It may either be a known value of some argument, its structure (i.e. the length of an input list) or, in case of a relational program, --- the direction in which it is intended to be run.
 An earlier paper~\cite{lozov2019relational} showed that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk{} programs.
@@
 The ideas described there are aimed at left-to-right evaluation strategy of \pro{}.
-Since the search in \mk{} is complete, it is safe to reorder come relation calls within the goal for better performance.
+Since the search in \mk{} is complete, it is safe to reorder some relation calls within the goal for better performance.
 While sometimes conjunctive partial deduction gives great performance boost, sometimes it does not behave as well as it could have.
 
-In this paper we show on examples some issues which conjunctive partial deduction faces.
+In this paper, we show on examples some issues which conjunctive partial deduction faces.
 We also describe a novel approach to partial deduction of a relational programming language \mk{}.

### 2020-07-04 papers/miniKanren/ocanren/maxlen/build.sh
Conjunction is made to be right-associative

@@ -1,2 +1,2 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml maxlen.ml ideal.ml ideal2.ml len.ml ecce.ml run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml maxlen.ml ideal.ml ideal2.ml len.ml ecce.ml ecce_right.ml run.ml
 

### 2020-07-04 src/Syntax.hs

@@ -87,14 +87,14 @@ call = Invoke
 
 disj :: [G a] -> Maybe (G a)
 disj [] = Nothing
-disj xs = Just $ foldl1 (:\/:) xs
+disj xs = Just $ foldr1 (:\/:) xs
 
 conj :: [G a] -> Maybe (G a)
 conj [] = Nothing
-conj xs = Just $ foldl1 (:/\:) xs
+conj xs = Just $ foldr1 (:/\:) xs
 
 unsafeConj :: [G a] -> G a
-unsafeConj = foldl1 (:/\:)
+unsafeConj = foldr1 (:/\:)
 
 successName :: String
 successName = "success"

### 2020-07-04 src/Util/ToProlog.hs

@@ -280,7 +280,7 @@ ruleToG v ((name, a ), b ) =
   let conjs1 = map (\(v,t) -> V v === t) $ zip v a in
   let conjs2 = map (\(n,a) -> Invoke n a) b in
   let conjs  = conjs1 ++ conjs2 in
-  let g      = if null conjs then success else foldl1 (&&&) conjs in
+  let g      = if null conjs then success else foldr1 (&&&) conjs in
   fresh (fvg g \\ v) g
 
 {-------------------------------------------}
@@ -288,7 +288,7 @@ rulesToDef :: Rules -> Def
 rulesToDef rs@(((n,a),_):_) =
   let v     = map (('z':) . show) [1..length a] in
   let disjs = map (ruleToG v) rs in
-  Def n v $ foldl1 (|||) disjs
+  Def n v $ foldr1 (|||) disjs
 
 {-------------------------------------------}
 prologToG :: String -> (G X, [String], [Def])

### 2020-07-04 papers/aplas/text/evaluation.tex

--- 2020-07-03 papers/aplas/evaluation.tex
@@
 
-\begin{align*}
-  &type \ term = \\
-  &| \ BConst \ of \ Bool \\
-  &| \ IConst \ of \ Int \\
-  &| \ Var \ of \ Int \\
-  &| \ term + term \\
-  &| \ term * term \\
-  &| \ term = term \\
-  &| \ term < term \\
-  &| \ \underline{let} \ term \ \underline{in} \ term \\
-  &| \ \underline{if} \ term \ \underline{then} \ term \ \underline{else} \ term
-\end{align*}
+%% \begin{align*}
+%%   &type \ term = \\
+%%   &| \ BConst \ of \ Bool \\
+%%   &| \ IConst \ of \ Int \\
+%%   &| \ Var \ of \ Int \\
+%%   &| \ term + term \\
+%%   &| \ term * term \\
+%%   &| \ term = term \\
+%%   &| \ term < term \\
+%%   &| \ \underline{let} \ term \ \underline{in} \ term \\
+%%   &| \ \underline{if} \ term \ \underline{then} \ term \ \underline{else} \ term
+%% \end{align*}
+
+\[\begin{array}{lllll}
+  type \ term = &\ BConst \ of \ Bool &| \ IConst \ of \ Int &| \ Var \ of \ Int & \\
+  & | \ term + term &| \ term * term &| \ term = term &| \ term < term \\
+  &| \ \underline{let} \ term \ \underline{in} \ term 
+  &\multicolumn{2}{l}{| \ \underline{if} \ term \ \underline{then} \ term \ \underline{else} \ term} &
+\end{array}\]
 

### 2020-07-05 papers/aplas/text/abstract.tex
Evaluation intro section rewritten

\begin{abstract}
  We explore partial deduction for \mk: a specialization technique aimed at improving the performance of a relation in the given direction.
  We describe a novel approach to specialization of \mk based on partial deduction and supercompilation.
  On several examples, we demonstrate issues which arise during partial deduction.
\end{abstract}

### 2020-07-05 papers/aplas/text/evaluation.tex
Evaluation intro section rewritten

--- 2020-07-03 papers/aplas/evaluation.tex
@@
 \section{Evaluation}
-In our study we compared the CPD adaptation for \mk{} and the new non-conjunctive partial deduction.
-For some programs we have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check whether it can improve the quality of the specialization.
-We measured the execution time of each specialized program for some queries and compared it with the execution time of the original program.
+In our study we compared the new non-conjunctive partial deduction with the \ecce partial deduction system.
+\ecce is designed for \pro programming language and cannot be directly applied to \mk.
+To be able to compare our approach with \ecce, we converted each input program first to the pure subset of \pro, then transformed it with \ecce, and then we converted the result back to the pure subset of \pro.
+The coversion to \pro is a simple syntactic conversion.
+In conversion from \pro to \mk, a conjunction is generated by each horn clause in which unifications are placed before any relation call.
+This mimics in \mk the execution order of \pro program within a conjunction.
 
-We used the following programs to test the specializers on.
-\begin{itemize}
-  \item A program to compute a concatenation of three lists.
-  \item A program to compute both the length of the list and its maximal element.
-  \item Two implementations of an evaluator of logic formulas.
-  \item Two implementations of a typechecker for a small expression language.
-  \item A program to compute a unifier of two terms.
-  \item A program to search for paths of a specific length in a graph.
-\end{itemize}
+The set of benchmarks consist of both examples from the literature on partial deduction and programs of our own creation.
+The first two programs are well-known in conjunctive partial deduction literature~\cite{de1999conjunctive}.
+The second two are introduced in the paper on relational interpreters~\cite{lozov2019relational}.
+The last two are examples of relational interpreters created by us which are observable, but non-trivial.
+In the last two examples we explore how different implementations of the same relation affects the quality of transformations.
 
-The last two relations are described in~\cite{lozov2019relational} thus we will only briefly describe them in this paper.
+The \lstinline{doubleAppend$^o$} program concatenates three lists of numbers by computing a conjunction of two calls of a relation which concatenates two lists.
+This example is prominent in CPD literature, since it demonstrates \emph{deforestation} --- a transformation which gets rid of intermidiate datastructures.
+This transformation cannot be done with the earlier partial deduction approach, since the relation uses a variable shared between two relation calls within a conjunction which requires the transformation to treat conjunctions as a whole.
+We explore how the fact that our approach sometimes splits conjunctions ignoring the variable sharing affects the quality of transformation.
 
-We focused on these particular examples because they are examples of relational interpreters which are natural and observable.
+The \lstinline{maxLength$^o$} program computes the maximum element of the list as well as its length.
+In this program an input list is shared between the call of the relation to compute the maximum element and the relation to compute the length of the list.
+This sharing leads to the list being traversed twice during execution.
+This example is used to demonstrate \emph{tupling}: the transformation in which multiple traversals of the same data structure are replaced with a single traversal which computes all the necessary results simultaneously.
+CPD performes tupling more often than our approach, and we explore how it affects the performance of the transformed program.
+
+The \lstinline{isPath} and \lstinline{unify} relations demonstrate the relation interpretation approach in which a relation is generated from a functional program.
+Partial deduction is capable to improve the execution time of the generated relations in the given direction.
+These two relations are described in~\cite{lozov2019relational} thus we will only briefly describe them in this paper.
+Here the aim is to study if our approach can help to achieve comparable performance improvement as \ecce.
+
+The \lstinline{eval$^o$} relation implements an evaluator of a subset of propositional formulas.
+We consider four different implementations of this relation to explore how the way program is implemented can affect the quality of transformation.
+Depending on the implementation, \ecce generates programs of varying performance, while the execution time of the program generated by our approach is similar.
+
+The \lstinline{typecheck$^o$} relation implements a typechecker for a tiny expression language.
+We consider two different implementations of this relation: one written by hand and the other generated from the functional program.
+We demonstrate how much these implementations differ in terms of performance before and after transformations.
+
+In this study we only measured the execution time for the sample queries, averaging them over multiple runs.
+All examples of \mk relations in this paper are written in \oc\footnote{\oc: statically typed \mk embedding in \ocaml.
+The repository of the project: \url{https://github.com/JetBrains-Research/OCanren}}.
+The queries were run on a laptop running Ubuntu 18.04 with quad core Intel Core i5 2.30GHz CPU and 8 GB of RAM.
+
+The tables and graphs use the following denotations.
+\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program transformed with \ecce; \emph{Non-CPD} --- of the program transformed by our approach.
+The first two examples also contain \emph{Ideal} row which contain the execution time of the ideal implementations of the relations presented in the paper~\cite{de1999conjunctive}.
 
@@
 
-The better implementation of this relation (see fig.~\ref{doubleApp:etalon}) does not traverse the first list twice.
-This etalon implementation is generated with the conjunctive partial deduction.
+The better implementation of this relation (see fig.~\ref{doubleApp:ideal}) does not traverse the first list twice.
+This ideal implementation is generated with the conjunctive partial deduction.
 
@@
   \begin{minipage}{0.6\textwidth}
-    \begin{lstlisting}[label={doubleApp:etalon}, caption={Etalon implementation of concatenation of three lists}, captionpos=b, frame=tb]
+    \begin{lstlisting}[label={doubleApp:ideal}, caption={Ideal implementation of concatenation of three lists}, captionpos=b, frame=tb]
   let doubleAppend$^o$ xs ys zs res = conde [
@@
 There is no significant difference in the execution time in the backward direction between different specialized versions~\ref{tbl:doubleApp}.
-However the execution time in the backward direction differs: the program generated by Ecce shows the same speedup as the etalon program, while non-conjunctive partial deduction slows the program down.
+However the execution time in the backward direction differs: the program generated by \ecce shows the same speedup as the ideal program, while non-conjunctive partial deduction slows the program down.
 \todo{WHY?}
@@
   Original         & 0.0040s & 0.0195s \\ \hline
-  Ecce             & 0.0031s & 0.0201s \\ \hline
+  \ecce             & 0.0031s & 0.0201s \\ \hline
   Non-CPD          & 0.0055s & 0.0202s \\ \hline
-  Etalon           & 0.0031s & 0.0203s \\
+  Ideal           & 0.0031s & 0.0203s \\
   \end{tabular}
@@
 
-The etalon implementation is shown in fig.~\ref{etalon:maxandlength}.
+The ideal implementation is shown in fig.~\ref{ideal:maxandlength}.
 It uses a recursive relation to compute the length and maximum element simultaneously while traversing the list once.
@@
   \begin{minipage}{0.7\textwidth}
-\begin{lstlisting}[label={etalon:maxandlength}, caption={Etalon implementation of maxlengtho}, captionpos=b, frame=tb]
+\begin{lstlisting}[label={ideal:maxandlength}, caption={Ideal implementation of maxlengtho}, captionpos=b, frame=tb]
   let maxLength$^o$ xs m l = maxLength$_1^o$ xs m zero l
@@
 Note, that the second disjunct in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
-Thus among others we compared two etalon implentations: one includes these disjuncts, while the other removes them (see table~\ref{tbl:maxlen})
+Thus among others we compared two ideal implentations: one includes these disjuncts, while the other removes them (see table~\ref{tbl:maxlen})
 
@@
   Original         & 2.286ms  \\ \hline
-  Etalon           & 4.623ms  \\ \hline
-  Etalon removed   & 3.133ms  \\ \hline
-  Ecce             & 1.701ms  \\ \hline
+  Ideal           & 4.623ms  \\ \hline
+  Ideal removed   & 3.133ms  \\ \hline
+  \ecce             & 1.701ms  \\ \hline
   Non-CPD          & 2.578ms
@@
 
-Surprising enough, the execution time of the etalon program is worse than of the original.
-Ecce succeeds at improving the program performance, while non-conjunctive partial deduction worsens it a little.
+Surprising enough, the execution time of the ideal program is worse than of the original.
+\ecce succeeds at improving the program performance, while non-conjunctive partial deduction worsens it a little.
 \todo{why}
@@
 
-All examples of \mk{} relations in this paper are written in \oc{}\footnote{\oc{}: statically typed \mk{} embedding in \ocaml{}.
-The repository of the project: \url{https://github.com/JetBrains-Research/OCanren}} syntax.
 We specialize the \lstinline{eval$^o$} relation to synthesize formulas which evaluate to \lstinline{^true}\footnote{An arrow lifts ordinary values to the logic domain}.
@@
 
-One possible implementation of the evaluator in the syntax of \oc{} is presented in listing~\ref{eval:last}.
+One possible implementation of the evaluator in the syntax of \oc is presented in listing~\ref{eval:last}.
 Here the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
@@
 CPD fails to recognize this, and thus unfolds recursive calls of \lstinline{eval$^o$} applied to fresh variables.
-It leads to over-unfolding, big residual programs and less than optimal performance.
+It leads to over-unfolding, big residual programs and poor performance.
 
@@
   \hline
-  CPD              & 1.721s  & 0.168s \\
-  \hline
-  Ecce             & 0.205s  & 0.264s \\
+  \ecce             & 0.205s  & 0.264s \\
   \hline
   Non-CPD          & 0.132s  & 0.137s \\
-  \hline
-  Branching        & 0.468s  & 0.179s \\
   \end{tabular}
@@
 The \lstinline{unify} relation was executed to find a unifier of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
-The original \mk{} program fails to terminate on this goal in 30 seconds.
+The original \mk program fails to terminate on this goal in 30 seconds.
 On this example, the most performant is the program generated by CPD (0.352 seconds) while the program generated by adding branching heuristic also fails to terminate in 30 seconds.
@@
 Each column corresponds to the relation being run as described above.
-The row marked ``Original'' contains the running time of the original \mk{} relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
+The row marked ``Original'' contains the running time of the original \mk relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
 
@@
   \centering
-  \begin{tabular}{c||c|c|c||c}
-                   & \multicolumn{3}{c||}{unify} & isPath \\
+  \begin{tabular}{c||c|c|c}
+    \multirow{ 2}{*}{Terms} &
+    f(X, a) & f(a \% b \% nil, c \% d \% nil, L) & f(X, X, g(Z, t))  \\
+    \cline{2-4} &
+    f(a, X) & f(X \% XS, YS, X \% ZS) & f(g(p, L), Y, Y)  \\
+    \hline\hline
+  Original         & $0.247*10^{-4}s$ & $>30s$  & $>30s$   \\
   \hline
-                   & first  & second & third     &        \\
-  \hline\hline
-  Original         & $0.247*10^{-4}s$ & $>30s$ (oum?) & $>30s$ (oum?) & $23.743s$ \\
+  Non-CPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  \\
   \hline
-  CPD              & $0.081*10^{-4}s$ & $0.017s$ & $0.343s$  & $2.999s$  \\
-  \hline
-  Non-CPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  & $1.786s$  \\
-  \hline
-  Ecce             & $0.128*10^{-4}s$ & $0.026s$ & $1.340s$  & $1.292s$ \\
-  \hline
-  Branching        & $0.205*10^{-4}s$ & $0.101s$ & $>30s$ (oum?) & N/A \\
+  \ecce             & $0.128*10^{-4}s$ & $0.026s$ & $1.340s$ \\
   \hline
@@
 \end{table}
+
+% \begin{table}
+%   \centering
+%   \begin{tabular}{c||c|c|c||c}
+%                    & \multicolumn{3}{c||}{unify} & isPath \\
+%   \hline
+%                    & first  & second & third     &        \\
+%   \hline\hline
+%   Original         & $0.247*10^{-4}s$ & $>30s$  & $>30s$  & $23.743s$ \\
+%   \hline
+%   Non-CPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  & $1.786s$  \\
+%   \hline
+%   \ecce             & $0.128*10^{-4}s$ & $0.026s$ & $1.340s$  & $1.292s$ \\
+%   \hline
+%   \end{tabular}
+
+%   \caption{Evaluation results of unification and path search}
+%   \label{tbl:unify}
+% \end{table}
 
@@
 
-\begin{align*}
-  &type \ term = \\
-  &| \ BConst \ of \ Bool \\
-  &| \ IConst \ of \ Int \\
-  &| \ Var \ of \ Int \\
-  &| \ term + term \\
-  &| \ term * term \\
-  &| \ term = term \\
-  &| \ term < term \\
-  &| \ \underline{let} \ term \ \underline{in} \ term \\
-  &| \ \underline{if} \ term \ \underline{then} \ term \ \underline{else} \ term
-\end{align*}
+\[\begin{array}{lllll}
+  type \ term = &\ BConst \ of \ Bool &| \ IConst \ of \ Int &| \ Var \ of \ Int & \\
+  & | \ term + term &| \ term * term &| \ term = term &| \ term < term \\
+  &| \ \underline{let} \ term \ \underline{in} \ term
+  &\multicolumn{2}{l}{| \ \underline{if} \ term \ \underline{then} \ term \ \underline{else} \ term} &
+\end{array}\]
 
@@
 The first one is obtained by unnesting of the functional program as described in~\cite{lozov2019relational}.
-The second program is hand-written in \oc{}.
-Each implementation has been transformed with non-conjunctive partial deducer and by Ecce.
+The second program is hand-written in \oc.
+Each implementation has been transformed with non-conjunctive partial deducer and by \ecce.
 
 We measured the time needed to generate 100 terms of type integer (averaged over 10 iterations).
-Ecce significantly worsens the execution time for both implementations.
+\ecce significantly worsens the execution time for both implementations.
 Non-conjunctive partial deduction improves the first implementation a little bit, while worsening the performance of the second implementation 3 times.
@@
   Non-CPD     & 0.448s           & 0.154s           \\ \hline
-  Ecce        & 1.518s           & 2.543s           \\
+  \ecce        & 1.518s           & 2.543s           \\
   \end{tabular}

### 2020-07-05 papers/aplas/text/intro.tex
Evaluation intro section rewritten

@@
 
-The core feature of the family of relational programming languages \mk{}\footnote{\mk{} language web site: \url{http://minikanren.org}} is their ability to run a program in different directions.
+The core feature of the family of relational programming languages \mk\footnote{\mk language web site: \url{http://minikanren.org}} is their ability to run a program in different directions.
 Having specified a relation for adding two numbers, one can also compute the subtraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
 Program synthesis can be done by running \emph{backwards} a relational interpreter for some language.
-In general, it is possible to create a solver for a recognizer by translating it into \mk{} and running in the appropriate direction~\cite{lozov2019relational}.
+In general, it is possible to create a solver for a recognizer by translating it into \mk and running in the appropriate direction~\cite{lozov2019relational}.
 
-The search employed in \mk{} is complete which means that every answer will be found, although it may take a long time.
-The promise of \mk{} falls short when speaking of performance.
-The running time of a program in \mk{} is highly unpredictable and varies greatly for different directions.
+The search employed in \mk is complete which means that every answer will be found, although it may take a long time.
+The promise of \mk falls short when speaking of performance.
+The running time of a program in \mk is highly unpredictable and varies greatly for different directions.
 What is even worse, it depends on the order of the relation calls within a program.
@@
 It may either be a known value of some argument, its structure (i.e. the length of an input list) or, in case of a relational program, --- the direction in which it is intended to be run.
-An earlier paper~\cite{lozov2019relational} showed that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk{} programs.
+An earlier paper~\cite{lozov2019relational} showed that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk programs.
 Unfortunately, it may also not affect the running time of a program or even make it slower.
 
-Control issues in partial deduction of logic programming language \pro{} have been studied before~\cite{leuschel2002logic}.
-The ideas described there are aimed at left-to-right evaluation strategy of \pro{}.
-Since the search in \mk{} is complete, it is safe to reorder some relation calls within the goal for better performance.
+Control issues in partial deduction of logic programming language \pro have been studied before~\cite{leuschel2002logic}.
+The ideas described there are aimed at left-to-right evaluation strategy of \pro.
+Since the search in \mk is complete, it is safe to reorder some relation calls within the goal for better performance.
 While sometimes conjunctive partial deduction gives great performance boost, sometimes it does not behave as well as it could have.
@@
 In this paper, we show on examples some issues which conjunctive partial deduction faces.
-We also describe a novel approach to partial deduction of a relational programming language \mk{}.
-We compare it to the existing specialization algorithms on several programs and discuss why some \mk{} programs run slower after specialization.
+We also describe a novel approach to partial deduction of a relational programming language \mk.
+We compare it to the existing specialization algorithms on several programs and discuss why some \mk programs run slower after specialization.

### 2020-07-05 papers/aplas/text/nonconj.tex
Move APLAS draft to another folder
Evaluation intro section rewritten

--- 2020-05-23 papers/miniKanren/text/nonconj.tex
@@
 This approach draws inspiration from both conjunctive partial deduction and supercompilation.
-The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve the performance of the input programs.
+The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk to improve the performance of the input programs.
 
@@
 Other disjuncts take the form of possibly empty conjunction of relation calls accompanied with a substitution computed from unifications.
-Any \mk{} term can be trivially transformed into the described form.
+Any \mk term can be trivially transformed into the described form.
 In Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization.
@@
   \includegraphics[width=0.85\textwidth]{algo_pseudo-crop.pdf}
-  \caption{\db{Non-conjunctive Partial Deduction Pseudo Code}}
+  \caption{Non-conjunctive Partial Deduction Pseudo Code}
   \label{fig:ncpd-pseudo}
@@
 If the selected call does not suit the criteria, the results of its unfolding are not propagated onto other relation calls within the conjunction, instead, the next suitable call is selected (line 22).
-According to the denotational semantics of \mk{} it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
+According to the denotational semantics of \mk it is safe to compute individual conjuncts in any order, thus it is ok to drive any call and then propagate its results onto the other calls.
 

### 2020-07-05 papers/aplas/text/related.tex
Evaluation intro section rewritten

--- 2020-06-29 papers/aplas/related.tex
@@
 
-Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro{}.
+Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro.
 CPD constructs a tree which models goal evaluation and is similar to SLD-NF tree, then a residual program is generated from this tree.
@@
 If there is more than one suitable head, then several branches are added into the tree which corresponds to the disjunction in the residualized program.
-An adaptation of CPD for the \mk{} programming language is described in~\cite{lozov2019relational}.
+An adaptation of CPD for the \mk programming language is described in~\cite{lozov2019relational}.
 
-The most well-behaved strategy of local control in CPD for \pro{} is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
+The most well-behaved strategy of local control in CPD for \pro is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
 An atom is unfolded only if only one suitable clause head exists for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
 This means that if a non-deterministic atom is the leftmost within conjunction, it is most likely to be unfolded and introduce many new relation calls within the conjunction.
-We believe this is the core problem with CPD which limits its power when applied to \mk{}.
-The strategy of unfolding atoms from left to right is reasonable in the context of \pro{} because it mimics the way programs in \pro{} execute.
+We believe this is the core problem with CPD which limits its power when applied to \mk.
+The strategy of unfolding atoms from left to right is reasonable in the context of \pro because it mimics the way programs in \pro execute.
 But it often leads to larger global control trees and, as a result, bigger, less efficient programs.
-The evaluation result of a \mk{} program does not depend on the order of atoms (relation calls) within a conjunction, thus we believe a better result can be achieved by selecting a relation call which can restrict the number of branches in the tree.
+The evaluation result of a \mk program does not depend on the order of atoms (relation calls) within a conjunction, thus we believe a better result can be achieved by selecting a relation call which can restrict the number of branches in the tree.
 We describe our approach which implements this idea in the next section.

### 2020-07-05 papers/aplas/text/evaluation.tex
Evaluation: double appendo

@@
 \section{Evaluation}
-In our study we compared the new non-conjunctive partial deduction with the \ecce partial deduction system.
+In our study we compared the new conservative partial deduction with the \ecce partial deduction system.
 \ecce is designed for \pro programming language and cannot be directly applied to \mk.
-To be able to compare our approach with \ecce, we converted each input program first to the pure subset of \pro, then transformed it with \ecce, and then we converted the result back to the pure subset of \pro.
+To be able to compare our approach with \ecce, we converted each input program first to the pure subset of \pro, then transformed it with \ecce, and then we converted the result back to \mk.
 The coversion to \pro is a simple syntactic conversion.
@@
 The tables and graphs use the following denotations.
-\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program transformed with \ecce; \emph{Non-CPD} --- of the program transformed by our approach.
+\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program transformed with \ecce with default conjunctive control setting; \emph{ConsCPD} --- of the program transformed by our approach.
 The first two examples also contain \emph{Ideal} row which contain the execution time of the ideal implementations of the relations presented in the paper~\cite{de1999conjunctive}.
@@
 The relation \lstinline{doubleAppend$^o$} concatenates three lists by a conjunction of two calls of the \lstinline{append$^o$} relation (see fig.~\ref{doubleApp}).
-These two calls share a variable --- an intermediate list \lstinline{ts} --- which should be treated with care during specialization.
-Partial deduction does ignore the sharing of the variables which renders the approach to be unproductive for this kind of relations.
-This is why this relation is particularly well known in the conjunctive partial deduction work.
-
-The first list gets traversed twice to construct the result.
-First, when the intermidiate list \lstinline{ts} is constructed during the first \lstinline{append$^o$} call and then, when \lstinline{ts} is read during the second call.
-This double traversal negatively impacts the execution time.
+These two calls share a variable --- an intermediate list \lstinline{ts}.
+The list \lstinline{xs} is traversed twice to construct the result: first when \lstinline{ts} is constructed and then, when \lstinline{ts} is traversed during the second call.
+This double traversal negatively impacts the execution time, but it can be removed by deforestation.
 
@@
 The better implementation of this relation (see fig.~\ref{doubleApp:ideal}) does not traverse the first list twice.
-This ideal implementation is generated with the conjunctive partial deduction.
+It first recursively copies \lstinline{xs} into the beginning of \lstinline{res}, and only when \lstinline{xs} is fully consumed, it calls the \lstinline{append$^o$} to concatenate \lstinline{ys} and \lstinline{zs}.
+This implementation is provided as the ideal implementation of \lstinline{doubleAppend$^o$} in~\cite{de1999conjunctive}.
 
@@
 
-We run all translated programs in the forward and the backward direction.
+To automatically transform the initial program into the ideal implementation, the transformer should treat the conjuncts which share variables with care.
+Partial deduction splits any conjunction into its subconjunctions, regardless variable sharing, while CPD only splits conjunctions when potential non-termination is detected.
+Conservative partial deduction processes subconjunctions in isolation and then joins them back together, if it can help to restrict the search space.
+In this example, ConsPD fails to perform deforestation: ConsPD first splits the conjunction \lstinline{append$^o$ xs ys ts /\ append$^o$ ts zs rs} into two relation calls of \lstinline{append$^o$}, then unfolds the first call, but since the second call is a renaming of the first one, it is not unfolded.
+The generated program is almost the same as the original: the only difference being that the fresh variables are introduced on the toplevel.
+
+We run all programs in the forward and the backward direction.
 The forward direction \lstinline{doubleAppend$^o$ x y z ?} concatenates three given lists: the first one of length 700, the second and the third --- of length 1.
-When run in the backward direction \lstinline{doubleAppend$^o$ ? ? ? r} , it searches for the first 100 list triples whose concatenation gives the given list of length 700.
-
-There is no significant difference in the execution time in the backward direction between different specialized versions~\ref{tbl:doubleApp}.
-However the execution time in the backward direction differs: the program generated by \ecce shows the same speedup as the ideal program, while non-conjunctive partial deduction slows the program down.
-\todo{WHY?}
+The backward direction \lstinline{doubleAppend$^o$ ? ? ? r} searches for the first 100 list triples whose concatenation gives the given list of length 700.
 
@@
   \hline\hline
-  Original         & 0.0040s & 0.0195s \\ \hline
-  \ecce             & 0.0031s & 0.0201s \\ \hline
-  Non-CPD          & 0.0055s & 0.0202s \\ \hline
-  Ideal           & 0.0031s & 0.0203s \\
+  Original        & 34.8ms & 1.6ms \\ \hline
+  \ecce           & 26.7ms & 1.7ms \\ \hline
+  Ideal           & 26.9ms & 1.7ms \\ \hline
+  ConsCPD         & 47.6ms & 1.7ms
   \end{tabular}
@@
 \end{table}
+
+\includegraphics{graphs/doubleApp.pdf}
+
+There is no significant difference in the execution time in the backward direction between different specialized versions (see table~\ref{tbl:doubleApp}).
+However the execution time in the backward direction differs: the program generated by \ecce shows the same speedup as the ideal program, while conservative partial deduction slows the program down.
+The reason behind the slowdown is the premature introduction of fresh variables which is significant for the input data of this size.
 
@@
   \ecce             & 1.701ms  \\ \hline
-  Non-CPD          & 2.578ms
+  ConsCPD          & 2.578ms
   \end{tabular}
@@
 Surprising enough, the execution time of the ideal program is worse than of the original.
-\ecce succeeds at improving the program performance, while non-conjunctive partial deduction worsens it a little.
+\ecce succeeds at improving the program performance, while conservative partial deduction worsens it a little.
 \todo{why}
@@
 
-The non-conjunctive partial deduction first unfolds those calls which are selected with the heuristic.
+The conservative partial deduction first unfolds those calls which are selected with the heuristic.
 Since exploring boolean operations makes more sense, they are unfolded before recursive calls of \lstinline{eval$^o$}.
-The way non-conjunctive partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in listing~\ref{eval:fst}.
+The way conservative partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in listing~\ref{eval:fst}.
 This program is easier for CPD to transform which demonstrates how unequal is the behaviour of CPD for similar programs.
@@
 It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
-Non-conjunctive partial deduction drives the selected call until it derives useful substitutions for the variables involved while CPD with deterministic unfolding may fail to do so.
+Conservative partial deduction drives the selected call until it derives useful substitutions for the variables involved while CPD with deterministic unfolding may fail to do so.
 
@@
   \hline
-  Non-CPD          & 0.132s  & 0.137s \\
+  ConsCPD          & 0.132s  & 0.137s \\
   \end{tabular}
@@
 On this example, the most performant is the program generated by CPD (0.352 seconds) while the program generated by adding branching heuristic also fails to terminate in 30 seconds.
-The non-conjunctive partial deduction shows some improvement with the residual program terminated within 1.947 seconds.
-While driving this program, the non-conjunctive partial deduction does too much unfolding which negatively impacts the running time as compared to CPD.
+The conservative partial deduction shows some improvement with the residual program terminated within 1.947 seconds.
+While driving this program, the conservative partial deduction does too much unfolding which negatively impacts the running time as compared to CPD.
 
 The last test executed \lstinline{isPath} relation to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges.
-On this program, non-conjunctive partial deduction showed better transformation results than CPD, although the difference is not that drastic.
+On this program, conservative partial deduction showed better transformation results than CPD, although the difference is not that drastic.
 
@@
 Each column corresponds to the relation being run as described above.
-The row marked ``Original'' contains the running time of the original \mk relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
+The row marked ``Original'' contains the running time of the original \mk relation before specialization, ``CPD'' and ``ConsCPD'' correspond to conjunctive and conservative partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
 
@@
   \hline
-  Non-CPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  \\
+  ConsCPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  \\
   \hline
@@
 %   \hline
-%   Non-CPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  & $1.786s$  \\
+%   ConsCPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  & $1.786s$  \\
 %   \hline
@@
 The second program is hand-written in \oc.
-Each implementation has been transformed with non-conjunctive partial deducer and by \ecce.
+Each implementation has been transformed with conservative partial deducer and by \ecce.
 
@@
 \ecce significantly worsens the execution time for both implementations.
-Non-conjunctive partial deduction improves the first implementation a little bit, while worsening the performance of the second implementation 3 times.
+Conservative partial deduction improves the first implementation a little bit, while worsening the performance of the second implementation 3 times.
 
@@
   Original    & 0.464s           & 0.057s           \\ \hline
-  Non-CPD     & 0.448s           & 0.154s           \\ \hline
+  ConsCPD     & 0.448s           & 0.154s           \\ \hline
   \ecce        & 1.518s           & 2.543s           \\

### 2020-07-05 papers/aplas/text/evaluation.tex
Maxlen subsection

@@
 
-The relation \lstinline{maxLenght$^o$} computes both the length of the list and its maximum element (see fig.~\ref{cpd:maxandlength}) by concatenation of two relation calls.
-This relation traverses the list \lstinline{xs} twice: first to compute the maximum and then to compute the length of the list.
-It can be transformed in such a way that the list is only traversed once while computing both components of the result simultaneously.
-This transformation is called \emph{tupling} and can be achieved with conjunctive partial deduction.
+The relation \lstinline{maxLenght$^o$} computes the maximum element of the input list and its length (see fig.~\ref{cpd:maxandlength}) by conjunction of two relation calls.
+The list \lstinline{xs} is traversed twice: first to compute the maximum in the relation \lstinline{max$^o$} and then to compute the length in the relation \lstinline{length$^o$}.
+The input list contain Peano numbers which are compared by relations less-or-equal (\lstinline{le$^o$}) and greater-than (\lstinline{gt$^o$}).
+Note that the comparison relations have three arguments with the last one representing the result of comparison.
 
@@
   let maxLength$^o$ xs m l = max$^o$ xs m /\ length$^o$ xs l
-  let rec length$^o$ xs l = conde [
+  let rec length$^o$ xs l =
+    fresh (h t m) ( conde [
       (xs === [] /\ l === zero);
-    (fresh (h t m) (
-      xs === h % t /\ l === succ m /\ length$^o$ t m))]
+      (xs === h % t /\ l === succ m /\ length$^o$ t m)])
+
   let max$^o$ xs m = max$_1^o$ xs zero m
-  let rec max$_1^o$ xs n m = conde [
+  let rec max$_1^o$ xs n m =
+    fresh (h t) ( conde [
       (xs === [] /\ m === n);
-    (fresh (h t) (
       (xs === h % t) /\
@@
         (le$^o$ h n ^true /\ max$_1^o$ t n m);
-        (gt$^o$ h n ^true /\ max$_1^o$ t h m)])))]
-  let rec le$^o$ x y b = conde [
+        (gt$^o$ h n ^true /\ max$_1^o$ t h m)])])
+
+  let rec le$^o$ x y b =
+    fresh (x$_1$ y$_1$) ( conde [
       (x === zero /\ b === ^true);
-    (fresh (x$_1$) (x === succ x$_1$ /\ y === zero /\ b === ^false));
-    (fresh (x$_1$ y$_1$) (x === succ x$_1$ /\ y === succ y$_1$ /\ le$^o$ x$_1$ y$_1$ b))]
-  let rec gt$^o$ x y b = conde [
+      (x === succ x$_1$ /\ y === zero /\ b === ^false);
+      (x === succ x$_1$ /\ y === succ y$_1$ /\ le$^o$ x$_1$ y$_1$ b)])
+  let rec gt$^o$ x y b =
+    fresh (x$_1$ y$_1$) ( conde [
       (x === zero /\ b === ^false);
-    (fresh (x$_1$) (x === succ x$_1$ /\ y === zero /\ b === ^false));
-    (fresh (x$_1$ y$_1$) (x === succ x$_1$ /\ y === succ y$_1$ /\ gt$^o$ x$_1$ y$_1$ b))]
+      (x === succ x$_1$ /\ y === zero /\ b === ^true);
+      (x === succ x$_1$ /\ y === succ y$_1$ /\ gt$^o$ x$_1$ y$_1$ b)])
   \end{lstlisting}
@@
 
-The ideal implementation is shown in fig.~\ref{ideal:maxandlength}.
-It uses a recursive relation to compute the length and maximum element simultaneously while traversing the list once.
+The ideal implementation provided in the CPD literature~\cite{de1999conjunctive} is shown in fig.~\ref{ideal:maxandlength}.
+This implementation traverses the input list \lstinline{xs} only once and can be generated by tupling.
+Note, that the disjunct which unifies the third argument with \lstinline{^false} in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
+Such disjuncts are removed by both \ecce and conservative partial deduction, thus we consider two ideal implementions: one with the original comparison relations (\emph{Ideal}) and another in which these disjuncts are deleted (\emph{Ideal removed}).
 
@@
   let maxLength$^o$ xs m l = maxLength$_1^o$ xs m zero l
-  let rec maxLength$_1^o$ xs m n l = conde [
+  let rec maxLength$_1^o$ xs m n l =
+    fresh (h t l$_1$) ( conde [
       (xs === [] /\ m === n /\ l === zero);
-    (fresh (h t l$_1$)
-       (xs === h % t) /\
+      ((xs === h % t) /\
        (l === succ l$_1$) /\
@@
          (le$^o$ h n ^true /\ maxLength$_1^o$ t m n l);
-         (gt$^o$ h n ^true /\ maxLength$_1^o$ t m h l)]))]
+         (gt$^o$ h n ^true /\ maxLength$_1^o$ t m h l)]))])
   \end{lstlisting}
@@
 We measured the execution time of \lstinline{maxLength$^o$ lst m l}, where \lstinline{lst} is a list of Peano numbers from $1$ to $100$.
-Note, that the second disjunct in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
-Thus among others we compared two ideal implentations: one includes these disjuncts, while the other removes them (see table~\ref{tbl:maxlen})
+The results are presented in table~\ref{tbl:maxlen}.
 
@@
                    & [1..100] \\ \hline\hline
-  Original         & 2.286ms  \\ \hline
-  Ideal           & 4.623ms  \\ \hline
-  Ideal removed   & 3.133ms  \\ \hline
-  \ecce             & 1.701ms  \\ \hline
-  ConsCPD          & 2.578ms
+  Original         & 48.5ms  \\ \hline
+  Ideal            & 32.2ms  \\ \hline
+  Ideal removed    & 19.6ms  \\ \hline
+  \ecce            & 23.7ms  \\ \hline
+  ConsCPD          & 35.7ms
   \end{tabular}
@@
 
-Surprising enough, the execution time of the ideal program is worse than of the original.
-\ecce succeeds at improving the program performance, while conservative partial deduction worsens it a little.
-\todo{why}
+\includegraphics{graphs/maxlen.pdf}
+
+Both \ecce and ConsPD improve performance of \lstinline{maxLength$^o$} as compared to the original program.
+Removing the disjuncts with the result unified with \lstinline{^false} in the comparison relations improves the ideal program by about 40\%.
+ConsPD does not perform tupling, but it does remove disjuncts which never succeed, thus its result is not as performant as the ideal program.
 

### 2020-07-05 papers/aplas/text/evaluation.tex
Evaluation: prop, unify, path

@@
 The tables and graphs use the following denotations.
-\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program transformed with \ecce with default conjunctive control setting; \emph{ConsCPD} --- of the program transformed by our approach.
+\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program transformed with \ecce with default conjunctive control setting; \emph{ConsPD} --- of the program transformed by our approach.
 The first two examples also contain \emph{Ideal} row which contain the execution time of the ideal implementations of the relations presented in the paper~\cite{de1999conjunctive}.
@@
   Ideal           & 26.9ms & 1.7ms \\ \hline
-  ConsCPD         & 47.6ms & 1.7ms
+  ConsPD         & 47.6ms & 1.7ms
   \end{tabular}
@@
 This implementation traverses the input list \lstinline{xs} only once and can be generated by tupling.
-Note, that the disjunct which unifies the third argument with \lstinline{^false} in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
+Note, that the disjunct which unifies the third argument with \lstinline{^false}\footnote{An arrow lifts ordinary values to the logic domain} in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
 Such disjuncts are removed by both \ecce and conservative partial deduction, thus we consider two ideal implementions: one with the original comparison relations (\emph{Ideal}) and another in which these disjuncts are deleted (\emph{Ideal removed}).
@@
   \ecce            & 23.7ms  \\ \hline
-  ConsCPD          & 35.7ms
+  ConsPD          & 35.7ms
   \end{tabular}
@@
 
-The relation \lstinline{eval$^o$} describes an evaluation of a subset of first-order logic formulas in a given substitution.
-It has 3 arguments.
+The relation \lstinline{eval$^o$} describes evaluation of a subset of first-order logic formulas in a given substitution.
+The relation has 3 arguments.
 The first argument is a list of boolean values which serves as a substitution.
@@
 
-We specialize the \lstinline{eval$^o$} relation to synthesize formulas which evaluate to \lstinline{^true}\footnote{An arrow lifts ordinary values to the logic domain}.
+We specialize the \lstinline{eval$^o$} relation to synthesize formulas which evaluate to \lstinline{^true}.
 To do so, we run the specializer for the goal with the last argument fixed to \lstinline{^true}, while the first two arguments remain free variables.
@@
 
-One possible implementation of the evaluator in the syntax of \oc is presented in listing~\ref{eval:last}.
+One possible implementation of the evaluator is presented in listing~\ref{eval:last}.
 Here the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
@@
 Note, that the calls to boolean relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} are placed last within each conjunction.
-This poses a challenge to the CPD-based specializers.
+This poses a challenge to the CPD-based specializers such as \ecce.
 Conjunctive partial deduction unfolds relation calls from left to right, so when specializing this relation for running backwards (i.e.
@@
 \subsubsection{Evaluation Results}
-In our study, we considered two implementations of \lstinline{eval$^o$}, one we call \lstinline{plain} and the other --- \lstinline{last}, and compared how specializers behave on them.
-The \lstinline{plain} relation uses table-based boolean operations and places them further to the left in each conjunction.
-The relation \lstinline{last} employs boolean operations implemented via \lstinline{nand$^o$} and place them at the end of each conjunction.
-These two programs are complete opposites from the standpoint of CPD.
+In our study we considered four implementations of \lstinline{eval$^o$}:
+\begin{itemize}
+  \item \emph{FirstPlain} in which boolean operations are \textbf{table-based} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
+  \item \emph{LastPlain} in which boolean operations are \textbf{table-based} and are placed \textbf{after} recursive calls to \lstinline{eval$^o$}
+  \item \emph{FirstNando} in which boolean operations are implemented via \lstinline{nand$^o$} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
+  \item \emph{LastNando} in which boolean operations are implemented via \lstinline{nand$^o$} andare placed \textbf{after} recursive calls to \lstinline{eval$^o$}
+
+\end{itemize}
+
+These four implementations are very different from the standpoint of \ecce.
 We measured the time necessary to generate $1000$ formulas over two variables which evaluate to \lstinline{^true} (averaged over 10 runs).
+The results are presented in table~\ref{tbl:eval}.
 
@@
   \centering
-  \begin{tabular}{c||c||c}
-                   & last    & plain  \\
-  \hline\hline
-  Original         & 0.695s  & 1.426s \\
-  \hline
-  \ecce             & 0.205s  & 0.264s \\
-  \hline
-  ConsCPD          & 0.132s  & 0.137s \\
+  \begin{tabular}{c||c|c|c|c}
+                   & FirstPlain & FirstNando & LastPlain & LastNando \\ \hline\hline
+  Original         & 14.8ms     & 14.5ms     & 7.4ms     & 10.1ms    \\ \hline
+  \ecce            & 16.0ms     & 22.3ms     & 14.0ms    & 14.9ms    \\ \hline
+  ConsPD          & 9.1ms      & 9.7ms      & 9.1ms     & 9.9ms     \\
   \end{tabular}
-
-  \caption{Evaluation results}
+  \caption{Execution time of evalo}
   \label{tbl:eval}
 \end{table}
+
+\includegraphics{graphs/prop.pdf}
+
+ConsPD generates programs with comparable performance for all four implementations, while the quality of \ecce transformations differs significantly.
+\ecce worsens performance for every implementation as compared to the original program.
+ConsPD worsens performance of only the fastest implementation \emph{LastPlain}, while improving it for others.
 
@@
 
-The \lstinline{unify} relation was executed to find a unifier of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
-The original \mk program fails to terminate on this goal in 30 seconds.
-On this example, the most performant is the program generated by CPD (0.352 seconds) while the program generated by adding branching heuristic also fails to terminate in 30 seconds.
-The conservative partial deduction shows some improvement with the residual program terminated within 1.947 seconds.
-While driving this program, the conservative partial deduction does too much unfolding which negatively impacts the running time as compared to CPD.
-
-The last test executed \lstinline{isPath} relation to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges.
-On this program, conservative partial deduction showed better transformation results than CPD, although the difference is not that drastic.
-
-All evaluation results are presented in the table~\ref{tbl:unify}.
-Each column corresponds to the relation being run as described above.
-The row marked ``Original'' contains the running time of the original \mk relation before specialization, ``CPD'' and ``ConsCPD'' correspond to conjunctive and conservative partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
+The \lstinline{unify} relation was executed to find a unifier of three pairs of terms (see table~\ref{tbl:unify}) which vary in their complexity.
+The original program failed to terminate on two last unification problems in under 30s, while both ConsPD and \ecce generate programs which are able to find the unifiers.
 
@@
   \centering
-  \begin{tabular}{c||c|c|c}
-    \multirow{ 2}{*}{Terms} &
-    f(X, a) & f(a \% b \% nil, c \% d \% nil, L) & f(X, X, g(Z, t))  \\
+  \begin{tabular}{c||c|c|c||c}
+    & \multicolumn{3}{c||}{unify} & \\
+    \multirow{2}{*}{Terms} & f(X, a) & f(a \% b \% nil, c \% d \% nil, L) & f(X, X, g(Z, t)) & \multirow{2}{*}{isPath}  \\
     \cline{2-4} &
@@
     \hline\hline
-  Original         & $0.247*10^{-4}s$ & $>30s$  & $>30s$   \\
-  \hline
-  ConsCPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  \\
-  \hline
-  \ecce             & $0.128*10^{-4}s$ & $0.026s$ & $1.340s$ \\
-  \hline
+  Original          & 0.23ms &  ---   &  ---    & 24.9ms \\ \hline
+  ConsPD            & 0.07ms & 16.5ms & 397.6ms & 1.6ms  \\ \hline
+  \ecce             & 0.12ms & 18.8ms & 352.0ms & 1.1ms  \\ \hline
   \end{tabular}
@@
 
-% \begin{table}
-%   \centering
-%   \begin{tabular}{c||c|c|c||c}
-%                    & \multicolumn{3}{c||}{unify} & isPath \\
-%   \hline
-%                    & first  & second & third     &        \\
-%   \hline\hline
-%   Original         & $0.247*10^{-4}s$ & $>30s$  & $>30s$  & $23.743s$ \\
-%   \hline
-%   ConsCPD          & $0.084*10^{-4}s$ & $0.021s$ & $2.198s$  & $1.786s$  \\
-%   \hline
-%   \ecce             & $0.128*10^{-4}s$ & $0.026s$ & $1.340s$  & $1.292s$ \\
-%   \hline
-%   \end{tabular}
-
-%   \caption{Evaluation results of unification and path search}
-%   \label{tbl:unify}
-% \end{table}
+The relation \lstinline{isPath} was executed to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges (see table~\ref{tbl:unify}).
+Both transformations significantly improve the performance as compared to the original program, although ConsPD is about 30\% worse than \ecce.
 
@@
   Original    & 0.464s           & 0.057s           \\ \hline
-  ConsCPD     & 0.448s           & 0.154s           \\ \hline
+  ConsPD     & 0.448s           & 0.154s           \\ \hline
   \ecce        & 1.518s           & 2.543s           \\

### 2020-07-05 papers/aplas/text/evaluation.tex
Evaluation: typechecker

@@
 \ecce is designed for \pro programming language and cannot be directly applied to \mk.
-To be able to compare our approach with \ecce, we converted each input program first to the pure subset of \pro, then transformed it with \ecce, and then we converted the result back to \mk.
+To be able to compare our approach with \ecce, we converted each input program first to the pure subset of \pro, then spelialized it with \ecce, and then we converted the result back to \mk.
 The coversion to \pro is a simple syntactic conversion.
@@
 The last two are examples of relational interpreters created by us which are observable, but non-trivial.
-In the last two examples we explore how different implementations of the same relation affects the quality of transformations.
+In the last two examples we explore how different implementations of the same relation affects the quality of specialization.
 
@@
 This example is used to demonstrate \emph{tupling}: the transformation in which multiple traversals of the same data structure are replaced with a single traversal which computes all the necessary results simultaneously.
-CPD performes tupling more often than our approach, and we explore how it affects the performance of the transformed program.
+CPD performes tupling more often than our approach, and we explore how it affects the performance of the specialized program.
 
@@
 The \lstinline{eval$^o$} relation implements an evaluator of a subset of propositional formulas.
-We consider four different implementations of this relation to explore how the way program is implemented can affect the quality of transformation.
+We consider four different implementations of this relation to explore how the way program is implemented can affect the quality of specialization.
 Depending on the implementation, \ecce generates programs of varying performance, while the execution time of the program generated by our approach is similar.
@@
 We consider two different implementations of this relation: one written by hand and the other generated from the functional program.
-We demonstrate how much these implementations differ in terms of performance before and after transformations.
+We demonstrate how much these implementations differ in terms of performance before and after specialization.
 
@@
 The tables and graphs use the following denotations.
-\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program transformed with \ecce with default conjunctive control setting; \emph{ConsPD} --- of the program transformed by our approach.
+\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program specialized by \ecce with default conjunctive control setting; \emph{ConsPD} --- of the program specialized by our approach.
 The first two examples also contain \emph{Ideal} row which contain the execution time of the ideal implementations of the relations presented in the paper~\cite{de1999conjunctive}.
@@
 There is no significant difference in the execution time in the backward direction between different specialized versions (see table~\ref{tbl:doubleApp}).
-However the execution time in the backward direction differs: the program generated by \ecce shows the same speedup as the ideal program, while conservative partial deduction slows the program down.
+However the execution time in the forward direction differs: the program generated by \ecce shows the same speedup as the ideal program, while conservative partial deduction slows the program down.
 The reason behind the slowdown is the premature introduction of fresh variables which is significant for the input data of this size.
@@
 The way conservative partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in listing~\ref{eval:fst}.
-This program is easier for CPD to transform which demonstrates how unequal is the behaviour of CPD for similar programs.
+This program is easier for \ecce to specialize which demonstrates how unequal is the behaviour of CPD for similar programs.
 
@@
 
-ConsPD generates programs with comparable performance for all four implementations, while the quality of \ecce transformations differs significantly.
+ConsPD generates programs with comparable performance for all four implementations, while the quality of \ecce specialization differs significantly.
 \ecce worsens performance for every implementation as compared to the original program.
@@
 
-Besides evaluator of logic formulas we also run the transformers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
+Besides evaluator of logic formulas we also run the specializers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
 These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
@@
 The relation \lstinline{isPath} was executed to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges (see table~\ref{tbl:unify}).
-Both transformations significantly improve the performance as compared to the original program, although ConsPD is about 30\% worse than \ecce.
+Both specializers significantly improve the performance as compared to the original program, although ConsPD is about 30\% worse than \ecce.
 
@@
 
-This relation implements a typechecker for a tiny language.
+This relation implements a typechecker for a tiny expression language.
 Being executed in the backward direction it serves as a generator of terms of the given type.
@@
 Boolean and integer constants have the corresponding types regardless of the environment.
-Only terms of type integer can be summed up, multiplied or checked for being less or equal.
+Only terms of type integer can be summed up, multiplied or compared by less-then operator.
 Any terms of the same type can be checked for equality.
+Addition and multiplication of two terms of suitable types have integer type, while comparisons have boolean type.
 If-then-else expression typechecks only if its condition is of type boolean, while both then- and else-branches have the same type.
@@
 To typecheck a let-binding, first, the term being bound is typechecked and is added in the beginning of the environment $\Gamma$, and then the body is typechecked in the context of the new environment.
-Typechecking a variable with the index $i$ boils down to getting a $i$-th element of the list.
+Typechecking a variable with the index $i$ boils down to getting an $i$-th element of the list.
 
@@
     \infer[]{\Gamma \vdash \underline{let} \ v \ b : \tau}{\Gamma \vdash v : \tau_v, \ (\tau_v :: \Gamma) \vdash b : \tau} &
-    \infer[\Gamma \lbrack v \rbrack = \tau]{\Gamma \vdash Var \ v : \tau}{} \vspace{0.5cm} \\
+    \infer[\Gamma \lbrack v \rbrack \equiv \tau]{\Gamma \vdash Var \ v : \tau}{} \vspace{0.5cm} \\
 
@@
     }
-
   \end{tabular}
@@
 
-We compared two possible implementations of these typing rules.
-The first one is obtained by unnesting of the functional program as described in~\cite{lozov2019relational}.
-The second program is hand-written in \oc.
-Each implementation has been transformed with conservative partial deducer and by \ecce.
-
-We measured the time needed to generate 100 terms of type integer (averaged over 10 iterations).
-\ecce significantly worsens the execution time for both implementations.
-Conservative partial deduction improves the first implementation a little bit, while worsening the performance of the second implementation 3 times.
+We compared two implementations of these typing rules.
+The first one is obtained by unnesting of the functional program as described in~\cite{lozov2019relational} (\emph{Generated}).
+The second program is hand-written in \oc (\emph{Hand-written}).
+Each implementation has been specialized with ConsPD and \ecce.
+We measured the time needed to generate 1000 closed terms of type integer (averaged over 10 iterations).
 
@@
   \begin{tabular}{c||c||c}
-              & Implementation 1 & Implementation 2 \\ \hline\hline
-  Original    & 0.464s           & 0.057s           \\ \hline
-  ConsPD     & 0.448s           & 0.154s           \\ \hline
-  \ecce        & 1.518s           & 2.543s           \\
+              & Hand-written & Generated \\ \hline\hline
+  Original    & 916.9ms      & 11458.6ms \\ \hline
+  ConsPD      & 335.9ms      & 289.8ms   \\ \hline
+  \ecce       & 218.6ms      & 382.0ms   \\
   \end{tabular}
-
-  \caption{Running time of generating 100 closed terms of type Int}
+  \caption{Running time of generating 1000 closed terms of type Int}
   \label{tbl:eval}
@@
 
-\todo{why}
+Note that the generated program is far slower than the hand-written.
+The principal difference between these two implementations is that the generated program is redundant.
+For example, typechecking of the sum of two terms in the hand-written implementation is implemented by a single conjunction (see fig.~\ref{type:hand}) while the generated program is far more complicated and also uses a special relation \lstinline{typeEq$^o$} to compare types (see fig.~\ref{type:gen}).
+
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.95\textwidth}
+    \begin{lstlisting}[label={type:hand}, caption={A fragment of hand-written typechecker}, captionpos=b, frame=tb]
+  let rec typecheck$^o$ gamma term res = conde [
+    ...
+    fresh (x y) (
+      (term === x + y /\
+       typecheck$^o$ gamma x ^(Some Integer) /\
+       typecheck$^o$ gamma y ^(Some Integer) /\
+       res === ^(Some Integer)));
+    ...]
+    \end{lstlisting}
+  \end{minipage}
+\end{figure*}
+
+\begin{figure*}[!h]
+  \centering
+  \begin{minipage}{0.95\textwidth}
+    \begin{lstlisting}[label={type:gen}, caption={A fragment of generated typechecker}, captionpos=b, frame=tb]
+  let rec typecheck$^o$ gamma term res = conde [
+    ...
+    fresh (x y t1 t2) (
+      (term === x + y /\
+      conde [
+        typecheck$^o$ gamma x ^None /\
+          res === ^None;
+        typecheck$^o$ gamma x ^(Some t1) /\
+          typecheck$^o$ gamma y ^None /\
+          res === ^None;
+        typecheck$^o$ gamma x ^(Some t1) /\
+          typecheck$^o$ gamma y ^(Some t2) /\
+          typeEq$^o$ t1 Integer ^true /\
+          typeEq$^o$ t2 Integer ^true /\
+          res === ^(Some Integer);
+      ])
+    ...]
+    \end{lstlisting}
+  \end{minipage}
+\end{figure*}
+
+Most of the redundancy of the generated program is removed by specialization with respect to the known type of the term.
+This is why both implementations have comparable speed after specialization.
+The specialized programs differ too much and it is not clear why one implementation is better specialized by \ecce and the other by ConsPD.

### 2020-07-06 papers/aplas/text/evaluation.tex
Gnuplot graphs are used instead of google sheets ones
Evaluation edits from dboulytchev

@@
 \section{Evaluation}
+
 In our study we compared the new conservative partial deduction with the \ecce partial deduction system.
-\ecce is designed for \pro programming language and cannot be directly applied to \mk.
+\ecce is designed for \pro programming language and cannot be directly applied for programs, written in \mk.
 To be able to compare our approach with \ecce, we converted each input program first to the pure subset of \pro, then spelialized it with \ecce, and then we converted the result back to \mk.
-The coversion to \pro is a simple syntactic conversion.
-In conversion from \pro to \mk, a conjunction is generated by each horn clause in which unifications are placed before any relation call.
-This mimics in \mk the execution order of \pro program within a conjunction.
+The coversion to \pro is a simple syntactic conversion. In the conversion from \pro to \mk, for each Horn clause a conjunction is generated in which unifications are placed before any relation call.
+
+%This mimics in \mk the execution order of \pro program within conjunction.
 
@@
 The second two are introduced in the paper on relational interpreters~\cite{lozov2019relational}.
-The last two are examples of relational interpreters created by us which are observable, but non-trivial.
-In the last two examples we explore how different implementations of the same relation affects the quality of specialization.
+The last two are examples of non-trivial relational interpreters which at the same time are observable.
+In the last two examples we explore how different implementations of the same relation affect the quality of specialization.
 
 The \lstinline{doubleAppend$^o$} program concatenates three lists of numbers by computing a conjunction of two calls of a relation which concatenates two lists.
-This example is prominent in CPD literature, since it demonstrates \emph{deforestation} --- a transformation which gets rid of intermidiate datastructures.
-This transformation cannot be done with the earlier partial deduction approach, since the relation uses a variable shared between two relation calls within a conjunction which requires the transformation to treat conjunctions as a whole.
+This example is prominent in CPD literature, since it demonstrates \emph{deforestation}~--- a transformation which gets rid of intermidiate datastructures.
+This transformation cannot be done with earlier partial deduction approachs, since the relation uses a variable shared between two relation calls within a
+conjunction which requires the transformation to treat conjunctions as a whole.
 We explore how the fact that our approach sometimes splits conjunctions ignoring the variable sharing affects the quality of transformation.
 
-The \lstinline{maxLength$^o$} program computes the maximum element of the list as well as its length.
-In this program an input list is shared between the call of the relation to compute the maximum element and the relation to compute the length of the list.
+The \lstinline{maxLength$^o$} program computes the maximal element of a list as well as this list length.
+In this program an input list is shared between the call of the relation to compute the maximal element and the relation to compute the length of the list.
 This sharing leads to the list being traversed twice during execution.
-This example is used to demonstrate \emph{tupling}: the transformation in which multiple traversals of the same data structure are replaced with a single traversal which computes all the necessary results simultaneously.
-CPD performes tupling more often than our approach, and we explore how it affects the performance of the specialized program.
+This example is used to demonstrate \emph{tupling}: the transformation in which multiple traversals of the same data structure are replaced with a single traversal which
+computes all the necessary results simultaneously. CPD performes tupling more often than our approach, and we explore how it affects the performance of the specialized program.
 
-The \lstinline{isPath} and \lstinline{unify} relations demonstrate the relation interpretation approach in which a relation is generated from a functional program.
-Partial deduction is capable to improve the execution time of the generated relations in the given direction.
-These two relations are described in~\cite{lozov2019relational} thus we will only briefly describe them in this paper.
+The \lstinline{isPath} and \lstinline{unify} relations demonstrate the relational interpretation approach in which a relation is generated from a functional program.
+Partial deduction is capable of improving the execution time of the generated relations in the given direction.
+These two relations are described in~\cite{lozov2019relational}, thus we will only briefly describe them in this paper.
 Here the aim is to study if our approach can help to achieve comparable performance improvement as \ecce.
@@
 We consider four different implementations of this relation to explore how the way program is implemented can affect the quality of specialization.
-Depending on the implementation, \ecce generates programs of varying performance, while the execution time of the program generated by our approach is similar.
+Depending on the implementation, \ecce generates programs of varying performance, while the execution times of the programs generated by our approach are similar.
 
@@
 The tables and graphs use the following denotations.
-\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE} --- of the program specialized by \ecce with default conjunctive control setting; \emph{ConsPD} --- of the program specialized by our approach.
-The first two examples also contain \emph{Ideal} row which contain the execution time of the ideal implementations of the relations presented in the paper~\cite{de1999conjunctive}.
+\emph{Original} represents the execution time of a program before any transformations were applied; \emph{ECCE}~--- of the program specialized by \ecce with default conjunctive control setting; \emph{ConsPD}~--- of the program specialized by our approach.
+The first two examples also contain \emph{Ideal} row which contains the execution time of the ideal implementations of the relations presented in the paper~\cite{de1999conjunctive}.
 
@@
 The relation \lstinline{doubleAppend$^o$} concatenates three lists by a conjunction of two calls of the \lstinline{append$^o$} relation (see fig.~\ref{doubleApp}).
-These two calls share a variable --- an intermediate list \lstinline{ts}.
-The list \lstinline{xs} is traversed twice to construct the result: first when \lstinline{ts} is constructed and then, when \lstinline{ts} is traversed during the second call.
+These two calls share a variable~--- an intermediate list \lstinline{ts}.
+The list \lstinline{xs} is traversed twice to construct the result: first when \lstinline{ts} is constructed and then when \lstinline{ts} is traversed during the second call.
 This double traversal negatively impacts the execution time, but it can be removed by deforestation.
@@
 In this example, ConsPD fails to perform deforestation: ConsPD first splits the conjunction \lstinline{append$^o$ xs ys ts /\ append$^o$ ts zs rs} into two relation calls of \lstinline{append$^o$}, then unfolds the first call, but since the second call is a renaming of the first one, it is not unfolded.
-The generated program is almost the same as the original: the only difference being that the fresh variables are introduced on the toplevel.
+The generated program is almost the same as the original: the only difference is that the fresh variables are introduced on the toplevel.
 
 We run all programs in the forward and the backward direction.
-The forward direction \lstinline{doubleAppend$^o$ x y z ?} concatenates three given lists: the first one of length 700, the second and the third --- of length 1.
+The forward direction \lstinline{doubleAppend$^o$ x y z ?} concatenates three given lists: the first one of length 700, the second and the third~--- of length 1.
 The backward direction \lstinline{doubleAppend$^o$ ? ? ? r} searches for the first 100 list triples whose concatenation gives the given list of length 700.
@@
 
-\includegraphics{graphs/doubleApp.pdf}
+\includegraphics{../data/doubleAppendo/da700log.pdf}
 
@@
 
-\subsection{Maximum Element and Length of a List}
+\subsection{Maximal Element and Length of a List}
 
-The relation \lstinline{maxLenght$^o$} computes the maximum element of the input list and its length (see fig.~\ref{cpd:maxandlength}) by conjunction of two relation calls.
+The relation \lstinline{maxLenght$^o$} computes the maximal element of the input list and its length (see fig.~\ref{cpd:maxandlength}) by conjunction of two relation calls.
 The list \lstinline{xs} is traversed twice: first to compute the maximum in the relation \lstinline{max$^o$} and then to compute the length in the relation \lstinline{length$^o$}.
@@
 
-The ideal implementation provided in the CPD literature~\cite{de1999conjunctive} is shown in fig.~\ref{ideal:maxandlength}.
+The ideal implementation provided in the CPD literature~\cite{de1999conjunctive} is shown on fig.~\ref{ideal:maxandlength}.
 This implementation traverses the input list \lstinline{xs} only once and can be generated by tupling.
-Note, that the disjunct which unifies the third argument with \lstinline{^false}\footnote{An arrow lifts ordinary values to the logic domain} in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
+Note, the disjunct which unifies the third argument with \lstinline{^false}\footnote{An arrow lifts ordinary values to the logic domain} in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
 Such disjuncts are removed by both \ecce and conservative partial deduction, thus we consider two ideal implementions: one with the original comparison relations (\emph{Ideal}) and another in which these disjuncts are deleted (\emph{Ideal removed}).
@@
 
-\includegraphics{graphs/maxlen.pdf}
+\includegraphics{../data/maxLengtho/max.pdf}
 
-Both \ecce and ConsPD improve performance of \lstinline{maxLength$^o$} as compared to the original program.
+Both \ecce and ConsPD improve the performance of \lstinline{maxLength$^o$} as compared to the original program.
 Removing the disjuncts with the result unified with \lstinline{^false} in the comparison relations improves the ideal program by about 40\%.
-ConsPD does not perform tupling, but it does remove disjuncts which never succeed, thus its result is not as performant as the ideal program.
+ConsPD removes disjuncts which never succeed, but does not perform tupling, thus its result is not as performant as the ideal program.
 
@@
 
-The relation \lstinline{eval$^o$} describes evaluation of a subset of first-order logic formulas in a given substitution.
-The relation has 3 arguments.
-The first argument is a list of boolean values which serves as a substitution.
+The relation \lstinline{eval$^o$} describes an evaluation of a propositional formula under given variable assignments.
+The relation has thee arguments. The first argument is a list of boolean values which plays a role of variable assignments.
 The $i$-th value of the substitution is the value of the $i$-th variable.
@@
 A formula is either a \emph{variable} represented with a Peano number, a \emph{negation} of a formula, a \emph{conjunction} of two formulas or a \emph{disjunction} of two formulas.
-The third argument is the value of the formula in the given substitution.
+The third argument is the value of the formula under the given assignment.
 
@@
 Here the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
-The relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} encode corresponding boolean operations.
+The relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} encode corresponding boolean connectives.
 
@@
 
-Note, that the calls to boolean relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} are placed last within each conjunction.
-This poses a challenge to the CPD-based specializers such as \ecce.
+Note, the calls to boolean relations \lstinline{and$^o$}, \lstinline{or$^o$}, and \lstinline{not$^o$} are placed last within each conjunction.
+This poses a challenge for the CPD-based specializers such as \ecce.
 Conjunctive partial deduction unfolds relation calls from left to right, so when specializing this relation for running backwards (i.e.
@@
 There are three possible options for these variables in the call \lstinline{or$^o$ v w res} and one for the call \lstinline{not$^o$}.
-These variables are used in recursive calls of \lstinline{eval$^o$} and thus restrict the result of driving them.
+These variables are used in recursive calls of \lstinline{eval$^o$} and thus restrict the result of driving.
 CPD fails to recognize this, and thus unfolds recursive calls of \lstinline{eval$^o$} applied to fresh variables.
-It leads to over-unfolding, big residual programs and poor performance.
+It leads to over-unfolding, large residual programs and poor performance.
 
-The conservative partial deduction first unfolds those calls which are selected with the heuristic.
-Since exploring boolean operations makes more sense, they are unfolded before recursive calls of \lstinline{eval$^o$}.
-The way conservative partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in listing~\ref{eval:fst}.
+The conservative partial deduction first unfolds those calls which are selected according to the heuristics.
+Since exploring the implementations of boolean connectives makes more sense, they are unfolded before recursive calls of \lstinline{eval$^o$}.
+The way conservative partial deduction treats this program is the same as it treats the other implementation in which boolean connectives
+are moved to the left, as shown in listing~\ref{eval:fst}.
 This program is easier for \ecce to specialize which demonstrates how unequal is the behaviour of CPD for similar programs.
@@
 Partial deduction tries to unfold every relation call unless it is unsafe, but not all relation calls serve to restrict the search space and thus should be unfolded.
-In the implementation of \lstinline{eval$^o$} boolean operations can effectively restrict variables within the conjunctions and should be unfolded until they do.
+In the implementation of \lstinline{eval$^o$} boolean connectives can effectively restrict variables within the conjunctions and should be unfolded until they do.
 But depending on the way they are implemented, the different number of driving steps should be performed for that.
-The simplest way to implement these relations is with a table as demonstrated with the implementation of \lstinline{not$^o$} in listing~\ref{not:table}.
+The simplest way to implement these relations is by mimicking a truth tables as demonstrated by the implementation of \lstinline{not$^o$} in listing~\ref{not:table}.
 It is enough to unfold such relation calls once to derive useful information about variables.
@@
 
-The other way to implement boolean operations is via one basic boolean relation such as \lstinline{nand$^o$} which is, in turn, has a table-based implementation (see listing~\ref{not:nando}).
-It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
+The other way to implement boolean connectives is to express them using a single basic boolean relation such as \lstinline{nand$^o$} which is, in turn, has a table-based
+implementation (see listing~\ref{not:nando}). It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should
+be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
 Conservative partial deduction drives the selected call until it derives useful substitutions for the variables involved while CPD with deterministic unfolding may fail to do so.
@@
 \begin{itemize}
-  \item \emph{FirstPlain} in which boolean operations are \textbf{table-based} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
-  \item \emph{LastPlain} in which boolean operations are \textbf{table-based} and are placed \textbf{after} recursive calls to \lstinline{eval$^o$}
-  \item \emph{FirstNando} in which boolean operations are implemented via \lstinline{nand$^o$} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
-  \item \emph{LastNando} in which boolean operations are implemented via \lstinline{nand$^o$} andare placed \textbf{after} recursive calls to \lstinline{eval$^o$}
+  \item \emph{FirstPlain} in which the implementations of boolean connectives are \textbf{table-based} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
+  \item \emph{LastPlain} in which the implementations of boolean connectives are \textbf{table-based} and are placed \textbf{after} recursive calls to \lstinline{eval$^o$}
+  \item \emph{FirstNando} in which boolean connectives are implemented via \lstinline{nand$^o$} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
+  \item \emph{LastNando} in which boolean connectives are implemented via \lstinline{nand$^o$} andare placed \textbf{after} recursive calls to \lstinline{eval$^o$}
 
@@
 
-These four implementations are very different from the standpoint of \ecce.
+These four implementations are very different from \ecce standpoint.
 We measured the time necessary to generate $1000$ formulas over two variables which evaluate to \lstinline{^true} (averaged over 10 runs).
@@
 
-\includegraphics{graphs/prop.pdf}
+\includegraphics{../data/propEval/prop.pdf}
 
@@
 
-Besides evaluator of logic formulas we also run the specializers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
-These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
+Besides evaluator of logic formulas we also run the specializers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath}
+specialized to search for paths in the graph. These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
 
@@
 
-Note that the generated program is far slower than the hand-written.
-The principal difference between these two implementations is that the generated program is redundant.
+Note that the generated program is far slower than hand-written.
+The principal difference between these two implementations is that the generated program is contains a certain redundancy.
 For example, typechecking of the sum of two terms in the hand-written implementation is implemented by a single conjunction (see fig.~\ref{type:hand}) while the generated program is far more complicated and also uses a special relation \lstinline{typeEq$^o$} to compare types (see fig.~\ref{type:gen}).

### 2020-07-06 papers/aplas/text/evaluation.tex
Small corrections to evaluation

@@
 
-The \lstinline{isPath} and \lstinline{unify} relations demonstrate the relational interpretation approach in which a relation is generated from a functional program.
+The \lstinline{isPath$^o$} and \lstinline{unify$^o$} relations demonstrate the relational interpretation approach in which a relation is generated from a functional program.
 Partial deduction is capable of improving the execution time of the generated relations in the given direction.
@@
 The relation \lstinline{eval$^o$} describes an evaluation of a propositional formula under given variable assignments.
-The relation has thee arguments. The first argument is a list of boolean values which plays a role of variable assignments.
+The relation has three arguments. The first argument is a list of boolean values which plays a role of variable assignments.
 The $i$-th value of the substitution is the value of the $i$-th variable.
@@
 
-\subsubsection{The Order of Relation Calls}
+\subsubsection{The Order of Relation Calls.}
 
-One possible implementation of the evaluator is presented in listing~\ref{eval:last}.
+One possible implementation of the evaluator is presented in fig.~\ref{eval:last}.
 Here the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
@@
 The way conservative partial deduction treats this program is the same as it treats the other implementation in which boolean connectives
-are moved to the left, as shown in listing~\ref{eval:fst}.
+are moved to the left, as shown in fig.~\ref{eval:fst}.
 This program is easier for \ecce to specialize which demonstrates how unequal is the behaviour of CPD for similar programs.
@@
 
-\subsubsection{Unfolding of Complex Relations}
+\subsubsection{Unfolding of Complex Relations.}
 
@@
 But depending on the way they are implemented, the different number of driving steps should be performed for that.
-The simplest way to implement these relations is by mimicking a truth tables as demonstrated by the implementation of \lstinline{not$^o$} in listing~\ref{not:table}.
+The simplest way to implement these relations is by mimicking a truth tables as demonstrated by the implementation of \lstinline{not$^o$} in fig.~\ref{not:table}.
 It is enough to unfold such relation calls once to derive useful information about variables.
@@
 The other way to implement boolean connectives is to express them using a single basic boolean relation such as \lstinline{nand$^o$} which is, in turn, has a table-based
-implementation (see listing~\ref{not:nando}). It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should
+implementation (see fig.~\ref{not:nando}). It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should
 be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
@@
 
-\subsubsection{Evaluation Results}
+\subsubsection{Evaluation Results.}
 In our study we considered four implementations of \lstinline{eval$^o$}:
@@
 
-\subsection{Unification and Path Search}
-
-Besides evaluator of logic formulas we also run the specializers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath}
-specialized to search for paths in the graph. These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
-
-The \lstinline{unify} relation was executed to find a unifier of three pairs of terms (see table~\ref{tbl:unify}) which vary in their complexity.
-The original program failed to terminate on two last unification problems in under 30s, while both ConsPD and \ecce generate programs which are able to find the unifiers.
-
-\begin{table}
-  \centering
-  \begin{tabular}{c||c|c|c||c}
-    & \multicolumn{3}{c||}{unify} & \\
-    \multirow{2}{*}{Terms} & f(X, a) & f(a \% b \% nil, c \% d \% nil, L) & f(X, X, g(Z, t)) & \multirow{2}{*}{isPath}  \\
-    \cline{2-4} &
-    f(a, X) & f(X \% XS, YS, X \% ZS) & f(g(p, L), Y, Y)  \\
-    \hline\hline
-  Original          & 0.23ms &  ---   &  ---    & 24.9ms \\ \hline
-  ConsPD            & 0.07ms & 16.5ms & 397.6ms & 1.6ms  \\ \hline
-  \ecce             & 0.12ms & 18.8ms & 352.0ms & 1.1ms  \\ \hline
-  \end{tabular}
-
-  \caption{Evaluation results of unification and path search}
-  \label{tbl:unify}
-\end{table}
-
-The relation \lstinline{isPath} was executed to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges (see table~\ref{tbl:unify}).
-Both specializers significantly improve the performance as compared to the original program, although ConsPD is about 30\% worse than \ecce.
-
 \subsection{Typechecker-Term Generator}
@@
 Boolean and integer constants have the corresponding types regardless of the environment.
-Only terms of type integer can be summed up, multiplied or compared by less-then operator.
+Only terms of type integer can be summed up, multiplied or compared by less-than operator.
 Any terms of the same type can be checked for equality.
@@
 Note that the generated program is far slower than hand-written.
-The principal difference between these two implementations is that the generated program is contains a certain redundancy.
+The principal difference between these two implementations is that the generated program contains a certain redundancy.
 For example, typechecking of the sum of two terms in the hand-written implementation is implemented by a single conjunction (see fig.~\ref{type:hand}) while the generated program is far more complicated and also uses a special relation \lstinline{typeEq$^o$} to compare types (see fig.~\ref{type:gen}).
@@
 This is why both implementations have comparable speed after specialization.
-The specialized programs differ too much and it is not clear why one implementation is better specialized by \ecce and the other by ConsPD.
+\todo{The specialized programs differ too much and it is not clear why one implementation is better specialized by \ecce and the other by ConsPD.
+}
+
+\subsection{Unification and Path Search}
+
+Besides evaluator of logic formulas we also run the specializers on the relation \lstinline{unify$^o$} which searches for a unifier of two terms and a relation \lstinline{isPath$^o$}
+specialized to search for paths in the graph.
+Both relations were generated from functional programs by unnesting.
+These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
+
+The \lstinline{unify$^o$} relation was executed to find a unifier of three pairs of terms (see table~\ref{tbl:unify}) which vary in their complexity.
+The original program failed to terminate on two last unification problems in under 30s, while both ConsPD and \ecce generate programs which are able to find the unifiers.
+
+\begin{table}
+  \centering
+  \begin{tabular}{c||c|c|c||c}
+    & \multicolumn{3}{c||}{unify$^o$} & \\
+    \multirow{2}{*}{Terms} & f(X, a) & f(a \% b \% nil, c \% d \% nil, L) & f(X, X, g(Z, t)) & \multirow{2}{*}{isPath$^o$}  \\
+    \cline{2-4} &
+    f(a, X) & f(X \% XS, YS, X \% ZS) & f(g(p, L), Y, Y)  \\
+    \hline\hline
+  Original          & 0.23ms &  ---   &  ---    & 24.9ms \\ \hline
+  ConsPD            & 0.07ms & 16.5ms & 397.6ms & 1.6ms  \\ \hline
+  \ecce             & 0.12ms & 18.8ms & 352.0ms & 1.1ms  \\ \hline
+  \end{tabular}
+
+  \caption{Evaluation results of unification and path search}
+  \label{tbl:unify}
+\end{table}
+
+The relation \lstinline{isPath$^o$} was executed to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges (see table~\ref{tbl:unify}).
+Both specializers significantly improve the performance as compared to the original program, although ConsPD is about 30\% worse than \ecce.

### 2020-07-06 papers/aplas/text/nonconj.tex
Small edits

@@
 
-\section{Non-conjunctive Partial Deduction}
+\section{Conservative Partial Deduction}
 
@@
 
-The algorithm pseudocode is shown in Fig.~\ref{fig:ncpd-pseudo}.
+The algorithm pseudocode is shown in fig.~\ref{fig:ncpd-pseudo}.
 For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} which describe how to process disjunctions and conjunctions respectively.
@@
 Any \mk term can be trivially transformed into the described form.
-In Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization.
+The function \code{normalize} in fig.~\ref{fig:ncpd-pseudo}  is assumed to perform term normalization.
 The code is omitted for brevity.
@@
   \centering
-  \includegraphics[width=0.85\textwidth]{algo_pseudo-crop.pdf}
-  \caption{Non-conjunctive Partial Deduction Pseudo Code}
+  \includegraphics[width=\textwidth]{algo_pseudo-crop.pdf}
+  \caption{Conservative partial deduction pseudo code}
   \label{fig:ncpd-pseudo}
@@
 The first is to select an arbitrary relation to unfold, not necessarily the leftmost which is safe.
-The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time (line 14; \code{heuristically\_select\_a\_call} stands for heuristics combination, see section~\ref{sec:heurictic} for details).
+The second idea is to use a heuristics which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time (line 14; \code{heuristically\_select\_a\_call} stands for heuristics combination, see section~\ref{sec:heurictic} for details).
 If those contradictions are found, then they are exposed by considering the conjunction as a whole and replacing the selected relation call with the result of its unfolding thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD (lines 15--22).
-Joining instead of splitting is why we call our transformer \emph{non-conjunctive} partial deduction.
-Finally, if the heuristic fails to select a potentially good call, then the conjunction is split into individual calls which are driven in isolation and are never joined (line 23).
+Joining instead of splitting is why we call our transformer \emph{conservative} partial deduction.
+Finally, if the heuristics fails to select a potentially good call, then the conjunction is split into individual calls which are driven in isolation and are never joined (line 23).
 
-When the heuristic selects a call to unfold (line 15), a process tree is constructed for the selected call \emph{in isolation} (line 16).
+When the heuristics selects a call to unfold (line 15), a process tree is constructed for the selected call \emph{in isolation} (line 16).
 The leaves of the computed tree are examined.
@@
 The former is maybe even worse than the latter.
-We believe that the following heuristic provides a reasonable approach to unfolding control.
+We believe that the following heuristics provides a reasonable approach to unfolding control.
 
-\subsection{Heuristic}
+\subsection{Less Branching Heuristics}
 \label{sec:heurictic}
 
-This heuristic is aimed at selecting the relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
+This heuristics is aimed at selecting the relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
 The unsafe unfolding leads to an uncontrollable increase in the number of relation calls in a conjunction.
@@
 
-Those relation calls which are neither static nor deterministic are examined with what we call the \emph{less-branching} heuristic.
+Those relation calls which are neither static nor deterministic are examined with what we call the \emph{less-branching} heuristics.
 It identifies the case when the unfolded relation contains fewer disjuncts than it could possibly have.
 This means that we found some contradiction, some computations were gotten rid of, and thus the answer set was restricted, which is desirable when unfolding.
-To compute this heuristic we precompute the maximum possible number of disjuncts in each relation and compare this number with the number of disjuncts when unfolding a concrete relation call.
+To compute this heuristics we precompute the maximum possible number of disjuncts in each relation and compare this number with the number of disjuncts when unfolding a concrete relation call.
 The maximum number of disjuncts is computed by unfolding the body of the relation in which all relation calls were replaced by a unification which always succeeds.
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{heuristic-crop.pdf}
+  \includegraphics[width=\textwidth]{heuristic-crop.pdf}
   \caption{Heuristic selection pseudocode}
@@
 
-The pseudocode describing heuristic is shown in fig.~\ref{fig:heu-pseudo}.
+The pseudocode describing heuristics is shown in fig.~\ref{fig:heu-pseudo}.
 Selecting a good relation call can fail (line 1).
 The implementation works such that we first select those relation calls which are static, and only if there are none, we proceed to consider deterministic unfoldings and then we search for those which are less branching.
-We believe this heuristic provides a good balance in unfolding.
+We believe this heuristics provides a good balance in unfolding.
 % The final heuristic selects the first conjunct which suites either of the following cases.

### 2020-07-06 papers/aplas/text/conclusion.tex
Evaluation intro section rewritten
Conclusion rewritten

--- 2020-05-21 papers/miniKanren/text/conclusion.tex
@@
 
-We compared some of the specialization techniques for \mk{}.
+In this paper we discussed some issues which arise in the area of partial deduction techniques for relational programming language \mk.
+We presented a novel approach to partial deduction --- conservative partial deduction --- which uses a heuristics to select a suitable relation call to unfold at each step of driving.
+We compared this approach with the most sophisticated implementation of conjunctive partial deduction --- \ecce partial deduction system --- on 10 relations which solve 6 different problems.
 
-There seem to be no one good technique.
+Our specializer improved the execution time of 10 queries to the  relations, while worsening 2 queries and not affecting one query.
+\ecce worsened the performance of all 4 implementations of the propositional evaluator relation, while improving other queries.
+Conservative PD is more stable with regrads to the order of relation calls than \ecce which is demonstrated by the similar performance of all 4 implementations of the evaluator of logic formulas.
+
+Some queries to the same relation were improved better by ConsPD, while others --- by \ecce.
+We conclude that there is still no one good technique which definitively speeds up every relational program.
+More research is needed to develop models capable of predicting the performance of a relation which can be used in specialization.

### 2020-07-06 papers/aplas/text/figures/heuristic.tex
Conclusion rewritten

--- 2020-05-23 papers/miniKanren/text/figures/heuristic.tex
@@
 heuristically_select_a_call :: Conjunction @*$\to$@* Maybe Call
-heuristically_select_a_call C =
-  find heuristic C
+heuristically_select_a_call C = find heuristics C
 
-heuristic :: Call -> Bool
-heuristic r =
-  isStatic r || isDeterministic r || isLessBranching r
+heuristics :: Call @*$\to$@* Bool
+heuristics r = isStatic r || isDeterministic r || isLessBranching r
 \end{PseudoCode}

### 2020-07-06 papers/aplas/text/intro.tex
Small edits
Intro about miniKanren added

@@
 
-The core feature of the family of relational programming languages \mk\footnote{\mk language web site: \url{http://minikanren.org}} is their ability to run a program in different directions.
+A family of embedded domain-specific languages \mk\footnote{\mk language web site: \url{http://minikanren.org}} impelement relational programming~---~a paradigm closely related to the pure logic programming.
+The main primitive in \mk is \emph{term unification}.
+Unifications are combined into \emph{conjunctions} and \emph{disjunctions} which form \emph{relations}.
+Relations can call other relations and be recursive.
+A program in \mk consists of a set of relations.
+To run a \mk program, one specifies a \emph{goal}: a \mk expression with some free variables.
+\mk searches for the values of the free variables which satisfy the given goal.
+
+The core feature of relational programming is the ability to run a program in different directions by executing goals with different free variables.
 Having specified a relation for adding two numbers, one can also compute the subtraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
@@
 The promise of \mk falls short when speaking of performance.
-The running time of a program in \mk is highly unpredictable and varies greatly for different directions.
+The execution time of a program in \mk is highly unpredictable and varies greatly for different directions.
 What is even worse, it depends on the order of the relation calls within a program.
@@
 An earlier paper~\cite{lozov2019relational} showed that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk programs.
-Unfortunately, it may also not affect the running time of a program or even make it slower.
+Unfortunately, it may also not affect the execution time of a program or even make it slower.
 
@@
 We also describe a novel approach to partial deduction of a relational programming language \mk.
-We compare it to the existing specialization algorithms on several programs and discuss why some \mk programs run slower after specialization.
+We compare it to the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.

### 2020-07-06 papers/aplas/text/intro.tex
Intro edits

@@
 To run a \mk program, one specifies a \emph{goal}: a \mk expression with some free variables.
-\mk searches for the values of the free variables which satisfy the given goal.
+\mk searches for the values of the free variables which satisfy given goal.
 
-The core feature of relational programming is the ability to run a program in different directions by executing goals with different free variables.
+The core feature of relational programming is the ability to run a program in various directions by executing goals with free variables.
+The distribution of free variable occurrences determines the direction of relational search.
 Having specified a relation for adding two numbers, one can also compute the subtraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
@@
 The promise of \mk falls short when speaking of performance.
-The execution time of a program in \mk is highly unpredictable and varies greatly for different directions.
+The execution time of a program in \mk is highly unpredictable and varies greatly for various directions.
 What is even worse, it depends on the order of the relation calls within a program.

### 2020-07-06 papers/aplas/text/evaluation.tex

@@
 
-The relation \lstinline{doubleAppend$^o$} concatenates three lists by a conjunction of two calls of the \lstinline{append$^o$} relation (see fig.~\ref{doubleApp}).
-These two calls share a variable~--- an intermediate list \lstinline{ts}.
-The list \lstinline{xs} is traversed twice to construct the result: first when \lstinline{ts} is constructed and then when \lstinline{ts} is traversed during the second call.
-This double traversal negatively impacts the execution time, but it can be removed by deforestation.
-
-\begin{figure*}[!h]
+\begin{figure*}[!t]
   \centering
-  \begin{minipage}{0.6\textwidth}
+  \begin{minipage}{\textwidth}
     \begin{lstlisting}[label={doubleApp}, caption={Concatenation of three lists}, captionpos=b, frame=tb]
   let doubleAppend$^o$ xs ys zs res =
-    fresh (ts) (
-      append$^o$ xs ys ts /\ append$^o$ ts zs res)
+    fresh (ts) (append$^o$ xs ys ts /\ append$^o$ ts zs res)
 
@@
     (xs === [] /\ ys === zs);
-    fresh (h t r) (
-      xs === (h % t) /\
-      zs === (h % r) /\
-      append$^o$ t ys r)]
+    fresh (h t r) (xs === (h % t) /\ zs === (h % r) /\ append$^o$ t ys r)]
+    \end{lstlisting}
+  \end{minipage}
+
+  \begin{minipage}{\textwidth}
+    \begin{lstlisting}[label={doubleApp:ideal}, caption={Ideal implementation of concatenation of three lists}, captionpos=b, frame=tb]
+      let doubleAppend$^o$ xs ys zs res = conde [
+        (xs === [] /\ append$^o$ ys zs res);
+        fresh (h t r) (xs === h % t /\ res === h % r /\ doubleAppend$^o$ t ys zs r)]
     \end{lstlisting}
@@
 
-The better implementation of this relation (see fig.~\ref{doubleApp:ideal}) does not traverse the first list twice.
-It first recursively copies \lstinline{xs} into the beginning of \lstinline{res}, and only when \lstinline{xs} is fully consumed, it calls the \lstinline{append$^o$} to concatenate \lstinline{ys} and \lstinline{zs}.
-This implementation is provided as the ideal implementation of \lstinline{doubleAppend$^o$} in~\cite{de1999conjunctive}.
-
-\begin{figure*}[!h]
-  \centering
-  \begin{minipage}{0.6\textwidth}
-    \begin{lstlisting}[label={doubleApp:ideal}, caption={Ideal implementation of concatenation of three lists}, captionpos=b, frame=tb]
-  let doubleAppend$^o$ xs ys zs res = conde [
-    (xs === [] /\ append$^o$ ys zs res);
-    fresh (h t r) (
-      xs === h % t /\
-      res === h % r /\
-      doubleAppend$^o$ t ys zs r)]
-    \end{lstlisting}
-  \end{minipage}
-\end{figure*}
-
-To automatically transform the initial program into the ideal implementation, the transformer should treat the conjuncts which share variables with care.
-Partial deduction splits any conjunction into its subconjunctions, regardless variable sharing, while CPD only splits conjunctions when potential non-termination is detected.
-Conservative partial deduction processes subconjunctions in isolation and then joins them back together, if it can help to restrict the search space.
-In this example, ConsPD fails to perform deforestation: ConsPD first splits the conjunction \lstinline{append$^o$ xs ys ts /\ append$^o$ ts zs rs} into two relation calls of \lstinline{append$^o$}, then unfolds the first call, but since the second call is a renaming of the first one, it is not unfolded.
-The generated program is almost the same as the original: the only difference is that the fresh variables are introduced on the toplevel.
-
-We run all programs in the forward and the backward direction.
-The forward direction \lstinline{doubleAppend$^o$ x y z ?} concatenates three given lists: the first one of length 700, the second and the third~--- of length 1.
-The backward direction \lstinline{doubleAppend$^o$ ? ? ? r} searches for the first 100 list triples whose concatenation gives the given list of length 700.
-
-\begin{table}
+\begin{figure}[!t]
+  \begin{subfigure}[c]{0.35\textwidth}
     \centering
@@
     \label{tbl:doubleApp}
-\end{table}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.6\textwidth}
+    \includegraphics[width=\textwidth]{../data/doubleAppendo/da700log.pdf}
+    \caption{???}
+    %\label{tbl:doubleApp}
+  \end{subfigure}
+  \caption{???}
+\end{figure}
 
-\includegraphics{../data/doubleAppendo/da700log.pdf}
+The relation \lstinline{doubleAppend$^o$} concatenates three lists by a conjunction of two calls of the \lstinline{append$^o$} relation (see fig.~\ref{doubleApp}).
+These two calls share a variable~--- an intermediate list \lstinline{ts}.
+The list \lstinline{xs} is traversed twice to construct the result: first when \lstinline{ts} is constructed and then when \lstinline{ts} is traversed during the second call.
+This double traversal negatively impacts the execution time, but it can be removed by deforestation.
+
+The better implementation of this relation (see fig.~\ref{doubleApp:ideal}) does not traverse the first list twice.
+It first recursively copies \lstinline{xs} into the beginning of \lstinline{res}, and only when \lstinline{xs} is fully consumed, it calls the \lstinline{append$^o$} to concatenate \lstinline{ys} and \lstinline{zs}.
+This implementation is provided as the ideal implementation of \lstinline{doubleAppend$^o$} in~\cite{de1999conjunctive}.
+
+To automatically transform the initial program into the ideal implementation, the transformer should treat the conjuncts which share variables with care.
+Partial deduction splits any conjunction into its subconjunctions, regardless variable sharing, while CPD only splits conjunctions when potential non-termination is detected.
+Conservative partial deduction processes subconjunctions in isolation and then joins them back together, if it can help to restrict the search space.
+In this example, ConsPD fails to perform deforestation: ConsPD first splits the conjunction \lstinline{append$^o$ xs ys ts /\ append$^o$ ts zs rs} into two relation calls of \lstinline{append$^o$}, then unfolds the first call, but since the second call is a renaming of the first one, it is not unfolded.
+The generated program is almost the same as the original: the only difference is that the fresh variables are introduced on the toplevel.
+
+We run all programs in the forward and the backward direction.
+The forward direction \lstinline{doubleAppend$^o$ x y z ?} concatenates three given lists: the first one of length 700, the second and the third~--- of length 1.
+The backward direction \lstinline{doubleAppend$^o$ ? ? ? r} searches for the first 100 list triples whose concatenation gives the given list of length 700.
 
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!t]
   \centering
@@
   let maxLength$^o$ xs m l = max$^o$ xs m /\ length$^o$ xs l
-  let rec length$^o$ xs l =
-    fresh (h t m) ( conde [
+  let rec length$^o$ xs l = fresh (h t m) ( conde [
       (xs === []    /\ l === zero);
@@
   let max$^o$ xs m = max$_1^o$ xs zero m
-  let rec max$_1^o$ xs n m =
-    fresh (h t) ( conde [
+  let rec max$_1^o$ xs n m = fresh (h t) ( conde [
       (xs === []     /\  m === n);
-      (xs === h % t) /\
-      (conde [
+      (xs === h % t) /\ (conde [
         (le$^o$ h n ^true /\  max$_1^o$ t n m);
@@
 
-  let rec le$^o$ x y b =
-    fresh (x$_1$ y$_1$) ( conde [
+  let rec le$^o$ x y b = fresh (x$_1$ y$_1$) ( conde [
       (x === zero    /\ b === ^true);
@@
       (x === succ x$_1$ /\  y === succ y$_1$ /\  le$^o$ x$_1$ y$_1$ b)])
-  let rec gt$^o$ x y b =
-    fresh (x$_1$ y$_1$) ( conde [
+  let rec gt$^o$ x y b = fresh (x$_1$ y$_1$) ( conde [
       (x === zero    /\ b === ^false);
@@
 \end{minipage}
-\end{figure*}
-
-The ideal implementation provided in the CPD literature~\cite{de1999conjunctive} is shown on fig.~\ref{ideal:maxandlength}.
-This implementation traverses the input list \lstinline{xs} only once and can be generated by tupling.
-Note, the disjunct which unifies the third argument with \lstinline{^false}\footnote{An arrow lifts ordinary values to the logic domain} in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
-Such disjuncts are removed by both \ecce and conservative partial deduction, thus we consider two ideal implementions: one with the original comparison relations (\emph{Ideal}) and another in which these disjuncts are deleted (\emph{Ideal removed}).
-
-\begin{figure*}[!h]
-  \centering
-  \begin{minipage}{0.7\textwidth}
+  \begin{minipage}{0.8\textwidth}
 \begin{lstlisting}[label={ideal:maxandlength}, caption={Ideal implementation of maxlengtho}, captionpos=b, frame=tb]
   let maxLength$^o$ xs m l = maxLength$_1^o$ xs m zero l
-  let rec maxLength$_1^o$ xs m n l =
-    fresh (h t l$_1$) ( conde [
-      (xs === [] /\ m === n /\ l === zero);
-      ((xs === h % t) /\
-       (l === succ l$_1$) /\
-       (conde [
+  let rec maxLength$_1^o$ xs m n l = fresh (h t l$_1$) ( conde [
+      ( xs === []     /\  m === n        /\ l === zero);
+      ((xs === h % t) /\ (l === succ l$_1$) /\  (conde [
          (le$^o$ h n ^true /\  maxLength$_1^o$ t m n l);
@@
 
-We measured the execution time of \lstinline{maxLength$^o$ lst m l}, where \lstinline{lst} is a list of Peano numbers from $1$ to $100$.
-The results are presented in table~\ref{tbl:maxlen}.
-
-\begin{table}
+\begin{figure}[!t]
+  \begin{subfigure}[c]{0.4\textwidth}
     \centering
@@
     \label{tbl:maxlen}
-\end{table}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{../data/maxLengtho/max.pdf}
+    \caption{???}
+    %\label{tbl:doubleApp}
+  \end{subfigure}
+  \caption{???}
+\end{figure}
 
-\includegraphics{../data/maxLengtho/max.pdf}
+The ideal implementation provided in the CPD literature~\cite{de1999conjunctive} is shown on fig.~\ref{ideal:maxandlength}.
+This implementation traverses the input list \lstinline{xs} only once and can be generated by tupling.
+Note, the disjunct which unifies the third argument with \lstinline{^false}\footnote{An arrow lifts ordinary values to the logic domain} in the implementation of both \lstinline{le$^o$} and \lstinline{gt$^o$} can never contribute into the result of \lstinline{maxLength$^o$} execution.
+Such disjuncts are removed by both \ecce and conservative partial deduction, thus we consider two ideal implementions: one with the original comparison relations (\emph{Ideal}) and another in which these disjuncts are deleted (\emph{Ideal removed}).
+
+We measured the execution time of \lstinline{maxLength$^o$ lst m l}, where \lstinline{lst} is a list of Peano numbers from $1$ to $100$.
+The results are presented in table~\ref{tbl:maxlen}.
 
@@
     \begin{lstlisting}[label={eval:last}, caption={Evaluator of formulas with boolean operation last}, captionpos=b, frame=tb]
-  let rec eval$^o$ subst fm res = conde [
-    fresh (x y z v w) (
+  let rec eval$^o$ subst fm res = conde [fresh (x y z v w) (
       (fm === var v    /\ elem$^o$ subst v res);
@@
       (fm === neg x    /\ eval$^o$ st x v /\  not$^o$ v res))]
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.95\textwidth}
+    \begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b, frame=tb]
+  let rec eval$^o$ subst fm res = conde [fresh (x y z v w) (
+      (fm === var v    /\ elem$^o$ subst v res);
+      (fm === conj x y /\ and$^o$ v w res /\  eval$^o$ st x v /\  eval$^o$ st y w);
+      (fm === disj x y /\ or$^o$   v w res /\ eval$^o$ st x v /\  eval$^o$ st y w);
+      (fm === neg x    /\ not$^o$ v res   /\ eval$^o$ st x v))]
     \end{lstlisting}
@@
 
-\begin{figure*}[!h]
-  \centering
-  \begin{minipage}{0.95\textwidth}
-    \begin{lstlisting}[label={eval:fst}, caption={Evaluator of formulas with boolean operation second}, captionpos=b, frame=tb]
-  let rec eval$^o$ subst fm res = conde [
-    fresh (x y z v w) (
-      (fm === var v /\ elem$^o$ subst v res);
-      (fm === conj x y /\ and$^o$ v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
-      (fm === disj x y /\ or$^o$ v w res /\ eval$^o$ st x v /\ eval$^o$ st y w);
-      (fm === neg x /\ not$^o$ v res /\ eval$^o$ st x v))]
-    \end{lstlisting}
-  \end{minipage}
-\end{figure*}
-
 \subsubsection{Unfolding of Complex Relations.}
@@
 
-\begin{figure*}[!h]
+\begin{figure*}[!t]
   \centering
-  \begin{minipage}{0.45\textwidth}
+  \begin{minipage}{0.7\textwidth}
     \begin{lstlisting}[label={not:table}, caption={Implementation of boolean \lstinline{not} as a table}, captionpos=b, frame=tb]
@@
       x === ^false /\ y === ^true)]
+    \end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{0.9\textwidth}
+    \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b, frame=tb]
+  let not$^o$   x y = nand$^o$ x x y
+  let or$^o$   x y z = nand$^o$ x x xx /\  nand$^o$ y y yy /\ nand$^o$ xx yy z
+  let and$^o$ x y z = nand$^o$ x y xy /\   nand$^o$ xy xy z
+  let nand$^o$ a b c = conde [
+    ( a === ^false /\ b === ^false /\ c === ^true );
+    ( a === ^false /\ b === ^true  /\ c === ^true );
+    ( a === ^true  /\ b === ^false /\ c === ^true );
+    ( a === ^true  /\ b === ^true  /\ c === ^false)]
     \end{lstlisting}
@@
 Conservative partial deduction drives the selected call until it derives useful substitutions for the variables involved while CPD with deterministic unfolding may fail to do so.
-
-\begin{figure*}[!h]
-  \centering
-  \begin{minipage}{0.85\textwidth}
-    \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b, frame=tb]
-  let not$^o$ x y = nand$^o$ x x y
-
-  let or$^o$ x y z = nand$^o$ x x xx /\ nand$^o$ y y yy /\ nand$^o$ xx yy z
-
-  let and$^o$ x y z = nand$^o$ x y xy /\ nand$^o$ xy xy z
-
-  let nand$^o$ a b c = conde [
-    ( a === ^false /\ b === ^false /\ c === ^true );
-    ( a === ^false /\ b === ^true /\ c === ^true );
-    ( a === ^true /\ b === ^false /\ c === ^true );
-    ( a === ^true /\ b === ^true /\ c === ^false )]
-    \end{lstlisting}
-  \end{minipage}
-\end{figure*}
 
@@
 
-\begin{table}
+\begin{figure}[!t]
+  \centering
+  \begin{subfigure}[c]{0.6\textwidth}
     \centering
@@
     \label{tbl:eval}
-\end{table}
+  \end{subfigure}
 
-\includegraphics{../data/propEval/prop.pdf}
+  \begin{subfigure}[c]{0.8\textwidth}
+    \includegraphics[width=\textwidth]{../data/propEval/prop.pdf}
+    \caption{???}
+  \end{subfigure}
+  \caption{???}
+\end{figure}
 
@@
 The variables are represented with de Bruijn indices, thus let-binding does not specify which variable is being bound.
-
-% \begin{align*}
-%   &type \ term = \\
-%   &| \ BConst \ of \ Bool \\
-%   &| \ IConst \ of \ Int \\
-%   &| \ Var \ of \ Int \\
-%   &| \ Plus \ of \ term * term \\
-%   &| \ Mult \ of \ term * term \\
-%   &| \ Eq \ of \ term * term \\
-%   &| \ Lt \ of \ term * term \\
-%   &| \ Let \ of \ term * term \\
-%   &| \ If \ of \ term * term * term
-% \end{align*}
 
@@
     ...
-    fresh (x y) (
-      (term === x + y /\
+    fresh (x y) ((term === x + y /\
        typecheck$^o$ gamma x ^(Some Integer) /\
@@
   \end{minipage}
-\end{figure*}
 
-\begin{figure*}[!h]
-  \centering
   \begin{minipage}{0.95\textwidth}
@@
     ...
-    fresh (x y t1 t2) (
-      (term === x + y /\
+    fresh (x y t1 t2) ((term === x + y /\
       conde [
-        typecheck$^o$ gamma x ^None /\
-          res === ^None;
+        typecheck$^o$ gamma x ^None       /\ res === ^None;
         typecheck$^o$ gamma x ^(Some t1) /\
-          typecheck$^o$ gamma y ^None /\
-          res === ^None;
-        typecheck$^o$ gamma x ^(Some t1) /\
-          typecheck$^o$ gamma y ^(Some t2) /\
-          typeEq$^o$ t1 Integer ^true /\
-          typeEq$^o$ t2 Integer ^true /\
+          typecheck$^o$ gamma y ^None     /\ res === ^None;
+        typecheck$^o$ gamma x ^(Some t1) /\  typecheck$^o$ gamma y ^(Some t2) /\
+          typeEq$^o$ t1 Integer ^true    /\ typeEq$^o$ t2 Integer ^true /\
           res === ^(Some Integer);

### 2020-07-06 papers/aplas/text/evaluation.tex
attempt to compress fig 5

@@
   \centering
-  \begin{subfigure}[c]{0.6\textwidth}
+  \begin{subfigure}[c]{0.35\textwidth}
     \centering
-    \begin{tabular}{c||c|c|c|c}
-      & FirstPlain & FirstNando & LastPlain & LastNando \\ \hline\hline
-      Original         & 14.8ms     & 14.5ms     & 7.4ms     & 10.1ms    \\ \hline
-      \ecce            & 16.0ms     & 22.3ms     & 14.0ms    & 14.9ms    \\ \hline
-      ConsPD          & 9.1ms      & 9.7ms      & 9.1ms     & 9.9ms     \\
+    %% \begin{tabular}{c||c|c|c|c}
+    %%   & FirstPlain & FirstNando & LastPlain & LastNando \\ \hline\hline
+    %%   Original         & 14.8ms     & 14.5ms     & 7.4ms     & 10.1ms    \\ \hline
+    %%   \ecce            & 16.0ms     & 22.3ms     & 14.0ms    & 14.9ms    \\ \hline
+    %%   ConsPD          & 9.1ms      & 9.7ms      & 9.1ms     & 9.9ms     \\
+    %% \end{tabular}
+    \begin{tabular}{e{1cm}||c|c|c}
+      & Original & \ecce & ConsPD \\ \hline\hline
+      FirstPlain & 14.8ms & 16.0ms & 9.1ms \\ \hline
+      FirstNando & 14.5ms & 22.3ms & 9.7ms \\ \hline
+      LastPlain & 7.4ms & 14.0ms & 9.1ms \\ \hline
+      LastNando & 10.1ms & 14.9ms & 9.9ms
     \end{tabular}
@@
   \end{subfigure}
-
-  \begin{subfigure}[c]{0.8\textwidth}
+  \hfill
+  \begin{subfigure}[c]{0.6\textwidth}
     \includegraphics[width=\textwidth]{../data/propEval/prop.pdf}

### 2020-07-06 papers/aplas/text/related.tex
Small edits
ECCE introduction

@@
 Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro.
-CPD constructs a tree which models goal evaluation and is similar to SLD-NF tree, then a residual program is generated from this tree.
+CPD constructs a tree which models goal evaluation and is similar to SLDNF tree, then a residual program is generated from this tree.
 Partial deduction itself resembles driving in supercompilation~\cite{gluck1994partial}.
@@
 At the local level, CPD examines a conjunction of atoms by considering each atom one-by-one from left to right.
-An atom is \emph{unfolded} if it is deemed safe, i.e. a whistle based on homeomorphic does not signal for the atom.
+An atom is \emph{unfolded} if it is deemed safe, i.e. a whistle based on homeomorphic embedding does not signal for the atom.
 When an atom is unfolded, a clause whose head can be unified with the atom is found, and a new node is added into the tree where the atom in the conjunction is replaced with the body of that clause.
@@
 
-The most well-behaved strategy of local control in CPD for \pro is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
+\ecce partial deduction system~\cite{leuschel1997ecce} is the most mature implementation of CPD for \pro.
+\ecce provides various implementations of both local and global control as well as several degrees of post-processing.
+Unfortunately there is neither an automatic procedure to choose what control setting is likely to improve input programs the most nor any informal recommendations on how to choose the best settings.
+The choice of the proper control is left to the user.
+
+An empitical study shown that the most well-behaved strategy of local control in CPD for \pro is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
 An atom is unfolded only if only one suitable clause head exists for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.

### 2020-07-06 papers/aplas/text/evaluation.tex
remove subfig captions
compress even more
fix indent in listing 1.10
fix refs

@@
 
-\begin{figure}[!t]
-  \begin{subfigure}[c]{0.35\textwidth}
-    \centering
-    \begin{tabular}{c||c||c}
-      & forward & backward \\
-      \hline\hline
-      Original       & 34.8ms & 1.6ms \\ \hline
-      \ecce          & 26.7ms & 1.7ms \\ \hline
-      Ideal          & 26.9ms & 1.7ms \\ \hline
-      ConsPD         & 47.6ms & 1.7ms
-    \end{tabular}
-    \caption{Evaluation results for doubleAppendo}
-    \label{tbl:doubleApp}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[c]{0.6\textwidth}
-    \includegraphics[width=\textwidth]{../data/doubleAppendo/da700log.pdf}
-    \caption{???}
-    %\label{tbl:doubleApp}
-  \end{subfigure}
-  \caption{???}
-\end{figure}
-
-The relation \lstinline{doubleAppend$^o$} concatenates three lists by a conjunction of two calls of the \lstinline{append$^o$} relation (see fig.~\ref{doubleApp}).
+The relation \lstinline{doubleAppend$^o$} concatenates three lists by a conjunction of two calls of the \lstinline{append$^o$} relation (see Listing~\ref{doubleApp}).
 These two calls share a variable~--- an intermediate list \lstinline{ts}.
@@
 
-The better implementation of this relation (see fig.~\ref{doubleApp:ideal}) does not traverse the first list twice.
+The better implementation of this relation (see Listing~\ref{doubleApp:ideal}) does not traverse the first list twice.
 It first recursively copies \lstinline{xs} into the beginning of \lstinline{res}, and only when \lstinline{xs} is fully consumed, it calls the \lstinline{append$^o$} to concatenate \lstinline{ys} and \lstinline{zs}.
@@
 
-There is no significant difference in the execution time in the backward direction between different specialized versions (see table~\ref{tbl:doubleApp}).
+There is no significant difference in the execution time in the backward direction between different specialized versions (see Fig.~\ref{tbl:doubleApp}).
 However the execution time in the forward direction differs: the program generated by \ecce shows the same speedup as the ideal program, while conservative partial deduction slows the program down.
@@
 
+\begin{figure}[!t]
+  \begin{subfigure}[c]{0.35\textwidth}
+    \centering
+    \begin{tabular}{c||c||c}
+      & forward & backward \\
+      \hline\hline
+      Original       & 34.8ms & 1.6ms \\ \hline
+      \ecce          & 26.7ms & 1.7ms \\ \hline
+      Ideal          & 26.9ms & 1.7ms \\ \hline
+      ConsPD         & 47.6ms & 1.7ms
+    \end{tabular}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.6\textwidth}
+    \includegraphics[width=\textwidth]{../data/doubleAppendo/da700log.pdf}
+  \end{subfigure}
+    \caption{Evaluation results for doubleAppendo}
+    \label{tbl:doubleApp}
+\end{figure}
+
 \subsection{Maximal Element and Length of a List}
-
-The relation \lstinline{maxLenght$^o$} computes the maximal element of the input list and its length (see fig.~\ref{cpd:maxandlength}) by conjunction of two relation calls.
-The list \lstinline{xs} is traversed twice: first to compute the maximum in the relation \lstinline{max$^o$} and then to compute the length in the relation \lstinline{length$^o$}.
-The input list contain Peano numbers which are compared by relations less-or-equal (\lstinline{le$^o$}) and greater-than (\lstinline{gt$^o$}).
-Note that the comparison relations have three arguments with the last one representing the result of comparison.
 
@@
 
-\begin{figure}[!t]
+\begin{figure}[!h]
   \begin{subfigure}[c]{0.4\textwidth}
@@
     \end{tabular}
-
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{../data/maxLengtho/max.pdf}
+  \end{subfigure}
   \caption{Execution time of maxlengtho}
   \label{tbl:maxlen}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[c]{0.5\textwidth}
-    \includegraphics[width=\textwidth]{../data/maxLengtho/max.pdf}
-    \caption{???}
-    %\label{tbl:doubleApp}
-  \end{subfigure}
-  \caption{???}
 \end{figure}
 
-The ideal implementation provided in the CPD literature~\cite{de1999conjunctive} is shown on fig.~\ref{ideal:maxandlength}.
+The relation \lstinline{maxLenght$^o$} computes the maximal element of the input list and its length (see Listing~\ref{cpd:maxandlength}) by conjunction of two relation calls.
+The list \lstinline{xs} is traversed twice: first to compute the maximum in the relation \lstinline{max$^o$} and then to compute the length in the relation \lstinline{length$^o$}.
+The input list contain Peano numbers which are compared by relations less-or-equal (\lstinline{le$^o$}) and greater-than (\lstinline{gt$^o$}).
+Note that the comparison relations have three arguments with the last one representing the result of comparison.
+
+The ideal implementation provided in the CPD literature~\cite{de1999conjunctive} is shown on Listing~\ref{ideal:maxandlength}.
 This implementation traverses the input list \lstinline{xs} only once and can be generated by tupling.
@@
 We measured the execution time of \lstinline{maxLength$^o$ lst m l}, where \lstinline{lst} is a list of Peano numbers from $1$ to $100$.
-The results are presented in table~\ref{tbl:maxlen}.
+The results are presented in Fig.~\ref{tbl:maxlen}.
 
@@
 
-One possible implementation of the evaluator is presented in fig.~\ref{eval:last}.
+One possible implementation of the evaluator is presented in Listing~\ref{eval:last}.
 Here the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
@@
 The way conservative partial deduction treats this program is the same as it treats the other implementation in which boolean connectives
-are moved to the left, as shown in fig.~\ref{eval:fst}.
+are moved to the left, as shown in Listing~\ref{eval:fst}.
 This program is easier for \ecce to specialize which demonstrates how unequal is the behaviour of CPD for similar programs.
@@
 But depending on the way they are implemented, the different number of driving steps should be performed for that.
-The simplest way to implement these relations is by mimicking a truth tables as demonstrated by the implementation of \lstinline{not$^o$} in fig.~\ref{not:table}.
+The simplest way to implement these relations is by mimicking a truth tables as demonstrated by the implementation of \lstinline{not$^o$} in Listing~\ref{not:table}.
 It is enough to unfold such relation calls once to derive useful information about variables.
@@
 The other way to implement boolean connectives is to express them using a single basic boolean relation such as \lstinline{nand$^o$} which is, in turn, has a table-based
-implementation (see fig.~\ref{not:nando}). It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should
+implementation (see Listing~\ref{not:nando}). It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should
 be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
@@
 We measured the time necessary to generate $1000$ formulas over two variables which evaluate to \lstinline{^true} (averaged over 10 runs).
-The results are presented in table~\ref{tbl:eval}.
+The results are presented in Fig.~\ref{tbl:eval}.
 
@@
     \end{tabular}
-    \caption{Execution time of evalo}
-    \label{tbl:eval}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]{../data/propEval/prop.pdf}
-    \caption{???}
   \end{subfigure}
-  \caption{???}
+  \caption{Execution time of evalo}
+  \label{tbl:eval}
 \end{figure}
@@
 
-\begin{table}
-  \setlength{\tabcolsep}{0.5cm}
-  \centering
-  \begin{tabular}{c c}
-    \infer[]{\Gamma \vdash IConst \ i : Int}{} &
-    \infer[]{\Gamma \vdash BConst \ b : Bool}{} \vspace{0.5cm} \\
-
-    \infer[]{\Gamma \vdash t + s : Int}{\Gamma \vdash t : Int, \Gamma \vdash  s : Int} &
-    \infer[]{\Gamma \vdash t * s : Int}{\Gamma \vdash t : Int, \Gamma \vdash  s : Int} \vspace{0.5cm} \\
-
-    \infer[]{\Gamma \vdash t = s : Bool}{\Gamma \vdash t : \tau, \Gamma \vdash  s : \tau} &
-    \infer[]{\Gamma \vdash t < s : Bool}{\Gamma \vdash t : Int, \Gamma \vdash  s : Int} \vspace{0.5cm} \\
-
-    \infer[]{\Gamma \vdash \underline{let} \ v \ b : \tau}{\Gamma \vdash v : \tau_v, \ (\tau_v :: \Gamma) \vdash b : \tau} &
-    \infer[\Gamma \lbrack v \rbrack \equiv \tau]{\Gamma \vdash Var \ v : \tau}{} \vspace{0.5cm} \\
-
-    \multicolumn{2}{c}{
-      \infer[]{\Gamma \vdash \underline{if} \ c \ \underline{then} \ t \ \underline{else} \ s : \tau}{\Gamma \vdash c : Bool, \Gamma \vdash t : \tau, \Gamma \vdash s : \tau}
-    }
-  \end{tabular}
-\end{table}
+\begin{figure*}[!h]
+  \includegraphics[width=\textwidth]{typechecker-crop.pdf}
+\end{figure*}
 
@@
 Each implementation has been specialized with ConsPD and \ecce.
-We measured the time needed to generate 1000 closed terms of type integer (averaged over 10 iterations).
+We measured the time needed to generate 1000 closed terms of type integer (averaged over 10 iterations; see Table~\ref{tbl:eval}).
 
@@
 The principal difference between these two implementations is that the generated program contains a certain redundancy.
-For example, typechecking of the sum of two terms in the hand-written implementation is implemented by a single conjunction (see fig.~\ref{type:hand}) while the generated program is far more complicated and also uses a special relation \lstinline{typeEq$^o$} to compare types (see fig.~\ref{type:gen}).
+For example, typechecking of the sum of two terms in the hand-written implementation is implemented by a single conjunction (see Listing~\ref{type:hand}) while the generated program is far more complicated and also uses a special relation \lstinline{typeEq$^o$} to compare types (see Listing~\ref{type:gen}).
 
@@
 
-The relation \lstinline{isPath$^o$} was executed to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges (see table~\ref{tbl:unify}).
+The relation \lstinline{isPath$^o$} was executed to search for 3 paths of length 7 in the graph with 20 vertices and 30 edges (see Table~\ref{tbl:unify}).
 Both specializers significantly improve the performance as compared to the original program, although ConsPD is about 30\% worse than \ecce.

### 2020-07-07 papers/aplas/text/evaluation.tex
Typing rules prettified
Better explanation for different degrees of specialization of typechecko

@@
   \end{tabular}
-  \caption{Running time of generating 1000 closed terms of type Int}
+  \caption{Execution  time of generating 1000 closed terms of type Integer}
   \label{tbl:eval}
@@
 This is why both implementations have comparable speed after specialization.
-\todo{The specialized programs differ too much and it is not clear why one implementation is better specialized by \ecce and the other by ConsPD.
-}
+\ecce shows bigger speedup for the hand-written than ConsPD and vice versa for the generated implementation.
+We believe that this difference can be explained with too much unfolding.
+\ecce performs a lot of excessive unfolding for the generated program and only barely changes the hand-written program.
+At the same time ConsPD specializes both implementations to comparable programs performing average amount of unfolding.
+This shows that the heuristics we presented gives more stable, although not the best, results.
 

### 2020-07-07 papers/aplas/text/evaluation.tex
Edits

@@
 are moved to the left, as shown in Listing~\ref{eval:fst}.
-This program is easier for \ecce to specialize which demonstrates how unequal is the behaviour of CPD for similar programs.
+This program is easier for \ecce to specialize which demonstrates how unequal the behaviour of CPD for similar programs is.
 
@@
 \ecce shows bigger speedup for the hand-written than ConsPD and vice versa for the generated implementation.
-We believe that this difference can be explained with too much unfolding.
+We believe that this difference can be explained by too much unfolding.
 \ecce performs a lot of excessive unfolding for the generated program and only barely changes the hand-written program.
@@
 
-Besides evaluator of logic formulas we also run the specializers on the relation \lstinline{unify$^o$} which searches for a unifier of two terms and a relation \lstinline{isPath$^o$}
-specialized to search for paths in the graph.
+Besides evaluator of logic formulas we also run the specializers on the relation \lstinline{unify$^o$}, which searches for a unifier of two terms, and a relation \lstinline{isPath$^o$},
+used to search for paths in the graph.
 Both relations were generated from functional programs by unnesting.

### 2020-07-07 papers/aplas/text/intro.tex
Edits

@@
 We also describe a novel approach to partial deduction of a relational programming language \mk.
-We compare it to the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.
+We compare the described approach to the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.

### 2020-07-07 papers/aplas/text/nonconj.tex
Conclusion rewritten
Edits

@@
 
-The algorithm pseudocode is shown in fig.~\ref{fig:ncpd-pseudo}.
+The algorithm pseudocode is shown in Fig.~\ref{fig:ncpd-pseudo}.
 For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} which describe how to process disjunctions and conjunctions respectively.
-Driving itself is a trivial combination of presented functions (line 2).
+Driving itself is a trivial combination of the functions provided (line 2).
 
@@
 The nodes of the process tree include a \emph{configuration} which describes the state of program evaluation at some point.
-In our case configuration is a conjunction of relation calls.
+In our case a configuration is a conjunction of relation calls.
 The substitution computed at each step is also stored in the tree node, although it is not included in the configuration.
@@
 Those disjuncts in which unifications fail are removed.
-Other disjuncts take the form of possibly empty conjunction of relation calls accompanied with a substitution computed from unifications.
+Each other disjunct takes the form of a possibly empty conjunction of relation calls accompanied with a substitution computed from unifications.
 Any \mk term can be trivially transformed into the described form.
-The function \code{normalize} in fig.~\ref{fig:ncpd-pseudo}  is assumed to perform term normalization.
+The function \code{normalize} in Fig.~\ref{fig:ncpd-pseudo} is assumed to perform term normalization.
 The code is omitted for brevity.
@@
 This process creates branchings whenever a disjunction is examined (lines 4--5).
-At each step, we make sure that we do not start driving conjunction which we have already examined.
+At each step, we make sure that we do not start driving a conjunction which we have already examined.
 To do this, we check if the current conjunction is a renaming of any other configuration in the tree (line 11).
@@
 
-We decided not to perform generalization in this approach in the same fashion as CPD or supercompilation does.
+In this approach, we decided not to generalize in the same fashion as CPD or supercompilation.
 Our conjunctions are always split into individual calls and are joined back together only if it is sensible.
 If the need for generalization arises, i.e. homeomorphic embedding of conjunctions~\cite{de1999conjunctive} is detected, then we immediately stop driving this conjunction (line 12).
-When residualizing such conjunction, we just generate a conjunction of calls to the input program before specialization.
+When residualizing such a conjunction, we just generate a conjunction of calls to the input program before specialization.
 
@@
 The unfolding itself is straightforward however it is not always clear what to unfold and when to \emph{stop} unfolding.
-Unfolding in functional programming languages specialization, as well as inlining in imperative one, is usually considered to be safe from the residual program efficiency point of view.
+Unfolding in the context of specialization of functional programming languages, as well as inlining in specialization of imperative languages, is usually considered to be safe from the residual program efficiency point of view.
 It may only lead to code explosion or code duplication which is mostly left to a target program compiler optimization or even is out of consideration at all if a specializer is considered as a standalone tool~\cite{jonesbook}.
 
-Unfortunately, this is not the case for the specialization of relational programming language.
-Unlike functional and imperative, in logic and relational programming language unfolding may easily affect the target program efficiency~\cite{leuschel2002logic}.
-Unfolding too much may create extra unifications, which is by itself a costly operation, or even introduce duplicated computations by propagating the unfolding results onto neighbouring conjuncts.
+Unfortunately, this is not the case for the specialization of a relational programming language.
+Unlike in functional and imperative languages, in logic and relational programming languages unfolding may easily affect the target program's efficiency~\cite{leuschel2002logic}.
+Unfolding too much may create extra unifications, which is by itself a costly operation, or even introduce duplicated computations by propagating the results of unfolding onto neighbouring conjuncts.
 
@@
 
-This heuristics is aimed at selecting the relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
-The unsafe unfolding leads to an uncontrollable increase in the number of relation calls in a conjunction.
+This heuristics is aimed at selecting a relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
+An unsafe unfolding leads to an uncontrollable increase in the number of relation calls in a conjunction.
 It is best to first unfold those relation calls which can be fully computed up to substitutions.
@@
   \includegraphics[width=\textwidth]{heuristic-crop.pdf}
-  \caption{Heuristic selection pseudocode}
+  \caption{Heuristics selection pseudocode}
   \label{fig:heu-pseudo}
@@
 
-The pseudocode describing heuristics is shown in fig.~\ref{fig:heu-pseudo}.
+The pseudocode describing our heuristics is shown in Fig.~\ref{fig:heu-pseudo}.
 Selecting a good relation call can fail (line 1).

### 2020-07-07 papers/aplas/text/related.tex
Edits

@@
 Specialization is an attractive technique aimed to improve the performance of a program if some of its arguments are known statically.
-It is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation} and many more.
+Specialization is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation} and many more.
 
 The heart of supercompilation-based techniques is \emph{driving}~---~a symbolic execution of a program through all possible execution paths.
-The result of driving is so-called \emph{process tree} where nodes correspond to \emph{configurations} which present computation state, for example, a term in case of pure functional programming languages.
+The result of driving is so-called \emph{process tree} where nodes correspond to \emph{configurations} which represent computation state.
+For example, in the case of pure functional programming languages, the computational state might be a term.
 Each path in the tree corresponds to some concrete program execution.
-The two main sources for supercompilation optimizations are aggressive information propagation about variable values, equalities, and disequalities and precomputing of all deterministic semantic evaluation steps, i.e.
-combining of consecutive process tree nodes with no branching, also known as \emph{deforestation}~\cite{deforestation}.
+The two main sources for supercompilation optimizations are aggressive information propagation about variables' values, equalities and disequalities, and precomputing of all deterministic semantic evaluation steps.
+The latter process, also known as \emph{deforestation}~\cite{deforestation}, means  combining of consecutive process tree nodes with no branching.
 When the tree is constructed, the resulting, or \emph{residual}, program can be extracted from the process tree by the process called \emph{residualization}.
 Of course, process tree can contain infinite branches.
-\emph{Whistles} --- heuristics to identify possibly infinite branches --- are used to ensure termination in supercompilation.
-If the whistle signalled during the construction of some branch, then something should be done to ensure termination.
-The most common approaches include either stopping driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or folding the process tree to a \emph{process graph}.
-The main instrument to perform such a folding is \emph{generalization}, i.e. abstracting away some computed data about the current term which makes folding possible.
+\emph{Whistles} --- heuristics to identify possibly infinite branches --- are used to ensure supercompilation terminates.
+If a whistle signalls during the construction of some branch, then something should be done to ensure termination.
+The most common approaches include stop driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or fold the process tree to a \emph{process graph}.
+The main instrument to perform such a folding is \emph{generalization}.
+Generalization, abstracting away some computed data about the current term, makes folding possible.
+%  i.e. abstracting away some computed data about the current term which makes folding possible.
 One source of infinite branches is consecutive recursive calls to the same function with an accumulating parameter: by unfolding such a call further one can only increase the term size which leads to nontermination.
 The accumulating parameter can be removed by replacing the call with its generalization.
-There are several ways to ensure process correctness and termination, the most common being \emph{homeomorphic embedding}~\cite{Higman52,Kruskal60} used as a whistle and most-specific generalization of terms.
+There are several ways to ensure process correctness and termination, most-specific generalization
+(anti-unification) and using \emph{homeomorphic embedding}~\cite{Higman52,Kruskal60} as a
+whistle being the most common.
+%the most common being \emph{homeomorphic embedding}~\cite{Higman52,Kruskal60} used as a whistle and most-specific generalization of terms.
 % When two dangerously similar nodes are encountered,
@@
 While supercompilation generally improves the behaviour of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in the general.
-What is worse, they rarely consider the residual program efficiency from the target language evaluator point of view.
+What is worse, the efficiency of residual program from the target language evaluator point of view is rarely considered in the literature.
 The main optimization source is computing in advance all possible intermediate and statically-known semantics steps at program transformation-time.
@@
 It is known that supercompilation may adversely affect GHC optimizations yielding standalone compilation more powerful~\cite{SCBE,TCES} and cause code explosion~\cite{SCHC}.
-Moreover, it may be hard to predict the real speedup of any given program on concrete examples even disregarding problems above because of the complexity of the transformation algorithm.
-The worst-case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization indeed improved its performance~\cite{jonesbook,bulyonkov84}.
-There is lack of research in determining the classes of programs which transformers would definitely speed up.
+Moreover, it may be hard to predict the real speedup of any given program on concrete examples even disregarding the problems above because of the complexity of the transformation algorithm.
+The worst-case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization indeed improves its performance~\cite{jonesbook,bulyonkov84}.
+There is a lack of research in determining the classes of programs which transformers would definitely speed up.
 
-Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro.
-CPD constructs a tree which models goal evaluation and is similar to SLDNF tree, then a residual program is generated from this tree.
+Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for the left-to-right evaluation strategy of \pro.
+CPD constructs a tree which models goal evaluation and is similar to a SLDNF tree, then a residual program is generated from this tree.
 Partial deduction itself resembles driving in supercompilation~\cite{gluck1994partial}.
@@
 An empitical study shown that the most well-behaved strategy of local control in CPD for \pro is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
-An atom is unfolded only if only one suitable clause head exists for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
-This means that if a non-deterministic atom is the leftmost within conjunction, it is most likely to be unfolded and introduce many new relation calls within the conjunction.
+An atom is unfolded only if precisely one suitable clause head exists for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
+This means that if a non-deterministic atom is the leftmost within a conjunction, it is most likely to be unfolded and to introduce many new relation calls within the conjunction.
 We believe this is the core problem with CPD which limits its power when applied to \mk.

### 2020-07-07 papers/aplas/text/intro.tex
Edits to intro

--- 2020-07-06 papers/aplas/text/intro.tex
@@
 The distribution of free variable occurrences determines the direction of relational search.
-Having specified a relation for adding two numbers, one can also compute the subtraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
-Program synthesis can be done by running \emph{backwards} a relational interpreter for some language.
+For example, having specified a relation for adding two numbers, one can also compute the subtraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
+Program synthesis can be done by running a relational interpreter for some language \emph{backwards}.
 In general, it is possible to create a solver for a recognizer by translating it into \mk and running in the appropriate direction~\cite{lozov2019relational}.
@@
 It may either be a known value of some argument, its structure (i.e. the length of an input list) or, in case of a relational program, --- the direction in which it is intended to be run.
-An earlier paper~\cite{lozov2019relational} showed that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk programs.
+An earlier paper~\cite{lozov2019relational} has shown that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk programs.
 Unfortunately, it may also not affect the execution time of a program or even make it slower.
@@
 Control issues in partial deduction of logic programming language \pro have been studied before~\cite{leuschel2002logic}.
-The ideas described there are aimed at left-to-right evaluation strategy of \pro.
-Since the search in \mk is complete, it is safe to reorder some relation calls within the goal for better performance.
-While sometimes conjunctive partial deduction gives great performance boost, sometimes it does not behave as well as it could have.
+Unlike Prolog, where atoms in the rhs of a clause cannot be arbitrarily reordered without changing the semantics of a program, in miniKanren the subgoals of conjunction/disjunction can be freely switched.
+This opens a yet another possibility for optimization, not taken into account by approaches initially developed in the context of conventional logic programming.
+% The ideas described there make use of the deterministic search strategy of \pro.
+% However in \mk it is allowed to reorder some relation calls within the goal for better performance.
+% While sometimes conjunctive partial deduction gives great performance boost, sometimes it does not behave as well as it could have.
 
-In this paper, we show on examples some issues which conjunctive partial deduction faces.
-We also describe a novel approach to partial deduction of a relational programming language \mk.
-We compare it to the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.
+In this paper, we study issues which conjunctive partial deduction faces being applied for \mk.
+We also describe a novel approach to partial deduction for relational programming, \emph{conservative partial deduction}.
+We implement this approach and compare it with the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.

### 2020-07-07 papers/aplas/text/intro.tex
Intro edits

@@
 An earlier paper~\cite{lozov2019relational} has shown that \emph{conjunctive partial deduction}~\cite{de1999conjunctive} can sometimes improve the performance of \mk programs.
-Unfortunately, it may also not affect the execution time of a program or even make it slower.
+Depending on the particular \emph{control} decisions, it may also not affect the execution time of a program or even make it slower.
 
@@
 We also describe a novel approach to partial deduction for relational programming, \emph{conservative partial deduction}.
-We implement this approach and compare it with the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.
+We implemented this approach and compared it with the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.

### 2020-07-07 papers/aplas/text/intro.tex
Intro edits

@@
 We also describe a novel approach to partial deduction for relational programming, \emph{conservative partial deduction}.
-We implemented this approach and compared it with the existing specialization system (\ecce) on several programs and discuss why some \mk programs run slower after specialization.
+We implemented this approach and compared it with the existing specialization system (\ecce) for several programs.
+We report here the result of the comparison and discuss why some \mk programs run slower after specialization.

### 2020-07-07 papers/aplas/text/abstract.tex
Abstract

@@
 \begin{abstract}
-  We explore partial deduction for \mk: a specialization technique aimed at improving the performance of a relation in the given direction.
-  We describe a novel approach to specialization of \mk based on partial deduction and supercompilation.
-  On several examples, we demonstrate issues which arise during partial deduction.
+  We study conjunctive partial deduction, an advanced specialization technique aimed at improving the performance of logical programs, in the context of relational programming language \mk.
+We identify a number of issues, caused by  \mk pecularities, and describe a novel approach to specialization based on partial deduction and supercompilation.
+While the results of the evaluation do not yet demonstrate the advantages of our approach in all cases, we nevertheless consider it as the first step towards an efficient optimization framework for \mk.
 \end{abstract}

### 2020-07-07 papers/aplas/text/intro.tex
Abstract
Edits

@@
 Control issues in partial deduction of logic programming language \pro have been studied before~\cite{leuschel2002logic}.
-Unlike Prolog, where atoms in the rhs of a clause cannot be arbitrarily reordered without changing the semantics of a program, in miniKanren the subgoals of conjunction/disjunction can be freely switched.
+Unlike Prolog, where atoms in the rhs of a clause cannot be arbitrarily reordered without changing the semantics of a program, in \mk the subgoals of conjunction/disjunction can be freely switched.
 This opens a yet another possibility for optimization, not taken into account by approaches initially developed in the context of conventional logic programming.
@@
 We implemented this approach and compared it with the existing specialization system (\ecce) for several programs.
-We report here the result of the comparison and discuss why some \mk programs run slower after specialization.
+We report here the results of the comparison and discuss why some \mk programs run slower after specialization.

### 2020-07-07 papers/aplas/text/related.tex
Abstract
Edits

@@
 If a whistle signalls during the construction of some branch, then something should be done to ensure termination.
-The most common approaches include stop driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or fold the process tree to a \emph{process graph}.
+The most common approaches are either to stop driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or to fold the process tree to a \emph{process graph}.
 The main instrument to perform such a folding is \emph{generalization}.
@@
 
-While supercompilation generally improves the behaviour of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in the general.
+While supercompilation generally improves the behaviour of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in general.
 What is worse, the efficiency of residual program from the target language evaluator point of view is rarely considered in the literature.
@@
 The worst-case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization indeed improves its performance~\cite{jonesbook,bulyonkov84}.
-There is a lack of research in determining the classes of programs which transformers would definitely speed up.
+There is a lack of research in determining the classes of programs which transformers would definitely speed~up.
 
@@
 This means that if a non-deterministic atom is the leftmost within a conjunction, it is most likely to be unfolded and to introduce many new relation calls within the conjunction.
-We believe this is the core problem with CPD which limits its power when applied to \mk.
+We believe this is the core problem of CPD which limits its power when applied to \mk.
 The strategy of unfolding atoms from left to right is reasonable in the context of \pro because it mimics the way programs in \pro execute.

### 2020-07-07 papers/aplas/text/evaluation.tex
Added histogram for ltypecheck
Done

@@
 
-In our study we compared the new conservative partial deduction with the \ecce partial deduction system.
+We implemented the conservative partial deduction for \mk and compared it with the \ecce partial deduction system.
 \ecce is designed for \pro programming language and cannot be directly applied for programs, written in \mk.
@@
 The \lstinline{doubleAppend$^o$} program concatenates three lists of numbers by computing a conjunction of two calls of a relation which concatenates two lists.
-This example is prominent in CPD literature, since it demonstrates \emph{deforestation}~--- a transformation which gets rid of intermidiate datastructures.
-This transformation cannot be done with earlier partial deduction approachs, since the relation uses a variable shared between two relation calls within a
+This example is prominent in CPD literature, since it demonstrates \emph{deforestation}~--- a transformation which gets rid of intermidiate data structures.
+This transformation cannot be done with earlier partial deduction approaches, since the relation uses a variable shared between two relation calls within a
 conjunction which requires the transformation to treat conjunctions as a whole.
@@
 \subsubsection{Evaluation Results.}
-In our study we considered four implementations of \lstinline{eval$^o$}:
+In our study we considered 4 implementations of \lstinline{eval$^o$}:
 \begin{itemize}
-  \item \emph{FirstPlain} in which the implementations of boolean connectives are \textbf{table-based} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
-  \item \emph{LastPlain} in which the implementations of boolean connectives are \textbf{table-based} and are placed \textbf{after} recursive calls to \lstinline{eval$^o$}
-  \item \emph{FirstNando} in which boolean connectives are implemented via \lstinline{nand$^o$} and are placed \textbf{before} recursive calls to \lstinline{eval$^o$}
-  \item \emph{LastNando} in which boolean connectives are implemented via \lstinline{nand$^o$} andare placed \textbf{after} recursive calls to \lstinline{eval$^o$}
+  \item \emph{FirstPlain} in which the implementations of boolean connectives are \textbf{table-based} and are placed \textbf{before} the recursive calls to \lstinline{eval$^o$}
+  \item \emph{LastPlain} in which the implementations of boolean connectives are \textbf{table-based} and are placed \textbf{after} the recursive calls to \lstinline{eval$^o$}
+  \item \emph{FirstNando} in which boolean connectives are implemented via \lstinline{nand$^o$} and are placed \textbf{before} the recursive calls to \lstinline{eval$^o$}
+  \item \emph{LastNando} in which boolean connectives are implemented via \lstinline{nand$^o$} and are placed \textbf{after} the recursive calls to \lstinline{eval$^o$}
 
@@
 
+\vspace{-15pt}
 \begin{figure*}[!h]
+  \centering
   \includegraphics[width=\textwidth]{typechecker-crop.pdf}
 \end{figure*}
+\vspace{-15pt}
 
@@
 Each implementation has been specialized with ConsPD and \ecce.
-We measured the time needed to generate 1000 closed terms of type integer (averaged over 10 iterations; see Table~\ref{tbl:eval}).
+We measured the time needed to generate 1000 closed terms of type integer (see Fig.~\ref{tbl:type}).
 
-\begin{table}
+\begin{figure}[!h]
+  \begin{subfigure}[c]{0.4\textwidth}
+    \centering
     \centering
@@
     \end{tabular}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[c]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{../data/lTypecheck/ltypelog.pdf}
+  \end{subfigure}
   \caption{Execution  time of generating 1000 closed terms of type Integer}
-  \label{tbl:eval}
-\end{table}
+  \label{tbl:type}
+\end{figure}
 
-Note that the generated program is far slower than hand-written.
+Note that the generated program is far slower than the hand-written.
 The principal difference between these two implementations is that the generated program contains a certain redundancy.
@@
 This is why both implementations have comparable speed after specialization.
-\ecce shows bigger speedup for the hand-written than ConsPD and vice versa for the generated implementation.
+\ecce shows bigger speedup for the hand-written program than ConsPD and vice versa for the generated implementation.
 We believe that this difference can be explained by too much unfolding.
@@
 The \lstinline{unify$^o$} relation was executed to find a unifier of three pairs of terms (see table~\ref{tbl:unify}) which vary in their complexity.
-The original program failed to terminate on two last unification problems in under 30s, while both ConsPD and \ecce generate programs which are able to find the unifiers.
+The original program failed to terminate on two last unification problems in under 30 seconds, while both ConsPD and \ecce generate programs which are able to find the unifiers.
 

### 2020-07-09 papers/aplas/text/intro.tex
Brushed up

@@
 To run a \mk program, one specifies a \emph{goal}: a \mk expression with some free variables.
-\mk searches for the values of the free variables which satisfy given goal.
+\mk then searches for the values of the free variables which satisfy given goal.
 
@@
 Program synthesis can be done by running a relational interpreter for some language \emph{backwards}.
-In general, it is possible to create a solver for a recognizer by translating it into \mk and running in the appropriate direction~\cite{lozov2019relational}.
+In general, it is possible to create a solver from a recognizer by translating it into \mk and running in the appropriate direction~\cite{lozov2019relational}.
 
@@
 What is even worse, it depends on the order of the relation calls within a program.
-One order can be good for one direction, but slow down the computation drastically in the other direction.
+One order can be good for one direction, but slow down the computation dramatically in the other direction.
 
@@
 Control issues in partial deduction of logic programming language \pro have been studied before~\cite{leuschel2002logic}.
-Unlike Prolog, where atoms in the rhs of a clause cannot be arbitrarily reordered without changing the semantics of a program, in \mk the subgoals of conjunction/disjunction can be freely switched.
+Unlike Prolog, where atoms in the right-hand side of a clause cannot be arbitrarily reordered without changing the semantics of a program, in \mk the subgoals of conjunction/disjunction can be freely switched.
 This opens a yet another possibility for optimization, not taken into account by approaches initially developed in the context of conventional logic programming.

### 2020-07-09 papers/aplas/text/nonconj.tex
compress
Brushed up

@@
 This approach draws inspiration from both conjunctive partial deduction and supercompilation.
-The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk to improve the performance of the input programs.
+The aim was to create a specialization algorithm which would be simpler than conjunctive partial deduction and use properties of \mk to improve the performance of the input programs.
 
@@
 % Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
-The unfolding itself is straightforward however it is not always clear what to unfold and when to \emph{stop} unfolding.
+The unfolding itself is straightforward; however it is not always clear what to unfold and when to \emph{stop} unfolding.
 Unfolding in the context of specialization of functional programming languages, as well as inlining in specialization of imperative languages, is usually considered to be safe from the residual program efficiency point of view.
@@
 Unfortunately, this is not the case for the specialization of a relational programming language.
-Unlike in functional and imperative languages, in logic and relational programming languages unfolding may easily affect the target program's efficiency~\cite{leuschel2002logic}.
+Unlike functional and imperative languages, in logic and relational programming languages unfolding may easily affect the target program's efficiency~\cite{leuschel2002logic}.
 Unfolding too much may create extra unifications, which is by itself a costly operation, or even introduce duplicated computations by propagating the results of unfolding onto neighbouring conjuncts.
@@
 This heuristics is aimed at selecting a relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
-An unsafe unfolding leads to an uncontrollable increase in the number of relation calls in a conjunction.
+An unsafe unfolding leads to an uncontrollable increase of the number of relation calls in a conjunction.
 It is best to first unfold those relation calls which can be fully computed up to substitutions.
@@
 
-\begin{figure}[h!]
+\begin{figure}[!t]
   \centering

### 2020-07-09 papers/aplas/text/related.tex
Brushed up

@@
 Specialization is an attractive technique aimed to improve the performance of a program if some of its arguments are known statically.
-Specialization is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation} and many more.
+Specialization is studied for functional, imperative, and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation}, and many others.
 
@@
 Of course, process tree can contain infinite branches.
-\emph{Whistles} --- heuristics to identify possibly infinite branches --- are used to ensure supercompilation terminates.
+\emph{Whistles} --- heuristics to identify possibly infinite branches --- are used to ensure supercompilation termination.
 If a whistle signalls during the construction of some branch, then something should be done to ensure termination.
@@
 There are several ways to ensure process correctness and termination, most-specific generalization
-(anti-unification) and using \emph{homeomorphic embedding}~\cite{Higman52,Kruskal60} as a
+(anti-unification) and \emph{homeomorphic embedding}~\cite{Higman52,Kruskal60} as a
 whistle being the most common.
@@
 It is known that supercompilation may adversely affect GHC optimizations yielding standalone compilation more powerful~\cite{SCBE,TCES} and cause code explosion~\cite{SCHC}.
-Moreover, it may be hard to predict the real speedup of any given program on concrete examples even disregarding the problems above because of the complexity of the transformation algorithm.
+Moreover, it may be hard to predict the real speedup of any given program using concrete benchmarks even disregarding the problems above because of the complexity of the transformation algorithm.
 The worst-case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization indeed improves its performance~\cite{jonesbook,bulyonkov84}.
@@
 
-An empitical study shown that the most well-behaved strategy of local control in CPD for \pro is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
+An empitical study has shown that the most well-behaved strategy of local control in CPD for \pro is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
 An atom is unfolded only if precisely one suitable clause head exists for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
-This means that if a non-deterministic atom is the leftmost within a conjunction, it is most likely to be unfolded and to introduce many new relation calls within the conjunction.
+This means that if a non-deterministic atom is the leftmost one within a conjunction, it is most likely to be unfolded, introducing many new relation calls within the conjunction.
 We believe this is the core problem of CPD which limits its power when applied to \mk.
@@
 But it often leads to larger global control trees and, as a result, bigger, less efficient programs.
-The evaluation result of a \mk program does not depend on the order of atoms (relation calls) within a conjunction, thus we believe a better result can be achieved by selecting a relation call which can restrict the number of branches in the tree.
-We describe our approach which implements this idea in the next section.
+On the contrary, the results of evaluation of a \mk program do not depend on the order of atoms (relation calls) within conjunctions, thus we believe a better result can be achieved by selecting a relation call which can restrict the number of branches in the tree.
+We describe our approach, which implements this idea, in the next section.

### 2020-07-10 papers/aplas/text/nonconj.tex
Final

@@
 
-In this approach, we decided not to generalize in the same fashion as CPD or supercompilation.
-Our conjunctions are always split into individual calls and are joined back together only if it is sensible.
+In this approach, we do not generalize in the same fashion as CPD or supercompilation.
+Our conjunctions are always split into individual calls and are joined back together only if it is meaningful.
 If the need for generalization arises, i.e. homeomorphic embedding of conjunctions~\cite{de1999conjunctive} is detected, then we immediately stop driving this conjunction (line 12).

### 2020-07-17 papers/miniKanren/benchmarks/path/build.sh
Updated benchmarks

@@ -1,1 +1,1 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml conspd.ml cpd.ml branch.ml run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml conspd.ml cpd.ml branch.ml cpd_new.ml branch_new.ml run.ml

### 2020-07-17 papers/miniKanren/benchmarks/plain/build.sh

@@ -1,1 +1,1 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml conspd.ml cpd.ml branch.ml conspd_elem.ml cpd_old.ml ecce_old.ml conspd_inline.ml conspd_fresh.ml run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml conspd.ml cpd.ml branch.ml cpd_new.ml branch_new.ml run.ml

### 2020-07-17 papers/miniKanren/benchmarks/unify/build.sh

@@ -1,2 +1,2 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml conspd.ml cpd.ml branch.ml run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml conspd.ml cpd.ml branch.ml cpd_new.ml branch_new.ml run.ml
 

### 2020-07-17 papers/miniKanren/text/abstract.tex
Camera-ready miniKanren'20

@@
 \begin{abstract}
-  We explore partial deduction for \mk{}: a specialization technique aimed at improving the performance of a relation in the given direction.
+  We explore partial deduction, an advanced specialization technique aimed at improving the performance of a relation in the given direction, in the context of \mk{}.
+  On several examples, we demonstrate issues which arise during partial deduction of relational programs.
   We describe a novel approach to specialization of \mk{} based on partial deduction and supercompilation.
-  On several examples, we demonstrate issues which arise during partial deduction.
+  Although the proposed approach does not give the best results in all cases, we view it as a stepping stone towards the efficient optimization of \mk{}.
+
 \end{abstract}

### 2020-07-17 papers/miniKanren/text/conclusion.tex
spell- and grammarly- checked
Camera-ready miniKanren'20

--- 2020-05-21 papers/miniKanren/text/conclusion.tex
@@
 
-We compared some of the specialization techniques for \mk{}.
+In this paper, we discussed some issues which arise in partial deduction of a relational programming language, \mk{}.
+We presented a novel approach to partial deduction which uses less-branching heuristic to select the most suitable relation call to unfold at each step of driving.
+We compared this approach to the earlier implementation of conjunctive partial deduction and the implementation of CPD equipped with the new less-branching heuristic.
 
-There seem to be no one good technique.
+The conservative partial deduction improved the execution time of all relations while the implementations of CPD degraded the performance of some of them.
+However, CPD equipped with the less-branching heuristic improved the execution time of one relation the most compared with the other specializers.
+We conclude that there is still no one good technique which definitely speeds up every relational program.
+More research is needed to develop models to predict performance of relations, these models can further be used in specialization.

### 2020-07-17 papers/miniKanren/text/evaluation.tex
Camera-ready miniKanren'20

@@
 
-In our study we compared the CPD adaptation for \mk{} and the new non-conjunctive partial deduction.
-We have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check wether it can improve the quality of the specialization.
+We implemented the new conservative partial deduction\footnote{The repository of the \mk{} specialization project: \url{https://github.com/kajigor/uKanren_transformations}.
+Access date: 17.07.2020.} and compared it with the CPD adaptation for \mk{} of~\cite{lozov2019relational}.
+We have also employed the branching heuristic instead of the deterministic unfolding in the CPD to check whether it can improve the quality of the specialization.
 
-We used the following 4 programs to test the specializers on.
+We used the following programs to test the specializers.
 \begin{itemize}
@@
 All examples of \mk{} relations in this paper are written in \oc{}\footnote{\oc{}: statically typed \mk{} embedding in \ocaml{}.
-The repository of the project: \url{https://github.com/JetBrains-Research/OCanren}} syntax.
-We specialize the \lstinline{eval$^o$} relation to synthesize formulas which evaluate to \lstinline{^true}\footnote{An arrow lifts ordinary values to the logic domain}.
+The repository of the project: \url{https://github.com/JetBrains-Research/OCanren}.
+Access date: 17.07.2020.} syntax.
+We specialize the \lstinline{eval$^o$} relation to synthesize formulas which evaluate to \lstinline{^true}\footnote{An arrow lifts ordinary values to the logic domain.}.
 To do so, we run the specializer for the goal with the last argument fixed to \lstinline{^true}, while the first two arguments remain free variables.
@@
 
-One possible implementation of the evaluator in the syntax of \oc{} is presented in listing~\ref{eval:last}.
+One possible implementation of the evaluator in the syntax of \oc{} is presented in Listing~\ref{eval:last}.
 Here the relation \lstinline{elem$^o$ subst v res} unifies \lstinline{res} with the value of the variable \lstinline{v} in the list \lstinline{subst}.
@@
 CPD fails to recognize this, and thus unfolds recursive calls of \lstinline{eval$^o$} applied to fresh variables.
-It leads to over-unfolding, big residual programs and less than optimal performance.
+It leads to over-unfolding, big residual programs and poor performance.
 
-The non-conjunctive partial deduction first unfolds those calls which are selected with the heuristic.
+The conservative partial deduction first unfolds those calls which are selected with the heuristic.
 Since exploring boolean operations makes more sense, they are unfolded before recursive calls of \lstinline{eval$^o$}.
-The way non-conjunctive partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in listing~\ref{eval:fst}.
-This program is easier for CPD to transform which demonstrates how unequal is the behavior of CPD for the similar programs.
+The way conservative partial deduction treats this program is the same as it treats the other implementation in which boolean operations are moved to the left, as shown in Listing~\ref{eval:fst}.
+This program is easier for CPD to transform which demonstrates how unequal the behaviour of CPD for similar programs is.
 
@@
 
-Depending on the way a relation is implemented, it may take different number of driving steps to reach the point when any useful information is derived through its unfolding.
-Partial deduction tries to unfold every relation call unless it is unsafe, but not all relation calls serve to restrict the search space and thus should be unfolded.
+Depending on the way a relation is implemented, it may take a different number of driving steps to reach the point when any useful information is derived through its unfolding.
+Partial deduction tries to unfold every relation call unless it is unsafe, but not all relation calls serve to restrict the search space and thus not every relation call should be unfolded.
 In the implementation of \lstinline{eval$^o$} boolean operations can effectively restrict variables within the conjunctions and should be unfolded until they do.
-But depending on the way they are implemented, different number of driving steps should be performed for that.
-The simplest way to implement these relations is with a table as demonstrated with the implementation of \lstinline{not$^o$} in listing~\ref{not:table}.
+But depending on the way boolean operations are implemented, different number of driving steps should be performed for that.
+The simplest way to implement these relations is with a table as demonstrated with the implementation of \lstinline{not$^o$} in Listing~\ref{not:table}.
 It is enough to unfold such relation calls once to derive useful information about variables.
@@
 
-The other way to implement boolean operations is via one basic boolean relation such as \lstinline{nand$^o$} which is in turn has a table-based implementation (see listing~\ref{not:nando}).
+The other way to implement boolean operations is via one basic boolean relation such as \lstinline{nand$^o$} which has, in turn, a table-based implementation (see Listing~\ref{not:nando}).
 It will take several sequential unfoldings to derive that variables \lstinline{v} and \lstinline{w} should be \lstinline{^true} when considering a call \lstinline{and$^o$ v w ^true} implemented via a basic relation.
-Non-conjunctive partial deduction drives the selected call until it derives useful substitutions for the variables involved while CPD with deterministic unfolding may fail to do so.
+Conservative partial deduction drives the selected call until it derives useful substitutions for the variables involved.
+CPD with deterministic unfolding may fail to derive useful substitutions.
 
@@
   \begin{minipage}{0.7\textwidth}
-    \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operation via \lstinline{nand}}, captionpos=b, frame=tb]
+    \begin{lstlisting}[label={not:nando}, caption={Implementation of boolean operations via \lstinline{nand}}, captionpos=b, frame=tb]
   let not$^o$ x y = nand$^o$ x x y
@@
 
-In our study we considered two implementations of \lstinline{eval$^o$}, one we call \lstinline{plain} and the other --- \lstinline{last}, and compared how specializers behave on them.
+In our study, we considered two implementations of \lstinline{eval$^o$}, one we call \lstinline{plain} and the other --- \lstinline{last}, and compared how specializers behave on them.
 The \lstinline{plain} relation uses table-based boolean operations and places them further to the left in each conjunction.
-The relation \lstinline{last} employs boolean operations implemented via \lstinline{nand$^o$} and place them at the end of each conjunction.
+The relation \lstinline{last} employs boolean operations implemented via \lstinline{nand$^o$} and places them at the end of each conjunction.
 These two programs are complete opposites from the standpoint of CPD.
 
-We measured time necessary to generate $10000$ formulas over two variables which evaluate to \lstinline{^true}.
-We compared the results of specialization of the goal \lstinline{eval$^o$ subst fm ^true} by our implementation of CPD, the new non-conjunctive partial deduction, and the CPD modified with the branching heuristic.
-Our evaluation confirmed that CPD behaves very differently on these two implementations of the same relation: the running time of the specialized \lstinline{last} relation is almost 6 times as high as the running time of the specialized \lstinline{plain} relation.
-The running time of two programs generated with our novel non-conjunctive partial deduction is very close and it is a little bit better than the best by CPD.
-CPD with the branching heuristic still constructs residual programs of different quality.
+We measured the time necessary to generate $1000$ formulas over two variables which evaluate to \lstinline{^true}.
+We compared the results of specialization of the goal \lstinline{eval$^o$ subst fm ^true} by our implementation of CPD, the new conservative partial deduction, and the CPD modified with the less-branching heuristic.
+Our evaluation confirmed that CPD behaves very differently on these two implementations of the same relation.
+CPD improves the execution time of the \lstinline{plain} relation, however CPD performs too much unfolding of the  \lstinline{last} relation which is why the specialized relation \lstinline{last} fails to terminate in under 10 seconds.
+The execution time of two programs generated with the novel conservative partial deduction is very similar and it is a little bit better than the best by CPD.
+CPD with the less-branching heuristic constructs residual programs of different quality, worsening the execution time for both implementations.
 The results are shown in table~\ref{tbl:eval}.
 
-Besides evaluator of logic formulas we also run the transformers on the relation \lstinline{unify} which searches for a unifier of two terms and a relation \lstinline{isPath} specialized to search for paths in the graph.
-These two relations are described in paper~\cite{lozov2019relational} so we will not go into too much details here.
+Besides the evaluator of logic formulas we also run the transformers on the relation \lstinline{unify}, which searches for a unifier of two terms, and the relation \lstinline{isPath} specialized to search for paths in a graph.
+These two relations are described in paper~\cite{lozov2019relational} so we will not go into too many details here.
 
-The \lstinline{unify} relation was executed to find a unifier of two terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
-The original \mk{} program fails to terminate on this goal in 300 seconds.
-On this example the most performant is the program generated by CPD (2.35 seconds) while the program generated by adding branching heuristic also fails to terminate in 300 seconds.
-The non-conjunctive partial deduction shows some improvement with the residual program terminated within 15 seconds.
-While driving this program, non-conjunctive partial deduction does too much unfolding which negatively impact the running time as compared to CPD.
+The \lstinline{unify} relation was executed to find a unifier of the terms $f(X, X, g(Z, t))$ and $f(g(p, L), Y, Y)$.
+The original \mk{} program fails to terminate on this goal in 30 seconds.
+On this example, the most performant is the program generated by conservative partial deduction (0.96 seconds).
 
-The last test executed \lstinline{isPath} relation to search for 5 paths in the graph with 20 vertices and 30 edges.
-On this program non-conjunctive partial deduction showed better transformation results than CPD, although the difference is not that drastic.
+The last test executed the \lstinline{isPath} relation to search for 5 paths in a graph with 20 vertices and 30 edges.
+The original \mk{} program fails to terminate on this goal in 30 seconds.
+On this program, CPD with branching heuristic showed much better transformation result than both CPD and conservative partial deduction, although all specialized versions show improvement as compared with the original relation.
 
 All evaluation results are presented in the table~\ref{tbl:eval}.
-Each column correspond to the relation being run as described above.
-The row marked ``Original'' contains the running time of the original \mk{} relation before specialization, ``CPD'' and ``Non-CPD'' correspond to conjunctive and non-conjunctive partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
+Each column corresponds to the relation being run as described above.
+The row marked ``Original'' contains the execution time of the original \mk{} relation before specialization, ``CPD'' and ``ConsPD'' correspond to conjunctive and conservative partial deduction respectively while ``Branching'' is for the CPD modified with the branching heuristic.
 
@@
   \begin{tabular}{c||c|c||c||c}
-                   & last & plain & unify & isPath \\
-  \hline\hline
-  Original         & >60.00s & >60.00s & >300.00s & 19.86s \\
-  \hline
-  CPD              & 31.31s & 5.46s & 2.35s & 4.66s \\
-  \hline
-  Non-CPD          & 4.99s  & 5.05s & 14.90s & 3.00s \\
-  \hline
-  Branching        & 17.21s  & 6.17s & >300.00s & N/A \\
-  \hline
+                   & last  & plain & unify  & isPath \\ \hline\hline
+  Original         & 1.06s & 1.84s & ---    & ---    \\ \hline
+  CPD              & ---   & 1.13s & 14.12s & 3.62s  \\ \hline
+  ConsPD           & 0.93s & 0.99s & 0.96s  & 2.51s  \\ \hline
+  Branching        & 3.11s & 7.53s & 3.53s  & 0.54s  \\ \hline
   \end{tabular}
@@
 \end{table}
+
+% \begin{table}
+%   \centering
+%   \begin{tabular}{c||c|c||c||c}
+%                    & last   & plain  & unify    & isPath  \\ \hline\hline
+%   Original         & 1.29s & 1.61s & >30.00s & 25.84s \\ \hline
+%   CPD              & 1.77s & 0.17s & 0.34s   & 3.54s  \\ \hline
+%   Non-CPD          & 1.07s & 0.98s & 0.41s   & 1.68s  \\ \hline
+%   Branching        & 0.51s & 0.18s & >30.00s & N/A     \\ \hline
+%   \end{tabular}
+
+%   \caption{Evaluation results}
+%   \label{tbl:eval}
+% \end{table}
+
+% \begin{table}
+%   \centering
+%   \begin{tabular}{c||c|c||c||c}
+%                    & last & plain & unify & isPath \\
+%   \hline\hline
+%   Original         & >60.00s & >60.00s & >300.00s & 19.86s \\
+%   \hline
+%   CPD              & 31.31s & 5.46s & 2.35s & 4.66s \\
+%   \hline
+%   Non-CPD          & 4.99s  & 5.05s & 14.90s & 3.00s \\
+%   \hline
+%   Branching        & 17.21s  & 6.17s & >300.00s & N/A \\
+%   \hline
+%   \end{tabular}
+
+%   \caption{Evaluation results}
+%   \label{tbl:eval}
+% \end{table}
 
@@
 % The details on this benchmark can be found in the earlier paper CITE.
-% This example demonstrates how too much unfolding can be introduced with the non-conjunctive partial deduction.
+% This example demonstrates how too much unfolding can be introduced with the conservative partial deduction.
 

### 2020-07-17 papers/miniKanren/text/intro.tex
Related works
Camera-ready miniKanren'20

@@
 
-The core feature of the family of relational programming languages \mk{}\footnote{\mk{} language web site: \url{http://minikanren.org}} is their ability to run a program in different directions.
-Having specified a relation for adding two numbers, one can also compute the substraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
+The core feature of the family of relational programming languages \mk{}\footnote{\mk{} language web site: \url{http://minikanren.org}.
+Access date: 17.07.2020.} is their ability to run a program in different directions.
+Having specified a relation for adding two numbers, one can also compute the subtraction of two numbers or find all pairs of numbers which can be summed up to get the given one.
 Program synthesis can be done by running \emph{backwards} a relational interpreter for some language.
@@
 The ideas described there are aimed at left-to-right evaluation strategy of \pro{}.
-Since the search in \mk{} is complete, it is safe to reorder come relation calls within the goal for better performance.
+Since the search in \mk{} is complete, it is safe to reorder some relation calls within the goal ahead-of-time for better performance.
 While sometimes conjunctive partial deduction gives great performance boost, sometimes it does not behave as well as it could have.
 
-In this paper we show on examples some issues which conjunctive partial deduction faces.
-We also describe a novel approach to partial deduction of a relational programming language \mk{}.
+In this paper, we show on examples some issues which conjunctive partial deduction faces.
+We also describe \emph{conservative partial deduction} --- a novel specialization approach for the relational programming language \mk{}.
 We compare it to the existing specialization algorithms on several programs and discuss why some \mk{} programs run slower after specialization.

### 2020-07-17 papers/miniKanren/text/related.tex
Camera-ready miniKanren'20

--- 2020-05-23 papers/miniKanren/text/related.tex
@@
 
-Specialization is an attractive technique aimed to improve the performance of a program if some of its arguments are known statically.
-It is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation} and many more.
+Specialization is an attractive technique aimed to improve the~performance of a~ program if some of its arguments are known statically.
+Specialization is studied for functional, imperative and logic programing and comes in different forms: partial evaluation~\cite{jonesbook} and partial deduction~\cite{lloyd1991partial}, supercompilation~\cite{soerensen1996positive}, distillation~\cite{hamilton2007distillation}, and many more.
 
-The heart of supercompilation-based techniques is \emph{driving}~---~a symbolic execution of a program through all possible execution paths.
-The result of driving is so-called \emph{process tree} where nodes correspond to \emph{configurations} which present computation state, for example, a term in case of pure functional programming languages.
-Each path in the tree corresponds to some concrete program execution.
-The two main sources for supercompilation optimizations are aggressive information propagation about variable values, equalities, and disequalities and precomputing of all deterministic semantic evaluation steps, i.e.
-combining of consecutive process tree nodes with no branching, also known as \emph{deforestation}~\cite{TODO}.
-When the tree is constructed, the resulting, or \emph{residual}, program can be extracted from the process tree by the process called \emph{residualization}.
+The heart of supercompilation-based techniques is \emph{driving}~---~a~ symbolic execution of a~ program through all possible execution paths.
+The result of driving is a~ \emph{process tree} where nodes correspond to \emph{configurations} which represent computation state.
+For example, in the~case of pure functional programming languages, the~computational state might be a~ term.
+Each path in the~tree corresponds to some concrete program execution.
+The two main sources for supercompilation optimizations are aggressive information propagation about variables' values, equalities and disequalities, and precomputing of all deterministic semantic evaluation steps.
+The latter process, also known as \emph{deforestation}, means combining of consecutive process tree nodes with no branching.
+When the~tree is constructed, the~resulting, or \emph{residual}, program can be extracted from the~process tree by the~process called \emph{residualization}.
 Of course, process tree can contain infinite branches.
-\emph{Whistles} --- heuristics to identify possibly infinite branches --- are used to ensure termination in supercompilation.
-If the whistle signalled during the construction of some branch, then something should be done to ensure termination.
-The most common approaches include either stopping driving the infinite branch completely (no specialization is done in this case and the source code is blindly copied into the residual program) or folding the process tree to fold in a \emph{process graph}.
-The main instrument to perform such a folding is \emph{generalization}, i.e. abstracting away some computed data about the current term which makes folding possible.
-One source of infinite branches is consecutive recursive calls to the same function with an accumulating parameter: by unfolding such a call further one can only increase the term size which leads to nontermination.
-The accumulating parameter can be removed by replacing the call with its generalization.
-There are several ways to ensure process correctness and termination, the most common being \emph{homeomorphic embedding}~\cite{TODO} used as a whistle and most-specific generalization of terms.
+\emph{Whistles} --- heuristics to identify possibly infinite branches --- are used to ensure supercompilation termination.
+If a~ whistle signals during the~construction of some branch, then something should be done to ensure termination.
+The most common approaches are either to stop driving the~infinite branch completely (no specialization is done in this case and the~source code is blindly copied into the~residual program) or to fold the~process tree to a~ \emph{process graph}.
+The main instrument to perform such a~ folding is \emph{generalization}.
+Generalization, abstracting away some computed data about the~current term, makes folding possible.
+One source of infinite branches is consecutive recursive calls to the~same function with an accumulating parameter: by unfolding such a~ call further one can only increase the~term size which leads to nontermination.
+The accumulating parameter can be removed by replacing the~call with its generalization.
+There are several ways to ensure process correctness and termination,
+most-specific generalization (anti-unification) and \emph{homeomorphic embedding} ~\cite{Higman52,Kruskal60} as a~ whistle being the~most common.
 % When two dangerously similar nodes are encountered,
-% For example, two consecutive recursive calls of a function with accumulating parameter, and the first is not an instance of the second, then one may construct a new, generalized, node such that both of original nodes are instances of the last one, then one of the initial nodes is replaced with the generalized one.
+% For example, two consecutive recursive calls of a~ function with accumulating parameter, and the~first is not an instance of the~second, then one may construct a~ new, generalized, node such that both of original nodes are instances of the~last one, then one of the~initial nodes is replaced with the~generalized one.
 
-While supercompilation generally improves the behavior of input programs and distillation can even provide superlinear speedup, there are no ways to predict the effect of specialization on a given program in general case.
-What is worse, they rarely consider the residual program efficiency from the point of view of the target language evaluator.
-The main optimization source is computing in advance all possible intermediate and statically-known semantics steps at transformation-time.
-Other criteria like the size of the generated program or possible optimizations and execution cost of different language constructions by the target language evaluator are usually out of consideration~\cite{jonesbook}.
+While supercompilation generally improves the~behaviour of input programs and distillation can even provide superlinear speedup, there are no ways to predict the~effect of specialization on a~ given program in the~general.
+What is worse, the~efficiency of residual program from the~target language evaluator point of view is rarely considered in the~literature.
+The main optimization source is computing in advance all possible intermediate and statically-known semantics steps at program transformation-time.
+Other criteria, like the~size of the~generated program or possible optimizations and execution cost of different language constructions by the~target language evaluator, are usually out of consideration~\cite{jonesbook}.
 It is known that supercompilation may adversely affect GHC optimizations yielding standalone compilation more powerful~\cite{SCBE,TCES} and cause code explosion~\cite{SCHC}.
-Moreover, it may be hard to predict the real speedup of any given program on concrete examples even disregarding problems above because of the complexity of the transformation algorithm.
-The worst case for partial evaluation is when all static variables are used in a dynamic context, and there is some advice on how to implement a partial evaluator as well as a target program so that specialization really improved its performance~\cite{jonesbook,bulyonkov84}.
-There is lack of research in determining the classes of programs which transformers would definitely speed up.
+Moreover, it may be hard to predict the~real speedup of any given program on concrete examples even disregarding the~problems above because of the~complexity of the~transformation algorithm.
+The worst-case for partial evaluation is when all static variables are used in a~ dynamic context, and there is some advice on how to implement a~ partial evaluator as well as a~ target program so that specialization indeed improves its performance~\cite{jonesbook,bulyonkov84}.
+There is a~ lack of research in determining the~classes of programs which transformers would definitely speed up.
 
-Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for left-to-right evaluation strategy of \pro{}.
-CPD constructs a tree which models goal evaluation and is similar to SLD-NF tree, then a residual program is generated from this tree.
+Conjunctive partial deduction~\cite{de1999conjunctive} makes an effort to provide reasonable control for the~left-to-right evaluation strategy of \pro{}.
+CPD constructs a~ tree which models goal evaluation and is similar to a~ SLDNF tree, then a~ residual program is generated from this tree.
 Partial deduction itself resembles driving in supercompilation~\cite{gluck1994partial}.
-The specialization is done in two levels of control: the local control determines the shape of the residual programs, while the global control ensures that every relation which can be called in the residual program is indeed defined.
-The leaves of local control trees become nodes of the global control tree.
-CPD analyses these nodes at the global level and runs local control for all those which are new.
+The specialization is done in two levels of control: the~local control determines the~shape of the~residual programs, while the~global control ensures that every relation which can be called in the~residual program is indeed defined.
+The leaves of local control trees become nodes of the~global control tree.
+CPD analyses these nodes at the~global level and runs local control for all those which are new.
 
-At the local level CPD examines a conjunction of atoms by considering each atom one-by-one from left to right.
-An atom is \emph{unfolded} if it is deemed safe, i.e. a whistle based on homeomorphic does not signal for the atom.
-When an atom is unfolded, a clause whose head can be unified with the atom is found, and a new node is added into the tree where the atom in the conjunction is replaced with the body of that clause.
-If there are more than one suitable head, then several branches are added into the tree which correspond to the disjunction in the residualized program.
-An adaptation of CPD for the \mk{} programming language is described in~\cite{lozov2019relational}.
+At the~local level, CPD examines a~ conjunction of atoms by considering each atom one-by-one from left to right.
+An atom is \emph{unfolded} if it is deemed safe, i.e. a~ whistle based on homeomorphic embedding does not signal for the~atom.
+When an atom is unfolded, a~ clause whose head can be unified with the~atom is found, and a~ new node is added into the~tree where the~atom in the~conjunction is replaced with the~body of that clause.
+If there is more than one suitable head, then several branches are added into the~tree which corresponds to the~disjunction in the~residualized program.
+An adaptation of CPD for the~\mk{} programming language is described in~\cite{lozov2019relational}.
 
 The most well-behaved strategy of local control in CPD for \pro{} is \emph{deterministic unfolding}~\cite{leuschel1997advanced}.
-An atom is unfolded only if only one suitable clause head exist for it with the one exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
-This means that if a non-deterministic atom is the leftmost within a conjunction, it is most likely to be unfoldled and introduce many new relation calls within the conjunction.
-We believe this is the core problem with CPD which limits its power when applied to \mk{}.
-The strategy of unfolding atoms from left to right is reasonable in the context of \pro{} because it mimics the way programs in \pro{} execute.
-But it often leads to larger global control trees and, as a result, bigger, less efficient programs.
-The evaluation result of a \mk{} program does not depend on the order of atoms (relation calls) within a conjunction, thus we believe a better result can be achieved by selecting a relation call which can restrict the number of branches in the tree.
-We describe our approach which implements this idea in the next section.
+An atom is unfolded only if precisely one suitable clause head exists for it with the~single exception: it is allowed to unfold an atom non-deterministically once for one local control tree.
+This means that if a~ non-deterministic atom is the~leftmost within a~ conjunction, it is most likely to be unfolded and to introduce many new relation calls within the~conjunction.
+We believe this is the~core problem with CPD which limits its power when applied to \mk{}.
+The strategy of unfolding atoms from left to right is reasonable in the~context of \pro{} because it mimics the~way programs in \pro{} execute.
+But it often leads to larger global control trees and, as a~ result, bigger, less efficient programs.
+The evaluation result of a~ \mk{} program does not depend on the~order of atoms (relation calls) within a~ conjunction, thus we believe a~ better result can be achieved by selecting a~ relation call which can restrict the~number of branches in the~tree.
+We describe our approach which implements this idea in the~next section.

### 2020-07-17 app/Main.hs
Fixed CPD residualization

@@ -1,6 +1,7 @@
 module Main where
 
 import           CPD.LocalControl       (Heuristic (..))
+import           Debug.Trace            (traceM)
 import           NonConjunctive.Unfold
 import qualified Program.Bottles
 import qualified Program.Bridge
@@ -139,19 +140,32 @@ runPropEval = do
 
 main :: IO ()
 main = do
+    Transformer.CPD.transform "branchPlainFirst" Program.PropEval.plainFirstQuery Nothing Branching
+    Transformer.CPD.transform "cpdPlainFirst" Program.PropEval.plainFirstQuery Nothing Deterministic
+    Transformer.CPD.transform "branchNandoLast" Program.PropEval.nandoLastQuery Nothing Branching
+    Transformer.CPD.transform "cpdNandoLast" Program.PropEval.nandoLastQuery Nothing Deterministic
+
+    Transformer.CPD.transform "branchUnify" Program.Unify.query Nothing Branching
+    Transformer.CPD.transform "cpdUnify" Program.Unify.query Nothing Deterministic
+
+    Transformer.CPD.transform "cpdPath" Program.Path.query1 Nothing Deterministic
+    traceM "\n==============================\nstarted transforming path:  branching\n\n"
+    Transformer.CPD.transform "branchPath" Program.Path.query1 Nothing Branching
+
+    runPath
+    runUnify
+
     -- runL'
 
     -- -- runPropEval
     -- -- runDoubleApp
     -- -- runMaxlen
     -- -- runL
-    -- -- runPath
-    -- -- runUnify
 
     -- Transformer.PrologToMk.transform  "test/out/nc/da/ecce.pl"
     -- Transformer.PrologToMk.transform  "test/out/nc/maxlen/ecce.pl"
     -- Transformer.PrologToMk.transform  "test/out/nc/llang1/ecce.pl"
-    Transformer.PrologToMk.transform  "test/out/nc/llangPeter/ecce.pl"
+    -- Transformer.PrologToMk.transform  "test/out/nc/llangPeter/ecce.pl"
     -- Transformer.PrologToMk.transform  "test/out/nc/path/ecce.pl"
     -- Transformer.PrologToMk.transform  "test/out/nc/unify/ecce.pl"
     -- Transformer.PrologToMk.transform  "test/out/nc/propFirstPlain/ecce.pl"

### 2020-07-17 src/CPD/Residualization.hs

@@ -142,7 +142,7 @@ residualizeSldTree rootGoals tree definitions = do
                     resultants
   let defArgs = map Res.vident rootVars
   let body = Eval.postEval defArgs $
-                foldl1 (|||) (reverse goals)
+                unsafeDisj (reverse goals)
 
   if null goals
   then fail (printf "No resultants in the sld tree for %s" (show rootGoals))

### 2020-07-17 src/Syntax.hs

@@ -96,6 +96,9 @@ conj xs = Just $ foldr1 (:/\:) xs
 unsafeConj :: [G a] -> G a
 unsafeConj = foldr1 (:/\:)
 
+unsafeDisj :: [G a] -> G a
+unsafeDisj = foldr1 (:\/:)
+
 successName :: String
 successName = "success"
 

### 2020-07-22 papers/miniKanren/text/conspd.tex
Rename nonconjunctive partial deduction to conservative partial deduction

--- 2020-05-23 papers/miniKanren/text/nonconj.tex
@@
 
-\section{Non-conjunctive Partial Deduction}
+\section{Conservative Partial Deduction}
 
-In this section, we describe a novel approach to relational programs specialization.
+In this section, we describe a~novel approach to relational programs specialization.
 This approach draws inspiration from both conjunctive partial deduction and supercompilation.
-The aim was to create a specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve the performance of the input programs.
+The aim was to create a~specialization algorithm which is simpler than conjunctive partial deduction and uses properties of \mk{} to improve the performance of the input programs.
 
@@
 For the sake of brevity and clarity, we provide functions \code{drive\_disj} and \code{drive\_conj} which describe how to process disjunctions and conjunctions respectively.
-Driving itself is a trivial combination of presented functions (line 2).
+Driving itself is a~trivial combination of the functions provided (line 2).
 
-A driving process creates a process tree, from which a residual program is later created.
+A driving process creates a~process tree, from which a~residual program is later created.
 The process tree is meant to mimic the execution of the input program.
-The nodes of the process tree include a \emph{configuration} which describes the state of program evaluation at some point.
-In our case configuration is a conjunction of relation calls.
+The nodes of the process tree include a~\emph{configuration} which describes the state of program evaluation at some point.
+In our case a~configuration is a~conjunction of relation calls.
 The substitution computed at each step is also stored in the tree node, although it is not included in the configuration.
 
-Hereafter, we consider all goals and relation bodies to be in \emph{canonical normal form}~---~a disjunction of conjunctions of either calls or unifications.
+Hereafter, we consider all goals and relation bodies to be in \emph{canonical normal form}~---~a~disjunction of conjunctions of either calls or unifications.
 Moreover, we assume all fresh variables to be introduced into the scope and all unifications to be computed at each step.
 Those disjuncts in which unifications fail are removed.
-Other disjuncts take the form of possibly empty conjunction of relation calls accompanied with a substitution computed from unifications.
+Each other disjunct takes the form of a~possibly empty conjunction of relation calls accompanied with a~substitution computed from unifications.
 Any \mk{} term can be trivially transformed into the described form.
-In Fig.~\ref{fig:ncpd-pseudo} function \code{normalize} is assumed to perform term normalization.
+In Fig.~\ref{fig:ncpd-pseudo} the function \code{normalize} is assumed to perform term normalization.
 The code is omitted for brevity.
@@
   \includegraphics[width=0.85\textwidth]{algo_pseudo-crop.pdf}
-  \caption{\db{Non-conjunctive Partial Deduction Pseudo Code}}
+  \caption{Conservative Partial Deduction Pseudo Code}
   \label{fig:ncpd-pseudo}
@@
 
-% The very first step of driving a conjunction is to apply a substitution to the variables in relation calls (line 10).
+% The very first step of driving a~conjunction is to apply a~substitution to the variables in relation calls (line 10).
 
@@
 The first is to select an arbitrary relation to unfold, not necessarily the leftmost which is safe.
-The second idea is to use a heuristic which decides if unfolding a relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time (line 14; \code{heuristically\_select\_a\_call} stands for heuristics combination, see section~\ref{sec:heurictic} for details).
-If those contradictions are found, then they are exposed by considering the conjunction as a whole and replacing the selected relation call with the result of its unfolding thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD (lines 15--22).
-Joining instead of splitting is why we call our transformer \emph{non-conjunctive} partial deduction.
-Finally, if the heuristic fails to select a potentially good call, then the conjunction is split into individual calls which are driven in isolation and are never joined (line 23).
+The second idea is to use a~heuristic which decides if unfolding a~relation call can lead to discovery of contradictions between conjuncts which in turn leads to restriction of the answer set at specialization-time (line 14; \code{heuristically\_select\_a\_call} stands for heuristics combination, see section~\ref{sec:heurictic} for details).
+If those contradictions are found, then they are exposed by considering the conjunction as a~whole and replacing the selected relation call with the result of its unfolding thus \emph{joining} the conjunction back together instead of using \emph{split} as in CPD (lines 15--22).
+Joining instead of splitting is why we call our transformer \emph{conservative} partial deduction.
+Finally, if the heuristic fails to select a~potentially good call, then the conjunction is split into individual calls which are driven in isolation and are never joined (line 23).
 
-When the heuristic selects a call to unfold (line 15), a process tree is constructed for the selected call \emph{in isolation} (line 16).
+When the heuristic selects a~call to unfold (line 15), a~process tree is constructed for the selected call \emph{in isolation} (line 16).
 The leaves of the computed tree are examined.
@@
 
-% Each time we examine a conjunction of calls, we \emph{split} them into separate nodes which are driven independently from each other.
+% Each time we examine a~conjunction of calls, we \emph{split} them into separate nodes which are driven independently from each other.
 % Among the relation calls we select one which is according to the heuristic is likely to narrow down the answer set \db{(line 15)}.
@@
 
-This process creates branchings whenever a disjunction is examined (lines 4--5).
-At each step, we make sure that we do not start driving conjunction which we have already examined.
-To do this, we check if the current conjunction is a renaming of any other configuration in the tree (line 11).
-If it is, then we fold the tree by creating a special node which then is residualized into a call to the corresponding relation.
+This process creates branchings whenever a~disjunction is examined (lines 4--5).
+At each step, we make sure that we do not start driving a~conjunction which we have already examined.
+To do this, we check if the current conjunction is a~renaming of any other configuration in the tree (line 11).
+If it is, then we fold the tree by creating a~special node which then is residualized into a~call to the corresponding relation.
 
-We decided not to perform generalization in this approach in the same fashion as CPD or supercompilation does.
-Our conjunctions are always split into individual calls and are joined back together only if it is sensible.
+In this approach, we decided not to generalize in the same fashion as CPD or supercompilation.
+Our conjunctions are always split into individual calls and are joined back together only if it meaningful.
 If the need for generalization arises, i.e. homeomorphic embedding of conjunctions~\cite{de1999conjunctive} is detected, then we immediately stop driving this conjunction (line 12).
-When residualizing such conjunction, we just generate a conjunction of calls to the input program before specialization.
+When residualizing such a~conjunction, we just generate a~conjunction of calls to the input program before specialization.
 
 % The generalization is used in supercompilation and partial deduction to ensure termination at the same time as some degree of specialization.
-% The generalization of two terms is usually a \emph{most-specific generalization}.
+% The generalization of two terms is usually a~\emph{most-specific generalization}.
 % Generalization is used to abstract away some information computed during driving.
@@
 % The generalization selects subconjuctions of two conjuncts which are similar (call to the same relation and their arguments have similar shape and distribution).
-% For the subconjunctions selected a most-specific generalization is computed.
+% For the subconjunctions selected a~most-specific generalization is computed.
 
-% In our approach we only do splitting of a conjunction into individual relation calls.
-% This makes any program with an accumulating parameter to be a problem.
-% Sometimes when there is a need to do a proper generalization, it is in reality just an instance of some other goal within the tree and we can simply create a call there \db{(line 13)}.
+% In our approach we only do splitting of a~conjunction into individual relation calls.
+% This makes any program with an accumulating parameter to be a~problem.
+% Sometimes when there is a~need to do a~proper generalization, it is in reality just an instance of some other goal within the tree and we can simply create a~call there \db{(line 13)}.
 % Otherwise we are unable to meaningfully specialize such goal, but we can always just include the initial program in the residual program and call the corresponding relation.
@@
 Unfolding in our case is done by substitution of some relation call by its body with simultaneous normalization and computation of unifications.
-% To unfold a relation call we do the following steps.
+% To unfold a~relation call we do the following steps.
 % \db{TODO: the following has alredy mentioned above.}
-% First, the formal arguments of a relation are substituted for the actual arguments of the call in the body.
+% First, the formal arguments of a~relation are substituted for the actual arguments of the call in the body.
 % All fresh variables get instantiated.
-% The body is transformed into a canonical form (disjunction of conjunctions of either calls or unifications).
+% The body is transformed into a~canonical form (disjunction of conjunctions of either calls or unifications).
 % All unifications are computed.
 % Those disjuncts in which unifications fails are removed.
-% Other disjuncts take form of a conjunction of relation calls accompanied with a substitution.
+% Other disjuncts take form of a~conjunction of relation calls accompanied with a~substitution.
 The unfolding itself is straightforward however it is not always clear what to unfold and when to \emph{stop} unfolding.
-Unfolding in functional programming languages specialization, as well as inlining in imperative one, is usually considered to be safe from the residual program efficiency point of view.
-It may only lead to code explosion or code duplication which is mostly left to a target program compiler optimization or even is out of consideration at all if a specializer is considered as a standalone tool~\cite{jonesbook}.
+Unfolding in the specialization of functional programming languages, as well as inlining in imperative languages, is usually considered to be safe from the residual program efficiency point of view.
+It may only lead to code explosion or code duplication which is mostly left to a~target program compiler optimization or even is out of consideration at all if a~specializer is considered as a~standalone tool~\cite{jonesbook}.
 
-Unfortunately, this is not the case for the specialization of relational programming language.
-Unlike functional and imperative, in logic and relational programming language unfolding may easily affect the target program efficiency~\cite{leuschel2002logic}.
-Unfolding too much may create extra unifications, which is by itself a costly operation, or even introduce duplicated computations by propagating the unfolding results onto neighbouring conjuncts.
+Unfortunately, this is not the case for the specialization of a~relational programming language.
+Unlike in functional and imperative languages, in logic and relational programming languages unfolding may easily affect the target program's efficiency~\cite{leuschel2002logic}.
+Unfolding too much may create extra unifications, which is by itself a~costly operation, or even introduce duplicated computations by propagating the unfolding's results onto neighbouring conjuncts.
 
-There is a fine edge between too much unfolding and not enough unfolding.
+There is a~fine edge between too much unfolding and not enough unfolding.
 The former is maybe even worse than the latter.
-We believe that the following heuristic provides a reasonable approach to unfolding control.
+We believe that the following heuristic provides a~reasonable approach to unfolding control.
 
-\subsection{Heuristic}
+\subsection{Less-Branching Heuristic}
 \label{sec:heurictic}
 
-This heuristic is aimed at selecting the relation call within a conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
-The unsafe unfolding leads to an uncontrollable increase in the number of relation calls in a conjunction.
+This heuristic is aimed at selecting a~relation call within a~conjunction which is both safe to unfold and may lead to discovering contradictions within the conjunction.
+An unsafe unfolding leads to an uncontrollable increase in the number of relation calls in a~conjunction.
 It is best to first unfold those relation calls which can be fully computed up to substitutions.
@@
 This means that we found some contradiction, some computations were gotten rid of, and thus the answer set was restricted, which is desirable when unfolding.
-To compute this heuristic we precompute the maximum possible number of disjuncts in each relation and compare this number with the number of disjuncts when unfolding a concrete relation call.
-The maximum number of disjuncts is computed by unfolding the body of the relation in which all relation calls were replaced by a unification which always succeeds.
+To compute this heuristic we precompute the maximum possible number of disjuncts in each relation and compare this number with the number of disjuncts when unfolding a~concrete relation call.
+The maximum number of disjuncts is computed by unfolding the body of the relation in which all relation calls were replaced by a~unification which always succeeds.
 
@@
 
-The pseudocode describing heuristic is shown in fig.~\ref{fig:heu-pseudo}.
-Selecting a good relation call can fail (line 1).
+The pseudocode describing our heuristic is shown in Fig.~\ref{fig:heu-pseudo}.
+Selecting a~good relation call can fail (line 1).
 The implementation works such that we first select those relation calls which are static, and only if there are none, we proceed to consider deterministic unfoldings and then we search for those which are less branching.
-We believe this heuristic provides a good balance in unfolding.
+We believe this heuristic provides a~good balance in unfolding.
 % The final heuristic selects the first conjunct which suites either of the following cases.
@@
 % Then those which are less branching.
-% The last to be unfolded are those calls, which unfold to a substitution with not conjunction.
+% The last to be unfolded are those calls, which unfold to a~substitution with not conjunction.
 
@@
 % Residualization is quite straightforward.
-% A branching in the process tree becomes a disjunction.
-% A split node becomes a conjunction.
-% Computed substitution is residualized as a conjunction of unifications.
-% A renaming node is just a call to a relation.
+% A branching in the process tree becomes a~disjunction.
+% A split node becomes a~conjunction.
+% Computed substitution is residualized as a~conjunction of unifications.
+% A renaming node is just a~call to a~relation.
 % Relations are created for configurations on which leaf nodes are renamed.
 
-% One other thing is that when some configuration is occurred within the tree which is an instance of a configuration for which a new relation is created, then we just create a call.
+% One other thing is that when some configuration is occurred within the tree which is an instance of a~configuration for which a~new relation is created, then we just create a~call.

### 2020-07-22 app/Main.hs

@@ -1,8 +1,8 @@
 module Main where
 
+import           ConsPD.Unfold
 import           CPD.LocalControl       (Heuristic (..))
 import           Debug.Trace            (traceM)
-import           NonConjunctive.Unfold
 import qualified Program.Bottles
 import qualified Program.Bridge
 import qualified Program.Bridge2
@@ -20,9 +20,9 @@ import           Program.Stlc           (evalo)
 import qualified Program.Typing
 import qualified Program.Unify
 import           Syntax
+import qualified Transformer.ConsPD
 import qualified Transformer.CPD
 import qualified Transformer.JustUnfold
-import qualified Transformer.NonConj
 import qualified Transformer.PrologToMk
 
 dA = Program doubleAppendo $ fresh ["x", "y", "z", "r"] (call "doubleAppendo" [V "x", V "y", V "z", V "r"])
@@ -33,7 +33,7 @@ lambda = Program evalo $ fresh ["m", "n"] (call "evalo" [V "m", V "n"])
 
 runJu = Transformer.JustUnfold.transform
 
-runNc = Transformer.NonConj.runNc
+runConsPD = Transformer.ConsPD.runConsPD
 
 runRep = do
     runJu 100 "rep" rep
@@ -41,20 +41,20 @@ runRep = do
     rep = Program Program.Programs.rep $ fresh ["n", "x"] (call "rep" [V "n", V "x"])
 
 runProp = do
-    -- runNc (-1) "prop"  prop
-    -- runNc (-1) "prop1" prop1
-    -- runNc (-1) "prop2" prop2
-    -- runNc (-1) "prop3" prop3
-    -- runNc (-1) "prop4" prop4
-    -- runNc (-1) "prop_" prop'
-    runNc (-1) "prop__" prop''
-    -- runNc (-1) "prop1___" prop1'''
-    -- runNc (-1) "prop2___" prop2'''
-    runNc (-1) "propPlain" propPlain
-    -- runNc (-1) "propPlain_" propPlain'
-    -- runNc (-1) "prop__1" prop''1
-    -- runNc (-1) "prop__2" prop''2
-    -- runNc (-1) "prop__3" prop''3
+    -- runConsPD (-1) "prop"  prop
+    -- runConsPD (-1) "prop1" prop1
+    -- runConsPD (-1) "prop2" prop2
+    -- runConsPD (-1) "prop3" prop3
+    -- runConsPD (-1) "prop4" prop4
[…]

### 2020-07-22 app/bench/Main.hs

@@ -1,4 +1,16 @@
 module Main where
 
+import           Program.Programs
+import           Syntax
+import qualified Transformer.ConsPD
+
+runConsPD = Transformer.ConsPD.runConsPD
+
+runDoubleAppendo =
+    runConsPD (-1) "doubleAppendo" prog
+  where
+    prog = Program doubleAppendo $ fresh ["x", "y", "z", "r"] (call "doubleAppendo" [V "x", V "y", V "z", V "r"])
+
 main :: IO ()
-main = return ()
+main =
+  runDoubleAppendo

### 2020-07-22 src/ConsPD/Residualization.hs

@@ -1,29 +1,28 @@
 {-# LANGUAGE TupleSections #-}
 
-module NonConjunctive.Residualization where
+module ConsPD.Residualization where
 
-import           Control.Applicative   ((<|>))
-import qualified CPD.LocalControl      as LC
-import qualified CPD.Residualization   as CpdR
-import           Data.List             (find, intercalate, nub, sort, union,
-                                        (\\))
-import           Data.Maybe            (catMaybes, fromJust, fromMaybe)
-import           Debug.Trace           (trace, traceM)
-import           Embed                 (isInst)
-import qualified Eval                  as E
-import           NonConjunctive.Unfold
-import qualified Residualize           as Res
+import           ConsPD.Unfold
+import           Control.Applicative ((<|>))
+import qualified CPD.LocalControl    as LC
+import qualified CPD.Residualization as CpdR
+import           Data.List           (find, intercalate, nub, sort, union, (\\))
+import           Data.Maybe          (catMaybes, fromJust, fromMaybe)
+import           Debug.Trace         (trace, traceM)
+import           Embed               (isInst)
+import qualified Eval                as E
+import qualified Residualize         as Res
 import           Syntax
-import           Text.Printf           (printf)
-import           Unfold                (normalize)
-import           Util.Miscellaneous    (fst3, show', snd3, trd3)
+import           Text.Printf         (printf)
+import           Unfold              (normalize)
+import           Util.Miscellaneous  (fst3, show', snd3, trd3)
 
 topLevel :: Program -> Program
 topLevel input =
-  residualize $ nonConjunctive (-1) input
+  residualize $ ConsPD.Unfold.topLevel (-1) input
 
 
-residualize :: (NCTree, G S, [S]) -> Program
+residualize :: (ConsPDTree, G S, [S]) -> Program
 residualize (Fail, goal, names) = Program [] (generateGoal goal names)
 residualize (tree, goal, names) =
   -- let restricted = restrictSubsts tree in
@@ -31,7 +30,7 @@ residualize (tree, goal, names) =
[…]

### 2020-07-22 src/ConsPD/Unfold.hs

@@ -1,6 +1,6 @@
 {-# LANGUAGE ScopedTypeVariables #-}
 
-module NonConjunctive.Unfold where
+module ConsPD.Unfold where
 
 import qualified CPD.LocalControl   as LC
 import           Data.Foldable      (foldlM)
@@ -20,18 +20,18 @@ import           Unfold             (findBestByComplexity, notMaximumBranches,
 import           Util.Check         (checkConj)
 import           Util.Miscellaneous (fst3, fst4, show', snd3)
 
-data NCTree = Fail
-            | Success E.Sigma E.Gamma
-            | Or [NCTree] (LC.Descend [G S]) E.Sigma
-            | Conj [NCTree] [G S] E.Sigma
-            | Gen NCTree [G S] [G S] Generalizer E.Sigma
-            | Leaf [G S] E.Sigma E.Gamma [G S] -- last argument is a goal renaming of which current node is
-            | Split [NCTree] [G S] E.Sigma
-            | Prune [G S] E.Sigma
-            -- deriving (Show, Eq)
+data ConsPDTree = Fail
+                | Success E.Sigma E.Gamma
+                | Or [ConsPDTree] (LC.Descend [G S]) E.Sigma
+                | Conj [ConsPDTree] [G S] E.Sigma
+                | Gen ConsPDTree [G S] [G S] Generalizer E.Sigma
+                | Leaf [G S] E.Sigma E.Gamma [G S] -- last argument is a goal renaming of which current node is
+                | Split [ConsPDTree] [G S] E.Sigma
+                | Prune [G S] E.Sigma
+                -- deriving (Show, Eq)
 
 
-instance Show NCTree where
+instance Show ConsPDTree where
   show Fail = "_|_"
   show (Success _ _) = "Success"
   show (Or ch (LC.Descend gs _) _) = printf "Or %s %s" (show gs) (show ch)
@@ -42,7 +42,7 @@ instance Show NCTree where
   show (Gen _ _ _ _ _) = "Gen"
 
 
-instance Eq NCTree where
+instance Eq ConsPDTree where
   Fail == Fail = True
   Success s _ == Success s' _ = s == s'
   Or ch ds s == Or ch' ds' s' = ch == ch' && ds == ds' && s == s'
[…]

### 2020-07-22 src/Printer/ConsPDTree.hs

@@ -1,25 +1,25 @@
 
-module Printer.NCTree where
+module Printer.ConsPDTree where
 
-import qualified CPD.LocalControl      as LC
-import           Data.List             ((\\))
-import           Debug.Trace           (trace)
-import qualified Eval                  as E
-import           NonConjunctive.Unfold (NCTree (..), restrictSubsts)
+import           ConsPD.Unfold    (ConsPDTree (..), restrictSubsts)
+import qualified CPD.LocalControl as LC
+import           Data.List        ((\\))
+import           Debug.Trace      (trace)
+import qualified Eval             as E
 import           Printer.Dot
 import           Syntax
 import           Text.Printf
 
-instance DotPrinter NCTree where
-  labelNode t@(Conj ch _ _)  = addChildren t ch
-  labelNode t@(Split ch _ _) = addChildren t ch
-  labelNode t@(Or ch _ _)    = addChildren t ch
-  labelNode t@(Gen ch _ _ _ _)   = addChild    t ch
-  labelNode t                = addLeaf     t
+instance DotPrinter ConsPDTree where
+  labelNode t@(Conj ch _ _)    = addChildren t ch
+  labelNode t@(Split ch _ _)   = addChildren t ch
+  labelNode t@(Or ch _ _)      = addChildren t ch
+  labelNode t@(Gen ch _ _ _ _) = addChild    t ch
+  labelNode t                  = addLeaf     t
 
   simplify = restrictSubsts
 
-instance Dot NCTree where
+instance Dot ConsPDTree where
   dot (Leaf gs s _ v) = printf "Leaf <BR/> %s <BR/> %s <BR/> %s" (dot gs) (E.dotSigma s) (dot v)
   dot Fail = "_|_"
   dot (Success s _) = printf "Success <BR/> %s" (E.dotSigma s)

### 2020-07-22 src/Transformer/ConsPD.hs

@@ -1,34 +1,34 @@
-module Transformer.NonConj where
+module Transformer.ConsPD where
 
 
-import           Control.Applicative            ((<|>))
-import           Control.Monad                  (guard)
-import           Debug.Trace                    (traceM)
-import           NonConjunctive.Residualization
-import qualified NonConjunctive.Unfold          as NC
-import qualified OCanrenize                     as OC
+import           ConsPD.Residualization
+import qualified ConsPD.Unfold          as ConsPD
+import           Control.Applicative    ((<|>))
+import           Control.Monad          (guard)
+import           Debug.Trace            (traceM)
+import qualified OCanrenize             as OC
 import           Printer.Dot
-import           Printer.NCTree                 ()
+import           Printer.ConsPDTree     ()
 import           Purification
-import           Residualize                    (vident)
+import           Residualize            (vident)
 import           Syntax
 import           System.Directory
-import           System.FilePath                ((</>), (<.>))
-import           System.Process                 (system)
+import           System.FilePath        ((<.>), (</>))
+import           System.Process         (system)
 import           Text.Printf
 import qualified Transformer.MkToProlog
-import           Util.Miscellaneous             (escapeTick)
+import           Util.Miscellaneous     (escapeTick)
 
 
 toOcanren fileName (Program defs goal) names =
   OC.topLevel fileName "topLevel" Nothing (goal, names, defs)
 
-runNc l = Transformer.NonConj.transform "test/out/nc" True Nothing (NC.nonConjunctive l)
+runConsPD l = Transformer.ConsPD.transform "test/out/consPD" True Nothing (ConsPD.topLevel l)
 
 transform dirName cleanDir env function filename goal@(Program definitions _) = (do
   let transformed@(tree, logicGoal, names) = function goal
-  let tree' = NC.simplify tree
-  -- traceM (printf "\n========================================\nBefore:\n%s\n\nAfter:\n%s\n========================================\n" (show tree) (show $ NC.simplify tree))
+  let tree' = ConsPD.simplify tree
+  -- traceM (printf "\n========================================\nBefore:\n%s\n\nAfter:\n%s\n========================================\n" (show tree) (show $ ConsPD.simplify tree))
 
   let path = dirName </> filename
   exists <- doesDirectoryExist path
@@ -41,7 +41,7 @@ transform dirName cleanDir env function filename goal@(Program definitions _) = 
[…]

### 2020-07-22 src/Transformer/JustUnfold.hs

@@ -1,6 +1,6 @@
 module Transformer.JustUnfold where
 
-import qualified Transformer.NonConj
-import qualified NonConjunctive.Unfold
+import qualified ConsPD.Unfold
+import qualified Transformer.ConsPD
 
-transform l = Transformer.NonConj.transform "test/out/nc" True Nothing (NonConjunctive.Unfold.justUnfold l)
+transform l = Transformer.ConsPD.transform "test/out/consPD" True Nothing (ConsPD.Unfold.justUnfold l)

### 2020-07-22 test/auto/Test/ConsPD.hs

@@ -1,0 +1,219 @@
+module Test.ConsPD where
+
+import           Test.Helper            (test, test2)
+
+import qualified ConsPD.Unfold          as ConsPD
+import           Printer.ConsPDTree     ()
+import qualified Program.Bottles
+import qualified Program.Bridge
+import qualified Program.Bridge2
+import qualified Program.Desert
+import           Program.List           (appendo, maxLengtho, nil, revAcco,
+                                         reverso, (%))
+import           Program.Path
+import           Program.Programs       (doubleAppendo, rep)
+import qualified Program.Prop
+import           Program.Stlc           (evalo)
+import qualified Program.Unify
+import           Syntax
+import qualified Transformer.ConsPD     as ConsPD
+import qualified Transformer.JustUnfold as JU
+import qualified Transformer.MkToProlog as Mk2Pl
+import qualified Transformer.PrologToMk as Pl2Mk
+
+dA = Program doubleAppendo $ fresh ["x", "y", "z", "r"] (call "doubleAppendo" [V "x", V "y", V "z", V "r"])
+revAcco' = Program revAcco $ fresh ["x", "y"] (call "revacco" [V "x", nil, V "y"])
+rev = Program reverso $ fresh ["x", "y"] (call "reverso" [V "x", V "y"])
+maxLen = Program maxLengtho $ fresh ["xs", "m", "l"] (call "maxLengtho" [V "xs", V "m", V "l"])
+lambda = Program evalo $ fresh ["m", "n"] (call "evalo" [V "m", V "n"])
+
+runJu = JU.transform
+
+runConsPD l = ConsPD.transform "test/out/nc" True Nothing (ConsPD.topLevel l)
+
+runRep = do
+    runJu 100 "rep" rep
+  where
+    rep = Program Program.Programs.rep $ fresh ["n", "x"] (call "rep" [V "n", V "x"])
+
+runAppendo = do
+    mapM (\d -> runJu d ("app" ++ show d) app) [1..15]
+    return ()
+  where
+    app = Program Program.List.appendo $ fresh ["x", "y", "z"] (call "appendo" [V "x", V "y", V "z"])
+
+runProp = do
[…]

### 2020-07-22 test/auto/Test/NonConjunctive.hs

@@ -1,219 +1,0 @@
-module Test.NonConjunctive  where
-
-import           Test.Helper            (test, test2)
-
-import qualified NonConjunctive.Unfold  as NC
-import           Printer.NCTree         ()
-import qualified Program.Bottles
-import qualified Program.Bridge
-import qualified Program.Bridge2
-import qualified Program.Desert
-import           Program.List           (appendo, maxLengtho, nil, revAcco,
-                                         reverso, (%))
-import           Program.Path
-import           Program.Programs       (doubleAppendo, rep)
-import qualified Program.Prop
-import           Program.Stlc           (evalo)
-import qualified Program.Unify
-import           Syntax
-import qualified Transformer.MkToProlog as Mk2Pl
-import qualified Transformer.NonConj    as NonConj
-import qualified Transformer.PrologToMk as Pl2Mk
-import qualified Transformer.JustUnfold as JU
-
-dA = Program doubleAppendo $ fresh ["x", "y", "z", "r"] (call "doubleAppendo" [V "x", V "y", V "z", V "r"])
-revAcco' = Program revAcco $ fresh ["x", "y"] (call "revacco" [V "x", nil, V "y"])
-rev = Program reverso $ fresh ["x", "y"] (call "reverso" [V "x", V "y"])
-maxLen = Program maxLengtho $ fresh ["xs", "m", "l"] (call "maxLengtho" [V "xs", V "m", V "l"])
-lambda = Program evalo $ fresh ["m", "n"] (call "evalo" [V "m", V "n"])
-
-runJu = JU.transform
-
-runNc l = NonConj.transform "test/out/nc" True Nothing (NC.nonConjunctive l)
-
-runRep = do
-    runJu 100 "rep" rep
-  where
-    rep = Program Program.Programs.rep $ fresh ["n", "x"] (call "rep" [V "n", V "x"])
-
-runAppendo = do
-    mapM (\d -> runJu d ("app" ++ show d) app) [1..15]
-    return ()
-  where
-    app = Program Program.List.appendo $ fresh ["x", "y", "z"] (call "appendo" [V "x", V "y", V "z"])
-
[…]

### 2020-07-22 test/auto/Test/SymExec.hs

@@ -12,7 +12,7 @@ dA = Program doubleAppendo $ fresh ["x", "y", "z", "r"] (call "doubleAppendo" [V
 revAcco' = Program revAcco $ fresh ["x", "y"] (call "revacco" [V "x", nil, V "y"])
 rev = Program reverso $ fresh ["x", "y"] (call "reverso" [V "x", V "y"])
 
-unit_nonConjunctiveTest = do
+unit_symExec = do
   transform 5 "da" dA
   transform 5 "rev" rev
   transform 5 "revAcco" revAcco'

### 2020-08-17 papers/miniKanren/slides/doubleAppPD.tex
Slides: many figures

\begin{tikzpicture}[
  processTree,
  level 3/.style={sibling distance=23em},
  level 4/.style={sibling distance=10em}]
  \node {\rel{doubleAppend}{$x \ y \ z \ r$}}
    child { node {\rel{append}{$x \ y \ t$} \and \rel{append}{$t \ z \ r$}}
      child { node[diamond] {\and}
        child { node (1) {\rel{append}{$x \ y \ t$} }
          child { node {\subst{$x \to [], y \to t$}} }
          child { node (12) {\rel{appendo}{$x' \ y \ t'$} \\ \subst{$x \to h :: x', t \to h :: t'$}}}}
        child { node (2) {\rel{appendo}{$t \ z \ r$} }
          child { node (21) {\subst{$t \to [], z \to r$}} }
          child { node (22) {\rel{appendo}{$t' \ y \ r'$} \\ \subst{$t \to h :: t', r \to h :: r'$}}}}}};
  \draw [dashed,<-] (1.east) to [out=10,in=30] (12.east);
  \draw [dashed,<-] (2.east) to [out=10,in=30] (22.east);
  \draw [red,<->] (12.south) to [out=300,in=240] (21.south);
\end{tikzpicture}

### 2020-08-17 papers/miniKanren/slides/propCPDunf3.tex
Slides: many figures

\begin{tikzpicture}[
  processTree,
  sibling distance=10em,
  level distance=8em]
  \node {
    \underline{\rel{eval}{$fm \ s \ true$}}}
    child { node[draw=none, fill=none] {...}}
    child { node {
      \rel{eval}{$x \ s \ a$} \and \\
      \underline{\rel{not}{$a \ true$}} \\
      \subst{$fm \to neg \ x$}}
      child { node {
        \rel{eval}{$x \ s \ false$} \\
        \subst{$a \to false$}}}}
    child { node[draw=none, fill=none] {...}}
    child { node[draw=none, fill=none] {...}}
  ;
\end{tikzpicture}

### 2020-08-18 papers/miniKanren/slides/doubleAppPD.tex
Actual slides

@@
   level 4/.style={sibling distance=10em}]
-  \node {\rel{doubleAppend}{$x \ y \ z \ r$}}
+  \node {\rel{double\_append}{$x \ y \ z \ r$}}
     child { node {\rel{append}{$x \ y \ t$} \and \rel{append}{$t \ z \ r$}}

### 2020-08-18 papers/miniKanren/slides/pd.tex
Actual slides

\begin{tikzpicture}[
  node distance = 10mm and 18 mm,
  decoration = {
    snake,
    pre length=2pt,
    post length=2pt
  }]
  \node (a) [program]
  {
    \adjustbox{scale=0.7}{
      \begin{minipage}[c]{0.47\textwidth}
        \input{doubleAppDef.tex}
        \input{appendoDef.tex}
      \end{minipage}
    }
  };

  \node (b) [goal,anchor=north east] at (a.south east)
  {
    \adjustbox{scale=0.7}{
      \begin{minipage}[c]{0.35\textwidth}
        \begin{lstlisting}
double_append$^o$ x y z r
        \end{lstlisting}
      \end{minipage}
    }
  };

  \node [transparent, anchor=south] at (a.north) {\footnotesize input};

  \node (c) [program, right=of a.north east, anchor=north west]
  {
    \adjustbox{scale=0.7}{
      \begin{minipage}[c]{0.52\textwidth}
        \input{doubleAppRes.tex}
        \input{appendoDef.tex}
      \end{minipage}
    }
  };

  \node [transparent, anchor=south] at (c.north) {\footnotesize output};

\end{tikzpicture}

### 2020-08-19 papers/miniKanren/slides/doubleAppPD.tex
More slides

--- 2020-08-17 papers/miniKanren/slides/doubleAppPD.tex
@@
   level 3/.style={sibling distance=23em},
-  level 4/.style={sibling distance=10em}]
-  \node {\rel{doubleAppend}{$x \ y \ z \ r$}}
+  level 4/.style={sibling distance=10em},
+  anchor=center]
+  \node
+    {\rel{double\_append}{$x \ y \ z \ r$}}
     child { node {\rel{append}{$x \ y \ t$} \and \rel{append}{$t \ z \ r$}}
-      child { node[diamond] {\and}
+      child { node (d) [diamond] {\and}
         child { node (1) {\rel{append}{$x \ y \ t$} }
@@
           child { node (22) {\rel{appendo}{$t' \ y \ r'$} \\ \subst{$t \to h :: t', r \to h :: r'$}}}}}};
+
+  \draw [densely dotted,->] ($(d.east)+(5,1.5)$) node[above,draw=none] {\Large split node} to [out=210,in=0] (d.east);
+
   \draw [dashed,<-] (1.east) to [out=10,in=30] (12.east);
   \draw [dashed,<-] (2.east) to [out=10,in=30] (22.east);
-  \draw [red,<->] (12.south) to [out=300,in=240] (21.south);
+
+  \draw [red,<->] ($(12.south)+(0.7,0.15)$) to [out=300,in=240] node[below, draw=none, black] {\Large substitutions conflict} ($(21.south)+(-0.65,0.15)$);
+
 \end{tikzpicture}

### 2020-08-25 papers/miniKanren/slides/evalSplit.tex
Slides reworked to have a single running example

\begin{tikzpicture}[
  processTree,
  anchor=center,
  level 2/.style={level distance=5em},
  level 3/.style={level distance=4em},
  sibling distance=10em]
  \node (root) {\rel{eval}{$fm \ s \ true$}}
      child { node {
        \rel{and}{$a \ b \ true$} \and
        \rel{eval}{$x \ s \ a$} \and
        \rel{eval}{$y \ s \ b$} \\
        \subst{$fm \to conj \ x \ y$}}
        child { node (d) [diamond] {\and}
          child { node {\rel{and}{$a \ b \ true$}}
            child { node (subst) {\subst{$b \to true, a \to true$}}}}
          child { node (x) {\rel{eval}{$x \ s \ a$}}
          child { node[draw=none, fill=none] {...}}
          }
          child { node (y) {\rel{eval}{$y \ s \ b$}}
            child { node[draw=none, fill=none] {...}}
        }
      }
    }
    ;

  \node[left=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.west) to [out=-170,in=10] (lookup.east);

  \node[right=4em of root, yshift=-0.5cm] (lookup) {...};
  \draw [->] (root.east) to [in=170,out=-10] (lookup.west);

  \draw [red,<->] ($(subst.south)+(0.5,0.15)$) to [out=300,in=-100] node[below, draw=none, red] {\Large data lost} ($(x.south)+(0.85,0.15)$);

  \draw [red,<->] ($(subst.south)+(-1.2,0.15)$) to [out=300,in=240] node[below, draw=none, red] {\Large data lost} ($(y.south)+(0.8,0.15)$);

  \draw [densely dotted,->] ($(d.east)+(5,0.5)$) node[above,draw=none] {\Large split node} to [out=210,in=0] (d.east);
\end{tikzpicture}

### 2020-08-25 papers/miniKanren/slides/evalUnfold.tex
Slides reworked to have a single running example

--- 2020-08-25 papers/miniKanren/slides/evalSplit.tex
@@
   processTree,
-  anchor=center,
-  level 2/.style={level distance=5em},
-  level 3/.style={level distance=4em},
-  sibling distance=10em]
+  anchor=center]
   \node (root) {\rel{eval}{$fm \ s \ true$}}
+    child { node {$fm \equiv conj \ x \ y $ \and \rel{and}{$a \ b \ true$} \and \rel{eval}{$x \ s \ a$} \and \rel{eval}{$y \ s \ b$}}
       child { node {
@@
         \subst{$fm \to conj \ x \ y$}}
-        child { node (d) [diamond] {\and}
-          child { node {\rel{and}{$a \ b \ true$}}
-            child { node (subst) {\subst{$b \to true, a \to true$}}}}
-          child { node (x) {\rel{eval}{$x \ s \ a$}}
-          child { node[draw=none, fill=none] {...}}
-          }
-          child { node (y) {\rel{eval}{$y \ s \ b$}}
-            child { node[draw=none, fill=none] {...}}
-        }
     }
@@
   \draw [->] (root.east) to [in=170,out=-10] (lookup.west);
-
-  \draw [red,<->] ($(subst.south)+(0.5,0.15)$) to [out=300,in=-100] node[below, draw=none, red] {\Large data lost} ($(x.south)+(0.85,0.15)$);
-
-  \draw [red,<->] ($(subst.south)+(-1.2,0.15)$) to [out=300,in=240] node[below, draw=none, red] {\Large data lost} ($(y.south)+(0.8,0.15)$);
-
-  \draw [densely dotted,->] ($(d.east)+(5,0.5)$) node[above,draw=none] {\Large split node} to [out=210,in=0] (d.east);
 \end{tikzpicture}

### 2020-08-25 papers/miniKanren/slides/propCPDNoSplit.tex
Slides reworked to have a single running example

--- 2020-08-25 papers/miniKanren/slides/evalSplit.tex
@@
       child { node {
-        \rel{and}{$a \ b \ true$} \and
+        \underline{\rel{and}{$a \ b \ true$}} \and
         \rel{eval}{$x \ s \ a$} \and
@@
         \subst{$fm \to conj \ x \ y$}}
-        child { node (d) [diamond] {\and}
-          child { node {\rel{and}{$a \ b \ true$}}
-            child { node (subst) {\subst{$b \to true, a \to true$}}}}
-          child { node (x) {\rel{eval}{$x \ s \ a$}}
-          child { node[draw=none, fill=none] {...}}
+        child { node {
+          \underline{\rel{eval}{$x \ s \ true$}} \and
+          \rel{eval}{$y \ s \ true$} \\
+          \subst{$a \to true, b \to true$}}
+          child {node {
+            \underline{\rel{and}{$x' \ y' \ true$}} \and
+            \rel{eval}{$x' \ s \ a'$} \and
+            \rel{eval}{$y' \ s \ b'$} \and
+            \rel{eval}{$y \ s \ b$} \\
+            \subst{$x \to conj \ x' \ y'$}
           }
-          child { node (y) {\rel{eval}{$y \ s \ b$}}
-            child { node[draw=none, fill=none] {...}}
+          child { node (grow) {
+            \underline{\rel{eval}{$x' \ s \ a'$}} \and
+            \rel{eval}{$y' \ s \ b'$} \and
+            \rel{eval}{$y \ s \ b$} \\
+            \subst{$a' \to true, b' \to true$}}
+            child {node[draw=none,fill=none] {...}}
           }
         }
-    }
+      }}
     ;
@@
 
-  \draw [red,<->] ($(subst.south)+(0.5,0.15)$) to [out=300,in=-100] node[below, draw=none, red] {\Large data lost} ($(x.south)+(0.85,0.15)$);
+  \draw [red,->] ($(grow.east)+(1,-2.5)$) node[below,draw=none,anchor=north east] {\Large uncontrollable growth} to [out=45,in=0] (grow.east);
 
-  \draw [red,<->] ($(subst.south)+(-1.2,0.15)$) to [out=300,in=240] node[below, draw=none, red] {\Large data lost} ($(y.south)+(0.8,0.15)$);
-
-  \draw [densely dotted,->] ($(d.east)+(5,0.5)$) node[above,draw=none] {\Large split node} to [out=210,in=0] (d.east);
 \end{tikzpicture}

### 2020-08-25 papers/miniKanren/slides/propCPDShort.tex
Slides reworked to have a single running example

--- 2020-08-25 papers/miniKanren/slides/propCPDNoSplit.tex
@@
           \subst{$a \to true, b \to true$}}
-          child {node {
-            \underline{\rel{and}{$x' \ y' \ true$}} \and
-            \rel{eval}{$x' \ s \ a'$} \and
-            \rel{eval}{$y' \ s \ b'$} \and
-            \rel{eval}{$y \ s \ b$} \\
-            \subst{$x \to conj \ x' \ y'$}
-          }
-          child { node (grow) {
-            \underline{\rel{eval}{$x' \ s \ a'$}} \and
-            \rel{eval}{$y' \ s \ b'$} \and
-            \rel{eval}{$y \ s \ b$} \\
-            \subst{$a' \to true, b' \to true$}}
           child {node[draw=none,fill=none] {...}}
@@
       }
-      }}
     ;
@@
 
-  \draw [red,->] ($(grow.east)+(1,-2.5)$) node[below,draw=none,anchor=north east] {\Large uncontrollable growth} to [out=45,in=0] (grow.east);
-
 \end{tikzpicture}

### 2020-08-25 papers/miniKanren/slides/propCPDSplit.tex
Slides reworked to have a single running example

--- 2020-08-25 papers/miniKanren/slides/propCPDShort.tex
@@
   level 3/.style={level distance=4em},
+  level 4/.style={level distance=3em},
   sibling distance=10em]
@@
         child { node {
-          \underline{\rel{eval}{$x \ s \ true$}} \and
+          \rel{eval}{$x \ s \ true$} \and
           \rel{eval}{$y \ s \ true$} \\
           \subst{$a \to true, b \to true$}}
-          child {node[draw=none,fill=none] {...}}
+          child { node [diamond] {\and}
+            child { node (x) {
+              \rel{eval}{$x \ s \ true$}}
+            }
+            child { node (y) {
+              \rel{eval}{$y \ s \ true$}}}
+          }
         }
@@
 
+  \draw[dashed,->] (x.west) to [out=165,in=-165] (root.west);
+  \draw[dashed,->] (y.east) to [out=15,in=-15] (root.east);
+
 \end{tikzpicture}

### 2020-08-25 papers/miniKanren/slides/propSimpl.tex
Slides reworked to have a single running example

\begin{tikzpicture}[
  processTree,
  sibling distance=7em,
  level distance=7em,
  level 2/.style={level distance=5em},
  anchor=center]
  \node {
    \rel{eval}{$fm \ s \ true$}}
    child { node[draw=none, fill=none] {...}}
    child { node {
      \rel{and}{$a \ b \ true$} \and  \\
      \rel{eval}{$x \ s \ a$} \and \\
      \rel{eval}{$y \ s \ b$} \\
      \subst{$fm \to conj \ x$}}
      child { node[draw=none, fill=none] {...}}
      }
    child { node[draw=none, fill=none] {...}}
  ;
\end{tikzpicture}

### 2020-08-25 papers/miniKanren/slides/propSimplFold.tex
Slides reworked to have a single running example

\begin{tikzpicture}[
  processTree,
  sibling distance=4em,
  level distance=3em,
  level 3/.style={level distance=4em,sibling distance=7em},
  anchor=center]
  \node (root) {
    \rel{eval}{$fm \ s \ true$}}
    child { node[draw=none, fill=none] {...}}
    child { node[draw=none, fill=none] {...}
      child { node {
        \rel{eval}{$x \ s \ a$} \and \\
        \rel{eval}{$y \ s \ b$}}
        child { node (1) {\rel{eval}{$x \ s \ a$}}}
        child { node (2) {\rel{eval}{$y \ s \ b$}}}
        }
    }
    child { node[draw=none, fill=none] {...}}
  ;

  \draw [dashed,->] (1.west) to [out=170,in=-150] (root.west);
  \draw [dashed,->] (2.east) to [out=10,in=-30] (root.east);

\end{tikzpicture}

### 2020-09-29 app/eval/Main.hs
Evaluation executable

@@ -1,0 +1,70 @@
+module Main where
+
+import Syntax
+import Stream
+import Eval
+import Driving
+import Tree
+import Program.List
+import Program.Num
+import Program.Bool
+import Program.Sort
+import Program.Programs
+import Prelude hiding (succ)
+
+reify :: Eq a => [(a, Term a)] -> Term a -> Term a
+reify s x@(V v) =
+  case lookup v s of
+    Nothing -> x
+    Just t  -> reify s t
+reify s (C n ts) = C n $ map (reify s) ts
+
+toplevel :: Integer -> (Term S -> String) -> Program -> [String]
+toplevel n printer program =
+  map (\s -> printer $ reify s (V 0)) $ takeS n $ run program
+
+main :: IO ()
+main =
+  do
+    print (toplevel 1 bool (Program nando $ fresh ["q"] (call "nando" [falso, falso, V "q"])))
+    print (toplevel 1 bool (Program nando $ fresh ["q"] (call "nando" [V "q", V "q", V "q"])))
+    print (toplevel 1 show (Program nando $ fresh ["q", "p", "r"] (V "q" === C "pair" [V "p", V "r"] &&& call "nando" [V "r", V "r", V "p"])))
+
+    print (toplevel 1 bool (Program noto $ fresh ["q"] (call "noto" [falso, V "q"])))
+    print (toplevel 1 bool (Program noto $ fresh ["q"] (call "noto" [V "q", V "q"])))
+    print (toplevel 2 show (Program noto $ fresh ["q", "p", "r"] (V "q" === C "pair" [V "p", V "r"] &&& call "noto" [V "p", V "r"])))
+
+    print (toplevel 3 show (Program oro $ fresh ["q", "p", "r"] (V "q" === C "pair" [V "p", V "r"] &&& call "oro" [V "p", V "r", C "true" []])))
+    print (toplevel 4 show (Program ando $ fresh ["q", "p", "r"] (V "q" === C "pair" [V "p", V "r"] &&& call "ando" [V "p", V "r", C "false" []])))
+
+
+    print (toplevel 4 show (Program leo $ fresh ["q", "p", "r"] (V "q" === C "pair" [V "p", V "r"] &&& call "leo" [V "p", V "p", V "r"] )))
[…]

### 2020-09-29 app/Main.hs
Added the ability to specify directory for transformer for easier execution

@@ -23,6 +23,7 @@ import           Syntax
 import qualified Transformer.ConsPD
 import qualified Transformer.CPD
 import qualified Transformer.JustUnfold
+import qualified Transformer.PD
 import qualified Transformer.PrologToMk
 
 dA = Program doubleAppendo $ fresh ["x", "y", "z", "r"] (call "doubleAppendo" [V "x", V "y", V "z", V "r"])
@@ -140,35 +141,53 @@ runPropEval = do
 
 main :: IO ()
 main = do
-    Transformer.CPD.transform "branchPlainFirst" Program.PropEval.plainFirstQuery Nothing Branching
-    Transformer.CPD.transform "cpdPlainFirst" Program.PropEval.plainFirstQuery Nothing Deterministic
-    Transformer.CPD.transform "branchNandoLast" Program.PropEval.nandoLastQuery Nothing Branching
-    Transformer.CPD.transform "cpdNandoLast" Program.PropEval.nandoLastQuery Nothing Deterministic
+    Transformer.PrologToMk.transform "experiments/propEval" "originalLastPlain.pl"
+    Transformer.PrologToMk.transform "experiments/propEval" "etalonLastPlain.pl"
+    Transformer.PrologToMk.transform "experiments/propEval" "originalLastPlainLimited.pl"
+    Transformer.PrologToMk.transform "experiments/propEval" "etalonLastPlainLimited.pl"
 
-    Transformer.CPD.transform "branchUnify" Program.Unify.query Nothing Branching
-    Transformer.CPD.transform "cpdUnify" Program.Unify.query Nothing Deterministic
 
-    Transformer.CPD.transform "cpdPath" Program.Path.query1 Nothing Deterministic
-    traceM "\n==============================\nstarted transforming path:  branching\n\n"
-    Transformer.CPD.transform "branchPath" Program.Path.query1 Nothing Branching
+    -- mapM_ (Transformer.PrologToMk.transform "/home/ev/prj/kajigor/cpd/examples")
+        --   [ "geoff_doubleApp.pl"
+        --   , "geoff_originalFirstNando.pl"
+        --   , "geoff_originalLastNando.pl"
+        --   , "geoff_path.pl"
+        --   , "geoff_maxLengtho.pl"
+        --   , "geoff_originalFirstPlain.pl"
[…]

### 2020-09-29 src/Transformer/PrologToMk.hs

@@ -1,15 +1,16 @@
 module Transformer.PrologToMk where
 
 import qualified OCanrenize      as OC
-import           System.FilePath ((<.>))
+import           System.FilePath (replaceExtension, (</>))
 import           System.IO
 import           Text.Printf
 import           Util.ToProlog
 
-transform fileName = do
+transform dirName file = do
+  let fileName = dirName </> file
   handle <- openFile fileName ReadMode
   contents <- hGetContents handle
-  let ocamlCodeFileName = fileName <.> "ml"
+  let ocamlCodeFileName = replaceExtension fileName "ml"
   let g = prologToG contents
   let env = Nothing
   OC.topLevel ocamlCodeFileName "topLevel" env g

### 2020-09-29 papers/miniKanren/slides/propSimplFold.tex
Finished slides for miniKanren-2020 talk

@@
   level distance=3em,
-  level 3/.style={level distance=4em,sibling distance=7em},
+  level 3/.style={level distance=4em,sibling distance=8em},
   anchor=center]
@@
       child { node {
-        \rel{eval}{$x \ s \ a$} \and \\
-        \rel{eval}{$y \ s \ b$}}
-        child { node (1) {\rel{eval}{$x \ s \ a$}}}
-        child { node (2) {\rel{eval}{$y \ s \ b$}}}
+        \rel{eval}{$x \ s \ true$} \and \\
+        \rel{eval}{$y \ s \ true$}}
+        child { node (1) {\rel{eval}{$x \ s \ true$}}}
+        child { node (2) {\rel{eval}{$y \ s \ true$}}}
         }

### 2020-09-29 experiments/propEval/build.sh
WIP experiments

@@ -1,0 +1,2 @@
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml originalLastPlain.ml etalonLastPlain.ml originalLastPlainLimited.ml etalonLastPlainLimited.ml run.ml
+

### 2020-09-29 experiments/propEval/etalonLastPlain.pl

@@ -1,0 +1,21 @@
+evaloT(St, conj(X, Y)) :- evaloT(St, X), evaloT(St, Y).
+evaloT(St, disj(X, Y)) :- evaloT(St, X), evaloT(St, Y).
+evaloT(St, disj(X, Y)) :- evaloT(St, X), evaloF(St, Y).
+evaloT(St, disj(X, Y)) :- evaloF(St, X), evaloT(St, Y).
+evaloT(St, neg(X)) :- evaloF(St, X).
+evaloT(St, var(Z)) :- elemoT(St, Z).
+
+evaloF(St, conj(X, Y)) :- evaloF(St, X), evaloF(St, Y).
+evaloF(St, conj(X, Y)) :- evaloT(St, X), evaloF(St, Y).
+evaloF(St, conj(X, Y)) :- evaloF(St, X), evaloT(St, Y).
+evaloF(St, disj(X, Y)) :- evaloF(St, X), evaloF(St, Y).
+evaloF(St, neg(X)) :- evaloT(St, X).
+evaloF(St, var(Z)) :- elemoF(St, Z).
+
+elemoT(cons(true, Q1), o).
+elemoT(cons(Q3, Q1), s(Q2)) :- elemoT(Q1, Q2).
+
+elemoF(cons(false, Q1), o).
+elemoF(cons(Q3, Q1), s(Q2)) :- elemoF(Q1, Q2).
+
+

### 2020-09-29 experiments/propEval/etalonLastPlainLimited.pl

@@ -1,0 +1,21 @@
+evaloT(St, conj(X, Y), s(N)) :- evaloT(St, X, N), evaloT(St, Y, N).
+evaloT(St, disj(X, Y), s(N)) :- evaloT(St, X, N), evaloT(St, Y, N).
+evaloT(St, disj(X, Y), s(N)) :- evaloT(St, X, N), evaloF(St, Y, N).
+evaloT(St, disj(X, Y), s(N)) :- evaloF(St, X, N), evaloT(St, Y, N).
+evaloT(St, neg(X), s(N)) :- evaloF(St, X, N).
+evaloT(St, var(Z), o) :- elemoT(St, Z).
+
+evaloF(St, conj(X, Y), s(N)) :- evaloF(St, X, N), evaloF(St, Y, N).
+evaloF(St, conj(X, Y), s(N)) :- evaloT(St, X, N), evaloF(St, Y, N).
+evaloF(St, conj(X, Y), s(N)) :- evaloF(St, X, N), evaloT(St, Y, N).
+evaloF(St, disj(X, Y), s(N)) :- evaloF(St, X, N), evaloF(St, Y, N).
+evaloF(St, neg(X), s(N)) :- evaloT(St, X, N).
+evaloF(St, var(Z), o) :- elemoF(St, Z).
+
+elemoT(cons(true, Q1), o).
+elemoT(cons(Q3, Q1), s(Q2)) :- elemoT(Q1, Q2).
+
+elemoF(cons(false, Q1), o).
+elemoF(cons(Q3, Q1), s(Q2)) :- elemoF(Q1, Q2).
+
+

### 2020-09-29 experiments/propEval/originalLastPlain.pl

@@ -1,0 +1,16 @@
+evalo(St, conj(X, Y), U) :- evalo(St, X, V), evalo(St, Y, W), ando(V, W, U).
+evalo(St, disj(X, Y), U) :- evalo(St, X, V), evalo(St, Y, W), oro(V, W, U).
+evalo(St, neg(X), U) :- evalo(St, X, V), noto(V, U).
+evalo(St, var(Z), U) :- elemo(Z, St, U).
+ando(true, true, true).
+ando(false, true, false).
+ando(true, false, false).
+ando(false, false, false).
+oro(true, true, true).
+oro(false, true, true).
+oro(true, false, true).
+oro(false, false, false).
+noto(true, false).
+noto(false, true).
+elemo(o, cons(H, T), H).
+elemo(s(N1), cons(H, T), V) :- elemo(N1, T, V).

### 2020-09-29 experiments/propEval/originalLastPlainLimited.pl

@@ -1,0 +1,16 @@
+evalo(St, conj(X, Y), U, s(N)) :- evalo(St, X, V, N), evalo(St, Y, W, N), ando(V, W, U).
+evalo(St, disj(X, Y), U, s(N)) :- evalo(St, X, V, N), evalo(St, Y, W, N), oro(V, W, U).
+evalo(St, neg(X), U, s(N)) :- evalo(St, X, V, N), noto(V, U).
+evalo(St, var(Z), U, o) :- elemo(Z, St, U).
+ando(true, true, true).
+ando(false, true, false).
+ando(true, false, false).
+ando(false, false, false).
+oro(true, true, true).
+oro(false, true, true).
+oro(true, false, true).
+oro(false, false, false).
+noto(true, false).
+noto(false, true).
+elemo(o, cons(H, T), H).
+elemo(s(N1), cons(H, T), V) :- elemo(N1, T, V).

### 2020-09-29 papers/aplas/benchmarks/lTypechecker/build.sh

@@ -1,1 +1,1 @@
-ocamlfind c -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml l.ml l_ocanren.ml ecce.ml eccePeter.ml nonconj.ml nonconjPeter.ml original.ml originalPeter.ml run.ml
+ocamlfind c -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml l.ml l_ocanren.ml ecce.ml eccePeter.ml nonconj.ml nonconjPeter.ml original.ml originalPeter.ml geoff.ml run.ml

### 2020-09-29 papers/aplas/benchmarks/propEval/build.sh

@@ -1,2 +1,2 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml ecceFirstNando.ml ecceFirstPlain.ml ecceLastNando.ml ecceLastPlain.ml originalFirstNando.ml originalFirstPlain.ml originalLastNando.ml originalLastPlain.ml nonconjFirstNando.ml nonconjFirstPlain.ml nonconjLastNando.ml nonconjLastPlain.ml geoff_originalLastPlain.ml geoff_originalLastNando.ml geoff_originalFirstPlain.ml geoff_originalFirstNando.ml  run.ml
+ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml ecceFirstNando.ml ecceFirstPlain.ml ecceLastNando.ml ecceLastPlain.ml originalFirstNando.ml originalFirstPlain.ml originalLastNando.ml originalLastPlain.ml nonconjFirstNando.ml nonconjFirstPlain.ml nonconjLastNando.ml nonconjLastPlain.ml geoff_originalLastPlain.ml geoff_originalLastNando.ml geoff_originalFirstPlain.ml geoff_originalFirstNando.ml geoff_reverse.ml geoff.ml etalonLastPlain.ml mergedLastPlain.ml run.ml
 

### 2020-09-29 papers/aplas/benchmarks/propEval/etalonLastPlain.pl

@@ -1,0 +1,21 @@
+evaloT(St, conj(X, Y)) :- evaloT(St, X), evaloT(St, Y).
+evaloT(St, disj(X, Y)) :- evaloT(St, X), evaloT(St, Y).
+evaloT(St, disj(X, Y)) :- evaloT(St, X), evaloF(St, Y).
+evaloT(St, disj(X, Y)) :- evaloF(St, X), evaloT(St, Y).
+evaloT(St, neg(X)) :- evaloF(St, X).
+evaloT(St, var(Z)) :- elemoT(St, Z). 
+
+evaloF(St, conj(X, Y)) :- evaloF(St, X), evaloF(St, Y).
+evaloF(St, conj(X, Y)) :- evaloT(St, X), evaloF(St, Y).
+evaloF(St, conj(X, Y)) :- evaloF(St, X), evaloT(St, Y).
+evaloF(St, disj(X, Y)) :- evaloF(St, X), evaloF(St, Y).
+evaloF(St, neg(X)) :- evaloT(St, X).
+evaloF(St, var(Z)) :- elemoF(St, Z).
+
+elemoT(cons(true, Q1), o).
+elemoT(cons(Q3, Q1), s(Q2)) :- elemoT(Q1, Q2). 
+
+elemoF(cons(false, Q1), o). 
+elemoF(cons(Q3, Q1), s(Q2)) :- elemoF(Q1, Q2).
+
+

### 2020-09-29 papers/aplas/benchmarks/propEval/mergedLastPlain.pl

@@ -1,0 +1,26 @@
+evaloT(St, conj(X, Y)) :- evaloTT(St, X, Y).
+evaloT(St, disj(X, Y)) :- evaloTT(St, X, Y).
+evaloT(St, disj(X, Y)) :- evaloTF(St, X, Y).
+evaloT(St, disj(X, Y)) :- evaloFT(St, X, Y).
+evaloT(St, neg(X)) :- evaloF(St, X).
+evaloT(St, var(Z)) :- elemoT(St, Z). 
+
+evaloF(St, conj(X, Y)) :- evaloFF(St, X, Y).
+evaloF(St, conj(X, Y)) :- evaloTF(St, X, Y).
+evaloF(St, conj(X, Y)) :- evaloFT(St, X, Y).
+evaloF(St, disj(X, Y)) :- evaloFF(St, X, Y).
+evaloF(St, neg(X)) :- evaloT(St, X).
+evaloF(St, var(Z)) :- elemoF(St, Z).
+
+evaloTT(St, X, Y) :- evaloT(St, X), evaloT(St, Y). 
+evaloTF(St, X, Y) :- evaloT(St, X), evaloF(St, Y). 
+evaloFT(St, X, Y) :- evaloF(St, X), evaloT(St, Y).
+evaloFF(St, X, Y) :- evaloF(St, X), evaloF(St, Y).
+
+elemoT(cons(true, Q1), o).
+elemoT(cons(Q3, Q1), s(Q2)) :- elemoT(Q1, Q2). 
+
+elemoF(cons(false, Q1), o). 
+elemoF(cons(Q3, Q1), s(Q2)) :- elemoF(Q1, Q2).
+
+

### 2020-09-29 papers/aplas/benchmarks/doubleAppendo/build.sh
Papers removed from the repository.
Papers themselves are now in the tex repository, and benchmarks are in mk-transofmerns-bench repository

@@ -1,3 +1,0 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml da.ml ideal.ml ecce.ml geoff.ml run.ml
-
-

### 2020-09-29 papers/aplas/benchmarks/maxLengtho/ecce.pl

@@ -1,46 +1,0 @@
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 60 ms */
-/* Unfolding time: 60 ms */
-/* Post-Processing time: 60 ms */
-
-/* Specialised Predicates: 
-maxLengtho__1(A,B,C) :- maxLengtho(A,B,C).
-maxo1_conj__2(A,B,C) :- maxo1(A,o,B), lengtho(A,C).
-maxo1_conj__3(A,B,C,D) :- maxo1(A,s(B),C), lengtho(A,D).
-leo_conj__4(A,B,C,D,E) :- leo(A,s(B),true), maxo1(C,s(B),D), lengtho(C,E).
-gto__5(A,B) :- gto(s(A),B,true).
-leo__6(A,B) :- leo(A,B,true).
-*/
-
-maxLengtho(A,B,C) :- 
-    maxLengtho__1(A,B,C).
-maxLengtho__1(nil,o,o).
-maxLengtho__1(cons(o,A),B,s(C)) :- 
-    maxo1_conj__2(A,B,C).
-maxLengtho__1(cons(s(A),B),C,s(D)) :- 
-    maxo1_conj__3(B,A,C,D).
-maxo1_conj__2(nil,o,o).
-maxo1_conj__2(cons(o,A),B,s(C)) :- 
-    maxo1_conj__2(A,B,C).
-maxo1_conj__2(cons(s(A),B),C,s(D)) :- 
-    maxo1_conj__3(B,A,C,D).
-maxo1_conj__3(nil,A,s(A),o).
-maxo1_conj__3(cons(A,B),C,D,s(E)) :- 
-    leo_conj__4(A,C,B,D,E).
-maxo1_conj__3(cons(s(s(A)),B),C,D,s(E)) :- 
-    gto__5(A,C), 
-    maxo1_conj__3(B,s(A),D,E).
-leo_conj__4(o,A,B,C,D) :- 
-    maxo1_conj__3(B,A,C,D).
-leo_conj__4(s(A),B,C,D,E) :- 
-    leo__6(A,B), 
[…]

### 2020-09-29 papers/aplas/benchmarks/maxLengtho/geoff.pl

@@ -1,15 +1,0 @@
-p(Ls,M,L) :- p_(M,Ls,L).
-p_(M,[o|T],s(Z)) :- p_(M,T,Z).
-p_(M,[s(Zz)|T],s(Z)) :- p__(T,Zz,M),p_____(T,Z).
-p_(o,[],o).
-p__([o|T__],Zz,M) :- p__(T__,Zz,M).
-p__([s(X1)|T__],Zz,M) :- p___(X1,Zz,T__,Zz,M).
-p__([s(X1)|T__],Zz,M) :- p____(X1,Zz,T__,X1,M).
-p__([],Zz,s(Zz)).
-p___(o,X_____,T__,X_______,M) :- p__(T__,X_______,M).
-p___(s(X1_),s(Y1_),T__,X_______,M) :- p___(X1_,Y1_,T__,X_______,M).
-p____(s(X1_),s(Y1_),T__,X_______,M) :- p____(X1_,Y1_,T__,X_______,M).
-p____(s(Zz_),o,T__,X_______,M) :- p__(T__,X_______,M).
-p_____([H__|T__],s(Z_)) :- p_____(T__,Z_).
-p_____([],o).
-<- p(Ls,M,L).

### 2020-09-29 papers/aplas/benchmarks/path/build.sh

@@ -1,1 +1,0 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml nonconj.ml ecce.ml geoff.ml run.ml

### 2020-09-29 papers/aplas/benchmarks/propEval/ecceFirstPlain.pl

@@ -1,57 +1,0 @@
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 120 ms */
-/* Unfolding time: 80 ms */
-/* Post-Processing time: 120 ms */
-
-/* Specialised Predicates: 
-evalo__1(A,B) :- evalo(A,B,true).
-oro_conj__2(A,B,C) :- oro(D1,E1,true), evalo(A,B,D1), evalo(A,C,E1).
-evalo__3(A,B) :- evalo(A,B,false).
-ando_conj__4(A,B,C) :- ando(D1,E1,false), evalo(A,B,D1), evalo(A,C,E1).
-elemo__5(A,B,C,D) :- elemo(A,cons(B,C),D).
-*/
-
-evalo(A,B,true) :- 
-    evalo__1(A,B).
-evalo__1(A,conj(B,C)) :- 
-    evalo__1(A,B), 
-    evalo__1(A,C).
-evalo__1(A,disj(B,C)) :- 
-    oro_conj__2(A,B,C).
-evalo__1(A,neg(B)) :- 
-    evalo__3(A,B).
-evalo__1(cons(A,B),var(C)) :- 
-    elemo__5(C,A,B,true).
-oro_conj__2(A,B,C) :- 
-    evalo__1(A,B), 
-    evalo__1(A,C).
-oro_conj__2(A,B,C) :- 
-    evalo__3(A,B), 
-    evalo__1(A,C).
-oro_conj__2(A,B,C) :- 
-    evalo__1(A,B), 
-    evalo__3(A,C).
-evalo__3(A,conj(B,C)) :- 
-    ando_conj__4(A,B,C).
-evalo__3(A,disj(B,C)) :- 
-    evalo__3(A,B), 
-    evalo__3(A,C).
-evalo__3(A,neg(B)) :- 
-    evalo__1(A,B).
-evalo__3(cons(A,B),var(C)) :- 
-    elemo__5(C,A,B,false).
-ando_conj__4(A,B,C) :- 
-    evalo__3(A,B), 
-    evalo__1(A,C).
-ando_conj__4(A,B,C) :- 
[…]

### 2020-09-29 papers/aplas/benchmarks/propEval/geoff_originalLastNando.pl

@@ -1,33 +1,0 @@
-p(St,conj(X,Y)) :- p(St,X),p(St,Y).
-p(St,disj(X,Y)) :- p(St,X),p(St,Y).
-p(St,disj(X,Y)) :- p(St,X),p___(St,Y).
-p(St,disj(X,Y)) :- p_(St,X),p(St,Y).
-p(St,neg(X)) :- p_____(St,X).
-p(St,var(Z)) :- p_______(Z,St).
-p_(St,conj(X___,Y_)) :- p(St,X___),p_(St,Y_).
-p_(St,conj(X___,Y_)) :- p_(St,X___),p(St,Y_).
-p_(St,conj(X___,Y_)) :- p_(St,X___),p_(St,Y_).
-p_(St,disj(X___,Y_)) :- p_(St,X___),p_(St,Y_).
-p_(St,neg(X___)) :- p(St,X___).
-p_(St,var(Z)) :- p__(Z,St).
-p__(o,[false|T]).
-p__(s(N1),[H|T]) :- p__(N1,T).
-p___(St,conj(X___,Y_)) :- p(St,X___),p___(St,Y_).
-p___(St,conj(X___,Y_)) :- p___(St,X___),p(St,Y_).
-p___(St,conj(X___,Y_)) :- p___(St,X___),p___(St,Y_).
-p___(St,disj(X___,Y_)) :- p___(St,X___),p___(St,Y_).
-p___(St,neg(X___)) :- p(St,X___).
-p___(St,var(Z)) :- p____(Z,St).
-p____(o,[false|T]).
-p____(s(N1),[H|T]) :- p____(N1,T).
-p_____(St,conj(X_,Y)) :- p(St,X_),p_____(St,Y).
-p_____(St,conj(X_,Y)) :- p_____(St,X_),p(St,Y).
-p_____(St,conj(X_,Y)) :- p_____(St,X_),p_____(St,Y).
-p_____(St,disj(X_,Y)) :- p_____(St,X_),p_____(St,Y).
-p_____(St,neg(X_)) :- p(St,X_).
-p_____(St,var(Z)) :- p______(Z,St).
-p______(o,[false|T]).
-p______(s(N1),[H|T]) :- p______(N1,T).
-p_______(o,[true|T]).
-p_______(s(N1),[H|T]) :- p_______(N1,T).
-<- p(St,Fm).

### 2020-09-29 papers/aplas/benchmarks/propEval/mergedLastPlain.pl

@@ -1,26 +1,0 @@
-evaloT(St, conj(X, Y)) :- evaloTT(St, X, Y).
-evaloT(St, disj(X, Y)) :- evaloTT(St, X, Y).
-evaloT(St, disj(X, Y)) :- evaloTF(St, X, Y).
-evaloT(St, disj(X, Y)) :- evaloFT(St, X, Y).
-evaloT(St, neg(X)) :- evaloF(St, X).
-evaloT(St, var(Z)) :- elemoT(St, Z). 
-
-evaloF(St, conj(X, Y)) :- evaloFF(St, X, Y).
-evaloF(St, conj(X, Y)) :- evaloTF(St, X, Y).
-evaloF(St, conj(X, Y)) :- evaloFT(St, X, Y).
-evaloF(St, disj(X, Y)) :- evaloFF(St, X, Y).
-evaloF(St, neg(X)) :- evaloT(St, X).
-evaloF(St, var(Z)) :- elemoF(St, Z).
-
-evaloTT(St, X, Y) :- evaloT(St, X), evaloT(St, Y). 
-evaloTF(St, X, Y) :- evaloT(St, X), evaloF(St, Y). 
-evaloFT(St, X, Y) :- evaloF(St, X), evaloT(St, Y).
-evaloFF(St, X, Y) :- evaloF(St, X), evaloF(St, Y).
-
-elemoT(cons(true, Q1), o).
-elemoT(cons(Q3, Q1), s(Q2)) :- elemoT(Q1, Q2). 
-
-elemoF(cons(false, Q1), o). 
-elemoF(cons(Q3, Q1), s(Q2)) :- elemoF(Q1, Q2).
-
-

### 2020-09-29 papers/aplas/benchmarks/propEval/nonconjLastNando.pl

@@ -1,22 +1,0 @@
-_evalo(conj(Q1, Q2), Y3) :- ___evaloEvalo(Y3, Q1, Q2).
-_evalo(disj(Q1, Q2), Y3) :- __evalo(Y3, Q1), _evalo(Q2, Y3).
-_evalo(disj(Q1, Q2), Y3) :- __evaloEvalo(Y3, Q1, Q2).
-_evalo(disj(Q1, Q2), Y3) :- ___evaloEvalo(Y3, Q1, Q2).
-_evalo(neg(Q1), Y3) :- __evalo(Y3, Q1).
-_evalo(var(o), cons(true, Q4)).
-_evalo(var(s(Q5)), cons(Q6, Q4)) :- _elemo(Q4, Q5).
-__evalo(Y4, conj(Q1, Q2)) :- _evaloEvalo(Y4, Q1, Q2).
-__evalo(Y4, conj(Q1, Q2)) :- evaloEvalo(Y4, Q1, Q2).
-__evalo(Y4, conj(Q1, Q2)) :- _evalo(Q1, Y4), __evalo(Y4, Q2).
-__evalo(Y4, disj(Q1, Q2)) :- _evaloEvalo(Y4, Q1, Q2).
-__evalo(Y4, neg(Q1)) :- _evalo(Q1, Y4).
-__evalo(cons(false, Q4), var(o)).
-__evalo(cons(Q6, Q4), var(s(Q5))) :- elemo(Q4, Q5).
-evaloEvalo(Y6, Y7, Y8) :- __evalo(Y6, Y7), _evalo(Y8, Y6).
-_evaloEvalo(Y9, Y10, Y11) :- __evalo(Y9, Y10), __evalo(Y9, Y11).
-elemo(cons(false, Q1), o).
-elemo(cons(Q3, Q1), s(Q2)) :- elemo(Q1, Q2).
-__evaloEvalo(Y14, Y15, Y16) :- _evalo(Y15, Y14), __evalo(Y14, Y16).
-___evaloEvalo(Y17, Y18, Y19) :- _evalo(Y18, Y17), _evalo(Y19, Y17).
-_elemo(cons(true, Q1), o).
-_elemo(cons(Q3, Q1), s(Q2)) :- _elemo(Q1, Q2).

### 2020-09-29 papers/aplas/benchmarks/propEval/nonconjLastPlain.pl

@@ -1,22 +1,0 @@
-_evalo(conj(Q1, Q2), Y3) :- evaloEvalo(Y3, Q1, Q2).
-_evalo(disj(Q1, Q2), Y3) :- evaloEvalo(Y3, Q1, Q2).
-_evalo(disj(Q1, Q2), Y3) :- __evalo(Y3, Q1), _evalo(Q2, Y3).
-_evalo(disj(Q1, Q2), Y3) :- ___evaloEvalo(Y3, Q1, Q2).
-_evalo(neg(Q1), Y3) :- __evalo(Y3, Q1).
-_evalo(var(o), cons(true, Q4)).
-_evalo(var(s(Q5)), cons(Q6, Q4)) :- _elemo(Q4, Q5).
-evaloEvalo(Y4, Y5, Y6) :- _evalo(Y5, Y4), _evalo(Y6, Y4).
-__evalo(Y7, conj(Q1, Q2)) :- _evaloEvalo(Y7, Q1, Q2).
-__evalo(Y7, conj(Q1, Q2)) :- _evalo(Q1, Y7), __evalo(Y7, Q2).
-__evalo(Y7, conj(Q1, Q2)) :- __evaloEvalo(Y7, Q1, Q2).
-__evalo(Y7, disj(Q1, Q2)) :- __evaloEvalo(Y7, Q1, Q2).
-__evalo(Y7, neg(Q1)) :- _evalo(Q1, Y7).
-__evalo(cons(false, Q4), var(o)).
-__evalo(cons(Q6, Q4), var(s(Q5))) :- elemo(Q4, Q5).
-_evaloEvalo(Y9, Y10, Y11) :- __evalo(Y9, Y10), _evalo(Y11, Y9).
-__evaloEvalo(Y12, Y13, Y14) :- __evalo(Y12, Y13), __evalo(Y12, Y14).
-elemo(cons(false, Q1), o).
-elemo(cons(Q3, Q1), s(Q2)) :- elemo(Q1, Q2).
-___evaloEvalo(Y17, Y18, Y19) :- _evalo(Y18, Y17), __evalo(Y17, Y19).
-_elemo(cons(true, Q1), o).
-_elemo(cons(Q3, Q1), s(Q2)) :- _elemo(Q1, Q2).

### 2020-09-29 papers/aplas/benchmarks/unify/original.pl

@@ -1,48 +1,0 @@
-eq_nat(z, z, true).
-eq_nat(s(Q39), z, false).
-eq_nat(z, s(Q41), false).
-eq_nat(s(X), s(Y), Q36) :- eq_nat(X, Y, Q36).
-get_term(Var, nil, none).
-get_term(z, cons(Q32, Xs), Q32).
-get_term(s(N), cons(X, Xs), Q32) :- get_term(N, Xs, Q32).
-check_uni(Subst, constr(N1, A1), constr(N2, A2), false) :- eq_nat(N1, N2, false), forall2(Subst, A1, A2, Q13).
-check_uni(Subst, constr(N1, A1), constr(N2, A2), Q13) :- eq_nat(N1, N2, true), forall2(Subst, A1, A2, Q13).
-check_uni(Subst, var_(V), constr(N, A), false) :- get_term(V, Subst, none).
-check_uni(Subst, var_(V), constr(N, A), Q31) :- get_term(V, Subst, some(T)), check_uni(Subst, T, constr(N, A), Q31).
-check_uni(Subst, constr(N, A), var_(V), false) :- get_term(V, Subst, none).
-check_uni(Subst, constr(N, A), var_(V), Q31) :- get_term(V, Subst, some(T)), check_uni(Subst, constr(N, A), T, Q31).
-check_uni(Subst, var_(V1), var_(V2), Q31) :- get_term(V1, Subst, some(T11)), check_uni(Subst, T11, var_(V2), Q31).
-check_uni(Subst, var_(V1), var_(V2), false) :- get_term(V1, Subst, none), get_term(V2, Subst, some(Q28)).
-check_uni(Subst, var_(V1), var_(V2), Q31) :- get_term(V1, Subst, none), get_term(V2, Subst, none), eq_nat(V1, V2, Q31).
-get_term(Var, nil, none).
-get_term(z, cons(Q32, Xs), Q32).
-get_term(s(N), cons(X, Xs), Q32) :- get_term(N, Xs, Q32).
-eq_nat(z, z, true).
-eq_nat(s(Q39), z, false).
-eq_nat(z, s(Q41), false).
-eq_nat(s(X), s(Y), Q36) :- eq_nat(X, Y, Q36).
-forall2(Subst, nil, nil, true).
-forall2(Subst, cons(X, Xs), cons(Y, Ys), false) :- check_uni(Subst, X, Y, false), forall2(Subst, Xs, Ys, Q4).
[…]

### 2020-09-29 papers/miniKanren/benchmarks/plain/build.sh

@@ -1,1 +1,0 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package mtime -package mtime.clock.os -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml conspd.ml cpd.ml branch.ml cpd_new.ml branch_new.ml run.ml

### 2020-09-29 papers/miniKanren/ecce/bottles/ecce.pl

@@ -1,102 +1,0 @@
-checkAnswer0(A,B,C,true) :- 
-    checkAnswer0__1(A,B,C).
-checkAnswer0__1([],A,o).
-checkAnswer0__1([pair(A,B)|C],D,E) :- 
-    checkStep_conj__2(A,B,D,C,E).
-checkStep_conj__2(fill,fst,pair(A,B),C,D) :- 
-    checkAnswer__4(A,o,C,A,B,D).
-checkStep_conj__2(fill,snd,pair(A,B),C,D) :- 
-    checkAnswer__4(o,B,C,A,B,D).
-checkStep_conj__2(empty,fst,pair(o,A),B,C) :- 
-    checkAnswer__3(B,o,A,C).
-checkStep_conj__2(empty,snd,pair(A,o),B,C) :- 
-    checkAnswer__3(B,A,o,C).
-checkAnswer__3([],A,B,o).
-checkAnswer__3([pair(A,B)|C],D,E,F) :- 
-    checkStep_conj__2(A,B,pair(D,E),C,F).
-checkAnswer__4(A,B,[],C,D,E) :- 
-    isFinishState__5(A,B,E).
-checkAnswer__4(A,B,[pair(C,D)|E],F,G,H) :- 
-    checkStep_conj__6(A,B,C,D,F,G,E,H).
-isFinishState__5(A,B,C) :- 
-    fancyEq__16(A,C), 
-    fancyEq__18(B,C).
-isFinishState__5(A,B,C) :- 
-    fancyEq_conj__17(A,C,B).
-checkStep_conj__6(o,A,fill,fst,B,C,D,E) :- 
-    checkAnswer__4(B,A,D,B,C,E).
-checkStep_conj__6(A,o,fill,snd,B,C,D,E) :- 
-    checkAnswer__4(A,C,D,B,C,E).
-checkStep_conj__6(A,B,empty,fst,C,D,E,F) :- 
-    fancyEq__16(A,C), 
-    checkAnswer__4(o,B,E,C,D,F).
-checkStep_conj__6(A,B,empty,snd,C,D,E,F) :- 
-    fancyEq__16(B,D), 
-    checkAnswer__4(A,o,E,C,D,F).
-checkStep_conj__6(s(A),B,pour,fst,C,D,E,F) :- 
-    fancyEq__7(B,D), 
-    doStep_conj__15(A,B,C,D,E,F).
[…]

### 2020-09-29 papers/miniKanren/ecce/prop/assoc.pl

@@ -1,652 +1,0 @@
-eval(St, var(X), U) :- assoc(X, St, U).
-eval(St, conj(X,Y), U)  :- eval(St, X, V), eval(St, Y, W), and(V, W, U).
-eval(St, disj(X,Y), U)  :- eval(St, X, V), eval(St, Y, W), or(V, W, U).
-eval(St, not(X), U) :- eval(St, X, neg(U)).
-
-assoc(X, [pair(X,V)|T], V).
-assoc(X, [pair(A,B)|T], V) :- assoc(X, T, V).
-
-nand(false, false, true).
-nand(false, true,  true).
-nand(true,  false, true).
-nand(true,  true,  false).
-
-neg(X, R) :- nand(X, X, R).
-
-or(X, Y, R) :- nand(X, X, Xx), nand(Y, Y, Yy), nand(Xx, Yy, R).
-
-and(X, Y, R) :- nand(X, Y, Xy), nand(Xy, Xy, R).
-
-
-/* Default Conjunctive */
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 580 ms */
-/* Unfolding time: 480 ms */
-/* Post-Processing time: 580 ms */
-
-/* Specialised Predicates:
-eval__1(A,B) :- eval(A,B,true).
-assoc__2(A,B,C,D,E) :- assoc(A,[pair(B,C)|D],E).
-eval_conj__3(A,B,C) :- eval(A,B,D1), eval(A,C,E1), nand(D1,D1,F1), nand(E1,E1,G1), nand(F1,G1,true).
-eval__4(A,B,C) :- eval(A,B,C).
-eval_conj__5(A,B,C,D,E) :- eval(A,B,neg(C)), eval(A,D,E).
-nand__6(A,B) :- nand(A,A,B).
-nand_conj__7(A,B) :- nand(A,A,C1), nand(B,C1,true).
-nand__8(A) :- nand(A,false,true).
-eval__9(A,B,C,D,E) :- eval([pair(A,B)|C],D,E).
-eval__10(A,B,C) :- eval(A,B,neg(C)).
[…]

### 2020-09-29 papers/miniKanren/ecce/prop/first.pl

@@ -1,660 +1,0 @@
-eval(St, var(X), U) :- elem(X, St, U).
-eval(St, conj(X,Y), U) :- and(V, W, U), eval(St, X, V), eval(St, Y, W).
-eval(St, disj(X,Y), U) :- or(V, W, U), eval(St, X, V), eval(St, Y, W).
-eval(St, not(X), U) :- eval(St, X, neg(U)).
-
-elem(zero, [H|T], H).
-elem(succ(N), [H|T], V) :- elem(N, T, V).
-
-nand(false, false, true).
-nand(false, true,  true).
-nand(true,  false, true).
-nand(true,  true,  false).
-
-neg(X, R) :- nand(X, X, R).
-
-or(X, Y, R) :- nand(X, X, Xx), nand(Y, Y, Yy), nand(Xx, Yy, R).
-
-and(X, Y, R) :- nand(X, Y, Xy), nand(Xy, Xy, R).
-
-/* Default conjunctive */
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 500 ms */
-/* Unfolding time: 340 ms */
-/* Post-Processing time: 500 ms */
-
-/* Specialised Predicates:
-eval__1(A,B) :- eval(A,B,true).
-nand_conj__2(A,B) :- nand(A,A,C1), nand(B,C1,true).
-eval_conj__3(A,B,C,D,E) :- eval(A,B,C), eval(A,D,E).
-eval__4(A,B,C) :- eval(A,B,neg(C)).
-elem__5(A,B,C,D) :- elem(A,[B|C],D).
-eval__6(A,B,C) :- eval(A,B,C).
-eval_conj__7(A,B,C,D,E) :- eval(A,B,neg(C)), eval(A,D,E).
-eval__8(A,B,C,D) :- eval([A|B],C,D).
-nand_conj__9(A,B,C,D,E) :- nand(F1,G1,H1), nand(H1,H1,A), eval([B|C],D,F1), eval([B|C],E,G1).
-nand__10(A,B) :- nand(A,A,B).
[…]

### 2020-09-29 papers/miniKanren/ecce/prop/last.pl

@@ -1,970 +1,0 @@
-eval(St, var(X), U) :- elem(X, St, U).
-eval(St, conj(X,Y), U)  :- eval(St, X, V), eval(St, Y, W), and(V, W, U).
-eval(St, disj(X,Y), U)  :- eval(St, X, V), eval(St, Y, W), or(V, W, U).
-eval(St, not(X), U) :- eval(St, X, V), neg(U, V).
-
-elem(zero, [H|T], H).
-elem(succ(N), [H|T], V) :- elem(N, T, V).
-
-nand(false, false, true).
-nand(false, true,  true).
-nand(true,  false, true).
-nand(true,  true,  false).
-
-neg(X, R) :- nand(X, X, R).
-
-or(X, Y, R) :- nand(X, X, Xx), nand(Y, Y, Yy), nand(Xx, Yy, R).
-
-and(X, Y, R) :- nand(X, Y, Xy), nand(Xy, Xy, R).
-
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 2480 ms */
-/* Unfolding time: 2020 ms */
-/* Post-Processing time: 2480 ms */
-
-/* Specialised Predicates:
-eval__1(A,B) :- eval(A,B,true).
-elem__2(A,B,C,D) :- elem(A,[B|C],D).
-eval_conj__3(A,B,C) :- eval(A,B,D1), eval(A,C,E1), nand(D1,D1,F1), nand(E1,E1,G1), nand(F1,G1,true).
-eval__4(A,B) :- eval(A,B,false).
-eval_conj__5(A,B,C,D,E) :- eval(A,B,C), eval(A,D,F1), nand(E,F1,true).
-eval__6(A,B,C) :- eval(A,B,C).
-nand__7(A,B,C) :- nand(A,B,C).
-nand__8(A,B) :- nand(A,A,B).
-nand_conj__9(A,B,C) :- nand(A,A,D1), nand(B,D1,C).
-nand__10(A,B) :- nand(A,true,B).
[…]

### 2020-09-29 papers/miniKanren/ecce/prop/simple.pl

@@ -1,319 +1,0 @@
-eval(St, var(X), R) :- elem(X, St, R).
-eval(St, conj(X,Y), true)  :- eval(St, X, true), eval(St, Y, true).
-eval(St, conj(X,Y), false) :- eval(St, X, false).
-eval(St, conj(X,Y), false) :- eval(St, Y, false).
-eval(St, disj(X,Y), false) :- eval(St, X, false), eval(St, Y, false).
-eval(St, disj(X,Y), true)  :- eval(St, X, true).
-eval(St, disj(X,Y), true)  :- eval(St, Y, true).
-eval(St, not(X), true) :- eval(St, X, false).
-eval(St, not(X), false) :- eval(St, X, true).
-
-elem(zero, [H|T], H).
-elem(succ(N), [H|T], V) :- elem(N, T, V).
-
-
-
-/* Specialised Predicates:
-eval__1(A,B) :- eval(A,B,true).
-elem__2(A,B,C,D) :- elem(A,[B|C],D).
-eval__3(A,B) :- eval(A,B,false).
-*/
-
-eval(A,B,true) :-
-  eval__1(A,B).
-eval__1([A|B],var(C)) :-
-  elem__2(C,A,B,true).
-eval__1(A,conj(B,C)) :-
-  eval__1(A,B),
-  eval__1(A,C).
-eval__1(A,disj(B,C)) :-
-  eval__1(A,B).
-eval__1(A,disj(B,C)) :-
-  eval__1(A,C).
-eval__1(A,not(B)) :-
-  eval__3(A,B).
-
-elem__2(zero,A,B,A).
-elem__2(succ(A),B,[C|D],E) :-
-  elem__2(A,C,D,E).
-
-/* Лишний анфолдинг (с другой стороны, у нас тоже) */
-eval__3([A|B],var(C)) :-
-  elem__2(C,A,B,false).
-eval__3(A,conj(B,C)) :-
-  eval__3(A,B).
-eval__3(A,conj(B,C)) :-
-  eval__3(A,C).
-eval__3(A,disj(B,C)) :-
-  eval__3(A,B),
-  eval__3(A,C).
-eval__3(A,not(B)) :-
-  eval__1(A,B).
-
-/* Conjunctive fast */
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
[…]

### 2020-09-29 papers/miniKanren/ocanren/bottles/clean.sh

@@ -1,1 +1,0 @@
-rm -f *.cmi *.cmx *.o *.cmo run

### 2020-09-29 papers/miniKanren/ocanren/bottles/ecce.pl

@@ -1,126 +1,0 @@
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 1300 ms */
-/* Unfolding time: 920 ms */
-/* Post-Processing time: 1300 ms */
-
-/* Specialised Predicates: 
-checkAnswer__1(A,B,C) :- checkAnswer(A,B,C,true).
-checkStep_conj__2(A,B,C,D,E) :- checkStep(pair(o,o),pair(A,B),C,true), doStep(pair(o,o),pair(A,B),C,F1), checkAnswer_0(F1,D,C,E,true).
-checkAnswer_0__3(A,B,C,D) :- checkAnswer_0(pair(o,o),A,pair(B,C),D,true).
-checkAnswer_0__4(A,B,C,D,E,F) :- checkAnswer_0(pair(A,B),C,pair(D,E),F,true).
-isFinishState__5(A,B,C) :- isFinishState(pair(A,B),C,true).
-checkStep_conj__6(A,B,C,D,E,F,G,H) :- checkStep(pair(A,B),pair(C,D),pair(E,F),true), doStep(pair(A,B),pair(C,D),pair(E,F),I1), checkAnswer_0(I1,G,pair(E,F),H,true).
-fancyEq__7(A,B) :- fancyEq(A,B,false).
-doStep_conj__8(A,B,C,D,E,F) :- doStep(pair(A,s(B)),pair(pour,snd),pair(C,D),G1), checkAnswer_0(G1,E,pair(C,D),F,true).
-add_conj__9(A,B,C,D,E,F) :- add(A,B,C), greater(D,E,false), add(A,B,F).
-greater__10(A,B) :- greater(A,B,false).
-greater__11(A,B) :- greater(s(A),B,true).
-add__12(A,B,C) :- add(A,B,C).
-sub_conj__13(A,B,C,D,E,F,G,H,I) :- sub(s(A),B,C), checkAnswer_0(pair(D,E),F,pair(G,H),I,true).
[…]

### 2020-09-29 papers/miniKanren/ocanren/desert/build.sh

@@ -1,2 +1,0 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg tester.ml lorry.ml lorry_run.ml
-

### 2020-09-29 papers/miniKanren/ocanren/desert/ecce.pl

@@ -1,343 +1,0 @@
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 16480 ms */
-/* Unfolding time: 11420 ms */
-/* Post-Processing time: 16480 ms */
-
-/* Specialised Predicates:
-checkAnswer__1(A,B,C,D) :- checkAnswer(A,B,C,some(D)).
-stations__3(A,B) :- stations(A,B).
-goe_conj__5(A,B,C,D,E) :- goe(A,B,true), goe(C,B,true), sub(D,B,E).
-calcFuel__6(A,B,C,D,E,F,G,H) :- calcFuel(st(A,B,cons(C,D)),E,s(F),s(G),true,some(H)).
-isMove_conj__7(A,B,C,D,E,F,G,H,I) :- isMove(A,false), checkStep(A,st(B,C,cons(D,E)),s(F),s(G),true), step(A,st(B,C,cons(D,E)),s(F),s(G),st(J1,K1,L1)), calcFuel(st(J1,K1,L1),H,s(F),s(G),false,some(M1)), getFuel(A,st(B,C,cons(D,E)),s(G),N1), add(N1,M1,I).
-checkStep_conj__8(A,B,C,D,E,F,G,H) :- checkStep(fill,st(A,B,cons(C,D)),s(E),s(F),true), step(fill,st(A,B,cons(C,D)),s(E),s(F),st(A,I1,J1)), calcFuel(st(A,I1,J1),G,s(E),s(F),false,some(K1)), getFuel(fill,st(A,B,cons(C,D)),s(F),L1), add(L1,K1,H).
-goe_conj__9(A,B,C) :- goe(A,B,true), sub(A,B,C).
-addForElem__10(A,B,C,D,E,F) :- addForElem(cons(A,B),C,s(D),cons(E,F)).
[…]

### 2020-09-29 papers/miniKanren/ocanren/len/build.sh

@@ -1,2 +1,0 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml len.ml lenBack.ml run.ml
-

### 2020-09-29 papers/miniKanren/ocanren/llang/ecce.pl

@@ -1,110 +1,0 @@
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 11820 ms */
-/* Unfolding time: 8720 ms */
-/* Post-Processing time: 11820 ms */
-
-/* Specialised Predicates:
-type___1(A,B) :- type_(A,B,some(integer_)).
-type__conj__2(A,B,C,D,E) :- type_(A,B,some(boolean_)), type_(C,B,some(D)), type_(E,B,some(D)).
-type___3(A,B,C) :- type_(A,B,some(C)).
-idx__4(A,B,C,D) :- idx(A,cons(B,C),some(D)).
-type__conj__5(A,B,C,D,E) :- type_(A,B,some(C)), type_(D,cons(E,B),some(integer_)).
-*/
-
-type_(A,B,some(integer_)) :-
-    type___1(A,B).
-type___1(iConst_(A),B).
-type___1(var_(A),cons(B,C)) :-
-    idx__4(A,B,C,integer_).
-type___1(plus_(A,B),C) :-
-    type___1(A,C),
-    type___1(B,C).
-type___1(mult_(A,B),C) :-
-    type___1(A,C),
-    type___1(B,C).
-type___1(let_(A,B),C) :-
-    type__conj__5(A,C,D,B,D).
-type___1(if_(A,B,C),D) :-
-    type__conj__2(A,D,B,integer_,C).
-type__conj__2(bConst_(A),B,C,D,E) :-
-    type___3(C,B,D),
-    type___3(E,B,D).
-type__conj__2(var_(A),cons(B,C),D,E,F) :-
-    idx__4(A,B,C,boolean_),
-    type___3(D,cons(B,C),E),
-    type___3(F,cons(B,C),E).
-type__conj__2(equal_(A,B),C,D,E,F) :-
-    type___3(A,C,G),
-    type___3(B,C,G),
-    type___3(D,C,E),
-    type___3(F,C,E).
-type__conj__2(less_(A,B),C,D,E,F) :-
-    type___1(A,C),
-    type___1(B,C),
-    type___3(D,C,E),
-    type___3(F,C,E).
[…]

### 2020-09-29 papers/miniKanren/ocanren/llang/llang.pl

@@ -1,32 +1,0 @@
-assoco(cons(pair(Y3, integer), Q1), Y3).
-assoco(cons(pair(Q2, Q3), Q1), Y3) :- assoco(Q1, Y3).
-_type(iConst(Q1), Y5).
-_type(var_(Q2), cons(pair(Q2, integer), Q3)).
-_type(var_(Q2), cons(pair(Q4, Q5), Q3)) :- assoco(Q3, Q2).
-_type(plus(Q6, Q7), Y5) :- typeType(Y5, Q6, Q7).
-_type(mult(Q6, Q7), Y5) :- typeType(Y5, Q6, Q7).
-_type(let_(Q8, Q9, Q10), Y5) :- __type(Y5, Q9, Q11), __type(cons(pair(Q8, Q11), Y5), Q10, integer).
-_type(if_(Q12, Q13, Q14), Y5) :- ___type(Y5, Q12), _type(Q13, Y5), _type(Q14, Y5).
-typeType(Y6, Y7, Y8) :- _type(Y7, Y6), _type(Y8, Y6).
-__type(Y9, bConst(Q1), boolean).
-__type(Y9, iConst(Q2), integer).
-__type(cons(pair(Q3, Y11), Q4), var_(Q3), Y11).
-__type(cons(pair(Q5, Q6), Q4), var_(Q3), Y11) :- _assoco(Q4, Y11, Q3).
-__type(Y9, plus(Q7, Q8), integer) :- typeType(Y9, Q7, Q8).
-__type(Y9, mult(Q7, Q8), integer) :- typeType(Y9, Q7, Q8).
-__type(Y9, eq(Q9, Q10), boolean) :- _typeType(Y9, Q9, Q10).
-__type(Y9, lt(Q9, Q10), boolean) :- _typeType(Y9, Q9, Q10).
-__type(Y9, let_(Q11, Q12, Q13), Y11) :- __type(Y9, Q12, Q14), __type(cons(pair(Q11, Q14), Y9), Q13, Y11).
-__type(Y9, if_(Q15, Q16, Q17), Y11) :- ___type(Y9, Q15), __type(Y9, Q16, Y11), __type(Y9, Q17, Y11).
-_assoco(cons(pair(Y14, Y13), Q1), Y13, Y14).
-_assoco(cons(pair(Q2, Q3), Q1), Y13, Y14) :- _assoco(Q1, Y13, Y14).
[…]

### 2020-09-29 papers/miniKanren/ocanren/llang/original.pl

@@ -1,11 +1,0 @@
-type(bConst(X), Gamma, boolean).
-type(iConst(Y), Gamma, integer).
-type(var_(V), Gamma, Ttype) :- assoco(V, Gamma, Ttype).
-type(plus(M, N), Gamma, integer) :- type(M, Gamma, integer), type(N, Gamma, integer).
-type(mult(M, N), Gamma, integer) :- type(M, Gamma, integer), type(N, Gamma, integer).
-type(eq(L, R), Gamma, boolean) :- type(L, Gamma, T), type(R, Gamma, T).
-type(lt(L, R), Gamma, boolean) :- type(L, Gamma, T), type(R, Gamma, T).
-type(let_(Var, Bound, Body), Gamma, Ttype) :- type(Bound, Gamma, Btype), type(Body, cons(pair(Var, Btype), Gamma), Ttype).
-type(if_(Cond, Thn, Els), Gamma, Ttype) :- type(Cond, Gamma, boolean), type(Thn, Gamma, Ttype), type(Els, Gamma, Ttype).
-assoco(X, cons(pair(X, V), Tl), V).
-assoco(X, cons(pair(A, B), Tl), V) :- assoco(X, Tl, V).

### 2020-09-29 papers/miniKanren/ocanren/llang2/ecce.pl

@@ -1,289 +1,0 @@
-
-/* Specialised program generated by ECCE 2.0 */
-/* PD Goal: A */
-/* Parameters: Abs:l InstCheck:v Msv:s NgSlv:g Part:e Prun:n Sel:t Whstl:f Raf:yesFar:yes Dce:yes Poly:y Dpu:yes ParAbs:yes Msvp:no Rrc:yes */
-/* Transformation time: 5560 ms */
-/* Unfolding time: 4340 ms */
-/* Post-Processing time: 5560 ms */
-
-/* Specialised Predicates:
-typecheck___1(A,B) :- typecheck_(A,B,some(integer)).
-nthOpt__2(A,B,C,D) :- nthOpt(cons(A,B),C,some(D)).
-typecheck__conj__3(A,B,C,D,E) :- typecheck_(A,B,some(C)), typecheck_(cons(D,A),E,some(integer)).
-nthOpt_conj__4(A,B,C,D,E,F) :- nthOpt(cons(A,B),C,some(D)), typecheck_(cons(E,cons(A,B)),F,some(integer)).
-typecheck___5(A,B,C) :- typecheck_(A,B,some(C)).
-typecheck__conj__6(A,B,C,D,E) :- typecheck_(A,B,some(boolean)), typecheck_(A,C,some(D)), typecheck_(A,E,some(D)), typeEq(D,D,true).
-nthOpt_conj__7(A,B,C,D,E,F) :- nthOpt(cons(A,B),C,some(boolean)), typecheck_(cons(A,B),D,some(E)), typecheck_(cons(A,B),F,some(E)), typeEq(E,E,true).
-typecheck__conj__8(A,B,C,D) :- typecheck_(A,B,some(boolean)), typecheck_(A,C,some(boolean)), typecheck_(A,D,some(boolean)).
-typeEq__9(A) :- typeEq(A,A,true).
-typecheck___10(A,B,C) :- typecheck_(cons(A,B),C,some(boolean)).
-nthOpt__11(A) :- nthOpt(cons(B1,A),s(C1),some(D1)).
-nthOpt_conj__12(A,B,C,D,E) :- nthOpt(cons(A,B),C,some(boolean)), typecheck_(cons(A,B),D,some(boolean)), typecheck_(cons(A,B),E,some(boolean)).
[…]

### 2020-09-29 papers/miniKanren/ocanren/maxlen/build.sh

@@ -1,2 +1,0 @@
-ocamlfind ocamlopt -rectypes -o run -package benchmark -package ocanren -package GT -syntax camlp5o -package GT.syntax.all -package ocanren.syntax -linkpkg helper.ml original.ml maxlen.ml ideal.ml ideal2.ml len.ml ecce.ml ecce_right.ml run.ml
-

### 2020-09-29 papers/miniKanren/ocanren/type/type.pl

@@ -1,15 +1,0 @@
-elemo(cons(arrow(Y5, Y6), Q1), Y5, Y6, o).
-elemo(cons(Q3, Q1), Y5, Y6, s(Q2)) :- elemo(Q1, Y5, Y6, Q2).
-_type(var(o), cons(arrow(Y10, Y11), Q2), Y10, Y11).
-_type(var(s(Q3)), cons(Q4, Q2), Y10, Y11) :- elemo(Q2, Y10, Y11, Q3).
-_type(abs(Q1, Q5, Q6), Y9, Y10, Y11) :- _type(Q6, cons(pair(Q1, Q5), Y9), Y10, Y11).
-_type(app(Q7, Q8), Y9, Y10, Y11) :- _type(Q7, Y9, Q9, Q10), __type(Y9, Q8, Q10).
-__type(Y12, bConst(Q1), boolean).
-__type(Y12, iConst(Q2), integer).
-__type(cons(Y14, Q4), var(o), Y14).
-__type(cons(Q6, Q4), var(s(Q5)), Y14) :- _elemo(Q4, Y14, Q5).
-__type(Y12, abs(Q3, Q7, Q8), Y14) :- __type(cons(pair(Q3, Q7), Y12), Q8, Y14).
-__type(Y12, app(Q9, Q10), Y14) :- typeType(Y12, Q9, Q10).
-_elemo(cons(Y16, Q1), Y16, o).
-_elemo(cons(Q3, Q1), Y16, s(Q2)) :- _elemo(Q1, Y16, Q2).
-typeType(Y18, Y19, Y20) :- _type(Y19, Y18, Q1, Q2), __type(Y18, Y20, Q2).

### 2020-09-29 papers/miniKanren/ocanren/unify/original.pl

@@ -1,48 +1,0 @@
-eq_nat(z, z, true).
-eq_nat(s(Q39), z, false).
-eq_nat(z, s(Q41), false).
-eq_nat(s(X), s(Y), Q36) :- eq_nat(X, Y, Q36).
-get_term(Var, nil, none).
-get_term(z, cons(Q32, Xs), Q32).
-get_term(s(N), cons(X, Xs), Q32) :- get_term(N, Xs, Q32).
-check_uni(Subst, constr(N1, A1), constr(N2, A2), false) :- eq_nat(N1, N2, false), forall2(Subst, A1, A2, Q13).
-check_uni(Subst, constr(N1, A1), constr(N2, A2), Q13) :- eq_nat(N1, N2, true), forall2(Subst, A1, A2, Q13).
-check_uni(Subst, var_(V), constr(N, A), false) :- get_term(V, Subst, none).
-check_uni(Subst, var_(V), constr(N, A), Q31) :- get_term(V, Subst, some(T)), check_uni(Subst, T, constr(N, A), Q31).
-check_uni(Subst, constr(N, A), var_(V), false) :- get_term(V, Subst, none).
-check_uni(Subst, constr(N, A), var_(V), Q31) :- get_term(V, Subst, some(T)), check_uni(Subst, constr(N, A), T, Q31).
-check_uni(Subst, var_(V1), var_(V2), Q31) :- get_term(V1, Subst, some(T19)), check_uni(Subst, T19, var_(V2), Q31).
-check_uni(Subst, var_(V1), var_(V2), false) :- get_term(V1, Subst, none), get_term(V2, Subst, some(Q28)).
-check_uni(Subst, var_(V1), var_(V2), Q31) :- get_term(V1, Subst, none), get_term(V2, Subst, none), eq_nat(V1, V2, Q31).
-get_term(Var, nil, none).
-get_term(z, cons(Q32, Xs), Q32).
-get_term(s(N), cons(X, Xs), Q32) :- get_term(N, Xs, Q32).
-eq_nat(z, z, true).
-eq_nat(s(Q39), z, false).
-eq_nat(z, s(Q41), false).
-eq_nat(s(X), s(Y), Q36) :- eq_nat(X, Y, Q36).
-forall2(Subst, nil, nil, true).
-forall2(Subst, cons(X, Xs), cons(Y, Ys), false) :- check_uni(Subst, X, Y, false), forall2(Subst, Xs, Ys, Q4).
[…]

### 2020-09-29 papers/miniKanren/prolog/app/app10.pl

@@ -1,14 +1,0 @@
-appendo(nil, Y2, Y2).
-appendo(cons(Q1, nil), Q3, cons(Q1, Q3)).
-appendo(cons(Q1, cons(Q4, nil)), Q6, cons(Q1, cons(Q4, Q6))).
-appendo(cons(Q1, cons(Q4, cons(Q7, nil))), Q9, cons(Q1, cons(Q4, cons(Q7, Q9)))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, nil)))), Q12, cons(Q1, cons(Q4, cons(Q7, cons(Q10, Q12))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, nil))))), Q15, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, Q15)))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, nil)))))), Q18, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, Q18))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, nil))))))), Q21, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, Q21)))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, nil)))))))), Q24, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, Q24))))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, cons(Q25, nil))))))))), Q27, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, cons(Q25, Q27)))))))))).
[…]

### 2020-09-29 papers/miniKanren/prolog/app/app11.pl

@@ -1,15 +1,0 @@
-appendo(nil, Y2, Y2).
-appendo(cons(Q1, nil), Q3, cons(Q1, Q3)).
-appendo(cons(Q1, cons(Q4, nil)), Q6, cons(Q1, cons(Q4, Q6))).
-appendo(cons(Q1, cons(Q4, cons(Q7, nil))), Q9, cons(Q1, cons(Q4, cons(Q7, Q9)))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, nil)))), Q12, cons(Q1, cons(Q4, cons(Q7, cons(Q10, Q12))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, nil))))), Q15, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, Q15)))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, nil)))))), Q18, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, Q18))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, nil))))))), Q21, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, Q21)))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, nil)))))))), Q24, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, Q24))))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, cons(Q25, nil))))))))), Q27, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, cons(Q25, Q27)))))))))).
[…]

### 2020-09-29 papers/miniKanren/prolog/app/app15.pl

@@ -1,19 +1,0 @@
-appendo(nil, Y2, Y2).
-appendo(cons(Q1, nil), Q3, cons(Q1, Q3)).
-appendo(cons(Q1, cons(Q4, nil)), Q6, cons(Q1, cons(Q4, Q6))).
-appendo(cons(Q1, cons(Q4, cons(Q7, nil))), Q9, cons(Q1, cons(Q4, cons(Q7, Q9)))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, nil)))), Q12, cons(Q1, cons(Q4, cons(Q7, cons(Q10, Q12))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, nil))))), Q15, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, Q15)))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, nil)))))), Q18, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, Q18))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, nil))))))), Q21, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, Q21)))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, nil)))))))), Q24, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, Q24))))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, cons(Q25, nil))))))))), Q27, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, cons(Q22, cons(Q25, Q27)))))))))).
[…]

### 2020-09-29 papers/miniKanren/prolog/app/app4.pl

@@ -1,8 +1,0 @@
-appendo(nil, Y2, Y2).
-appendo(cons(Q1, nil), Q3, cons(Q1, Q3)).
-appendo(cons(Q1, cons(Q4, nil)), Q6, cons(Q1, cons(Q4, Q6))).
-appendo(cons(Q1, cons(Q4, cons(Q7, nil))), Q9, cons(Q1, cons(Q4, cons(Q7, Q9)))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, nil)))), Q12, cons(Q1, cons(Q4, cons(Q7, cons(Q10, Q12))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, Q14))))), Y1, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, Q15)))))) :- _appendo(Y1, Q14, Q15).
-_appendo(Y5, nil, Y5).
-_appendo(Y3, cons(Q1, Q2), cons(Q1, Q3)) :- _appendo(Y3, Q2, Q3).

### 2020-09-29 papers/miniKanren/prolog/app/app6.pl

@@ -1,10 +1,0 @@
-appendo(nil, Y2, Y2).
-appendo(cons(Q1, nil), Q3, cons(Q1, Q3)).
-appendo(cons(Q1, cons(Q4, nil)), Q6, cons(Q1, cons(Q4, Q6))).
-appendo(cons(Q1, cons(Q4, cons(Q7, nil))), Q9, cons(Q1, cons(Q4, cons(Q7, Q9)))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, nil)))), Q12, cons(Q1, cons(Q4, cons(Q7, cons(Q10, Q12))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, nil))))), Q15, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, Q15)))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, nil)))))), Q18, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, Q18))))))).
-appendo(cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, Q20))))))), Y1, cons(Q1, cons(Q4, cons(Q7, cons(Q10, cons(Q13, cons(Q16, cons(Q19, Q21)))))))) :- _appendo(Y1, Q20, Q21).
-_appendo(Y5, nil, Y5).
-_appendo(Y3, cons(Q1, Q2), cons(Q1, Q3)) :- _appendo(Y3, Q2, Q3).

### 2020-09-29 papers/miniKanren/prolog/len/len.pl

@@ -1,40 +1,0 @@
-% This is prolog!
-
-len([],o). 
-len([_|T], s(R)) :- len(T,R).
-
-len1([],o).
-len1([_|T], S) :- len1(T,R), unify_with_occurs_check(S, s(R)).
-
-len2([],o).
-len2([_|T], S) :- unify_with_occurs_check(S, s(R)), len2(T,R).
-
-run(Fst, Snd, Trd, Lst) :-
-  test(Fst,Lst),
-  test(Snd,Lst),
-  test(Trd,Lst). 
-
-go :-
-  Lst = [0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0
[…]
